---
title: "R Notebook"
output: html_notebook
---

```{r}
library(Seurat)
library(Matrix)
library(rhdf5)
library(dplyr)
library(qs)
library(ggplot2)
```

```{r}
filtered_merged_caudate = qread("Cellbender_seurat/filtered_merged_caudate_clustered.qs")
filtered_merged_putamen = qread("Cellbender_seurat/filtered_merged_putamen_clustered.qs")
filtered_merged_caudate
filtered_merged_putamen
```

```{r}
head(filtered_merged_caudate)
head(caudate_sobj)
```


```{r}
caudate_sobj = qread("Cellbender_seurat/Bennet_caudate_clean.qs")
putamen_sobj = qread("Cellbender_seurat/Bennet_putamen_clean.qs")

caudate_sobj
putamen_sobj
```
```{r}
DimPlot(putamen_sobj, group.by = "Condition")
```


```{r}
DimPlot(caudate_sobj, group.by = "RNA_snn_res.0.5")
```




#subset just neurons --> get 5581 neurons
```{r}
caudate_neurons = subset(caudate_sobj, subset = cell_class == "neuron")
caudate_neurons
head(caudate_neurons)
```
```{r}
DimPlot(caudate_neurons, group.by = "library")
DimPlot(caudate_neurons, group.by = "donor_id")
DimPlot(caudate_neurons, group.by = "Condition")

DimPlot(caudate_neurons, group.by = "RNA_snn_res.0.2")
DimPlot(caudate_neurons, group.by = "RNA_snn_res.0.3")
DimPlot(caudate_neurons, group.by = "RNA_snn_res.0.4")
DimPlot(caudate_neurons, group.by = "RNA_snn_res.0.5")
```
#Putamen has 12518 neurons
```{r}
putamen_neurons = subset(putamen_sobj, subset = cell_class == "neuron")
putamen_neurons
head(putamen_neurons)

DimPlot(putamen_neurons, group.by = "library")
DimPlot(putamen_neurons, group.by = "donor_id")
DimPlot(putamen_neurons, group.by = "Condition")

DimPlot(putamen_neurons, group.by = "RNA_snn_res.0.2")
DimPlot(putamen_neurons, group.by = "RNA_snn_res.0.3")
DimPlot(putamen_neurons, group.by = "RNA_snn_res.0.4")
DimPlot(putamen_neurons, group.by = "RNA_snn_res.0.5")
```


```{r}
library(harmony)
putamen_neurons = normalizeScalePcaClusterUmap(putamen_neurons)
putamen_neurons = (putamen_neurons
    %>% RunHarmony(
        group.by = "donor_id")
    %>% FindNeighbors(reduction='harmony', dims=1:20)
    %>% FindClusters(res=0.2)
    %>% FindClusters(res=0.3)
    %>% FindClusters(res=0.4)
    %>% FindClusters(res=0.5)
    %>% RunUMAP(reduction="harmony", dims=1:20)
) 
```


```{r}
qsave(putamen_neurons, "Cellbender_seurat/putamen_neurons.qs")
```

```{r}
DimPlot(putamen_neurons, group.by = "library")
DimPlot(putamen_neurons, group.by = "donor_id")
DimPlot(putamen_neurons, group.by = "Condition")

DimPlot(putamen_neurons, group.by = "RNA_snn_res.0.2")
DimPlot(putamen_neurons, group.by = "RNA_snn_res.0.3")
DimPlot(putamen_neurons, group.by = "RNA_snn_res.0.4")
DimPlot(putamen_neurons, group.by = "RNA_snn_res.0.5")
```

```{r}
Idents(putamen_neurons) = putamen_neurons$RNA_snn_res.0.2
markers = FindAllMarkers(putamen_neurons, only.pos = TRUE, min.pct = 0.2, logfc.threshold = 0.5)

num_clusters = length(unique(Idents(putamen_neurons))) - 1

for (i in 0:num_clusters){
   printMarkersByCluster(markers, cluster=i)
}
```


```{r}
classes = c("D2 SPN_1", "D1 SPN_1", "eSPN", "glutamatergic", "D1 SPN_2", "interneuron_1", "interneuron_2", "interneuron_3", "D2 SPN_2", "D1 SPN_3?", "interneuron_4", "interneuron_5", "cholinergic", "interneuron_6")

#classes = c("SPN", "SPN", "SPN", "glutamatergic", "SPN", "interneuron", "interneuron", "interneuron", "SPN", "SPN", "interneuron", "interneuron", "cholinergic", "interneuron")

putamen_neurons= assignCellClasses(putamen_neurons, classes=classes, cluster_col="RNA_snn_res.0.2", class_col = "cell_class")

Idents(putamen_neurons) <- "cell_class"

DimPlot(putamen_neurons, label= TRUE)
```
```{r}
FeaturePlot(putamen_neurons, features = c("DRD1", "DRD2"))
FeaturePlot(putamen_neurons, features = c("EPHA4", "SEMA3E"))
FeaturePlot(putamen_neurons, features =c("CASZ1"))
FeaturePlot(putamen_neurons, features =c("CHAT"))
FeaturePlot(putamen_neurons, features =c("C1QA"))
FeaturePlot(putamen_neurons, features = c("PPP1R1B", "MT-CO1"))
```
```{r}
FeaturePlot(putamen_neurons, features = c("GAD1", "GAD2"))
FeaturePlot(putamen_neurons, features = c("VIP", "CALB2"))
FeaturePlot(putamen_neurons, features = c("SLC17A7", "SLC17A6"))

```


```{r}
table(putamen_neurons$donor_id)

```


```{r}
donor_list = unique(putamen_neurons$donor_id)
neurons_per_donor = as.data.frame(table(putamen_neurons$donor_id))

classes = c("D2 SPN_1", "D1 SPN_1", "eSPN", "glutamatergic", "D1 SPN_2", "interneuron_1", "interneuron_2", "interneuron_3", "D2 SPN_2", "D1 SPN_3?", "interneuron_4", "interneuron_5", "cholinergic", "interneuron_6")

putamen_neurons= assignCellClasses(putamen_neurons, classes=classes, cluster_col="RNA_snn_res.0.2", class_col = "cell_class")
Idents(putamen_neurons) <- "cell_class"


for (donor in donor_list) {
  single_donor_neurons = subset(putamen_neurons, subset = donor_id == donor)
  neuron_count = neurons_per_donor$Freq[neurons_per_donor$Var1 == donor]
  print(DimPlot(single_donor_neurons,  label = TRUE)+ ggtitle(paste(donor, "Number of neurons: ", neuron_count)))
}

```










```{r}
normalizeScalePcaClusterUmap = function(
    sobj,
    subset_col="donor_id",
    n_hvgs_orig=2500,
    n_dims_use=20,
    resolutions=c(0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1)){
    # this function takes a Seurat object, normalizes the data, scales the data, runs PCA, finds neighbors, clusters, and runs UMAP
    # it then returns the Seurat object
    hvgs = getSeuratVarFeatureIntersectByCol(sobj, subset_col=subset_col, original_nfeatures=n_hvgs_orig)
    sobj = (sobj
        %>% NormalizeData() # log normalizes raw counts
        %>% ScaleData(features=hvgs, split.by=subset_col) # within each cell, for each gene scale the data (i.e. subtract the donor's mean and divide by the donor's standard deviation of that gene)
        %>% RunPCA(features=hvgs, npcs=n_dims_use) # Reduce the dimensions to the n_dims_use dimensions that best explain the data https://en.wikipedia.org/wiki/Principal_component_analysis
        %>% FindNeighbors(dims = 1:n_dims_use) # Finds every cells closest neighbors in the PCA space
    )
    for(res in resolutions){
        sobj = sobj %>% FindClusters(resolution = res) # finds clusters at a variety of resolutions
    }
    sobj = sobj %>% RunUMAP(dims = 1:n_dims_use) # projects the PCA onto 2 dimensions, such that both "global structure" and "local structure" are optimally preserved
    return(sobj)
}

getSeuratVarFeatureIntersectByCol = function(
   seurat_obj,
   subset_col,
   original_nfeatures,
   n_subsets_to_cover=NULL
){
  
   hvgs = list()


   unique_ids = unique(seurat_obj@meta.data[[subset_col]])


   if (is.null(n_subsets_to_cover)){
       n_subsets_to_cover = floor(length(unique_ids)/2)
   }


   i=1
   for (id in unique_ids) {
       print(paste("Subset", id, "--", i, "of", length(unique_ids)))
       i = i + 1
      
       seurat_subset = seurat_obj[, seurat_obj[[subset_col]] == id]
      
       if (ncol(seurat_subset) < 2){next}
      
       suppressWarnings({
           seurat_subset = FindVariableFeatures(
               seurat_subset, nfeatures=original_nfeatures, verbose = FALSE)
       })
      
       hvgs[[id]] = getSeuratVarFeatures(seurat_subset)
      
   }


   common_hvgs = getCommonStrings(hvgs, n_subsets_to_cover)
   print(paste("Number of HVGs in common across", n_subsets_to_cover, "--", length(common_hvgs)))


   return(common_hvgs)
}

getSeuratVarFeatures = function(sobj){
    # the magical incantation that returns the slot of the attribute of the slot that actually holds the list of variable feature -_-
    return(sobj@assays$RNA@var.features)
}

getCommonStrings <- function(list_of_lists, n) {
  # Create an empty list to store string occurrences
  string_occurrences <- list()

  # Iterate over each sublist
  for (sublist in list_of_lists) {
    # For each string in the sublist, increment its count in string_occurrences
    for (string in unique(sublist)) {
      if (!is.null(string_occurrences[[string]])) {
        string_occurrences[[string]] <- string_occurrences[[string]] + 1
      } else {
        string_occurrences[[string]] <- 1
      }
    }
  }

  # Filter the strings that occur at least n times
  common_strings <- names(string_occurrences)[string_occurrences >= n]

  return(common_strings)
}
```


```{r}
library(harmony)
caudate_neurons = normalizeScalePcaClusterUmap(caudate_neurons)
caudate_neurons = (caudate_neurons
    %>% RunHarmony(
        group.by = "donor_id")
    %>% FindNeighbors(reduction='harmony', dims=1:20)
    %>% FindClusters(res=0.2)
    %>% FindClusters(res=0.3)
    %>% FindClusters(res=0.4)
    %>% FindClusters(res=0.5)
    %>% RunUMAP(reduction="harmony", dims=1:20)
) 
```



```{r}
head(putamen_neurons)
```


```{r}
library(qs)
qsave(caudate_neurons, "Cellbender_seurat/caudate_neurons.qs")
```

```{r}
DimPlot(caudate_neurons, group.by = "RNA_snn_res.0.2")
DimPlot(caudate_neurons, group.by = "RNA_snn_res.0.3")
DimPlot(caudate_neurons, group.by = "RNA_snn_res.0.4")
DimPlot(caudate_neurons, group.by = "RNA_snn_res.0.5")

DimPlot(caudate_neurons, group.by = "library")
DimPlot(caudate_neurons, group.by = "donor_id")
DimPlot(caudate_neurons, group.by = "Condition")
```

```{r}
FeaturePlot(caudate_neurons, features = c("DRD1", "DRD2"))
FeaturePlot(caudate_neurons, features = c("EPHA4", "SEMA3E"))
```



```{r}
caudate_neurons_preintergrated = subset(caudate_sobj, subset = cell_class == "neuron")
caudate_neurons_preintergrated
head(caudate_neurons_preintergrated)
```


```{r}
DimPlot(caudate_neurons_preintergrated, group.by = "RNA_snn_res.0.2")
DimPlot(caudate_neurons_preintergrated, group.by = "RNA_snn_res.0.3")
DimPlot(caudate_neurons_preintergrated, group.by = "RNA_snn_res.0.4")
DimPlot(caudate_neurons_preintergrated, group.by = "RNA_snn_res.0.5")

DimPlot(caudate_neurons_preintergrated, group.by = "library")
DimPlot(caudate_neurons_preintergrated, group.by = "donor_id")
DimPlot(caudate_neurons_preintergrated, group.by = "Condition")
```

```{r}
FeaturePlot(caudate_neurons_preintergrated, features = c("DRD1", "DRD2"))
FeaturePlot(caudate_neurons_preintergrated, features = c("EPHA4", "SEMA3E"))

```

#for tmr: Do the rest of the featureplots for pct mito, log10numi, GAPDH, UBB, TUBB2A
#make forloop to do umaps for every donor + maybe even for just neuronal level too 
#try running Findallmarkers here? 


```{r}
head(caudate_neurons_preintergrated)
```

```{r}
caudate_neurons_preintergrated@meta.data$log10nUMI = log10(caudate_neurons_preintergrated@meta.data$nUmi)
caudate_neurons@meta.data$log10nUMI = log10(caudate_neurons@meta.data$nUmi)

head(caudate_neurons)
head(caudate_neurons_preintergrated)
```




```{r}
FeaturePlot(caudate_neurons_preintergrated, features = c("pct_mito"))
FeaturePlot(caudate_neurons_preintergrated, features = c("log10nUMI"))
FeaturePlot(caudate_neurons_preintergrated, features = c("GAPDH", "UBB", "TUBB2A"))

FeaturePlot(caudate_neurons, features = c("pct_mito"))
FeaturePlot(caudate_neurons, features = c("log10nUMI"))
FeaturePlot(caudate_neurons, features = c("GAPDH", "UBB", "TUBB2A"))
```

```{r}
DimPlot(caudate_neurons_preintergrated, label = TRUE)
```

```{r}
#FeaturePlot(caudate_neurons, features =c("RBFOX3", "C1QA", "CX3CR1", "CB69"))
#FeaturePlot(caudate_neurons, features =c("EPHA4", "SEMA3E", "CASZ1", "CHAT"))
#FeaturePlot(caudate_neurons, features =c("MBP"))
FeaturePlot(caudate_neurons, features =c("CASZ1"))
FeaturePlot(caudate_neurons, features =c("CHAT"))
FeaturePlot(caudate_neurons, features =c("C1QA"))
```


```{r}
FeaturePlot(caudate_neurons, features =c("C1QA", "C1QB"))
FeaturePlot(caudate_neurons, features =c("CX3CR1", "P2RY12"))
FeaturePlot(caudate_neurons, features =c("PPP1R1B"))

```


#Find all markers on the intergrated neuron cah
```{r}
#only.pos = False

Idents(caudate_neurons) = caudate_neurons$RNA_snn_res.0.2
markers = FindAllMarkers(caudate_neurons, only.pos = TRUE, min.pct = 0.2, logfc.threshold = 1.25)

num_clusters = length(unique(Idents(caudate_neurons))) - 1

for (i in 0:num_clusters){
   printMarkersByCluster(markers, cluster=i)
}
```

```{r}
#classes = c("0?", "1?", "2?", "3?", "GAD1-2,STY1", "GAD1", "GAD1/SYT1", "microglia???", "SST/SYNPR", "LAMP5", "10?", "SLC17A8, SYT1,GAD1", "GAD1/ambig", "GAD1/MOBP")


#classes = c("SPN", "SPN", "SPN", "SPN", "interneuron", "interneuron", "interneuron", "SPN", "interneuron", "SPN", "SPN", "cholinergic", "interneuron", "interneuron")

classes = c("D1 SPN_1", "D2_SPN_1", "D1 SPN_2?", "eSPN", "interneuron_1", "interneuron_2", "interneuron_3", "SPN?", "interneuron_4", "D1 SPN_3", "D2 SPN_2", "cholinergic", "interneuron_6", "interneuron_7")

caudate_neurons= assignCellClasses(caudate_neurons, classes=classes, cluster_col="RNA_snn_res.0.2", class_col = "cell_class")

Idents(caudate_neurons) <- "cell_class"

DimPlot(caudate_neurons, label= TRUE)
```

```{r}
FeaturePlot(caudate_neurons, features = c("pct_mito"))
FeaturePlot(caudate_neurons, features = c("PPP1R1B", "MT-CO1"))
FeaturePlot(caudate_neurons, features = c("DRD1", "DRD2"))
```

```{r}
FeaturePlot(caudate_neurons, features = c("PPP1R1B"))
FeaturePlot(caudate_neurons, features = c("SST"))
```



```{r}
"EPHA4", "SEMA3E", "CASZ1", "CHAT", "ACHE"

```



#make forloop to do umaps for every donor + maybe even for just neuronal level too 
#try running Findallmarkers here? 




```{r}
donor_list = unique(caudate_neurons$donor_id)
neurons_per_donor = as.data.frame(table(caudate_neurons$donor_id))

#classes = c(" ")

caudate_neurons= assignCellClasses(caudate_neurons, classes=classes, cluster_col="RNA_snn_res.0.2", class_col = "cell_class")
Idents(caudate_neurons) <- "cell_class"


for (donor in donor_list) {
  single_donor_neurons = subset(caudate_neurons, subset = donor_id == donor)
  neuron_count = neurons_per_donor$Freq[neurons_per_donor$Var1 == donor]
  print(DimPlot(single_donor_neurons,  label = TRUE)+ ggtitle(paste(donor, "Number of neurons: ", neuron_count)))
}

```
```{r}
head(filtered_merged_caudate)
```





```{r}
donor_list = unique(filtered_merged_caudate_sub$donor_id)
cells_per_donor = as.data.frame(table(filtered_merged_caudate_sub$donor_id, filtered_merged_caudate_sub$cell_class))

classes = c("oligo", "oligo", "oligo", "microglia", "oligo", "oligo", "astrocyte", "astrocyte", "astrocyte", "astrocyte", "SPN", "opc", "oligo", "SPN", "SPN", "interneuron", "oligo", "oligo", "endothelial", "ependymal", "oligo", "interneuron", "interneuron", "astrocyte", "immune", "oligo")

filtered_merged_caudate_sub= assignCellClasses(filtered_merged_caudate_sub, classes=classes, cluster_col="RNA_snn_res.0.5", class_col = "cell_class")
Idents(filtered_merged_caudate_sub) <- "cell_class"


for (donor in donor_list) {
  single_donor_cells = subset(filtered_merged_caudate_sub, subset = donor_id == donor)
  cell_count = cells_per_donor$Freq[cells_per_donor$Var1 == donor]
  print(DimPlot(single_donor_cells,  label = TRUE)+ ggtitle(paste(donor, "Number of cells: ", cell_count)))
}
```

```{r}
cells_per_donor
```


#putamen
```{r}
donor_list = unique(putamen_sobj$donor_id)
cells_per_donor = as.data.frame(table(putamen_sobj$donor_id))

classes = c("oligo", "oligo", "astrocyte", "neuron", "oligo", "microglia", "neuron", "oligo", "opc", "oligo", "neuron", "neuron", "astrocyte", "neuron", "neuron", "oligo", "endo", "neuron", "neuron", "neuron", "oligo", "neuron")

putamen_sobj= assignCellClasses(putamen_sobj, classes=classes, cluster_col="RNA_snn_res.0.2", class_col = "cell_class")
Idents(putamen_sobj) <- "cell_class"


for (donor in donor_list) {
  single_donor_cells = subset(putamen_sobj, subset = donor_id == donor)
  cell_count = cells_per_donor$Freq[cells_per_donor$Var1 == donor]
  print(DimPlot(single_donor_cells,  label = TRUE)+ ggtitle(paste(donor, "Number of cells: ", cell_count)))
}
```

```{r}
head
```


```{r}
DimPlot(putamen_sobj, group.by = "donor_id")

```




```{r}
DimPlot(caudate_neurons, group.by = "RNA_snn_res.0.3")
```

```{r}
#only.pos = False

Idents(caudate_neurons) = caudate_neurons$RNA_snn_res.0.3
markers = FindAllMarkers(caudate_neurons, only.pos = TRUE, min.pct = 0.2, logfc.threshold = 1.25)

num_clusters = length(unique(Idents(caudate_neurons))) - 1

for (i in 0:num_clusters){
   printMarkersByCluster(markers, cluster=i)
}
```

```{r}
caudate_neurons_filtered
```


```{r}
DimPlot(caudate_neurons_filtered, group.by = "RNA_snn_res.0.3")
DimPlot(caudate_neurons_filtered, group.by = "RNA_snn_res.0.4")
DimPlot(caudate_neurons_filtered, group.by = "RNA_snn_res.0.5")
DimPlot(caudate_neurons_filtered, group.by = "RNA_snn_res.0.6")
DimPlot(caudate_neurons_filtered, group.by = "RNA_snn_res.0.7")
DimPlot(caudate_neurons_filtered, group.by = "RNA_snn_res.0.8")
DimPlot(caudate_neurons_filtered, group.by = "RNA_snn_res.0.9")
DimPlot(caudate_neurons_filtered, group.by = "RNA_snn_res.1")
        
```



#reclustering from 0.2 to 0.3
```{r}
#combine clusters 5 and 11 because they look similar
classes = c("D2_SPN_1", "D1_SPN_1", "D1_SPN_2","D1_SPN_4?", "eSPN", "interneuron_1", "interneuron_2", "interneuron_3", "D2_SPN_3", "interneuron_4", "D1_SPN_3", "interneuron_1", "D2_SPN_2", "cholinergic","interneuron_6","interneuron_7", "interneuron_8")
  
  
  #"D1 SPN_1", "D2_SPN_1", "D1 SPN_2?", "eSPN", "interneuron_1", "interneuron_2", "interneuron_3", "SPN?", "interneuron_4", "D1 SPN_3", "D2 SPN_2", "cholinergic", "interneuron_6", "interneuron_7")

caudate_neurons= assignCellClasses(caudate_neurons, classes=classes, cluster_col="RNA_snn_res.0.3", class_col = "cell_class")

Idents(caudate_neurons) <- "cell_class"

DimPlot(caudate_neurons, label= TRUE)
```

```{r}
caudate_neurons_filtered = subset(caudate_neurons, subset = cell_class != "interneuron_8")
caudate_neurons_filtered = subset(caudate_neurons, subset = cell_class != "D2_SPN_3")
caudate_neurons_filtered
```
```{r}
DimPlot(caudate_neurons_filtered, label = TRUE)
```



```{r}
FeaturePlot(caudate_neurons, features =c("CASZ1","CHAT"))
FeaturePlot(caudate_neurons, features =c("DRD1", "DRD2"))
FeaturePlot(caudate_neurons, features =c("PPP1R1B"))
FeaturePlot(caudate_neurons, features =c("SST", "VIP"))
FeaturePlot(caudate_neurons, features =c("GAD1", "GAD2"))
FeaturePlot(caudate_neurons, features = c("EPHA4", "SEMA3E"))
FeaturePlot(caudate_neurons, features =c("PVALB", "CALB2"))
FeaturePlot(caudate_neurons, features =c("SLC17A7", "SLC17A6"))
```
```{r}
FeaturePlot(caudate_neurons, features =c("CD163", "CD74"))
FeaturePlot(caudate_neurons, features =c("CRYAB", "TF"))
FeaturePlot(caudate_neurons, features =c("MBP", "MOBP"))
```




```{r}
Idents(putamen_neurons) = putamen_neurons$RNA_snn_res.0.3
markers = FindAllMarkers(putamen_neurons, only.pos = TRUE, min.pct = 0.2, logfc.threshold = 1.25)

num_clusters = length(unique(Idents(putamen_neurons))) - 1

for (i in 0:num_clusters){
   printMarkersByCluster(markers, cluster=i)
}
```


```{r}
classes = c("D2_SPN_1", "D1_SPN_1", "D1_SPN_2", "D2_SPN_2", "glutamatergic", "D1_SPN_3", "interneuron_1", "interneuron_2", "D2_SPN_3", "interneuron_3", "D1_SPN_4?", "interneuron_4", "interneuron_5", "cholinergic", "interneuron_6", "interneuron_7")
  
putamen_neurons= assignCellClasses(putamen_neurons, classes=classes, cluster_col="RNA_snn_res.0.3", class_col = "cell_class")

Idents(putamen_neurons) <- "cell_class"

DimPlot(putamen_neurons, label= TRUE)
```
```{r}
qsave(caudate_neurons, "Cellbender_seurat/new_caudate_neurons_integrated_subcluster0.3.qs")
qsave(putamen_neurons, "Cellbender_seurat/new_putamen_neurons_integrated_subcluster0.3.qs")
```



```{r}
FeaturePlot(putamen_neurons, features =c("CASZ1","CHAT"))
FeaturePlot(putamen_neurons, features =c("DRD1", "DRD2"))
FeaturePlot(putamen_neurons, features =c("PPP1R1B"))
FeaturePlot(putamen_neurons, features =c("SST", "VIP"))
FeaturePlot(putamen_neurons, features =c("GAD1", "GAD2"))
FeaturePlot(putamen_neurons, features = c("EPHA4", "SEMA3E"))
FeaturePlot(putamen_neurons, features =c("PVALB", "CALB2"))
FeaturePlot(putamen_neurons, features =c("SLC17A7", "SLC17A6"))
```













#from chatgpt, ask later about removing doublets
```{r}
FeaturePlot(filtered_merged_caudate, c("MOBP"))
```
```{r}
# Define marker genes
oligo_genes <- c("OLIG2", "MBP", "MOG")
microglia_genes <- c("CX3CR1", "P2RY12", "TMEM119")
astrocyte_genes <- c("GFAP", "S100B", "ALDH1L1")

# Load your Seurat object
# seurat_obj <- readRDS("path_to_your_seurat_object.rds")

# Add module scores
filtered_merged_caudate <- AddModuleScore(filtered_merged_caudate, features = list(oligo_genes), name = "Oligo_Score")
filtered_merged_caudate <- AddModuleScore(filtered_merged_caudate, features = list(microglia_genes), name = "Microglia_Score")
filtered_merged_caudate <- AddModuleScore(filtered_merged_caudate, features = list(astrocyte_genes), name = "Astrocyte_Score")

# Identify potential doublets
filtered_merged_caudate@meta.data <- filtered_merged_caudate@meta.data %>%
  mutate(Doublet_Score = Oligo_Score1 + Microglia_Score1 + Astrocyte_Score1)

# Define a threshold for doublet score
doublet_threshold <- quantile(filtered_merged_caudate@meta.data$Doublet_Score, 0.99)

# Filter out doublets
filtered_merged_caudate_filtered <- subset(filtered_merged_caudate, subset = Doublet_Score < doublet_threshold)

# Optionally, visualize the scores
FeaturePlot(filtered_merged_caudate, features = c("Oligo_Score1", "Microglia_Score1", "Astrocyte_Score1"))

# Save the filtered Seurat object
# saveRDS(seurat_obj_filtered, "path_to_filtered_seurat_object.rds")

```
