---
title: "R Notebook"
output: html_notebook
---

#OH MY GOD A REDO CAUSE I DIDNT USE THE RIGHT SUBCLUSTERS
#Okay not as big of a deal it's just the old ones switched patch and matrix + ambiguious SPN cluster is labeled as D2
```{r}
caudate_neuron = qread("UPDATED_caudate_neuron_subclustered.qs")
putamen_neuron= qread("UPDATED_caudate_neuron_subclustered.qs")
```

```{r}
unique(caudate_neuron$cell_class)
```


```{r}
caudate_neuron@meta.data$general_class = caudate_neuron@meta.data$cell_class
caudate_neuron$general_class[caudate_neuron$general_class == "D1_SPN_1"] = "D1_SPN"
caudate_neuron$general_class[caudate_neuron$general_class == "D2_SPN_1"] = "D2_SPN"
caudate_neuron$general_class[caudate_neuron$general_class == "D1_SPN_matrix_1"] = "D1_SPN"
caudate_neuron$general_class[caudate_neuron$general_class == "eSPN"] = "eSPN"
caudate_neuron$general_class[caudate_neuron$general_class == "D1_SPN_matrix_2"] = "D1_SPN"
caudate_neuron$general_class[caudate_neuron$general_class == "D2_SPN_matrix_1"] = "D2_SPN"
caudate_neuron$general_class[caudate_neuron$general_class == "D1_SPN_patch"] = "D1_SPN"
caudate_neuron$general_class[caudate_neuron$general_class == "D1_SPN_2"] = "D1_SPN"
caudate_neuron$general_class[caudate_neuron$general_class == "SPN"] = "D2_SPN" #idk maybe controversial.... 


caudate_neuron@meta.data
```

#subset for SPN, interneurons, XDP vs Control
```{r}
XDP_D1_SPN = subset(caudate_neuron, subset = general_class == "D1_SPN" & Condition == "XDP")
XDP_D2_SPN = subset(caudate_neuron, subset = general_class == "D2_SPN" & Condition == "XDP")
XDP_eSPN = subset(caudate_neuron, subset = general_class == "eSPN" & Condition == "XDP")

Control_D1_SPN= subset(caudate_neuron, subset = general_class == "D1_SPN" & Condition == "Control")
Control_D2_SPN= subset(caudate_neuron, subset = general_class == "D2_SPN" & Condition == "Control")
Control_eSPN= subset(caudate_neuron, subset = general_class == "eSPN" & Condition == "Control")

XDP_D1_SPN@meta.data
XDP_D2_SPN@meta.data
Control_eSPN@meta.data
XDP_eSPN@meta.data

XDP_D1_SPN
XDP_D2_SPN
XDP_eSPN
Control_D1_SPN
Control_D2_SPN
Control_eSPN
```




#subset matrix
```{r}
caudate_neuron@meta.data$sub_class = caudate_neuron@meta.data$cell_class
#caudate_neuron$sub_class[caudate_neuron$sub_class == "D1_SPN_1"] = "D1_SPN"
#caudate_neuron$sub_class[caudate_neuron$sub_class == "D2_SPN_1"] = "D2_SPN"
caudate_neuron$sub_class[caudate_neuron$sub_class == "D1_SPN_matrix_1"] = "D1_SPN_matrix"
caudate_neuron$sub_class[caudate_neuron$sub_class == "D1_SPN_matrix_2"] = "D1_SPN_matrix"
caudate_neuron$sub_class[caudate_neuron$sub_class == "D2_SPN_matrix_1"] = "D2_SPN_matrix"
caudate_neuron$sub_class[caudate_neuron$sub_class == "D1_SPN_patch"] = "D1_SPN_patch"
#caudate_neuron$sub_class[caudate_neuron$sub_class == "D1_SPN_2"] = "D1_SPN"
#caudate_neuron$sub_class[caudate_neuron$sub_class == "SPN"] = "D2_SPN" #idk maybe controversial.... 


caudate_neuron@meta.data
```

#subset for SPN, interneurons, XDP vs Control
```{r}
XDP_D1_SPN_matrix = subset(caudate_neuron, subset = sub_class == "D1_SPN_matrix" & Condition == "XDP")
XDP_D2_SPN_matrix = subset(caudate_neuron, subset = sub_class == "D2_SPN_matrix" & Condition == "XDP")
XDP_D1_SPN_patch = subset(caudate_neuron, subset = sub_class == "D1_SPN_patch" & Condition == "XDP")

Control_D1_SPN_matrix= subset(caudate_neuron, subset = sub_class == "D1_SPN_matrix" & Condition == "Control")
Control_D2_SPN_matrix= subset(caudate_neuron, subset = sub_class == "D2_SPN_matrix" & Condition == "Control")
Control_D1_SPN_patch= subset(caudate_neuron, subset = sub_class == "D1_SPN_patch" & Condition == "Control")

XDP_D1_SPN_matrix
XDP_D2_SPN_matrix
XDP_D1_SPN_patch
Control_D1_SPN_matrix
Control_D2_SPN_matrix
Control_D1_SPN_patch
```



#computer is not happy
```{r}

TAF1correlation = function(sobj, title, csv){

sobj_counts = as.matrix(sobj@assays$RNA@counts)
sobj_taf1 = FetchData(sobj, c("TAF1"))$TAF1

gene_names = rownames(sobj_counts)
correlations = numeric(length(gene_names))
p_values = numeric(length(gene_names))


for (i in seq_along(gene_names)) {
  gene_expr = sobj_counts[i, ]
  result = cor.test(gene_expr, sobj_taf1, method = "spearman")
  correlations[i] = result$estimate
  p_values[i] = result$p.value
}

sobj_corr_df = data.frame(
  gene = gene_names,
  spearman_correlation = correlations,
  p_value = p_values
)
# Display the data frame
print(sobj_corr_df)


#sobj_corr_df_filtered = sobj_corr_df[sobj_corr_df$p_value < 0.05,]
sobj_corr_df_filtered = na.omit(sobj_corr_df)
rownames(sobj_corr_df_filtered) = sobj_corr_df_filtered$gene 
sobj_avg_exp = rowMeans(sobj_counts)
sobj_avg_exp = as.data.frame(sobj_avg_exp)
sobj_avg_exp$gene = rownames(sobj_avg_exp)
sobj_df = merge(sobj_corr_df_filtered, sobj_avg_exp, by ="gene")
sobj_df <- sobj_df[sobj_df$gene != "TAF1",]
print(sobj_df)
rownames(sobj_df) = sobj_df$gene 


write.csv(sobj_df, csv)


 top_5 <- sobj_df[order(sobj_df$spearman_correlation, decreasing = TRUE), ][1:5, ]
 top_5

 ggplot(sobj_df, aes(x = spearman_correlation, y = sobj_avg_exp)) +
   geom_point() +
   geom_point(aes(x = spearman_correlation, y = sobj_avg_exp), color = "red") +
   labs(
     x = "Taf1 correlation",
     y = "Average gene expression", title= title) +
   geom_text(data = top_5, aes(label = gene), vjust = -1, color = "black")
 }
```


```{r}
TAF1correlation(XDP_D1_SPN_matrix, "XDP_D1_SPN_matrix", "GSEA/de/new_caudate_neurons_SPN/XDP_D1_SPN_matrix_df.csv")
TAF1correlation(XDP_D2_SPN_matrix, "XDP_D2_SPN_matrix","GSEA/de/new_caudate_neurons_SPN/XDP_D2_SPN_matrix_df.csv")
TAF1correlation(XDP_D1_SPN_patch,"XDP_D1_SPN_patch","GSEA/de/new_caudate_neurons_SPN/XDP_D1_SPN_patch_df.csv")
TAF1correlation(Control_D1_SPN_matrix,"Control_D1_SPN_matrix","GSEA/de/new_caudate_neurons_SPN/Control_D1_SPN_matrix_df.csv")
TAF1correlation(Control_D2_SPN_matrix,"Control_D2_SPN_matrix","GSEA/de/new_caudate_neurons_SPN/Control_D2_SPN_matrix_df.csv")
TAF1correlation(Control_D1_SPN_patch,"Control_D1_SPN_patch","GSEA/de/new_caudate_neurons_SPN/Control_D1_SPN_patch_df.csv")
```


```{r}
 # TAF1correlation(XDP_D1_SPN, "XDP_D1_SPN", "GSEA/de/new_caudate_neurons_SPN/XDP_D1_SPN_df.csv")
 # TAF1correlation(XDP_D2_SPN, "XDP_D2_SPN","GSEA/de/new_caudate_neurons_SPN/XDP_D2_SPN_df.csv")
 # TAF1correlation(XDP_eSPN,"XDP_eSPN","GSEA/de/new_caudate_neurons_SPN/XDP_eSPN_df.csv")
 # TAF1correlation(Control_D1_SPN,"Control_D1_SPN","GSEA/de/new_caudate_neurons_SPN/Control_D1_SPN_df.csv")
 # TAF1correlation(Control_D2_SPN,"Control_D2_SPN","GSEA/de/new_caudate_neurons_SPN/Control_D2_SPN_df.csv")
 # TAF1correlation(Control_eSPN,"Control_eSPN","GSEA/de/new_caudate_neurons_SPN/Control_eSPN_df.csv")
```


```{r}
XDP_D1_SPN_matrix = read.csv("~/GSEA/de/new_caudate_neurons_SPN_matrix/XDP_D1_SPN_matrix_df.csv")
XDP_D2_SPN_matrix = read.csv("~/GSEA/de/new_caudate_neurons_SPN_matrix/XDP_D2_SPN_matrix_df.csv")
XDP_D1_SPN_patch = read.csv("~/GSEA/de/new_caudate_neurons_SPN_matrix/XDP_D1_SPN_patch_df.csv")
Control_D1_SPN_matrix = read.csv("~/GSEA/de/new_caudate_neurons_SPN_matrix/Control_D1_SPN_matrix_df.csv")
Control_D2_SPN_matrix = read.csv("~/GSEA/de/new_caudate_neurons_SPN_matrix/Control_D2_SPN_matrix_df.csv")
Control_D1_SPN_patch = read.csv("~/GSEA/de/new_caudate_neurons_SPN_matrix/Control_D1_SPN_patch_df.csv")

XDP_D1_SPN_matrix$X = NULL
XDP_D2_SPN_matrix$X = NULL
XDP_D1_SPN_patch$X = NULL
Control_D1_SPN_matrix$X = NULL
Control_D2_SPN_matrix$X = NULL
Control_D1_SPN_patch$X = NULL

XDP_D1_SPN_matrix
XDP_D2_SPN_matrix
XDP_D1_SPN_patch
Control_D1_SPN_matrix
Control_D2_SPN_matrix
Control_D1_SPN_patch
```

#GSEA

```{r}
sobj_df = XDP_D1_SPN_matrix
title= "XDP D1 SPN Matrix"

top_5 <- sobj_df[order(sobj_df$spearman_correlation, decreasing = TRUE), ][1:5, ]
top_5

ggplot(sobj_df, aes(x = spearman_correlation, y = sobj_avg_exp)) +
  geom_point(aes(x = spearman_correlation, y = sobj_avg_exp), color = "red", alpha = 0.3) + 
  labs(
    x = "Taf1 correlation",
    y = "Average gene expression", title= title) +
  geom_text(data = top_5, aes(label = gene), vjust = -1, color = "black") + xlim(-0.15, 0.3) + ylim(0,500)

sobj_df = XDP_D2_SPN_matrix
title= "XDP D2 SPN Matrix"

top_5 <- sobj_df[order(sobj_df$spearman_correlation, decreasing = TRUE), ][1:5, ]
top_5

ggplot(sobj_df, aes(x = spearman_correlation, y = sobj_avg_exp)) +
  geom_point(aes(x = spearman_correlation, y = sobj_avg_exp), color = "red", alpha = 0.3) + 
  labs(
    x = "Taf1 correlation",
    y = "Average gene expression", title= title) +
  geom_text(data = top_5, aes(label = gene), vjust = -1, color = "black") + xlim(-0.15, 0.2) + ylim(0,500)

sobj_df = XDP_D1_SPN_patch
title= "XDP D1 SPN patch"

top_5 <- sobj_df[order(sobj_df$spearman_correlation, decreasing = TRUE), ][1:5, ]
top_5

ggplot(sobj_df, aes(x = spearman_correlation, y = sobj_avg_exp)) +
  geom_point(aes(x = spearman_correlation, y = sobj_avg_exp), color = "red", alpha = 0.3) + 
  labs(
    x = "Taf1 correlation",
    y = "Average gene expression", title= title) +
  geom_text(data = top_5, aes(label = gene), vjust = -1, color = "black") + xlim(-1, 1) + ylim(0,400)

sobj_df = Control_D1_SPN_matrix
title= "Control D1 SPN Matrix"

top_5 <- sobj_df[order(sobj_df$spearman_correlation, decreasing = TRUE), ][1:5, ]
top_5

ggplot(sobj_df, aes(x = spearman_correlation, y = sobj_avg_exp)) +
  geom_point(aes(x = spearman_correlation, y = sobj_avg_exp), color = "red", alpha = 0.3) + 
  labs(
    x = "Taf1 correlation",
    y = "Average gene expression", title= title) +
  geom_text(data = top_5, aes(label = gene), vjust = -1, color = "black") + xlim(-0.15, 0.30) + ylim(0,500)

sobj_df = Control_D2_SPN_matrix
title= "Control D2 SPN Matrix"

top_5 <- sobj_df[order(sobj_df$spearman_correlation, decreasing = TRUE), ][1:5, ]
top_5

ggplot(sobj_df, aes(x = spearman_correlation, y = sobj_avg_exp)) +
  geom_point(aes(x = spearman_correlation, y = sobj_avg_exp), color = "red", alpha = 0.3) + 
  labs(
    x = "Taf1 correlation",
    y = "Average gene expression", title= title) +
  geom_text(data = top_5, aes(label = gene), vjust = -1, color = "black") + xlim(-0.15, 0.20) + ylim(0,500)

sobj_df = Control_D1_SPN_patch
title= "Control D1 SPN patch"

top_5 <- sobj_df[order(sobj_df$spearman_correlation, decreasing = TRUE), ][1:5, ]
top_5

ggplot(sobj_df, aes(x = spearman_correlation, y = sobj_avg_exp)) +
  geom_point(aes(x = spearman_correlation, y = sobj_avg_exp), color = "red", alpha = 0.3) + 
  labs(
    x = "Taf1 correlation",
    y = "Average gene expression", title= title) +
  geom_text(data = top_5, aes(label = gene), vjust = -1, color = "black") + xlim(-1,1) + ylim(0,400)
```


#change around the logfc and stuff
```{r}
library(Seurat)
library(dplyr)
library(qs)

library(Matrix)
library(fgsea)


# when concatenated together, these should be the paths to your DE csv files
BASE_PATH="~/GSEA"
DE_SUBDIR= "de/new_caudate_neurons_SPN_matrix"


.extraHumanGeneAnnoAdderFn=function(inputGeneNames=NULL){
  #require(EnsDb.Hsapiens.v75)
  require(EnsDb.Hsapiens.v86)
  
  if(!dir.exists("~/serverFiles")){
    dir.create("~/serverFiles",recursive = T)
  }
  
  gns <- as.data.frame(genes(EnsDb.Hsapiens.v86))
  gns$gene_short_name=gns$gene_name
  gns$symbol=toupper(gns$symbol)
  gns$ensembl_gene_id=row.names(gns)
  
  if(!is.null(inputGeneNames)){
    rwNames=toupper(inputGeneNames)
    psCols=c("gene_short_name","ensembl_gene_id")
    slCounts=0
    slCol=""
    if(sum(grepl("\\.",rwNames)&grepl("^ENS",rwNames))>0){
      rwNames=strsplit(rwNames,"\\.")
      rwNames=unlist(lapply(rwNames,function(x)x[1]))
    }
    system(paste0("gsutil -m cp gs://fc-71ac3b81-2441-4171-8038-baf653634620/serverFiles/human_map_to_ensembl.rda ."))
    load("human_map_to_ensembl.rda")
    
    
    map_to_ensmbl$source=toupper(map_to_ensmbl$source)
    
    system(paste0("gsutil -m cp gs://fc-71ac3b81-2441-4171-8038-baf653634620/serverFiles/human_mapping_hg19.rda ."))
      
    
    load("human_mapping_hg19.rda")
    human_hg19$source=toupper(human_hg19$source)
    
    if(sum(toupper(rwNames) %in% human_hg19$source) > sum(toupper(rwNames) %in% map_to_ensmbl$source)){
      map_to_ensmbl=merge(human_hg19,data.frame(source=toupper(rwNames),stringsAsFactors = F),by="source",all.y=T)
    } else {
      map_to_ensmbl=merge(map_to_ensmbl,data.frame(source=toupper(rwNames),stringsAsFactors = F),by="source",all.y=T)
    }
    
    gns=merge(gns,map_to_ensmbl,by.x="ensembl_gene_id",by.y="target",all.y=T)
    gns=gns[match(rwNames,gns$source),]
    row.names(gns)=inputGeneNames
    gns$gene_id=inputGeneNames
    gns=gns[,-which(colnames(gns) %in% c("source","target"))]
  }
  
  return(gns)
}

.sconline.GSEA.readGMT=function (file,bkg_genes=NULL,min.gs.size=NULL,max.gs.size=NULL) {
  if (!grepl("\\.gmt$", file)[1]&F) {
    stop("Pathway information must be in a .gmt file format")
  }
  geneSetDB = readLines(file)
  geneSetDB = strsplit(geneSetDB, "\t")
  names(geneSetDB) = sapply(geneSetDB, "[", 1)
  geneSetDB = lapply(geneSetDB, "[", -1:-2)
  geneSetDB = lapply(geneSetDB, function(x) {
    x[which(x != "")]
  })
  
  if(!is.null(bkg_genes)){
    for(i in 1:length(geneSetDB)){
      tmp=geneSetDB[[i]]
      tmp=bkg_genes[which(toupper(bkg_genes) %in% toupper(tmp))]
      geneSetDB[[i]]=tmp
    }
  }
  
  if(!is.null(min.gs.size)){
    size.dist=unlist(lapply(geneSetDB,length))
    geneSetDB=geneSetDB[size.dist>=min.gs.size]
  }
  
  if(!is.null(max.gs.size)){
    size.dist=unlist(lapply(geneSetDB,length))
    geneSetDB=geneSetDB[size.dist<=max.gs.size]
  }
  
  return(geneSetDB)
}

.myfGSEAfn=function(rankedVec,gs,minSize  = 15,maxSize  = 250, scoreType='std'){
  require(fgsea)
  fgseaRes <- fgsea(pathways = gs, 
                    stats    = rankedVec,
                    minSize  = minSize,
                    maxSize  = maxSize,
                    scoreType= scoreType)
  fgseaRes=fgseaRes[order(fgseaRes$pval,decreasing = F),]
  fgseaRes=as.data.frame(fgseaRes)
  fgseaRes$leadingEdge=unlist(lapply(fgseaRes$leadingEdge,function(x) paste(x,collapse = ",")))
  
  return(fgseaRes)
}

runGSEA = function(
    de_df,
    gs_list_of_char,
    rank_col='logFC',
    gene_id_col='gene_short_name',
    desc=FALSE,
    abs=TRUE,
    scoreType='std'){

    de_df=de_df[!duplicated(de_df[[gene_id_col]]),]
    
    order_vector = de_df[[rank_col]]
    if(abs){
        order_vector = abs(order_vector)
    }
    
    ranked_vec=de_df[,rank_col]
    names(ranked_vec)=de_df[[gene_id_col]]
    ranked_vec=ranked_vec[order(order_vector,decreasing = desc)]
    
    print(head(ranked_vec))

    res_fGSEA=.myfGSEAfn(rankedVec=ranked_vec,gs=gs_list_of_char, scoreType=scoreType)
    res_fGSEA=res_fGSEA[order(res_fGSEA$padj,decreasing=F),]
    return(res_fGSEA)
}
```



#IDK why but gsea is not workign
```{r}
outpath = file.path(BASE_PATH, DE_SUBDIR, "gsea")
# if you're running into permission errors may need to delete directory and remake with mkdir on command line
dir.create(outpath, recursive=TRUE, showWarnings=FALSE)
de_files = list.files(file.path(BASE_PATH, DE_SUBDIR))
de_files = de_files[grep(".csv$", de_files)]

gene_sets = list(
    # add the paths to the gene sets you want to use here
    kegg_2021_human = "GSEA/genesets/KEGG_2021_Human.txt",
    go_process = "GSEA/genesets/GO_Biological_Process_2021.txt",
    go_function = "GSEA/genesets/GO_Molecular_Function_2021.txt",
    disgenet = "GSEA/genesets/DisGeNET.txt",
    msigdb_hallmark = "GSEA/genesets/MSigDB_Hallmark_2020.txt",
    #neural_activity = "GSEA/genesets/neural_activity_arranged.txt",
    wikipathways = "GSEA/genesets/WikiPathways_2019_Human.txt",
    #tf_perturbations = "GSEA/genesets/TF_Perturbations_Followed_by_Expression.txt",
    jensen_diseases = "GSEA/genesets/Jensen_DISEASES.txt"
    #X_chromsome_positional= "GSEA/genesets/allchrome.txt",  
    #Xlinked_recessive = "GSEA/genesets/Xlinked_RESS.txt",
    #Xlinked_dominant = "GSEA/genesets/Xlinked_DOM.txt",
    #similar_diseases = "GSEA/genesets/similardiseases.txt"
)

for (de_file in de_files){
    de_slogan = sub(".txt$", "", de_file)

    #if (! grepl("trend", de_slogan)){next}

    #if (strsplit(de_slogan, "__")[[1]][[2]] != "opc"){next}

    dataDE = read.csv(file.path(BASE_PATH, DE_SUBDIR, de_file))
    anno=.extraHumanGeneAnnoAdderFn(inputGeneNames=dataDE$gene)
    anno=anno[match(dataDE$gene,anno$gene_id),]
    dataDE=cbind(dataDE,anno)
    dataDE=dataDE[which(dataDE$gene_biotype=="protein_coding"),]

    gsea_list = list()

    for (gene_set in names(gene_sets)){
        path = gene_sets[[gene_set]]

        gs=.sconline.GSEA.readGMT(file=path,bkg_genes=dataDE$gene_short_name,min.gs.size=15,max.gs.size=500)
        gsea = runGSEA(dataDE, gs, rank_col="spearman_correlation", abs=FALSE, desc=TRUE)
        gsea$gene_set = gene_set
        gsea = gsea[which(gsea$padj<0.05),]

        if (nrow(gsea) > 0){
            gsea_list[[gene_set]] = gsea
        }
    }
    gsea_df = do.call(rbind, gsea_list)

    write.table(gsea_df, file.path(outpath, paste0(de_slogan, "__gsea_taf1correlation.tsv")), sep="\t", quote=FALSE, row.names=FALSE)
}
```

```{r}
caudate_neurons_types = list("XDP_D1_SPN_matrix_df","XDP_D1_SPN_patch_df", "Control_D1_SPN_matrix_df","Control_D1_SPN_patch_df","Control_D2_SPN_matrix_df")
caudate_neurons_gsea_list = list()

# Loop through each cell type and read the corresponding file into a dataframe
for (cell_type in caudate_neurons_types) {
  # Construct the file name
  gsea_file_name <- paste("~/GSEA/de/new_caudate_neurons_SPN_matrix/gsea/", cell_type, ".csv__gsea_taf1correlation.tsv", sep = "")
  
  # Print the file name to verify the path
  print(paste("Trying to read file:", gsea_file_name))
  
  # Check if the file exists
  if (file.exists(gsea_file_name)) {
    # Read the file into a dataframe
    gsea_file <- read.delim(gsea_file_name, header = TRUE)
    
    # Store the dataframe in the list with the cell type as the name
    caudate_neurons_gsea_list[[cell_type]] <- gsea_file
  } else {
    print(paste("File does not exist:", gsea_file_name))
  }
}
caudate_neurons_gsea_list
```



```{r}
XDP_D1_SPN_matrix_df = caudate_neurons_gsea_list[["XDP_D1_SPN_matrix_df"]]
XDP_D1_SPN_patch_df= caudate_neurons_gsea_list[["XDP_D1_SPN_patch_df"]]
Control_D1_SPN_matrix_df= caudate_neurons_gsea_list[["Control_D1_SPN_matrix_df"]]
Control_D1_SPN_patch_df = caudate_neurons_gsea_list[["Control_D1_SPN_patch_df"]]
Control_D2_SPN_matrix_df= caudate_neurons_gsea_list[["Control_D2_SPN_matrix_df"]]


XDP_D1_SPN_matrix_df
XDP_D1_SPN_patch_df
Control_D1_SPN_matrix_df
Control_D1_SPN_patch_df
Control_D2_SPN_matrix_df
```

#gsea with leading edge annotations
```{r}
plot_gsea_result_hdot = function(
    df,
    title=NULL,
    xlim=c(-4, 4),
    fig_filename = NULL,
    leading_edge_n=10,
    leading_edge_linebreak_n=5,
    top_n=10) {
    #takes in a df with the following columns: NES, pathway, size, leading_edge
    # along with a title, xlim, fig_filename (nullable)
    # leading_edge_n is the max number of leading_edge genes to be annotated
    # leading_edge_linebreak_n is the number of leading_edge genes to be displayed before a line break
    # top_n is the number of pathways to be displayed on the plot for each direction
    # (i.e. max rows is 2*top_n for up- and down-regulated genesets)
    # the df is filtered to the top_n pathways by NES above 0 and below 0
    # Y axis: pathway. every 30 characters, replace the next space with a \n and then start over
    # X axis: NES, appearing as an open circle. If NES > 0, this circle is blue, < 0 red. The size of the circle relates to the Gene Set Size.
    # The plot is annotated on the right with the leading_edge_top_n.
    # first, turn the leading edge into something we can read by taking the first `leading_edge_n` elemnets
    # separating them with `, ` and finally adding line breaks
    df_pos = df[df$NES > 0,]
    df_neg = df[df$NES < 0,]
    df_pos = df_pos[order(df_pos$NES, decreasing = TRUE),]
    df_neg = df_neg[order(df_neg$NES, decreasing = FALSE),]
    df = rbind(df_pos[1:top_n,], df_neg[1:top_n,])
    df = df[order(df$NES),]
    df = df[!is.na(df$NES),]
    # Function to insert line breaks every `n` entries
    insert_line_breaks = function(text, n = leading_edge_linebreak_n) {
        parts = strsplit(text, ",\\s*")[[1]]
        if (length(parts) <= n) {
            return(text)
        }
        new_text = ""
        for (i in seq_along(parts)) {
            new_text = paste0(new_text, parts[i])
            if (i < length(parts)) {
                new_text = paste0(new_text, ",")
            }
            if (i %% n == 0 && i != length(parts)) {
                new_text = paste0(new_text, "\n")
            } else if (i != length(parts)) {
                new_text = paste0(new_text, " ")
            }
        }
        return(new_text)
    }
    df$leading_edge_top_n = sapply(df$leadingEdge, function(x){
        edge_list = strsplit(x, ",")[[1]]
        len = length(edge_list)
        if (len > leading_edge_n){
            len=leading_edge_n
        }
        return(insert_line_breaks(paste(edge_list[1:len], collapse=", ")))
    })
    if (is.null(title)){
        title = "GSEA NES by Pathway"
    }
    # Format the pathway names to include newline every 30 characters
    df$pathway = gsub("(.{30}\\s)", "\\1\n", df$pathway, perl = TRUE)
    df$NES_direction = ifelse(df$NES > 0, "NES > 0", "NES < 0")
    # Create the plot
    p = ggplot(df, aes(x = NES, y = reorder(pathway, NES))) +
        geom_point(aes(size = size, color = NES_direction), shape = 1) +
        scale_color_manual(name = "NES Direction", values = c("NES > 0" = "blue", "NES < 0" = "red")) +
        scale_size_continuous(name = "Size") +
        geom_vline(xintercept = 0, linetype = "dotted", color = "black") +
        labs(title = title, x = "NES", y = "Pathway") +
        xlim(xlim) +
        theme(
            panel.grid.major = element_line(color = "gray", linetype = "dotted"),
            panel.grid.minor = element_blank(),
            legend.position = "left",
            plot.title = element_text(size = 14),
            plot.caption = element_text(hjust = 0.1),
            plot.title.position = "plot",
            axis.line = element_line(color = "black"),
            axis.ticks = element_line(color = "black"),
            axis.text = element_text(size = 13),
            axis.title = element_text(size = 13),
            axis.text.y.right = element_text(hjust = 1)
        )
        # Create a secondary y-axis with leading_edge_top_n labels
    p = p + geom_text(aes(y = reorder(pathway, NES), x = Inf, label = leading_edge_top_n),
                      hjust = -0.1, size = 4) +
        coord_cartesian(clip = 'off') +
        theme(
            plot.margin = margin(5.5, 340, 5.5, 5.5),  # Increase right margin to make space for secondary labels
            axis.text.y.right = element_text(hjust = 1)
        )
    # Save the plot to file if a filename is provided
    if (!is.null(fig_filename)) {
        ggsave(fig_filename, plot = p, width = 5, height = 2*nrow(df)/3)
    }
    # Print the plot to the R console
    print(p)
}
```



```{r}
plot_gsea_result_hdot(XDP_D1_SPN_matrix_df, title = "XDP_D1_SPN_matrix_df", leading_edge_n=5)
plot_gsea_result_hdot(XDP_D1_SPN_patch_df, title = "XDP_D1_SPN_patch_df", leading_edge_n=5)
plot_gsea_result_hdot(Control_D1_SPN_matrix_df, title = "Control_D1_SPN_matrix_df", leading_edge_n=5)
plot_gsea_result_hdot(Control_D1_SPN_patch_df, title = "Control_D1_SPN_patch_df", leading_edge_n=5)
plot_gsea_result_hdot(Control_D2_SPN_matrix_df, title = "Control_D2_SPN_matrix_df", leading_edge_n=5)

```



```{r}
hist(XDP_D1_SPN_matrix$spearman_correlation, main = "XDP D1 SPN Matrix TAF1 Correlation", breaks =20, xlab = "TAF1 Spearman Coefficient")
hist(XDP_D2_SPN_matrix$spearman_correlation, main = "XDP D2 SPN Matrix TAF1 Correlation", breaks =20, xlab = "TAF1 Spearman Coefficient")
hist(XDP_D1_SPN_patch$spearman_correlation, main = "XDP D1 SPN Patch TAF1 Correlation", breaks =20, xlab = "TAF1 Spearman Coefficient")
hist(Control_D1_SPN_matrix$spearman_correlation, main = "Control D1 SPN Matrix TAF1 Correlation", breaks =20, xlab = "TAF1 Spearman Coefficient")
hist(Control_D2_SPN_matrix$spearman_correlation, main = "Control D2 SPN Matrix TAF1 Correlation", breaks =20, xlab = "TAF1 Spearman Coefficient")
hist(Control_D1_SPN_patch$spearman_correlation, main = "Control D1 SPN Patch TAF1 Correlation", breaks =20, xlab = "TAF1 Spearman Coefficient")
```

```{r}
D1_SPN_matrix_DE = read.csv("GSEA/de/SPN_DE/D1_SPN_matrix_DE.csv") 
D2_SPN_matrix_DE = read.csv("GSEA/de/SPN_DE/D2_SPN_Matrix_DE.csv") 
D1_SPN_patch_DE = read.csv("GSEA/de/SPN_DE/D1_SPN_Patch_DE.csv")   

D1_SPN_matrix_DE
D2_SPN_matrix_DE
D1_SPN_patch_DE

```

```{r}
XDP_merged_D1_matrix = merge(XDP_D1_SPN_matrix, D1_SPN_matrix_DE, by = "gene")
XDP_merged_D2_matrix= merge(XDP_D2_SPN_matrix, D2_SPN_matrix_DE, by = "gene")
XDP_merged_D1_patch= merge(XDP_D1_SPN_patch, D1_SPN_patch_DE, by = "gene")

Control_merged_D1_matrix = merge(Control_D1_SPN_matrix, D1_SPN_matrix_DE, by = "gene")
Control_merged_D2_matrix= merge(Control_D2_SPN_matrix, D2_SPN_matrix_DE, by = "gene")
Control_merged_D1_patch= merge(Control_D1_SPN_patch, D1_SPN_patch_DE, by = "gene")

```


```{r}
XDP_merged_D1_matrix 
XDP_merged_D2_matrix
XDP_merged_D1_patch

Control_merged_D1_matrix
Control_merged_D2_matrix
Control_merged_D1_patch
```


#not happy with this because now the red can show nonsignificant taf1?
```{r}
degraphs = function(sobj_df, title){
 
  sobj_df$significant_DE = sobj_df$adj.P.Val < 0.05
 significant_genes = sobj_df[sobj_df$significant_DE, ]
 
 # Order the filtered dataframe by spearman correlation in decreasing order
 top_5 <- significant_genes[order(abs(significant_genes$spearman_correlation), decreasing = TRUE), ][1:5, ]
 print(top_5)
 
 ggplot(sobj_df, aes(x = spearman_correlation, y = logFC)) +
   geom_point(aes(x = spearman_correlation, y = logFC, color = significant_DE), alpha = 0.3) + scale_color_manual(values = c("FALSE" = "black", "TRUE"="red"))+
   labs(
     x = "Taf1 correlation",
     y = "DE logFC", title= title) +
   geom_text(data = top_5, aes(label = gene), vjust = -1, color = "black") 
 }
```


#use this if you filtered
```{r}
# degraphs = function(sobj_df, title){
# 
#  # Order the filtered dataframe by spearman correlation in decreasing order
#  top_5 <- sobj_df[order(sobj_df$spearman_correlation, decreasing = TRUE), ][1:5, ]
#  print(top_5)
#  
#  ggplot(sobj_df, aes(x = spearman_correlation, y = logFC)) +
#    geom_point(aes(x = spearman_correlation, y = logFC), alpha = 0.3) +
#    labs(
#      x = "Taf1 correlation",
#      y = "DE logFC", title= title) +
#    geom_text(data = top_5, aes(label = gene), vjust = -1, color = "black") 
#  }
```



```{r}
degraphs(XDP_merged_D1_matrix, "XDP D1 SPN Matrix") 
degraphs(XDP_merged_D2_matrix, "XDP D2 SPN Matrix") 
degraphs(XDP_merged_D1_patch, "XDP D1 SPN Patch")  
degraphs(Control_merged_D1_matrix, "Control D1 SPN Matrix") 
degraphs(Control_merged_D2_matrix, "Control D2 SPN Matrix") 
degraphs(Control_merged_D1_patch, "Control D1 SPN Patch")
```





