---
title: "R Notebook"
output: html_notebook
---

```{r}
All_XDP_Cohort = qread("All_XDP_Cohorts_CaH_Put_final.qs")
All_XDP_Cohort
```
```{r}
DimPlot(All_XDP_Cohort, repel = T, group.by = "final_cell_class_merged_harmony", label = T, raster = F)
```



```{r}
table(All_XDP_Cohort$class_label_name)
```
```{r}
All_cells_df
```


```{r}
library(dplyr)
library(ggplot2)
temp = subset(All_XDP_Cohort@meta.data, subset = class_label_name =="CN LGE GABA")

# Count total cells per donor
cells_per_donor <- as.data.frame(table(temp$donor_id))
colnames(cells_per_donor) <- c("donor_id", "total_cells")

# Count cells by class and donor
cell_class_df <- as.data.frame(table(temp$group_label_name, temp$donor_id))
colnames(cell_class_df) <- c("group_label_name", "donor_id", "Freq")

# Merge cell class counts with total cell counts
cell_class_df <- merge(cell_class_df, cells_per_donor, by = "donor_id")

# Calculate proportions
cell_class_df$cell_proportions <- cell_class_df$Freq / cell_class_df$total_cells

# Add donor-level metadata
cell_class_df$Cohort <- All_cells_df$Cohort[match(cell_class_df$donor_id, All_cells_df$donor_id)]
cell_class_df$Cohort_Condition <- All_cells_df$Cohort_Condition[match(cell_class_df$donor_id, All_cells_df$donor_id)]
cell_class_df$Sex <- All_cells_df$Sex[match(cell_class_df$donor_id, All_cells_df$donor_id)]
cell_class_df$Condition <- All_cells_df$Condition[match(cell_class_df$donor_id, All_cells_df$donor_id)]

# Identify controls and cases by donor_id suffix
controls <- cells_per_donor[grepl("CM$|CF$|CM2$|CM6$", cells_per_donor$donor_id), ]
cases <- cells_per_donor[!grepl("CM$|CF$|CM2$|CM6$", cells_per_donor$donor_id), ]

# Sort each group by descending cell count
controls <- controls[order(-controls$total_cells), ]
cases <- cases[order(-cases$total_cells), ]

# Combine sorted donor IDs
ordered_donors <- c(controls$donor_id, cases$donor_id)

# Set donor_id as a factor with correct order
cell_class_df$donor_id <- factor(cell_class_df$donor_id, levels = ordered_donors)

# Optional: re-sort full dataframe to follow donor order
cell_class_df <- cell_class_df %>%
  arrange(factor(donor_id, levels = ordered_donors))



# Now compute y_mid for neurons
astro_labels <- cell_class_df %>%
  filter(group_label_name == "STRd D2 Matrix") %>%
  mutate(y_mid = 0.5-(cell_proportions / 2) )


# PLOT 1: Absolute counts
ggplot(cell_class_df, aes(x = donor_id, y = Freq, fill = group_label_name)) +
  geom_bar(stat = "identity") +
  xlab("Donors") + ylab("Number of cells by Class Labels") +
  labs(fill = "Class Labels") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 

# PLOT 2: Proportions
ggplot(cell_class_df, aes(x = donor_id, y = cell_proportions, fill = group_label_name)) +
  geom_bar(stat = "identity", position = "stack") +
  xlab("Donors") + ylab("Class Label Proportion") +
  labs(fill = "Class Labels") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# PLOT 3: Faceted absolute counts
a <- ggplot(cell_class_df, aes(x = donor_id, y = Freq, fill = group_label_name)) +
  geom_bar(stat = "identity") +
  xlab("Donors") + ylab("Number of cells by Class Labels") +
  labs(fill = "Class Labels") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_grid(~ Cohort, scales = "free_x", space = "free_x") +
  geom_text(aes(y = total_cells, label = total_cells), vjust = -0.2, size = 3, nudge_y = 0.5)

# PLOT 4: Faceted proportions
b <- ggplot(cell_class_df, aes(x = donor_id, y = cell_proportions, fill = group_label_name)) +
  geom_bar(stat = "identity", position = "stack") +
  xlab("Donors") + ylab("Class Labels") +
  labs(fill = "Class Labels") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_grid(~ Cohort, scales = "free_x", space = "free_x")+ 
  geom_text( data = astro_labels,  
    aes(x = donor_id, y = y_mid, label = round(cell_proportions, 2)),  # Correctly centered text
    size = 3)

# Save plots
ggsave(a, filename = "pic.png", width = 18, height = 5)
ggsave(b, filename = "pic2.png", width = 18, height = 5)
```

```{r}
cell_class_df
```



```{r}
```{r}
All_XDP_Cohort <- AddModuleScore_UCell(All_XDP_Cohort, features = list(Score = bican_recon_astro_markers_top50_wm), name = 'bican_recon_astro_markers_top50_wm')
All_XDP_Cohort <- AddModuleScore_UCell(All_XDP_Cohort, features = list(Score = bican_recon_astro_markers_top50_gm), name = 'bican_recon_astro_markers_top50_gm')
```


```{r}
softmax_cols = function(df, cols){
    z_cols = paste0(cols, "__z")
    softmax_cols = paste0(cols, "__softmax")
    
    # Z-score normalization of the selected columns
    df[z_cols] = scale(df[cols])
    
    # Apply softmax row-wise: exp of z-scores divided by row sum of exp(z-scores)
    df[softmax_cols] =  t(apply(df[z_cols], 1, function(x) exp(x) / sum(exp(x))))
    
    return(df)
}


columns = c("ScoreNFKB", "antigen_processing_scores", "Scorebican_recon_astro_markers_top50_wm", "Scorebican_recon_astro_markers_top50_gm")
softmax_cols(All_XDP_Cohort@meta.data, columns)
```


```{r}
All_XDP_Cohort_meta = All_XDP_Cohort@meta.data
All_XDP_Cohort_meta

columns = c("ScoreNFKB", "antigen_processing_scores", "Scorebican_recon_astro_markers_top50_wm", "Scorebican_recon_astro_markers_top50_gm")

All_XDP_Cohort_meta = softmax_cols(All_XDP_Cohort_meta, columns)
All_XDP_Cohort_meta
```


```{r}
All_XDP_Cohort_meta$WM_GM <- ifelse(
  All_XDP_Cohort_meta$Scorebican_recon_astro_markers_top50_wm__softmax > All_XDP_Cohort_meta$Scorebican_recon_astro_markers_top50_gm__softmax, "White Matter", 
  ifelse(All_XDP_Cohort_meta$Scorebican_recon_astro_markers_top50_gm__softmax > All_XDP_Cohort_meta$Scorebican_recon_astro_markers_top50_wm__softmax, "Gray Matter", "Uncertain")
)
All_XDP_Cohort_meta
```


```{r}
p = ggplot(All_XDP_Cohort_meta, aes(x = Scorebican_recon_astro_markers_top50_wm__softmax, y = Scorebican_recon_astro_markers_top50_gm__softmax, color = WM_GM)) +
  geom_point(alpha = 0.5, size = 0.1) 
print(p)
p1 = ggMarginal(p, type="histogram" , xparams = list(  bins=100))
print(p1)
```


```{r}
temp = subset(All_XDP_Cohort_meta, subset = final_cell_class_merged_harmony == "astrocyte")

min =  min(temp$ScoreNFKB__softmax
)
max = max(temp$ScoreNFKB__softmax
)
step = (max-min)/100

limits = seq(min, max, step)


a = plot_overlapping_density_histogram(df = temp, 
                                          hist_col = temp$ScoreNFKB__softmax,
                                          fill_col = "Condition",
                                  colors = c("XDP" = "red", "Control" = "blue"),
                                          breaks = limits,
                                          title = paste0("Astrocyte"),
                                          xlab = "NFKB Score",
                                          fig_filename = NULL) + facet_wrap(~ Cohort) #+ WM_GM)
 print(a)
 #ggsave(a, filename = "pic.png", width = 10, height = 4)
```







```{r}
taf1_df <- FetchData(seurat_obj, vars = c("TAF1", "condition", "cohort", "region"))

```

```{r}
All_cells_df = qread("All_cells_df.qs")
All_cells_df
```
```{r}
All_cells_df_cohort1 = subset()
```


```{r}
adata <- readH5AD("/broad/macosko/kimkathl/XDP/QC_and_Clustering/files_needed/Atlas_Cortex_trim.h5ad")

# Inspect metadata
colnames(colData(adata))
```

```{r}
table(colData(adata)$leiden)  # count of cells per cell type
```


```{r}

library(SeuratDisk)

# First, convert .h5ad to .h5seurat
Convert("/broad/macosko/kimkathl/XDP/QC_and_Clustering/files_needed/Atlas_Cortex_trim.h5ad", dest = "h5seurat", overwrite = TRUE)

# Load into Seurat
ref <- LoadH5Seurat("/broad/macosko/kimkathl/XDP/QC_and_Clustering/files_needed/Atlas_Cortex_trim.h5seurat")

# View metadata
head(ref@meta.data)

```
```{r}
table(ref$cell.type)  # Check the cell types and how many cells in each class
```
```{r}
# Subset to 2000 cells per class (adjust this logic if needed)
set.seed(123)  # To ensure reproducibility
subsetted_ref <- ref[, unlist(lapply(unique(ref$cell.type), function(ct) {
  sample(which(ref$cell.type == ct), min(2000, length(which(ref$cell.type == ct))))
}))]
table(subsetted_ref$cell.type)

```
```{r}

seurat_obj <- as.Seurat(subsetted_ref)

# Now save the Seurat object as .h5ad
SaveH5Seurat(seurat_obj, filename = "/broad/macosko/kimkathl/XDP/QC_and_Clustering/subsetted_cortical.h5ad", overwrite = TRUE)
```





```{r}
# Load required libraries
library(Seurat)
library(SeuratDisk)

# Convert .h5ad to .h5seurat format
Convert("/broad/macosko/kimkathl/XDP/QC_and_Clustering/Atlas_Cortex_trim.h5ad", dest = "h5seurat", overwrite = TRUE)

# Load into Seurat
ref <- LoadH5Seurat("/broad/macosko/kimkathl/XDP/QC_and_Clustering/Atlas_Cortex_trim.h5seurat")

# Check the metadata to inspect the cell types
head(ref@meta.data)

```

```{r}
# Get the cell types
cell_types <- unique(ref$cell.type)  # Replace 'cell.type' with the actual column name

# Subset to 2000 cells per cell type
subset_cells <- lapply(cell_types, function(ct) {
  cells_of_type <- WhichCells(ref, expression = cell.type == ct)  # Get cells of this type
  sampled_cells <- sample(cells_of_type, size = min(2000, length(cells_of_type)), replace = FALSE)  # Randomly sample 2000 or all if fewer than 2000
  return(sampled_cells)
})

# Flatten the list into a vector of selected cells
selected_cells <- unlist(subset_cells)

# Subset the Seurat object to only the selected cells
subsetted_ref <- subset(ref, cells = selected_cells)

# Check the subsetted object
table(subsetted_ref$cell.type)  # Verify the number of cells per type
```


```{r}
# Save the subsetted Seurat object as .h5ad
SaveH5Seurat(subsetted_ref, filename = "/broad/macosko/kimkathl/XDP/QC_and_Clustering/subsetted_cortex.h5ad", overwrite = TRUE)

```






```{r}
xdp_recon = qread("xdp_recon_apr2025_sct_mask.qs")
xdp_recon@meta.data
```

```{r}
xdp_recon_neurons = subset(xdp_recon@meta.data, subset = Class_label_name == "CN LGE GABA")
xdp_recon_neurons
```

```{r}
spatial_scores_xdp = function(metadata, column, title, updown = -1){
a = ggplot(metadata, aes(x= x_um, y = y_um, color = !!sym(column))) + geom_point(size =0.6) + ggtitle(paste0(" ", title, " Scores")) + scale_color_viridis_c(option = "magma", name = "Expression", direction = updown) +   labs(color = paste0(title," Scores")) +
  theme_void() +
  theme(
    plot.title = element_text(size = 20),
    plot.subtitle = element_text(size = 20),
    strip.text = element_text(size = 20),
    plot.background = element_rect(fill = "white", color = NA),
     legend.text = element_text(size = 24),  # Increase legend text size
    legend.title = element_text(size = 26)
  ) +
  ylab(NULL) + facet_wrap(~ Group_label_name)

ggsave(a, filename = "pic4.png", width = 15, height = 10)
}
```

```{r}
spatial_scores_xdp(xdp_recon_neurons, "C_minus_scores", "C_minus")
spatial_scores_xdp(xdp_recon_neurons, "C_plus_scores", "C_plus")
```


```{r}
xdp_recon
```

