---
title: "SN Cohort 1 Seurat Object and Clustering Walkthrough"
output: html_notebook
---
Overview: Substantia Nigra (SN) for XDP Cohort 1 has 4 library reactions that should be merged together into one final seurat object. In this tutorial, you will learn to cluster seurat objects, clean them for heterotypic doublets, identify and label different cell classes, and run a MASC analysis to look for potential cell type specific depletion. 


Step 1: Download relevant cellranger (raw feature barcode matrix, molecule info), cellbender (filtered feature barcode matrix), and vireo (donor ids) outputs for each library from GS to your VM. I recommend creating one main folder for all cohort 1 libraries, and create a subfolder for each library named SN_rxn1, SN_rxn2, etc.

The library names for Cohort 1 SN are: 
- pXDPsHSrSNid240830rxn1
- pXDPsHSrSNid240830rxn2
- pXDPsHSrSNid240830rxn3
- pXDPsHSrSNid240830rxn4

1). raw feature barcode matrix: gs://fc-5f122a74-fee3-4ec5-8838-c12344109e6d/01_LIBS__PostCellBender/[LIBRARY_NAME]/outs/raw_feature_bc_matrix.h5

2). molecule info: gs://fc-5f122a74-fee3-4ec5-8838-c12344109e6d/01_LIBS__PostCellBender/[LIBRARY_NAME]/outs/molecule_info.h5 

3). filtered feature barcode matrix after cellbender: gs://fc-5f122a74-fee3-4ec5-8838-c12344109e6d/01_LIBS__PostCellBender/[LIBRARY_NAME]/filtered_feature_bc_matrix/pXDPsHSrSNid240830rxn2/pXDPsHSrSNid240830rxn2_out_filtered.h5 

4). vireo donor ids: gs://fc-5f122a74-fee3-4ec5-8838-c12344109e6d/RESULTS/[LIBRARY_NAME]/05_VIREO/vireo_final_out/donor_ids.tsv 
 
5). Web summary of cellranger (this can be useful to look at for a general sense of library quality and potential overloading): gs://fc-5f122a74-fee3-4ec5-8838-c12344109e6d/01_LIBS__PostCellBender/[LIBRARY_NAME]/outs/web_summary.html 


```{r}
#Load up libraries
library(ggplot2)
library(Seurat)
library(Matrix)
library(rhdf5)
library(dplyr)
library(qs)
```

This code will loop through each library in your list and create a seurat object for each using the cellranger/cellbender/vireo outputs you downloaded. 
```{r}
library_list = list("SN_rxn1", "SN_rxn2", "SN_rxn3", "SN_rxn4")
RAW_COUNTS_BASENAME = "raw_feature_bc_matrix.h5"
FILTERED_COUNTS_BASENAME = "out_filtered.h5"
MOLECULE_INFO_BASENAME = "molecule_info.h5"
DONOR_ID_BASENAME = "donor_ids.tsv"
this_base_path = "/broad/macosko/kimkathl/Cohort1/SN" #change this to your folder

raw_dgc_list = list()
filtered_dgc_list = list()
mol_df_list = list()
donor_ids_list = list()
metadata_df_list = list()
seurat_object_list = list()


#helper function for cellbender loading
sum_duplicate_rownames_of_dgc_matrix=function(dgc){   
    # some umi_data has repeated rownames which causes problems down the road
    # we will sum the umi counts in these rows
    # collect rows with duplicate rownames as a dataframe
    n_occur = data.frame(table(rownames(dgc)))
    repeated_rownames = n_occur[n_occur$Freq > 1,]$Var1
    duplicated_dgc = as.data.frame(dgc[rownames(dgc) %in% repeated_rownames,])
    duplicated_dgc['symbol'] = sub("\\.\\d$", "", rownames(duplicated_dgc))

    # sum across numeric columns
    numeric_cols = setdiff(colnames(duplicated_dgc), "symbol")
    summed_duped_umi_rows = duplicated_dgc %>%
        group_by(symbol) %>%
        summarise_at(vars(numeric_cols), sum, na.rm = TRUE) %>%
        ungroup() %>%
        as.data.frame()
    # set rownames to be symbol column, then remove it
    rownames(summed_duped_umi_rows) = summed_duped_umi_rows$symbol
    summed_duped_umi_rows = summed_duped_umi_rows[, !(names(summed_duped_umi_rows) %in% c('symbol'))]

    # Hack
    # each round of the loop will remove ONE of each duplicate row 
    for(i in 1:max(n_occur$Freq)){
        dgc = dgc[!rownames(dgc) %in% repeated_rownames,]
    }

    # finally, replace the deduped rows with the summed rows via an rbind
    dgc = rbind(dgc, as(summed_duped_umi_rows, "sparseMatrix"))

    return(dgc)
}

  # Read your data
for (name in library_list) {
  
cat(name)
cat("\n")

raw_counts_path = file.path(this_base_path, name, RAW_COUNTS_BASENAME)
filtered_counts_path = file.path(this_base_path, name, paste0("pXDPsHSrSNid240830_" , name, "_", FILTERED_COUNTS_BASENAME))
molecule_info_path = file.path(this_base_path, name, MOLECULE_INFO_BASENAME)
donor_id_path = file.path(this_base_path, name, DONOR_ID_BASENAME)

h5ls(filtered_counts_path)

file_path = filtered_counts_path # the path to the *_out_filtered.h5 file
    h5_data = h5read(file_path, "/matrix")
    counts = h5_data$data
    indices = h5_data$indices
    indptr = h5_data$indptr
    barcodes = h5_data$barcodes
    num_genes = length(h5_data$features$name)  # Number of genes
    num_cells = length(barcodes)  # Number of cells

    counts_matrix = sparseMatrix(i=indices + 1,  # +1 for 1-based indexing in R
                                    p=indptr,
                                    x=counts,
                                    dims=c(num_genes, num_cells))
    
 
    rownames(counts_matrix) = h5_data$features$name # set rownames to be genes
    colnames(counts_matrix) = barcodes # set colnames to be cell barcodes
    counts_matrix
    filtered_dgc = counts_matrix
    
       filtered_dgc_list[[name]] = filtered_dgc
       
       
print("READING RAW DGC .H5")
   raw_counts = Read10X_h5(raw_counts_path)
   raw_dgc_list[[name]] = raw_counts

   cd = data.frame(
       row.names=colnames(raw_counts),
       nUMI=colSums(raw_counts),
       log10_nUMI = log10(colSums(raw_counts) + 1),
       nGene=colSums(raw_counts > 0),
       is_in_filtered=colnames(raw_counts) %in% colnames(filtered_dgc_list[[name]])
   )

# load the molecule data such that you can map the barcodes to what came out of CellRanger
   # load umi_type to get whether or not the umi was intronic
   h5fetch = function(x){return(h5read(molecule_info_path, x))}
   mol_df = data.frame(
               barcode=h5fetch("barcodes")[h5fetch("barcode_idx")+1] %>% paste0("-1"),
               feature=h5fetch("features/name")[h5fetch("feature_idx")+1],
               umi=h5fetch("umi"),
               umi_type=h5fetch("umi_type"),
               count=h5fetch("count")
       )
      
   # only consider cells that are in the filtered umi data
   mol_df = mol_df[mol_df$barcode %in% colnames(filtered_dgc),]

mol_df_grouped = mol_df %>% group_by(barcode) %>%
       summarize(
           nUmi=n(),
           nRead=sum(count),
           pct_intronic=sum(umi_type==0)/nUmi)
   rownames(mol_df_grouped) = mol_df_grouped$barcode
   mol_df_list[[name]] = mol_df_grouped

donor_ids = read.table(donor_id_path, header=T)
  donor_ids = donor_ids[, c("cell", "donor_id", "prob_max")]
  donor_ids_list[[name]] = donor_ids

mol_df_rxn = mol_df_list[[name]] 
  metadata = merge(donor_ids, mol_df_rxn, by.x = "cell", by.y = "barcode", all = TRUE)
  rownames(metadata) <- metadata$cell
  
  metadata <- metadata[rownames(metadata) %in% colnames(filtered_dgc), ]
metadata

all(colnames(filtered_dgc) == rownames(metadata))
  metadata <- metadata[match(colnames(filtered_dgc), rownames(metadata)), ]
  all(colnames(filtered_dgc) == rownames(metadata))
  
metadata

metadata <- metadata[!is.na(metadata$donor_id), ]
metadata

file_path = file.path(this_base_path, name, paste0("pXDPsHSrSNid240830_" , name, "_", FILTERED_COUNTS_BASENAME))
    h5_data = h5read(file_path, "/matrix")
    counts = h5_data$data
    indices = h5_data$indices
    indptr = h5_data$indptr
    barcodes = h5_data$barcodes
    num_genes = length(h5_data$features$name)  # Number of genes
    num_cells = length(barcodes)  # Number of cells

    counts_matrix = sparseMatrix(i=indices + 1,  # add 1 for 1-based indexing in R
                                    p=indptr,
                                    x=counts,
                                    dims=c(num_genes, num_cells))
    

    rownames(counts_matrix) = h5_data$features$name # set rownames to be genes
    colnames(counts_matrix) = barcodes # set colnames to be cell barcodes

# now we want to look only at cells that are in BOTh the CellRanger and CellBender objects
    # use filtered_dgc from earlier in the loop
    cellranger_filtered = filtered_dgc[, colnames(filtered_dgc) %in% all_of(colnames(counts_matrix))]
    cellranger_filtered

counts_matrix_filtered = counts_matrix[, colnames(counts_matrix) %in% all_of(colnames(cellranger_filtered))]
    counts_matrix_filtered
    print(dim(counts_matrix_filtered))
    # for whatever reason, there can be duplicated rownames (i.e. genes) in the cellbender output, causing errors later on. Sum these together.
    counts_matrix_filtered = sum_duplicate_rownames_of_dgc_matrix(counts_matrix_filtered) 
    print(dim(counts_matrix_filtered))
    counts_matrix_filtered = counts_matrix_filtered[, match(colnames(cellranger_filtered), colnames(counts_matrix_filtered))]
    print(dim(counts_matrix_filtered))

# create your metadata here....
metadata = metadata[rownames(metadata) %in% colnames(counts_matrix_filtered), ]
metadata

metadata_df_list[[name]] = metadata 
    # add metadata argument to CreateSeuratObject

# Compare the first few entries directly
head(colnames(counts_matrix_filtered))
head(metadata$cell)

counts_matrix_filtered
metadata

extra_barcodes <- setdiff(colnames(counts_matrix_filtered), metadata$cell)
length(extra_barcodes)  # Number of extra barcodes
extra_barcodes  # List of extra barcodes
counts_matrix_filtered <- counts_matrix_filtered[, !(colnames(counts_matrix_filtered) %in% extra_barcodes)]

# Sort both and compare
all(sort(colnames(counts_matrix_filtered)) == sort(metadata$cell))
# Check types
str(colnames(counts_matrix_filtered))
str(metadata$cell)
# Convert both to character if needed
metadata$cell <- as.character(metadata$cell)
colnames(counts_matrix_filtered) <- as.character(colnames(counts_matrix_filtered))
# Check unique values
unique(colnames(counts_matrix_filtered))[1:10]  # First 10 in counts_matrix
unique(metadata$cell)[1:10]  # First 10 in metadata
metadata_ordered <- metadata[match(colnames(counts_matrix_filtered), metadata$cell), ]

cellbender_sobj = CreateSeuratObject(counts_matrix_filtered, meta.data = metadata)

seurat_object_list[[name]] = cellbender_sobj
}
seurat_object_list

qsave(seurat_object_list, "Cohort1_SN_list.qs") #save
```

```{r}
Cohort_1_SN_rxn1 = seurat_object_list[["SN_rxn1"]]
Cohort_1_SN_rxn2 = seurat_object_list[["SN_rxn2"]]
Cohort_1_SN_rxn3 = seurat_object_list[["SN_rxn3"]]
Cohort_1_SN_rxn4 = seurat_object_list[["SN_rxn4"]]

Cohort_1_SN_rxn1 #should have 14497 cells
Cohort_1_SN_rxn2 #should have 16234 cells
Cohort_1_SN_rxn3 #should have 16152 cells
Cohort_1_SN_rxn4 #should have 19377 cells
```

```{r}
rxn_names <- c("Cohort_1_SN_rxn1", "Cohort_1_SN_rxn2", "Cohort_1_SN_rxn3", "Cohort_1_SN_rxn4")

for (name in rxn_names) {

  obj <- get(name)
  
  # Remove cells that vireo identified as doublet/unassigned
  obj <- subset(obj, subset = donor_id != "doublet" & donor_id != "unassigned")
  
  # Convert pct_intronic to percentage
  obj@meta.data$pct_intronic <- obj$pct_intronic * 100
  
  # Calculate percent mitochondrial content
  mito_genes <- grep(pattern = "^MT-", x = rownames(obj), value = TRUE)
  mito_counts <- Matrix::colSums(GetAssayData(obj, assay = "RNA", slot = "counts")[mito_genes, ])
  total_counts <- Matrix::colSums(GetAssayData(obj, assay = "RNA", slot = "counts"))
  obj$pct_mito <- (mito_counts / total_counts) * 100
  
  assign(name, obj)
}

```

Now we'll merge SN libraries together
```{r}
sobj_list = list("SN_rxn1" = Cohort_1_SN_rxn1,
                 "SN_rxn2" =Cohort_1_SN_rxn2, 
                 "SN_rxn3" = Cohort_1_SN_rxn3, 
                 "SN_rxn4"= Cohort_1_SN_rxn4)

for(name in names(sobj_list)){
     sobj =sobj_list[[name]]
     sobj$library = name
     sobj = RenameCells(sobj, new.name = paste(sobj$library, colnames(sobj), sep="__"))
     sobj_list[[name]] = sobj
     
}

sobj_list

counts_list <- lapply(sobj_list, function(sobj) GetAssayData(sobj, slot = "counts"))


merged_counts <- do.call(cbind, counts_list)

metadata_list <- lapply(sobj_list, function(sobj) sobj@meta.data)
metadata_list

merged_metadata <- do.call(rbind, metadata_list)
merged_metadata
rownames(merged_metadata) <- unlist(lapply(metadata_list, rownames))
merged_metadata

stopifnot(all(rownames(merged_metadata) == colnames(merged_counts)))

# Create Seurat object
SN_Cohort1 <- CreateSeuratObject(counts = merged_counts, meta.data = merged_metadata)
SN_Cohort1
SN_Cohort1@meta.data$Region = "SN"
SN_Cohort1@meta.data 
```

```{r}
SN_Cohort1 #there should be 56166 cells all together

filtered_SN_Cohort1 = subset(SN_Cohort1, subset = pct_mito < 10)
filtered_SN_Cohort1
filtered_SN_Cohort1 = subset(filtered_SN_Cohort1, subset = pct_intronic >= 25) 
filtered_SN_Cohort1
filtered_SN_Cohort1 = subset(filtered_SN_Cohort1, subset = nUmi > 500)
filtered_SN_Cohort1
filtered_SN_Cohort1 = subset(filtered_SN_Cohort1, subset = prob_max > 0.9) 

filtered_SN_Cohort1 #there should be 52283 cells all together after filtering
```

Now add additional donor level metadata (such as condition, age of death, etc). You can download a .csv file with metadata for all donors included in cohorts 1-3 here: gs://macosko_data/XDP/XDP_Cohorts_1_2_3_metadata.csv

```{r}
cohort_metadata = read.csv("XDP_Cohorts_1_2_3_metadata.csv", header = T)
cohort_metadata

meta <- filtered_SN_Cohort1@meta.data %>%
  mutate(cell_id = rownames(.))

cohort_metadata_joined <- meta %>%
  left_join(cohort_metadata, by = c("donor_id" = "Donor"))

rownames(cohort_metadata_joined) <- cohort_metadata_joined$cell_id
cohort_metadata_joined$cell_id <- NULL  

filtered_SN_Cohort1@meta.data <- cohort_metadata_joined
filtered_SN_Cohort1@meta.data 
```

Now we are going to cluster our seurat object

You can also follow along this presentation for visual images for clustering and cleaning steps: https://docs.google.com/presentation/d/1B2KFKWiq59tlFCxHuIbjfPIBaqUO7imUqdbU7wRJZaA/edit?slide=id.g360b5160763_0_0#slide=id.g360b5160763_0_0 

```{r}
# this code should take 3-10 minutes
hvgs = getSeuratVarFeatureIntersectByCol(filtered_SN_Cohort1, subset_col="donor_id", original_nfeatures=2500)
n_dims_use=20
filtered_SN_Cohort1 = (filtered_SN_Cohort1
   %>% NormalizeData() # log normalizes raw counts
   %>% ScaleData(features=hvgs, split.by="donor_id") # within each cell, for each gene scale the data (i.e. subtract the donor's mean and divide by the donor's standard deviation of that gene)
   %>% RunPCA(features=hvgs, npcs=n_dims_use) # Reduce the dimensions to the n_dims_use dimensions that best explain the data https://en.wikipedia.org/wiki/Principal_component_analysis
   %>% FindNeighbors(dims = 1:n_dims_use) # Finds every cells closest neighbors in the PCA space
   %>% FindClusters(resolution = 0.2) # finds clusters at a variety of resolutions
   %>% FindClusters(resolution = 0.3) # after clustering, the cluster labels are accessible via filtered_Cohort1$RNA_snn_res.{resolution}
   %>% FindClusters(resolution = 0.4) # e.g.: filtered_Cohort1$RNA_snn_res.0.4 here
   %>% FindClusters(resolution = 0.5)
   %>% FindClusters(resolution = 0.6)
   %>% FindClusters(resolution = 0.7)
   %>% FindClusters(resolution = 0.8)
   %>% FindClusters(resolution = 0.9)
   %>% FindClusters(resolution = 1)
   %>% RunUMAP(dims = 1:n_dims_use) # projects the PCA onto 2 dimensions, such that both "global structure" and "local structure" are optimally preserved
)

setwh = function(w, h){
   # set the width of the ensuing figures to w, height to h
   options(repr.plot.width = w, repr.plot.height = h)
}

#With these Dimplots, you can see how cells cluster with different metadata and at different resolutions. You might see some donor-specific clustering happening, but it doesn't matter too much at this step since we'll be starting off with a very "crude" manual labelling.  

DimPlot(filtered_SN_Cohort1, group.by = "donor_id") 
DimPlot(filtered_SN_Cohort1, group.by = "Sex")
DimPlot(filtered_SN_Cohort1, group.by = "Condition")
DimPlot(filtered_SN_Cohort1, group.by = "library") 
DimPlot(filtered_SN_Cohort1, group.by = "RNA_snn_res.0.2", label=T) 
DimPlot(filtered_SN_Cohort1, group.by = "RNA_snn_res.0.3", label=T) 
DimPlot(filtered_SN_Cohort1, group.by = "RNA_snn_res.0.4", label=T) 
DimPlot(filtered_SN_Cohort1, group.by = "RNA_snn_res.0.5", label=T) 
DimPlot(filtered_SN_Cohort1, group.by = "RNA_snn_res.0.6", label=T) 
DimPlot(filtered_SN_Cohort1, group.by = "RNA_snn_res.0.7", label=T) 
DimPlot(filtered_SN_Cohort1, group.by = "RNA_snn_res.0.8", label=T)
```


You can look at specific marker genes and how they are expressed in the seurat object to help identify which clusters belong to which cell class. Right now, we'll be clustering just the major cell types: neuron, astrocyte, oligo, opc, microglia, endothelial. Depending on the brain region and how high your resolution is, you may also be able to identify smaller classes like immune and ependymal at this stage. 

For choosing a good resolution, I tend to stay on the lower end (0.2 - 0.4). How I personally choose it is if I can identify an immune cluster based on the marker genes, I will pick the lowest resolution that can separate out the immune cells as its own cluster. For this tutorial, I chose 0.2, but play around with other resolutions!

```{r}
#neuron marker genes (SYT1 and RBFOX3 are the main ones. Some clusters have both, others can be one or the other)
FeaturePlot(filtered_SN_Cohort1, features = c("SYT1", "RBFOX3", "GAD2", "SLC17A6"), raster = F) 

#astrocyte marker genes (AQP3 and GFAP are main ones)
FeaturePlot(filtered_SN_Cohort1, features = c("AQP4", "GINS3", "GFAP"), raster = F) 

#microglia marker genes
FeaturePlot(filtered_SN_Cohort1, features = c("C1QA", "C1QB", "CX3CR1", "P2RY12"), raster = F) 

#endothelial marker genes (you might see FLT1 and DCN expression on opposites ends of the endothelial cluster)
FeaturePlot(filtered_SN_Cohort1, features = c("FLT1", "DCN", "RGS5"), raster = F) 

#oligodendrocyte marker genes (MOG and MOBP are the main ones I look at. OLIG1 and 2 can be shared with OPCs)
FeaturePlot(filtered_SN_Cohort1, features = c("OLIG1","OLIG2", "MOG", "MOBP"), raster = F) 

#OPC marker genes (VCAN is often also expressed in some astrocytes, but the opc cluster should be obvious)
FeaturePlot(filtered_SN_Cohort1, features = c("VCAN"), raster = F)

#ependymal marker genes (often ependymal cluster hangs off the astrocyte cluster. This particular seurat object might not have very obvious ependymal cluster at this stage)
FeaturePlot(filtered_SN_Cohort1, features = c("ZBBX", "CFAP157", "CFAP299"), raster = F) 

#immune marker genes (often the immune cluster is either a completely separate dot OR sometimes it hangs off of the microglia cluster)
FeaturePlot(filtered_SN_Cohort1, features = c("CD96", "NKG7", "SKAP1"), raster = F) 

#general QC - pct_mito should cluster together within each cell type, nUmi are often much higher in neurons
FeaturePlot(filtered_SN_Cohort1, features = c("pct_mito", "pct_intronic", "nUmi"),raster=FALSE) 

#Bonus: For SN specifically, there are a subpopulation of neurons that are interesting called dopaminergic neurons which express TH. This however doesn't cluster out very cleaning in my experience... 
FeaturePlot(filtered_SN_Cohort1, features = c("TH"),raster=FALSE) 
```


Now we'll label our clusters based on these featureplots
```{r}
Idents(filtered_SN_Cohort1) = filtered_SN_Cohort1$RNA_snn_res.0.2

#First check how many clusters your chosen resolution has. Now assign each cluster a number and run this code block. After seeing which number each cluster is associated with, you can manually change the labels to the appropriate cell class.

classes = c("0", "1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12", "13", "14", "15", "16", "17", "18", "19", "20")

#Example: classes = c("oligo", "microglia", "astrocyte", "oligo", "opc", "neuron", "oligo", "oligo", "oligo", "endothelial", "oligo", "neuron", "neuron", "neuron", "microglia", "immune", "oligo", "neuron", "microglia", "endothelial", "oligo")

filtered_SN_Cohort1= assignCellClasses(filtered_SN_Cohort1, classes=classes, cluster_col="RNA_snn_res.0.2", class_col = "cell_class")

Idents(filtered_SN_Cohort1) = "cell_class"

DimPlot(filtered_SN_Cohort1, group.by = "cell_class" ,label = T, raster = FALSE)
```


This is an additional way to visualize which clusters express which marker genes
```{r}
marker_genes <- c("SYT1", "RBFOX3", "GAD2", "SLC17A6","AQP4", "GINS3", "GFAP","C1QA", "C1QB", "CX3CR1", "P2RY12","FLT1", "DCN", "RGS5", "OLIG1", "MOG", "MOBP", "OLIG2", "VCAN","ZBBX", "CFAP157", "CFAP299","CD96", "NKG7", "SKAP1")

Dotplot = DotPlot(object = filtered_SN_Cohort1, features = marker_genes, group.by = "RNA_snn_res.0.2")
Dotplot  <- Dotplot  + theme(axis.text.x = element_text(angle = 45, hjust = 1))+ coord_flip() 
print(Dotplot)
```


Bonus: If you cannot identify a cluster based on these marker genes, you can also examine individual clusters by running FindMarkers 
```{r}
#Example: this will find all markers from cluster 2. You can then look for genes that have a high pct.1 and low pct.2 (indictative of a good marker gene) and google these genes to see if they are associated with a particular cell type. 
cluster2.markers <- FindMarkers(filtered_SN_Cohort1, ident.1 = 2)
cluster2.markers
```


SN cleaning + integration
```{r}
#This will list out all cell class and how many of each 
table(filtered_SN_Cohort1$cell_class)
```

For each major cell class, you should create a temporary seurat object that subsets out just that cell class. Below is an example for just the astrocytes, but this should be done for all cell classes you labeled originally (A very tedious process sorry...)


```{r}
temp_sobj = subset(filtered_SN_Cohort1, subset = cell_class == "astrocyte")

#After subsetting, this will run through nearly the exact same steps as clustering the larger seurat object. By doing subclustering on individual cell classes, you will be able to separate out clusters containing heterotypic doublets that you weren't able to see in the original seurat object.  

hvgs = getSeuratVarFeatureIntersectByCol(temp_sobj, subset_col="donor_id", original_nfeatures=2500)
n_dims_use=20
temp_sobj = (temp_sobj
   %>% NormalizeData() 
   %>% ScaleData(features=hvgs, split.by="donor_id") 
   %>% RunPCA(features=hvgs, npcs=n_dims_use) 
   %>% FindNeighbors(dims = 1:n_dims_use) 
   %>% FindClusters(resolution = 0.2) 
   %>% FindClusters(resolution = 0.3) 
   %>% FindClusters(resolution = 0.4) 
   %>% FindClusters(resolution = 0.5)
   %>% FindClusters(resolution = 0.6)
   %>% FindClusters(resolution = 0.7)
   %>% FindClusters(resolution = 0.8)
   %>% FindClusters(resolution = 0.9)
   %>% FindClusters(resolution = 1)
   %>% RunUMAP(dims = 1:n_dims_use) 
)

setwh = function(w, h){
   options(repr.plot.width = w, repr.plot.height = h)
}

#This will integrate your dataset based on donors, that way you can ensure that single donors aren't clustering together based on donor identity

library(harmony)
temp_sobj = (temp_sobj
    %>% RunHarmony(
        group.by = "donor_id")
    %>% FindNeighbors(reduction='harmony', dims=1:20)
    %>% FindClusters(res=0.05)
    %>% FindClusters(res=0.1)
    %>% FindClusters(res=0.2)
    %>% FindClusters(res=0.3)
    %>% FindClusters(res=0.4)
    %>% FindClusters(res=0.5)
    %>% FindClusters(res=0.6)
    %>% FindClusters(res=0.7)
    %>% FindClusters(res=0.8)    
    %>% RunUMAP(reduction="harmony", dims=1:20)
)

#Here you can pick the resolution you want to work with and look at how each marker gene is expressed.

#On the UMAPs, you'll usually see the real doublets clusters as far away "islands" while non-doublet cells often cluster together as one cloud. I am a lot more confident that islands with coexpressing marker genes are doublets and will remove them, but would be more lenient to potential coexpression happening in this main cloud.  

DimPlot(temp_sobj, group.by = "donor_id") 
DimPlot(temp_sobj, group.by = "RNA_snn_res.0.2", label=T) 
DimPlot(temp_sobj, group.by = "RNA_snn_res.0.3", label=T) 
DimPlot(temp_sobj, group.by = "RNA_snn_res.0.4", label=T) 
DimPlot(temp_sobj, group.by = "RNA_snn_res.0.5", label=T) 
DimPlot(temp_sobj, group.by = "RNA_snn_res.0.6", label=T) 
DimPlot(temp_sobj, group.by = "RNA_snn_res.0.7", label=T) 
DimPlot(temp_sobj, group.by = "RNA_snn_res.0.8", label=T) 

#Ideally you should only see astrocyte markers strongly expressed, but in most cases, you'll see clusters that are also expressing marker genes they should not be expressing. The most popular ones you may see here are MOG and SYT1 (since there are more oligos and neurons) coexpressing with astrocyte marker genes. 

FeaturePlot(temp_sobj, features = c("SYT1", "RBFOX3", "GAD2", "SLC17A6"), raster = F) FeaturePlot(temp_sobj, features = c("AQP4", "GINS3", "GFAP"), raster = F) 
FeaturePlot(temp_sobj, features = c("C1QA", "C1QB", "CX3CR1", "P2RY12"), raster = F) 
FeaturePlot(temp_sobj, features = c("FLT1", "DCN", "RGS5"), raster = F) 
FeaturePlot(temp_sobj, features = c("OLIG1","OLIG2", "MOG", "MOBP"), raster = F) 
FeaturePlot(temp_sobj, features = c("VCAN"), raster = F)
FeaturePlot(temp_sobj, features = c("ZBBX", "CFAP157", "CFAP299"), raster = F) 
FeaturePlot(temp_sobj, features = c("CD96", "NKG7", "SKAP1"), raster = F) 
FeaturePlot(temp_sobj, features = c("pct_mito", "pct_intronic", "nUmi"),raster=FALSE) 
```

```{r}
#I personally prefer using this plot to visualize which clusters have doublets (they are a lot more obvious here) 

marker_genes <- c("SYT1", "RBFOX3", "GAD2", "SLC17A6","AQP4", "GINS3", "GFAP","C1QA", "C1QB", "CX3CR1", "P2RY12","FLT1", "DCN", "RGS5", "OLIG1", "MOG", "MOBP", "OLIG2", "VCAN","ZBBX", "CFAP157", "CFAP299","CD96", "NKG7", "SKAP1", "TH")

Dotplot = DotPlot(object = temp_sobj, features = marker_genes, group.by = "RNA_snn_res.0.2")
Dotplot  <- Dotplot  + theme(axis.text.x = element_text(angle = 45, hjust = 1))+ coord_flip() 
print(Dotplot)
```

```{r}
#Once you've identified which clusters are doublets, you can remove them but subsetting them out
temp_sobj
temp_sobj = subset(temp_sobj, subset = RNA_snn_res.0.2 != "4") 
temp_sobj
DimPlot(temp_sobj, group.by = "RNA_snn_res.0.2")

#make sure to keep track of how many cells you've removed for each cell class
```

```{r}
#assign your temporary seurat object to a new object. As you repeat this for other cell classes, you'll be rewriting the temp_sobj object (mostly to save space)
SN_astro = temp_sobj

# SN_endo
# SN_immune
# SN_micro
# SN_oligo
# SN_opc
# SN_neuron
```


Once you have your cleaned individual cell class seurat object, we can remerge them back together into one cleaned object
```{r}
sobj_list = list(SN_astro,
SN_endo,
SN_immune,
SN_micro,
SN_oligo,
SN_opc,
SN_neuron)

counts_list <- lapply(sobj_list, function(sobj) GetAssayData(sobj, slot = "counts"))

merged_counts <- do.call(cbind, counts_list)

metadata_list <- lapply(sobj_list, function(sobj) sobj@meta.data)

merged_metadata <- do.call(rbind, metadata_list)
merged_metadata

stopifnot(all(rownames(merged_metadata) == colnames(merged_counts)))

# Create Seurat object
cleaned_SN_Cohort1 <- CreateSeuratObject(counts = merged_counts, meta.data = merged_metadata)
cleaned_SN_Cohort1

hvgs = getSeuratVarFeatureIntersectByCol(cleaned_SN_Cohort1, subset_col="donor_id", original_nfeatures=2500)
n_dims_use=20
cleaned_SN_Cohort1 = (cleaned_SN_Cohort1
   %>% NormalizeData() 
   %>% ScaleData(features=hvgs, split.by="donor_id") 
   %>% RunPCA(features=hvgs, npcs=n_dims_use) 
   %>% FindNeighbors(dims = 1:n_dims_use) 
   %>% FindClusters(resolution = 0.2) 
   %>% FindClusters(resolution = 0.3) 
   %>% FindClusters(resolution = 0.4) 
   %>% FindClusters(resolution = 0.5)
   %>% FindClusters(resolution = 0.6)
   %>% FindClusters(resolution = 0.7)
   %>% FindClusters(resolution = 0.8)
   %>% FindClusters(resolution = 0.9)
   %>% FindClusters(resolution = 1)
   %>% RunUMAP(dims = 1:n_dims_use) 
)

setwh = function(w, h){
   options(repr.plot.width = w, repr.plot.height = h)
}

library(harmony)
cleaned_SN_Cohort1 = (cleaned_SN_Cohort1
    %>% RunHarmony(
        group.by = "donor_id")
    %>% FindNeighbors(reduction='harmony', dims=1:20)
    %>% FindClusters(res=0.05)
    %>% FindClusters(res=0.1)
    %>% FindClusters(res=0.2)
    %>% FindClusters(res=0.3)
    %>% FindClusters(res=0.4)
    %>% FindClusters(res=0.5)
    %>% FindClusters(res=0.6)
    %>% FindClusters(res=0.7)
    %>% FindClusters(res=0.8)    
    %>% RunUMAP(reduction="harmony", dims=1:20)
)

DimPlot(cleaned_SN_Cohort1, group.by = "donor_id") 
DimPlot(cleaned_SN_Cohort1, group.by = "library") 
DimPlot(cleaned_SN_Cohort1, group.by = "Condition") 
DimPlot(cleaned_SN_Cohort1, group.by = "RNA_snn_res.0.2", label=T) 
DimPlot(cleaned_SN_Cohort1, group.by = "RNA_snn_res.0.3", label=T) 
DimPlot(cleaned_SN_Cohort1, group.by = "RNA_snn_res.0.4", label=T) 
DimPlot(cleaned_SN_Cohort1, group.by = "RNA_snn_res.0.5", label=T) 
DimPlot(cleaned_SN_Cohort1, group.by = "RNA_snn_res.0.6", label=T) 
DimPlot(cleaned_SN_Cohort1, group.by = "RNA_snn_res.0.7", label=T) 
DimPlot(cleaned_SN_Cohort1, group.by = "RNA_snn_res.0.8", label=T)
DimPlot(cleaned_SN_Cohort1, group.by = "cell_class", label=T)

qsave(cleaned_SN_Cohort1, "cleaned_SN_Cohort1.qs") #save 
```


MASC analysis: https://pmc.ncbi.nlm.nih.gov/articles/PMC6448773/
This analysis will test how a covariate (in this case, case-control) influences the amount of each cell type after controlling for any technical or biological variation
```{r}
library(lme4)
sobj = cleaned_SN_Cohort1
model_cols = c("donor_id", "cell_class", "Condition", "Sex", "Age_of_Death") #include any other covariates of interest, replace cell_class with whatever column defines the cluster annotations
df = sobj@meta.data[,model_cols]
df = df[complete.cases(df),] # don't want any NAs
df$case_control_factor = as.factor(df$Condition)

masc_df = .sconline.MASCfn(
    dataset=df,
    cluster=df$cell_class, # cluster annotations
    contrast="case_control_factor", # name of contrast annotations (what you want to run the test for)
    random_effects="donor_id", # name of random effects annotations (not interested in these coefficients, but account for variability in probable sources ob batch effects, in this case donor
    fixed_effects = c("Age_of_Death", "Sex") # your covariates
)

masc_df

my_order <- unique(masc_df$cluster)
masc_df$cluster_name <- sub("cluster", "", masc_df$cluster) # for legibility
masc_df$log2_or = log2(masc_df$case_control_factorXDP.OR) # the names of the case_control_factor<whatever> columns will change from project to project depending on the unique values of your Condition column
masc_df$log2_or_ci_low = log2(masc_df$case_control_factorXDP.OR.95pct.ci.lower)
masc_df$log2_or_ci_high = log2(masc_df$case_control_factorXDP.OR.95pct.ci.upper)

# order the graph to put disease-enriched populations on top
masc_df = masc_df[order(-masc_df$log2_or), ] 
masc_df$cluster_name = factor(masc_df$cluster_name, levels = masc_df$cluster_name[order(masc_df$log2_or)])
masc_df


library(RColorBrewer)
library(ggplot2)

# Create the forest plot with ticks on error bars, axis lines with ticks, RdBu color map, and opaque white circles on top
ggplot(masc_df, aes(y = cluster_name, x = log2_or)) +
  ggtitle("XDP vs. Control: SN Cohort 1") +
  geom_vline(xintercept = 0, linetype = "dotted", color = "gray") +  # Add dotted vertical line at x=0
  geom_segment(aes(x = log2_or_ci_low, xend = log2_or_ci_high, y = cluster_name, yend = cluster_name, color = log2_or), size = 1) +  # Add horizontal error bars
  geom_point(size = 3, aes(color = log2_or), shape = 1) +  # Add points for effect sizes
  geom_point(size = 3, shape = 21, fill = "white") +  # Add opaque white circle on top of the error bar line
  scale_color_gradientn(colors = RColorBrewer::brewer.pal(10, "RdBu")) +  # Use RdBu color map
  theme_minimal() +  # Minimal theme
  labs(x = "log2(OR)", y = "Cell Classes") +  # Axis labels
  theme(
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(), 
    legend.position = "none",
    plot.title = element_text(size=16),
    axis.line = element_line(color = "black"),  # Add axis lines
    axis.ticks = element_line(color = "black"),  # Add axis ticks
    axis.text = element_text(size = 14),  # Increase tick label font size
    axis.title = element_text(size = 15)  # Increase axis label font size
  )
```

