---
title: "R Notebook"
output: html_notebook
---

```{r}
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install("fgsea")
```


```{r}
library(Seurat)
library(dplyr)
library(qs)

library(Matrix)
library(fgsea)


# when concatenated together, these should be the paths to your DE csv files
BASE_PATH="~/GSEA"
DE_SUBDIR="de/filtered_merged_putamen_clustered_clean_pseudocells__split_by_cell_class__grouped_by_donor_id__mean_size_30"
#DE_SUBDIR="de/filtered_merged_caudate_clustered_clean_pseudocells__split_by_cell_class__grouped_by_donor_id__mean_size_30"

.extraHumanGeneAnnoAdderFn=function(inputGeneNames=NULL){
  #require(EnsDb.Hsapiens.v75)
  require(EnsDb.Hsapiens.v86)
  
  if(!dir.exists("~/serverFiles")){
    dir.create("~/serverFiles",recursive = T)
  }
  
  gns <- as.data.frame(genes(EnsDb.Hsapiens.v86))
  gns$gene_short_name=gns$gene_name
  gns$symbol=toupper(gns$symbol)
  gns$ensembl_gene_id=row.names(gns)
  
  if(!is.null(inputGeneNames)){
    rwNames=toupper(inputGeneNames)
    psCols=c("gene_short_name","ensembl_gene_id")
    slCounts=0
    slCol=""
    if(sum(grepl("\\.",rwNames)&grepl("^ENS",rwNames))>0){
      rwNames=strsplit(rwNames,"\\.")
      rwNames=unlist(lapply(rwNames,function(x)x[1]))
    }
    system(paste0("gsutil -m cp gs://fc-71ac3b81-2441-4171-8038-baf653634620/serverFiles/human_map_to_ensembl.rda ."))
    load("human_map_to_ensembl.rda")
    
    
    map_to_ensmbl$source=toupper(map_to_ensmbl$source)
    
    system(paste0("gsutil -m cp gs://fc-71ac3b81-2441-4171-8038-baf653634620/serverFiles/human_mapping_hg19.rda ."))
      
    
    load("human_mapping_hg19.rda")
    human_hg19$source=toupper(human_hg19$source)
    
    if(sum(toupper(rwNames) %in% human_hg19$source) > sum(toupper(rwNames) %in% map_to_ensmbl$source)){
      map_to_ensmbl=merge(human_hg19,data.frame(source=toupper(rwNames),stringsAsFactors = F),by="source",all.y=T)
    } else {
      map_to_ensmbl=merge(map_to_ensmbl,data.frame(source=toupper(rwNames),stringsAsFactors = F),by="source",all.y=T)
    }
    
    gns=merge(gns,map_to_ensmbl,by.x="ensembl_gene_id",by.y="target",all.y=T)
    gns=gns[match(rwNames,gns$source),]
    row.names(gns)=inputGeneNames
    gns$gene_id=inputGeneNames
    gns=gns[,-which(colnames(gns) %in% c("source","target"))]
  }
  
  return(gns)
}

.sconline.GSEA.readGMT=function (file,bkg_genes=NULL,min.gs.size=NULL,max.gs.size=NULL) {
  if (!grepl("\\.gmt$", file)[1]&F) {
    stop("Pathway information must be in a .gmt file format")
  }
  geneSetDB = readLines(file)
  geneSetDB = strsplit(geneSetDB, "\t")
  names(geneSetDB) = sapply(geneSetDB, "[", 1)
  geneSetDB = lapply(geneSetDB, "[", -1:-2)
  geneSetDB = lapply(geneSetDB, function(x) {
    x[which(x != "")]
  })
  
  if(!is.null(bkg_genes)){
    for(i in 1:length(geneSetDB)){
      tmp=geneSetDB[[i]]
      tmp=bkg_genes[which(toupper(bkg_genes) %in% toupper(tmp))]
      geneSetDB[[i]]=tmp
    }
  }
  
  if(!is.null(min.gs.size)){
    size.dist=unlist(lapply(geneSetDB,length))
    geneSetDB=geneSetDB[size.dist>=min.gs.size]
  }
  
  if(!is.null(max.gs.size)){
    size.dist=unlist(lapply(geneSetDB,length))
    geneSetDB=geneSetDB[size.dist<=max.gs.size]
  }
  
  return(geneSetDB)
}

.myfGSEAfn=function(rankedVec,gs,minSize  = 15,maxSize  = 250, scoreType='std'){
  require(fgsea)
  fgseaRes <- fgsea(pathways = gs, 
                    stats    = rankedVec,
                    minSize  = minSize,
                    maxSize  = maxSize,
                    scoreType= scoreType)
  fgseaRes=fgseaRes[order(fgseaRes$pval,decreasing = F),]
  fgseaRes=as.data.frame(fgseaRes)
  fgseaRes$leadingEdge=unlist(lapply(fgseaRes$leadingEdge,function(x) paste(x,collapse = ",")))
  
  return(fgseaRes)
}

runGSEA = function(
    de_df,
    gs_list_of_char,
    rank_col='logFC',
    gene_id_col='gene_short_name',
    desc=FALSE,
    abs=TRUE,
    scoreType='std'){

    de_df=de_df[!duplicated(de_df[[gene_id_col]]),]
    
    order_vector = de_df[[rank_col]]
    if(abs){
        order_vector = abs(order_vector)
    }
    
    ranked_vec=de_df[,rank_col]
    names(ranked_vec)=de_df[[gene_id_col]]
    ranked_vec=ranked_vec[order(order_vector,decreasing = desc)]
    
    print(head(ranked_vec))

    res_fGSEA=.myfGSEAfn(rankedVec=ranked_vec,gs=gs_list_of_char, scoreType=scoreType)
    res_fGSEA=res_fGSEA[order(res_fGSEA$padj,decreasing=F),]
    return(res_fGSEA)
}
```



#example
```{r}
outpath = file.path(BASE_PATH, DE_SUBDIR, "gsea")
# if you're running into permission errors may need to delete directory and remake with mkdir on command line
dir.create(outpath, recursive=TRUE, showWarnings=FALSE)
de_files = list.files(file.path(BASE_PATH, DE_SUBDIR))
de_files = de_files[grep(".csv$", de_files)]

gene_sets = list(
    # add the paths to the gene sets you want to use here
    kegg_2021_human = "GSEA/genesets/KEGG_2021_Human.txt",
    go_process = "GSEA/genesets/GO_Biological_Process_2021.txt",
    go_function = "GSEA/genesets/GO_Molecular_Function_2021.txt",
    disgenet = "GSEA/genesets/DisGeNET.txt",
    msigdb_hallmark = "GSEA/genesets/MSigDB_Hallmark_2020.txt",
    neural_activity = "GSEA/genesets/neural_activity_arranged.txt",
    wikipathways = "GSEA/genesets/WikiPathways_2019_Human.txt",
    tf_perturbations = "GSEA/genesets/TF_Perturbations_Followed_by_Expression.txt",
    jensen_diseases = "GSEA/genesets/Jensen_DISEASES.txt"
)

for (de_file in de_files){
    de_slogan = sub(".txt$", "", de_file)

    if (! grepl("trend", de_slogan)){next}

    #if (strsplit(de_slogan, "__")[[1]][[2]] != "opc"){next}

    dataDE = read.csv(file.path(BASE_PATH, DE_SUBDIR, de_file))
    anno=.extraHumanGeneAnnoAdderFn(inputGeneNames=dataDE$gene)
    anno=anno[match(dataDE$gene,anno$gene_id),]
    dataDE=cbind(dataDE,anno)
    dataDE=dataDE[which(dataDE$gene_biotype=="protein_coding"),]

    gsea_list = list()

    for (gene_set in names(gene_sets)){
        path = gene_sets[[gene_set]]

        gs=.sconline.GSEA.readGMT(file=path,bkg_genes=dataDE$gene_short_name,min.gs.size=15,max.gs.size=500)
        gsea = runGSEA(dataDE, gs, rank_col="logFC", abs=FALSE, desc=TRUE)
        gsea$gene_set = gene_set
        gsea = gsea[which(gsea$padj<0.05),]

        if (nrow(gsea) > 0){
            gsea_list[[gene_set]] = gsea
        }
    }
    gsea_df = do.call(rbind, gsea_list)

    write.table(gsea_df, file.path(outpath, paste0(de_slogan, "__gsea_logfc.tsv")), sep="\t", quote=FALSE, row.names=FALSE)
}
```


```{r}
# Define cell types
caudate_cell_types <- list("astro", "endo", "ependymal", "immune", "mg", "neuron", "oligo", "opc")
putamen_cell_types <- list("astro", "endo", "mg", "neuron", "oligo", "opc")

# Initialize an empty named list to store the dataframes
#caudate_gsea_df_list <- list()
putamen_gsea_df_list <- list()

# Loop through each cell type and read the corresponding file into a dataframe
for (cell_type in putamen_cell_types) {
  # Construct the file name
  gsea_file_name <- paste("~/GSEA/de/filtered_merged_putamen_clustered_clean_pseudocells__split_by_cell_class__grouped_by_donor_id__mean_size_30/gsea/trend__", cell_type, "__20240529.csv__gsea_logfc.tsv", sep = "")
  
  # Print the file name to verify the path
  print(paste("Trying to read file:", gsea_file_name))
  
  # Check if the file exists
  if (file.exists(gsea_file_name)) {
    # Read the file into a dataframe
    gsea_file <- read.delim(gsea_file_name, header = TRUE)
    
    # Store the dataframe in the list with the cell type as the name
    putamen_gsea_df_list[[cell_type]] <- gsea_file
  } else {
    print(paste("File does not exist:", gsea_file_name))
  }
}

# Accessing the dataframes
# You can access each dataframe by its cell type name, for example:
if ("astro" %in% names(putamen_gsea_df_list)) {
  astro_df <- putamen_gsea_df_list[["astro"]]
}

```


```{r}
# Extract relevant data from GSEA results
data <- gsea_results@result
data$GeneRatio <- sapply(data$GeneRatio, function(x) eval(parse(text=x)))

# Define a new column for NES absolute value for dot circumference
data$NES_abs <- abs(data$NES)

# Define a new column for -log10(p.adjust) for coloring the dots
data$log10_pvalue <- -log10(data$p.adjust)

```



```{r}
df_ordered <- df[order(df$column_name), ]
```

```{r}
caudate_astro_df = caudate_gsea_df_list[["astro"]]
caudate_astro_df
caudate_astro_df = caudate_astro_df[order(caudate_astro_df$NES),]
caudate_astro_df
```


```{r}
caudate_astro_df = caudate_gsea_df_list[["astro"]]
caudate_endo_df= caudate_gsea_df_list[["endo"]]
caudate_ependymal_df= caudate_gsea_df_list[["ependymal"]]
caudate_immune_df =caudate_gsea_df_list[["immune"]]
caudate_microglia_df = caudate_gsea_df_list[["mg"]]
caudate_neuron_df = caudate_gsea_df_list[["neuron"]]
caudate_oligo_df=caudate_gsea_df_list[["oligo"]]
caudate_opc_df=caudate_gsea_df_list[["opc"]]
```

```{r}
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install(version = "3.12")
```


```{r}
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("clusterProfiler")
```

```{r}
# Load necessary libraries
library(clusterProfiler)
# Read GSEA results
gsea_results <- caudate_neuron_df

# Select top 10 pathways by adjusted p-value
top_gsea_results <- gsea_results[order(gsea_results$padj), ][1:10, ]

# Create a bar plot
ggplot(top_gsea_results, aes(x = reorder(Description, NES), y = NES)) +
    geom_bar(stat = "identity", fill = "steelblue") +
    coord_flip() +  # Flip coordinates for better readability
    labs(title = "Top 10 Enriched Pathways",
         x = "Pathway",
         y = "Normalized Enrichment Score (NES)") +
    theme_minimal()

```


```{r}
# Select the top 20 NES
top20_gsea <- caudate_astro_df[order(-abs(caudate_astro_df$NES)), ][1:20, ]
top20_gsea

```


```{r}
# Load ggplot2 library
library(ggplot2)

# Create a bar plot
ggplot(top20_gsea, aes(x = reorder(pathway, NES), y = NES, fill = NES > 0)) +
    geom_bar(stat = "identity") +
    coord_flip() +  # Flip coordinates for better readability
    scale_fill_manual(values = c("red", "blue")) +  # Custom colors for positive and negative NES
    theme_minimal() +
    labs(title = "Top 20 Gene Sets by Normalized Enrichment Score (NES)",
         x = "Gene Set",
         y = "NES",
         fill = "NES > 0")
```

```{r}
# Create a dot plot
ggplot(top20_gsea, aes(x = reorder(pathway, NES), y = NES, color = NES > 0)) +
    geom_point(aes(size = -log10(padj)), shape = 21, stroke = 1.5) +
    coord_flip() +  # Flip coordinates for better readability
    scale_color_manual(values = c("red", "blue")) +  # Custom colors for positive and negative NES
    theme_minimal() +
    labs(title = "Top 20 Gene Sets by Normalized Enrichment Score (NES)",
         x = "Gene Set",
         y = "NES",
         size = "-log10(adjusted p-value)",
         color = "NES > 0")

```



```{r}
merged_caudate_gsea <- intersect(caudate_astro_df$pathway, caudate_neuron_df$pathway)
length(merged_caudate_gsea)
merged_caudate_gsea<- intersect(merged_caudate_gsea, caudate_oligo_df$pathway)
length(merged_caudate_gsea)
merged_caudate_gsea<- intersect(merged_caudate_gsea, caudate_microglia_df$pathway)
length(merged_caudate_gsea)
merged_caudate_gsea<- intersect(merged_caudate_gsea, caudate_opc_df$pathway)
length(merged_caudate_gsea)

```
```{r}
ca =caudate_astro_df
```

```{r}
merged_caudate_gsea
```



```{r}
nrow(caudate_astro_df)
nrow(caudate_endo_df)
nrow(caudate_ependymal_df)
nrow(caudate_immune_df)
nrow(caudate_microglia_df)
nrow(caudate_oligo_df)
nrow(caudate_neuron_df)
nrow(caudate_opc_df)
```





```{r}
caudate_cell_types = list("astro", "endo", "ependymal", "immune", "mg", "neuron", "oligo", "opc")
putamen_cell_types = list("astro", "endo", "mg", "neuron", "oligo", "opc")
caudate_gsea_df_list = list()

for (cell_type in caudate_cell_types){
gsea_file_name = paste("~/GSEA/de/filtered_merged_caudate_clustered_clean_pseudocells__split_by_cell_class__grouped_by_donor_id__mean_size_30/gsea/trend__", cell_type, "__20240529.csv__gsea_logfc.tsv", sep="")

gsea_file = read.delim(gsea_file_name, header = TRUE)
#caudate_gsea_df_list[[cell_type]] = gsea_file
}

```




```{r}
caudate_gsea_df_list[["astro"]]
```



```{r}
df_ordered <- df[order(df$column_name), ]
```

```{r}
caudate_astro_df = caudate_gsea_df_list[["astro"]]
caudate_astro_df
caudate_astro_df = caudate_astro_df[order(caudate_astro_df$NES),]
caudate_astro_df
```


```{r}
caudate_astro_df = caudate_gsea_df_list[["astro"]]
caudate_astro_df = caudate_astro_df[caudate_astro_df$gene_set != "tf_perturbations",]
caudate_astro_df

caudate_endo_df= caudate_gsea_df_list[["endo"]]
caudate_endo_df = caudate_endo_df[caudate_endo_df$gene_set != "tf_perturbations",]
caudate_endo_df

caudate_ependymal_df= caudate_gsea_df_list[["ependymal"]]
caudate_ependymal_df = caudate_ependymal_df[caudate_ependymal_df$gene_set != "tf_perturbations",]
caudate_ependymal_df

caudate_immune_df =caudate_gsea_df_list[["immune"]]
caudate_immune_df = caudate_immune_df[caudate_immune_df$gene_set != "tf_perturbations",]
caudate_immune_df

caudate_microglia_df = caudate_gsea_df_list[["mg"]]
caudate_microglia_df = caudate_microglia_df[caudate_microglia_df$gene_set != "tf_perturbations",]
caudate_microglia_df

caudate_neuron_df = caudate_gsea_df_list[["neuron"]]
caudate_neuron_df = caudate_neuron_df[caudate_neuron_df$gene_set != "tf_perturbations",]
caudate_neuron_df

caudate_oligo_df=caudate_gsea_df_list[["oligo"]]
caudate_oligo_df = caudate_oligo_df[caudate_oligo_df$gene_set != "tf_perturbations",]
caudate_oligo_df

caudate_opc_df=caudate_gsea_df_list[["opc"]]
caudate_opc_df = caudate_opc_df[caudate_opc_df$gene_set != "tf_perturbations",]
caudate_opc_df
```

```{r}
merged_caudate_gsea <- intersect(caudate_oligo_df$pathway, caudate_neuron_df$pathway)
length(merged_caudate_gsea)
merged_caudate_gsea
```


```{r}
merged_caudate_gsea<- intersect(merged_caudate_gsea, caudate_astro_df$pathway)
length(merged_caudate_gsea)
merged_caudate_gsea
```


```{r}
merged_caudate_gsea<- intersect(merged_caudate_gsea, caudate_microglia_df$pathway)
length(merged_caudate_gsea)
merged_caudate_gsea<- intersect(merged_caudate_gsea, caudate_opc_df$pathway)
length(merged_caudate_gsea)
```





```{r}
Dot_plot = function(cell_type_df, title){

top10_gsea_pos <- cell_type_df[order(cell_type_df$NES, decreasing =TRUE), ][1:10, ]
top10_gsea_pos

top10_gsea_neg <- cell_type_df[order(cell_type_df$NES), ][1:10, ]
top10_gsea_neg

top20_gsea <- rbind(top10_gsea_pos, top10_gsea_neg)
top20_gsea

ggplot(top20_gsea, aes(x = reorder(pathway, NES), y = NES, color = NES > 0)) +
    geom_point(aes(size = padj), shape = 21, stroke = 1.5) +
    coord_flip() +  # Flip coordinates for better readability
    scale_color_manual(values = c("red", "blue")) +  # Custom colors for positive and negative NES
    theme_minimal() +
    theme(
        axis.text.x = element_text(size = 14),  # X-axis text size
        axis.text.y = element_text(size = 14),  # Y-axis text size
        axis.title.x = element_text(size = 16),  # X-axis title size
        axis.title.y = element_text(size = 16),  # Y-axis title size
        plot.title = element_text(size = 18, face = "bold"),  # Plot title size and bold
        legend.title = element_text(size = 16),  # Legend title size
        legend.text = element_text(size = 14)  # Legend text size
    ) +
    labs(title = title ,
         x = "Gene Set",
         y = "NES",
         size = "adjusted p-value",
         color = "NES > 0")
}
```

```{r}
Dot_plot(caudate_astro_df, "Astrocytes: Top 10 Gene Sets by NES +/-")
Dot_plot(caudate_endo_df, "Endo: Top 10 Gene Sets by NES +/-")
Dot_plot(caudate_ependymal_df, "Ependymal: Top 10 Gene Sets by NES +/-")
Dot_plot(caudate_immune_df, "Immune: Top 10 Gene Sets by NES +/-")
Dot_plot(caudate_microglia_df, "Microglia: Top 10 Gene Sets by NES +/-")
Dot_plot(caudate_neuron_df, "Neuron: Top 10 Gene Sets by NES +/-")
Dot_plot(caudate_oligo_df, "Oligo: Top 10 Gene Sets by NES +/-")
Dot_plot(caudate_opc_df, "OPC: Top 10 Gene Sets by NES +/-")

```

```{r}
putamen_gsea_df_list
```


```{r}
putamen_astro_df = putamen_gsea_df_list[["astro"]]

putamen_endo_df= putamen_gsea_df_list[["endo"]]
putamen_microglia_df = putamen_gsea_df_list[["mg"]]
putamen_neuron_df = putamen_gsea_df_list[["neuron"]]
putamen_oligo_df=putamen_gsea_df_list[["oligo"]]
putamen_opc_df=putamen_gsea_df_list[["opc"]]
```


```{r}
putamen_astro_df = putamen_gsea_df_list[["astro"]]
putamen_astro_df = putamen_astro_df[putamen_astro_df$gene_set != "tf_perturbations",]
putamen_astro_df

putamen_endo_df= putamen_gsea_df_list[["endo"]]
putamen_endo_df = putamen_endo_df[putamen_endo_df$gene_set != "tf_perturbations",]
putamen_endo_df

putamen_microglia_df = putamen_gsea_df_list[["mg"]]
putamen_microglia_df = putamen_microglia_df[putamen_microglia_df$gene_set != "tf_perturbations",]
putamen_microglia_df

putamen_neuron_df = putamen_gsea_df_list[["neuron"]]
putamen_neuron_df = putamen_neuron_df[putamen_neuron_df$gene_set != "tf_perturbations",]
putamen_neuron_df

putamen_oligo_df=putamen_gsea_df_list[["oligo"]]
putamen_oligo_df = putamen_oligo_df[putamen_oligo_df$gene_set != "tf_perturbations",]
putamen_oligo_df

putamen_opc_df=putamen_gsea_df_list[["opc"]]
putamen_opc_df = putamen_opc_df[putamen_opc_df$gene_set != "tf_perturbations",]
putamen_opc_df
```



```{r}
Dot_plot(putamen_astro_df, "Astrocytes: Top 10 Gene Sets by NES +/-")
Dot_plot(putamen_endo_df, "Endo: Top 10 Gene Sets by NES +/-")
Dot_plot(putamen_microglia_df, "Microglia: Top 10 Gene Sets by NES +/-")
Dot_plot(putamen_neuron_df, "Neuron: Top 10 Gene Sets by NES +/-")
Dot_plot(putamen_oligo_df, "Oligo: Top 10 Gene Sets by NES +/-")
Dot_plot(putamen_opc_df, "OPC: Top 10 Gene Sets by NES +/-")

```

#Try with just positional geneset: I'm using chrXq13 from msigdb because TAF1 is on Xq13.1?


```{r}
library(Seurat)
library(dplyr)
library(qs)

library(Matrix)
library(fgsea)


# when concatenated together, these should be the paths to your DE csv files
BASE_PATH="~/GSEA"
DE_SUBDIR="de/filtered_merged_putamen_clustered_clean_pseudocells__split_by_cell_class__grouped_by_donor_id__mean_size_30"
#DE_SUBDIR="de/filtered_merged_caudate_clustered_clean_pseudocells__split_by_cell_class__grouped_by_donor_id__mean_size_30"

.extraHumanGeneAnnoAdderFn=function(inputGeneNames=NULL){
  #require(EnsDb.Hsapiens.v75)
  require(EnsDb.Hsapiens.v86)
  
  if(!dir.exists("~/serverFiles")){
    dir.create("~/serverFiles",recursive = T)
  }
  
  gns <- as.data.frame(genes(EnsDb.Hsapiens.v86))
  gns$gene_short_name=gns$gene_name
  gns$symbol=toupper(gns$symbol)
  gns$ensembl_gene_id=row.names(gns)
  
  if(!is.null(inputGeneNames)){
    rwNames=toupper(inputGeneNames)
    psCols=c("gene_short_name","ensembl_gene_id")
    slCounts=0
    slCol=""
    if(sum(grepl("\\.",rwNames)&grepl("^ENS",rwNames))>0){
      rwNames=strsplit(rwNames,"\\.")
      rwNames=unlist(lapply(rwNames,function(x)x[1]))
    }
    system(paste0("gsutil -m cp gs://fc-71ac3b81-2441-4171-8038-baf653634620/serverFiles/human_map_to_ensembl.rda ."))
    load("human_map_to_ensembl.rda")
    
    
    map_to_ensmbl$source=toupper(map_to_ensmbl$source)
    
    system(paste0("gsutil -m cp gs://fc-71ac3b81-2441-4171-8038-baf653634620/serverFiles/human_mapping_hg19.rda ."))
      
    
    load("human_mapping_hg19.rda")
    human_hg19$source=toupper(human_hg19$source)
    
    if(sum(toupper(rwNames) %in% human_hg19$source) > sum(toupper(rwNames) %in% map_to_ensmbl$source)){
      map_to_ensmbl=merge(human_hg19,data.frame(source=toupper(rwNames),stringsAsFactors = F),by="source",all.y=T)
    } else {
      map_to_ensmbl=merge(map_to_ensmbl,data.frame(source=toupper(rwNames),stringsAsFactors = F),by="source",all.y=T)
    }
    
    gns=merge(gns,map_to_ensmbl,by.x="ensembl_gene_id",by.y="target",all.y=T)
    gns=gns[match(rwNames,gns$source),]
    row.names(gns)=inputGeneNames
    gns$gene_id=inputGeneNames
    gns=gns[,-which(colnames(gns) %in% c("source","target"))]
  }
  
  return(gns)
}

.sconline.GSEA.readGMT=function (file,bkg_genes=NULL,min.gs.size=NULL,max.gs.size=NULL) {
  if (!grepl("\\.gmt$", file)[1]&F) {
    stop("Pathway information must be in a .gmt file format")
  }
  geneSetDB = readLines(file)
  geneSetDB = strsplit(geneSetDB, "\t")
  names(geneSetDB) = sapply(geneSetDB, "[", 1)
  geneSetDB = lapply(geneSetDB, "[", -1:-2)
  geneSetDB = lapply(geneSetDB, function(x) {
    x[which(x != "")]
  })
  
  if(!is.null(bkg_genes)){
    for(i in 1:length(geneSetDB)){
      tmp=geneSetDB[[i]]
      tmp=bkg_genes[which(toupper(bkg_genes) %in% toupper(tmp))]
      geneSetDB[[i]]=tmp
    }
  }
  
  if(!is.null(min.gs.size)){
    size.dist=unlist(lapply(geneSetDB,length))
    geneSetDB=geneSetDB[size.dist>=min.gs.size]
  }
  
  if(!is.null(max.gs.size)){
    size.dist=unlist(lapply(geneSetDB,length))
    geneSetDB=geneSetDB[size.dist<=max.gs.size]
  }
  
  return(geneSetDB)
}

.myfGSEAfn=function(rankedVec,gs,minSize  = 15,maxSize  = 250, scoreType='std'){
  require(fgsea)
  fgseaRes <- fgsea(pathways = gs, 
                    stats    = rankedVec,
                    minSize  = minSize,
                    maxSize  = maxSize,
                    scoreType= scoreType)
  fgseaRes=fgseaRes[order(fgseaRes$pval,decreasing = F),]
  fgseaRes=as.data.frame(fgseaRes)
  fgseaRes$leadingEdge=unlist(lapply(fgseaRes$leadingEdge,function(x) paste(x,collapse = ",")))
  
  return(fgseaRes)
}

runGSEA = function(
    de_df,
    gs_list_of_char,
    rank_col='logFC',
    gene_id_col='gene_short_name',
    desc=FALSE,
    abs=TRUE,
    scoreType='std'){

    de_df=de_df[!duplicated(de_df[[gene_id_col]]),]
    
    order_vector = de_df[[rank_col]]
    if(abs){
        order_vector = abs(order_vector)
    }
    
    ranked_vec=de_df[,rank_col]
    names(ranked_vec)=de_df[[gene_id_col]]
    ranked_vec=ranked_vec[order(order_vector,decreasing = desc)]
    
    print(head(ranked_vec))

    res_fGSEA=.myfGSEAfn(rankedVec=ranked_vec,gs=gs_list_of_char, scoreType=scoreType)
    res_fGSEA=res_fGSEA[order(res_fGSEA$padj,decreasing=F),]
    return(res_fGSEA)
}
```



```{r}
print(gene_sets)
```


#example
```{r}
outpath = file.path(BASE_PATH, DE_SUBDIR, "positional_gsea_new")
# if you're running into permission errors may need to delete directory and remake with mkdir on command line
dir.create(outpath, recursive=TRUE, showWarnings=FALSE)
de_files = list.files(file.path(BASE_PATH, DE_SUBDIR))
de_files = de_files[grep(".csv$", de_files)]

gene_sets = list(
    # add the paths to the gene sets you want to use here
  Xchrome= "GSEA/genesets/allchrome.txt",  
  #Xq13 = "GSEA/genesets/Xq13geneset.txt",
    Xlinked_recessive = "GSEA/genesets/Xlinked_RESS.txt",
    Xlinked_dominant = "GSEA/genesets/Xlinked_DOM.txt")

for (de_file in de_files){
    de_slogan = sub(".txt$", "", de_file)

    if (! grepl("trend", de_slogan)){next}

    #if (strsplit(de_slogan, "__")[[1]][[2]] != "opc"){next}

    dataDE = read.csv(file.path(BASE_PATH, DE_SUBDIR, de_file))
    anno=.extraHumanGeneAnnoAdderFn(inputGeneNames=dataDE$gene)
    anno=anno[match(dataDE$gene,anno$gene_id),]
    dataDE=cbind(dataDE,anno)
    dataDE=dataDE[which(dataDE$gene_biotype=="protein_coding"),]

    gsea_list = list()

    for (gene_set in names(gene_sets)){
        path = gene_sets[[gene_set]]

        gs=.sconline.GSEA.readGMT(file=path,bkg_genes=dataDE$gene_short_name,min.gs.size=15,max.gs.size=500)
        gsea = runGSEA(dataDE, gs, rank_col="logFC", abs=FALSE, desc=TRUE)
        gsea$gene_set = gene_set
        gsea = gsea[which(gsea$padj<0.05),]

        if (nrow(gsea) > 0){
            gsea_list[[gene_set]] = gsea
        }
    }
    gsea_df = do.call(rbind, gsea_list)

    write.table(gsea_df, file.path(outpath, paste0(de_slogan, "__gsea_logfc.tsv")), sep="\t", quote=FALSE, row.names=FALSE)
}
```


```{r}
#caudate_cell_types = list("astro", "endo", "ependymal", "neuron", "oligo", "opc")
putamen_cell_types = list("astro")
putamen_gsea_df_list = list()

for (cell_type in putamen_cell_types){
gsea_file_name = paste("~/GSEA/de/filtered_merged_putamen_clustered_clean_pseudocells__split_by_cell_class__grouped_by_donor_id__mean_size_30/positional_gsea_new/trend__", cell_type, "__20240529.csv__gsea_logfc.tsv", sep="")

#gsea_file_name = paste("~/GSEA/de/filtered_merged_caudate_clustered_clean_pseudocells__split_by_cell_class__grouped_by_donor_id__mean_size_30/positional_gsea_new/trend__", cell_type, "__20240529.csv__gsea_logfc.tsv", sep="")
gsea_file = read.delim(gsea_file_name, header = TRUE)
putamen_gsea_df_list[[cell_type]] = gsea_file
}
```

```{r}
caudate_gsea_df_list
putamen_gsea_df_list
```

```{r}
putamen_astro_df = putamen_gsea_df_list[["astro"]]
```


```{r}
caudate_astro_df = caudate_gsea_df_list[["astro"]]
caudate_endo_df = caudate_gsea_df_list[["endo"]]
caudate_ependymal_df = caudate_gsea_df_list[["ependymal"]]
caudate_neuron_df = caudate_gsea_df_list[["neuron"]]
caudate_oligo_df = caudate_gsea_df_list[["oligo"]]
caudate_opc_df = caudate_gsea_df_list[["opc"]]
```

```{r}
Dot_plot(caudate_astro_df, "Astrocytes: Top 10 Gene Sets by NES +/-")
Dot_plot(caudate_endo_df, "Endo: Top 10 Gene Sets by NES +/-")
Dot_plot(caudate_ependymal_df, "Ependymal: Top 10 Gene Sets by NES +/-")
Dot_plot(caudate_neuron_df, "Neuron: Top 10 Gene Sets by NES +/-")
Dot_plot(caudate_oligo_df, "Oligo: Top 10 Gene Sets by NES +/-")
Dot_plot(caudate_opc_df, "OPC: Top 10 Gene Sets by NES +/-")
```






```{r}
# Define cell types
caudate_cell_types <- list("astro", "endo", "ependymal", "immune", "mg", "neuron", "oligo", "opc")
putamen_cell_types <- list("astro", "endo", "mg", "neuron", "oligo", "opc")

# Initialize an empty named list to store the dataframes
new_caudate_gsea_df_list <- list()
#new_putamen_gsea_df_list <- list()

# Loop through each cell type and read the corresponding file into a dataframe
for (cell_type in caudate_cell_types) {
  # Construct the file name
  gsea_file_name <- paste("~/GSEA/de/filtered_merged_caudate_clustered_clean_pseudocells__split_by_cell_class__grouped_by_donor_id__mean_size_30/positional_gsea_new/trend__", cell_type, "__20240529.csv__gsea_logfc.tsv", sep = "")
  
  # Print the file name to verify the path
  print(paste("Trying to read file:", gsea_file_name))
  
  # Check if the file exists
  if (file.exists(gsea_file_name)) {
    # Read the file into a dataframe
    gsea_file <- read.delim(gsea_file_name, header = TRUE)
    
    # Store the dataframe in the list with the cell type as the name
    new_caudate_gsea_df_list[[cell_type]] <- gsea_file
  } else {
    print(paste("File does not exist:", gsea_file_name))
  }
}

```


"

```{r}
caudate_astro_df$name = "CaH_astrocyte"
caudate_endo_df$name = "CaH_endothelial"
caudate_ependymal_df$name = "CaH_ependymal"
caudate_neuron_df$name = "CaH_neuron"
caudate_oligo_df$name = "CaH_oligo"
caudate_opc_df$name = "CaH_opc"
putamen_astro_df$name = "Put_astrocyte"
```

```{r}
caudate_astro_df
caudate_endo_df
caudate_neuron_df
caudate_ependymal_df
caudate_oligo_df
caudate_opc_df
putamen_astro_df
```


```{r}
merged_cah_positional = rbind(caudate_astro_df, caudate_endo_df, caudate_ependymal_df, caudate_neuron_df, caudate_oligo_df, caudate_opc_df, putamen_astro_df) 

merged_cah_positional
```





```{r}
data = merged_cah_positional

data$NES_abs <- abs(data$NES)
data$log10_padj <- -log10(data$padj)

# Create the dot plot with the specified color gradient
ggplot(data, aes(x = name, y = pathway)) +
  geom_point(aes(size = NES_abs, color = padj, fill = NES), shape = 21, stroke = 1) +
  scale_size_continuous(range = c(1, 10)) +  # Adjust the range as needed
  scale_color_gradient(low = "black", high = "white") +
  scale_fill_gradient2(low = "red", mid = "white", high = "green", midpoint = 0) +  # Gradient from blue to red
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    title = "Positional GSEA DotPlot for X chromosome",
    x = "Cell Type",
    y = "Gene Sets",
    size = "Absolute NES",
    color = "p-value", 
    fill = "NES"
  ) +
  guides(color = guide_legend(order = 1), size = guide_legend(order = 2))
```
```






























```{r}
data$NES_abs <- abs(data$NES)
data$log10_padj <- -log10(data$p.adjust)

# Create the dot plot with the specified color gradient
ggplot(data, aes(x = CellType, y = Description)) +
  geom_point(aes(size = NES_abs, color = p.adjust, fill = NES), shape = 21, stroke = 1) +
  scale_size_continuous(range = c(1, 10)) +  # Adjust the range as needed
  scale_color_gradient(low = "white", high = "black") +
  scale_fill_gradient2(low = "red", mid = "white", high = "green", midpoint = 0) +  # Gradient from blue to red
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    title = "GSEA DotPlot by Cell Type",
    x = "Cell Type",
    y = "Gene Sets",
    size = "Absolute NES",
    color = "p-value", 
    fill = "NES"
  ) +
  guides(color = guide_legend(order = 1), size = guide_legend(order = 2))

```

```{r}
caudate_astro_df = caudate_gsea_df_list[["astro"]]
caudate_endo_df= caudate_gsea_df_list[["endo"]]
caudate_ependymal_df= caudate_gsea_df_list[["ependymal"]]
caudate_immune_df= caudate_gsea_df_list[["immune"]]
caudate_microglia_df = caudate_gsea_df_list[["mg"]]
caudate_neuron_df = caudate_gsea_df_list[["neuron"]]
caudate_oligo_df=caudate_gsea_df_list[["oligo"]]
caudate_opc_df=caudate_gsea_df_list[["opc"]]
```


```{r}
merged_caudate_gsea <- intersect(caudate_astro_df$pathway, caudate_neuron_df$pathway)
length(merged_caudate_gsea)
merged_caudate_gsea<- intersect(merged_caudate_gsea, caudate_oligo_df$pathway)
length(merged_caudate_gsea)
merged_caudate_gsea<- intersect(merged_caudate_gsea, caudate_microglia_df$pathway)
length(merged_caudate_gsea)
merged_caudate_gsea<- intersect(merged_caudate_gsea, caudate_opc_df$pathway)
length(merged_caudate_gsea)
```
```{r}
library(ggplot2)

# Example data
data <- data.frame(
  CellType = rep(c("astro", "da", "endo", "immune", "mg", "nonda", "oligo", "opc"), times = 10),
  Description = rep(paste("GeneSet", 1:10, sep = ""), each = 8),
  NES = runif(80, -3, 3),
  p.adjust = runif(80, 0, 0.1)
)
data
```



```{r}
cah_astro_sub= subset(caudate_astro_df, caudate_astro_df$pathway %in% merged_caudate_gsea)
cah_endo_sub= subset(caudate_endo_df, caudate_endo_df$pathway %in% merged_caudate_gsea)
cah_ependymal_sub= subset(caudate_ependymal_df, caudate_ependymal_df$pathway %in% merged_caudate_gsea)
cah_immune_sub= subset(caudate_immune_df, caudate_immune_df$pathway %in% merged_caudate_gsea)
cah_microglia_sub= subset(caudate_microglia_df, caudate_microglia_df$pathway %in% merged_caudate_gsea)
cah_neuron_sub= subset(caudate_neuron_df, caudate_neuron_df$pathway %in% merged_caudate_gsea)
cah_oligo_sub= subset(caudate_oligo_df, caudate_oligo_df$pathway %in% merged_caudate_gsea)
cah_opc_sub= subset(caudate_opc_df, caudate_opc_df$pathway %in% merged_caudate_gsea)
```


```{r}
cah_astro_sub$cell_type = "Astrocyte"
cah_endo_sub$cell_type = "Endo"
cah_ependymal_sub$cell_type = "Ependymal"
#cah_immune_sub$cell_type = "Immune"
cah_microglia_sub$cell_type = "Microglia"
cah_neuron_sub$cell_type = "Neuron"
cah_oligo_sub$cell_type = "Oligo"
cah_opc_sub$cell_type = "OPC"
```


```{r}
cah_immune_sub
```

```{r}
merged_cah_df = rbind(cah_astro_sub, cah_endo_sub,cah_ependymal_sub, cah_microglia_sub,cah_neuron_sub,cah_oligo_sub, cah_opc_sub)
merged_cah_df
```



```{r}
data = merged_cah_df

data$NES_abs <- abs(data$NES)
data$log10_padj <- -log10(data$padj)

# Create the dot plot with the specified color gradient
ggplot(data, aes(x = cell_type, y = pathway)) +
  geom_point(aes(size = NES_abs, color = padj, fill = NES), shape = 21, stroke = 1) +
  scale_size_continuous(range = c(1, 10)) +  # Adjust the range as needed
  scale_color_gradient(low = "white", high = "black") +
  scale_fill_gradient2(low = "red", mid = "white", high = "green", midpoint = 0) +  # Gradient from blue to red
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    title = "GSEA DotPlot by Cell Type",
    x = "Cell Type",
    y = "Gene Sets",
    size = "Absolute NES",
    color = "p-value", 
    fill = "NES"
  ) +
  guides(color = guide_legend(order = 1), size = guide_legend(order = 2))
```



```{r}
putamen_astro_df = putamen_gsea_df_list[["astro"]]
putamen_endo_df= putamen_gsea_df_list[["endo"]]
putamen_microglia_df = putamen_gsea_df_list[["mg"]]
putamen_neuron_df = putamen_gsea_df_list[["neuron"]]
putamen_oligo_df=putamen_gsea_df_list[["oligo"]]
putamen_opc_df=putamen_gsea_df_list[["opc"]]
```


```{r}
merged_putamen_gsea <- intersect(putamen_astro_df$pathway, putamen_neuron_df$pathway)
length(merged_putamen_gsea)
merged_putamen_gsea<- intersect(merged_putamen_gsea, putamen_oligo_df$pathway)
length(merged_putamen_gsea)
merged_putamen_gsea<- intersect(merged_putamen_gsea, putamen_microglia_df$pathway)
length(merged_putamen_gsea)
merged_putamen_gsea<- intersect(merged_putamen_gsea, putamen_opc_df$pathway)
length(merged_putamen_gsea)
```


```{r}
```

