---
title: "R Notebook"
output: html_notebook
---
```{r}
clean_recon_sobj_with_neuron_subclusters@meta.data
```

```{r}
spatial_scores(clean_recon_sobj_with_neuron_subclusters@meta.data, "Scoreinterferon_alpha", "Interferon Alpha")
```
```{r}
celltypes = unique(clean_recon_sobj_with_neuron_subclusters$final_cell_class)
clean_recon_sobj_with_neuron_subclusters@meta.data$interferon_alpha_patch = "Other"
for (cell in celltypes) {
df = subset(clean_recon_sobj_with_neuron_subclusters@meta.data, subset = final_cell_class == cell)
threshold <- quantile(df$Scoreinterferon_alpha, 0.99)
cell_ids <- rownames(df[df$Scoreinterferon_alpha > threshold, ])

df$interferon_alpha_patch[df$Scoreinterferon_alpha > threshold] = "Interferon_Positive"
clean_recon_sobj_with_neuron_subclusters@meta.data[cell_ids, "interferon_alpha_patch"] <- "Interferon_Positive"
a = ggplot(data =df, aes(x = Scoreinterferon_alpha)) + geom_histogram(binwidth = 0.005) + ggtitle(paste0(cell)) + xlab("Interferon Alpha Score") +  geom_vline(xintercept = threshold, color = "red", linetype = "dashed")
print(a)
print(threshold)


b=ggplot(data = df, aes(x = x_um, y = y_um, color =interferon_alpha_patch)) + geom_point(size = 0.5) + scale_color_manual(values = c("Other" = "lightgrey", "Interferon_Positive"= "blue")) + ggtitle(paste0(cell, ", threshold: ", sprintf("%.2f", threshold))) + geom_rect(aes(xmin = 12000, xmax = 15000, ymin = 15500, ymax = 19000), color = "red", fill = NA, size = 0.5)
print(b)
                                                                                                              
}
```


```{r}
celltypes = unique(clean_recon_sobj_with_neuron_subclusters$final_cell_class)
for (cell in celltypes) {
df = subset(clean_recon_sobj_with_neuron_subclusters@meta.data, subset = final_cell_class == cell)


a = ggplot(df, aes(x= x_um, y = y_um, color =Scoreinterferon_alpha )) + geom_point(size =0.4) + ggtitle(paste0(cell, " Scores")) + scale_color_viridis_c(option = "magma", name = "Expression", direction = -1) +   labs(color = paste0(" Scores")) +
  theme_void() 

print(a)
}
```


```{r}
ggplot(data = clean_recon_sobj_with_neuron_subclusters@meta.data, aes(x = x_um, y = y_um, color =interferon_alpha_patch)) + geom_point(size = 0.1) + facet_wrap(~ final_cell_class)
```

```{r}
table(clean_recon_sobj_with_neuron_subclusters$interferon_alpha_patch, clean_recon_sobj_with_neuron_subclusters$final_cell_class)
clean_recon_sobj_with_neuron_subclusters$interferon_alpha_patch[clean_recon_sobj_with_neuron_subclusters$x_um > 15000 | clean_recon_sobj_with_neuron_subclusters$x_um <12000 | clean_recon_sobj_with_neuron_subclusters$y_um > 19000 | clean_recon_sobj_with_neuron_subclusters$y_um <15500] = "Other"
table(clean_recon_sobj_with_neuron_subclusters$interferon_alpha_patch, clean_recon_sobj_with_neuron_subclusters$final_cell_class)
```

```{r}
clean_recon_sobj_with_neuron_subclusters@meta.data$interferon_alpha_classes = paste0(clean_recon_sobj_with_neuron_subclusters$interferon_alpha_patch,"_", clean_recon_sobj_with_neuron_subclusters$final_cell_class) 

table(clean_recon_sobj_with_neuron_subclusters$interferon_alpha_classes)
```


```{r}
Idents(clean_recon_sobj_with_neuron_subclusters) = "interferon_alpha_classes"

astrocyte_roi_markers_new = FindMarkers(clean_recon_sobj_with_neuron_subclusters, ident.1 = "Interferon_Positive_astrocyte", ident.2 = "Other_astrocyte", min.pct = 0.1, logfc_threshold = 0.5, max.cells.per.ident = 10000)
astrocyte_roi_markers_new

endo_roi_markers_new = FindMarkers(clean_recon_sobj_with_neuron_subclusters, ident.1 = "Interferon_Positive_endothelial", ident.2 = "Other_endothelial", min.pct = 0.1, logfc_threshold = 0.5)
endo_roi_markers_new

micro_roi_markers_new = FindMarkers(clean_recon_sobj_with_neuron_subclusters, ident.1 = "Interferon_Positive_microglia", ident.2 = "Other_microglia", min.pct = 0.1, logfc_threshold = 0.5)
micro_roi_markers_new

neuron_roi_markers_new = FindMarkers(clean_recon_sobj_with_neuron_subclusters, ident.1 = "Interferon_Positive_neuron", ident.2 = "Other_neuron", min.pct = 0.1, logfc_threshold = 0.5, max.cells.per.ident = 10000)
neuron_roi_markers_new

oligo_roi_markers_new = FindMarkers(clean_recon_sobj_with_neuron_subclusters, ident.1 = "Interferon_Positive_oligo", ident.2 = "Other_oligo", min.pct = 0.1, logfc_threshold = 0.5, max.cells.per.ident = 10000)
oligo_roi_markers_new

opc_roi_markers_new = FindMarkers(clean_recon_sobj_with_neuron_subclusters, ident.1 = "Interferon_Positive_opc", ident.2 = "Other_opc", min.pct = 0.1, logfc_threshold = 0.5)
opc_roi_markers_new

```

```{r}
astrocyte_roi_markers_sig_new = subset(astrocyte_roi_markers_new , subset = p_val_adj < 0.05)
endo_roi_markers_sig_new   = subset(endo_roi_markers_new , subset = p_val_adj < 0.05)
oligo_roi_markers_sig_new   = subset(oligo_roi_markers_new , subset = p_val_adj < 0.05)
opc_roi_markers_sig_new   = subset(opc_roi_markers_new , subset = p_val_adj < 0.05)
neuron_roi_markers_sig_new   = subset(neuron_roi_markers_new , subset = p_val_adj < 0.05)
micro_roi_markers_sig_new   = subset(micro_roi_markers_new , subset = p_val_adj < 0.05)

astrocyte_roi_markers_sig_new 
endo_roi_markers_sig_new 
oligo_roi_markers_sig_new 
opc_roi_markers_sig_new 
neuron_roi_markers_sig_new 
micro_roi_markers_sig_new 
```

```{r}
write.csv(astrocyte_roi_markers_sig_new , "interferon_classes_markers_new/astrocyte_roi_markers_sig_new.csv")
write.csv(endo_roi_markers_sig_new , "interferon_classes_markers_new/endo_roi_markers_sig_new.csv")
write.csv(oligo_roi_markers_sig_new , "interferon_classes_markers_new/oligo_roi_markers_sig_new.csv")
write.csv(opc_roi_markers_sig_new , "interferon_classes_markers_new/opc_roi_markers_sig_new.csv")
write.csv(neuron_roi_markers_sig_new , "interferon_classes_markers_new/neuron_roi_markers_sig_new.csv")
write.csv(micro_roi_markers_sig_new , "interferon_classes_markers_new/micro_roi_markers_sig_new.csv")
```


```{r}
astrocyte_roi_genes = rownames(astrocyte_roi_markers_sig_new)
length(astrocyte_roi_genes)
astrocyte_roi_genes = intersect(rownames(clean_recon_sobj_with_neuron_subclusters),astrocyte_roi_genes)
astrocyte_roi_genes = intersect(rownames(xdp_cah_put),astrocyte_roi_genes)
length(astrocyte_roi_genes)


endo_roi_genes = rownames(endo_roi_markers_sig_new)
length(endo_roi_genes)
endo_roi_genes = intersect(rownames(clean_recon_sobj_with_neuron_subclusters),endo_roi_genes)
endo_roi_genes = intersect(rownames(xdp_cah_put),endo_roi_genes)
length(endo_roi_genes)

oligo_roi_genes = rownames(oligo_roi_markers_sig_new)
length(oligo_roi_genes)
oligo_roi_genes = intersect(rownames(clean_recon_sobj_with_neuron_subclusters),oligo_roi_genes)
oligo_roi_genes = intersect(rownames(xdp_cah_put),oligo_roi_genes)
length(oligo_roi_genes)

opc_roi_genes = rownames(opc_roi_markers_sig_new)
length(opc_roi_genes)
opc_roi_genes = intersect(rownames(clean_recon_sobj_with_neuron_subclusters),opc_roi_genes)
opc_roi_genes = intersect(rownames(xdp_cah_put),opc_roi_genes)
length(opc_roi_genes)

neuron_roi_genes = rownames(neuron_roi_markers_sig_new)
length(neuron_roi_genes)
neuron_roi_genes = intersect(rownames(clean_recon_sobj_with_neuron_subclusters),neuron_roi_genes)
neuron_roi_genes = intersect(rownames(xdp_cah_put),neuron_roi_genes)
length(neuron_roi_genes)

micro_roi_genes = rownames(micro_roi_markers_sig_new)
length(micro_roi_genes)
micro_roi_genes = intersect(rownames(clean_recon_sobj_with_neuron_subclusters),micro_roi_genes)
micro_roi_genes = intersect(rownames(xdp_cah_put),micro_roi_genes)
length(micro_roi_genes)
```
```{r}
qsave(astrocyte_roi_genes, "astrocyte_roi_genes.qs")
qsave(endo_roi_genes, "endo_roi_genes.qs")
qsave(oligo_roi_genes, "oligo_roi_genes.qs")
qsave(opc_roi_genes, "opc_roi_genes.qs")
qsave(neuron_roi_genes, "neuron_roi_genes.qs")
qsave(micro_roi_genes, "micro_roi_genes.qs")
```

```{r}
up_astrocyte_roi_genes = subset(astrocyte_roi_markers_sig_new, subset = avg_log2FC > 0)
up_endo_roi_genes = subset(endo_roi_markers_sig_new, subset = avg_log2FC > 0)
up_oligo_roi_genes = subset(oligo_roi_markers_sig_new, subset = avg_log2FC > 0)
up_opc_roi_genes = subset(opc_roi_markers_sig_new, subset = avg_log2FC > 0)
up_neuron_roi_genes = subset(neuron_roi_markers_sig_new, subset = avg_log2FC > 0)
up_micro_roi_genes = subset(micro_roi_markers_sig_new, subset = avg_log2FC > 0)

down_astrocyte_roi_genes = subset(astrocyte_roi_markers_sig_new, subset = avg_log2FC < 0)
down_endo_roi_genes = subset(endo_roi_markers_sig_new, subset = avg_log2FC < 0)
down_oligo_roi_genes = subset(oligo_roi_markers_sig_new, subset = avg_log2FC < 0)
down_opc_roi_genes = subset(opc_roi_markers_sig_new, subset = avg_log2FC < 0)
down_neuron_roi_genes = subset(neuron_roi_markers_sig_new, subset = avg_log2FC < 0)
down_micro_roi_genes = subset(micro_roi_markers_sig_new, subset = avg_log2FC < 0)


up_astrocyte_roi_genes = rownames(up_astrocyte_roi_genes)
length(up_astrocyte_roi_genes)
up_astrocyte_roi_genes = intersect(rownames(clean_recon_sobj_with_neuron_subclusters),up_astrocyte_roi_genes)
up_astrocyte_roi_genes = intersect(rownames(xdp_cah_put),up_astrocyte_roi_genes)
length(up_astrocyte_roi_genes)

up_endo_roi_genes = rownames(up_endo_roi_genes)
length(up_endo_roi_genes)
up_endo_roi_genes = intersect(rownames(clean_recon_sobj_with_neuron_subclusters),up_endo_roi_genes)
up_endo_roi_genes = intersect(rownames(xdp_cah_put),up_endo_roi_genes)
length(up_endo_roi_genes)

up_oligo_roi_genes = rownames(up_oligo_roi_genes)
length(up_oligo_roi_genes)
up_oligo_roi_genes = intersect(rownames(clean_recon_sobj_with_neuron_subclusters),up_oligo_roi_genes)
up_oligo_roi_genes = intersect(rownames(xdp_cah_put),up_oligo_roi_genes)
length(up_oligo_roi_genes)

up_opc_roi_genes = rownames(up_opc_roi_genes)
length(up_opc_roi_genes)
up_opc_roi_genes = intersect(rownames(clean_recon_sobj_with_neuron_subclusters),up_opc_roi_genes)
up_opc_roi_genes = intersect(rownames(xdp_cah_put),up_opc_roi_genes)
length(up_opc_roi_genes)

up_neuron_roi_genes = rownames(up_neuron_roi_genes)
length(up_neuron_roi_genes)
up_neuron_roi_genes = intersect(rownames(clean_recon_sobj_with_neuron_subclusters),up_neuron_roi_genes)
up_neuron_roi_genes = intersect(rownames(xdp_cah_put),up_neuron_roi_genes)
length(up_neuron_roi_genes)

up_micro_roi_genes = rownames(up_micro_roi_genes)
length(up_micro_roi_genes)
up_micro_roi_genes = intersect(rownames(clean_recon_sobj_with_neuron_subclusters),up_micro_roi_genes)
up_micro_roi_genes = intersect(rownames(xdp_cah_put),up_micro_roi_genes)
length(up_micro_roi_genes)
```

```{r}
down_astrocyte_roi_genes = rownames(down_astrocyte_roi_genes)
length(down_astrocyte_roi_genes)
down_astrocyte_roi_genes = intersect(rownames(clean_recon_sobj_with_neuron_subclusters),down_astrocyte_roi_genes)
down_astrocyte_roi_genes = intersect(rownames(xdp_cah_put),down_astrocyte_roi_genes)
length(down_astrocyte_roi_genes)

down_endo_roi_genes = rownames(down_endo_roi_genes)
length(down_endo_roi_genes)
down_endo_roi_genes = intersect(rownames(clean_recon_sobj_with_neuron_subclusters),down_endo_roi_genes)
down_endo_roi_genes = intersect(rownames(xdp_cah_put),down_endo_roi_genes)
length(down_endo_roi_genes)

down_oligo_roi_genes = rownames(down_oligo_roi_genes)
length(down_oligo_roi_genes)
down_oligo_roi_genes = intersect(rownames(clean_recon_sobj_with_neuron_subclusters),down_oligo_roi_genes)
down_oligo_roi_genes = intersect(rownames(xdp_cah_put),down_oligo_roi_genes)
length(down_oligo_roi_genes)

down_opc_roi_genes = rownames(down_opc_roi_genes)
length(down_opc_roi_genes)
down_opc_roi_genes = intersect(rownames(clean_recon_sobj_with_neuron_subclusters),down_opc_roi_genes)
down_opc_roi_genes = intersect(rownames(xdp_cah_put),down_opc_roi_genes)
length(down_opc_roi_genes)

down_neuron_roi_genes = rownames(down_neuron_roi_genes)
length(down_neuron_roi_genes)
down_neuron_roi_genes = intersect(rownames(clean_recon_sobj_with_neuron_subclusters),down_neuron_roi_genes)
down_neuron_roi_genes = intersect(rownames(xdp_cah_put),down_neuron_roi_genes)
length(down_neuron_roi_genes)

down_micro_roi_genes = rownames(down_micro_roi_genes)
length(down_micro_roi_genes)
down_micro_roi_genes = intersect(rownames(clean_recon_sobj_with_neuron_subclusters),down_micro_roi_genes)
down_micro_roi_genes = intersect(rownames(xdp_cah_put),down_micro_roi_genes)
length(down_micro_roi_genes)
```


```{r}
qsave(up_astrocyte_roi_genes, "up_astrocyte_roi_genes_new.qs")
qsave(up_endo_roi_genes, "up_endo_roi_genes_new.qs")
qsave(up_oligo_roi_genes, "up_oligo_roi_genes_new.qs")
qsave(up_opc_roi_genes, "up_opc_roi_genes_new.qs")
qsave(up_neuron_roi_genes, "up_neuron_roi_genes_new.qs")
qsave(up_micro_roi_genes, "up_micro_roi_genes_new.qs")

qsave(down_astrocyte_roi_genes, "down_astrocyte_roi_genes_new.qs")
qsave(down_oligo_roi_genes, "down_oligo_roi_genes_new.qs")
qsave(down_neuron_roi_genes, "down_neuron_roi_genes_new.qs")
qsave(down_micro_roi_genes, "down_micro_roi_genes_new.qs")
```




```{r}
clean_recon_sobj_with_neuron_subclusters@meta.data
```
```{r}
qsave(clean_recon_sobj_with_neuron_subclusters, "clean_recon_sobj_with_neuron_subclusters_sct.qs")
```


```{r}
xdp_cah_put <- AddModuleScore_UCell(xdp_cah_put,features = list(ROI_upreg_new = up_astrocyte_roi_genes),
                                                                 name = '_astrocyte'
)

xdp_cah_put <- AddModuleScore_UCell(xdp_cah_put, features = list(ROI_upreg_new = up_endo_roi_genes),
                                                                 name = '_endo'
)

xdp_cah_put <- AddModuleScore_UCell(xdp_cah_put, features = list(ROI_upreg_new = up_oligo_roi_genes),
                                                                 name = '_oligo'
)

xdp_cah_put <- AddModuleScore_UCell(xdp_cah_put,features = list(ROI_upreg_new = up_opc_roi_genes),
                                                                 name = '_opc'
)

xdp_cah_put <- AddModuleScore_UCell(xdp_cah_put,features = list(ROI_upreg_new = up_neuron_roi_genes),
                                                                 name = '_neuron'
)

xdp_cah_put <- AddModuleScore_UCell(xdp_cah_put,features = list(ROI_upreg_new = up_micro_roi_genes),
                                                                 name = '_mg'
)


#down
xdp_cah_put <- AddModuleScore_UCell(xdp_cah_put, features = list(ROI_downreg_new = down_astrocyte_roi_genes),
                                                                 name = '_astrocyte'
)

xdp_cah_put <- AddModuleScore_UCell(xdp_cah_put,features = list(ROI_downreg_new = down_oligo_roi_genes),
                                                                 name = '_oligo'
)


xdp_cah_put <- AddModuleScore_UCell(xdp_cah_put,features = list(ROI_downreg_new = down_neuron_roi_genes),
                                                                 name = '_neuron'
)

xdp_cah_put <- AddModuleScore_UCell(xdp_cah_put,features = list(ROI_downreg_new = down_micro_roi_genes),
                                                                 name = '_mg'
)

qsave(xdp_cah_put, "xdp_cah_put_sct_new.qs")



clean_recon_sobj_with_neuron_subclusters <- AddModuleScore_UCell(clean_recon_sobj_with_neuron_subclusters,features = list(ROI_upreg_new = up_astrocyte_roi_genes),
                                                                 name = '_astrocyte'
)

clean_recon_sobj_with_neuron_subclusters <- AddModuleScore_UCell(clean_recon_sobj_with_neuron_subclusters, features = list(ROI_upreg_new = up_endo_roi_genes),
                                                                 name = '_endo'
)

clean_recon_sobj_with_neuron_subclusters <- AddModuleScore_UCell(clean_recon_sobj_with_neuron_subclusters, features = list(ROI_upreg_new = up_oligo_roi_genes),
                                                                 name = '_oligo'
)

clean_recon_sobj_with_neuron_subclusters <- AddModuleScore_UCell(clean_recon_sobj_with_neuron_subclusters,features = list(ROI_upreg_new = up_opc_roi_genes),
                                                                 name = '_opc'
)

clean_recon_sobj_with_neuron_subclusters <- AddModuleScore_UCell(clean_recon_sobj_with_neuron_subclusters,features = list(ROI_upreg_new = up_neuron_roi_genes),
                                                                 name = '_neuron'
)

clean_recon_sobj_with_neuron_subclusters <- AddModuleScore_UCell(clean_recon_sobj_with_neuron_subclusters,features = list(ROI_upreg_new = up_micro_roi_genes),
                                                                 name = '_mg'
)


#down
clean_recon_sobj_with_neuron_subclusters <- AddModuleScore_UCell(clean_recon_sobj_with_neuron_subclusters, features = list(ROI_downreg_new = down_astrocyte_roi_genes),
                                                                 name = '_astrocyte'
)

clean_recon_sobj_with_neuron_subclusters <- AddModuleScore_UCell(clean_recon_sobj_with_neuron_subclusters,features = list(ROI_downreg_new = down_oligo_roi_genes),
                                                                 name = '_oligo'
)


clean_recon_sobj_with_neuron_subclusters <- AddModuleScore_UCell(clean_recon_sobj_with_neuron_subclusters,features = list(ROI_downreg_new = down_neuron_roi_genes),
                                                                 name = '_neuron'
)

clean_recon_sobj_with_neuron_subclusters <- AddModuleScore_UCell(clean_recon_sobj_with_neuron_subclusters,features = list(ROI_downreg_new = down_micro_roi_genes),
                                                                 name = '_mg'
)

qsave(clean_recon_sobj_with_neuron_subclusters, "clean_recon_sobj_with_neuron_subclusters_sct_new.qs")

```


```{r}
#clean_recon_sobj_with_neuron_subclusters = qread("clean_recon_sobj_with_neuron_subclusters_sct_new.qs")
clean_recon_sobj_with_neuron_subclusters
clean_recon_sobj_with_neuron_subclusters@meta.data
xdp_cah_put
xdp_cah_put@meta.data
```


```{r}
spatial_scores(clean_recon_sobj_with_neuron_subclusters@meta.data, "ScoreNFKB", "NFKB in Neurons")

histograms_by_all_celltype(xdp_cah_put@meta.data, "ScoreNFKB", "NFKB score", 20, 10)
```

```{r}
histograms_by_donor(xdp_cah_put@meta.data, "ScoreNFKB", "astrocyte" ,"NFKB score", 25, 15)
```

```{r}
neuron_roi = read.csv("~/interferon_classes_markers/neuron_roi_markers_sig.csv")
neuron_roi

neuron_int_roi = read.csv("~/interferon_classes_markers/neuron_roi_markers_sig.csv")
```







#SPN Matrix score
```{r}
clean_recon_sobj_with_neuron_subclusters@meta.data$interferon_alpha_subclasses = paste0(clean_recon_sobj_with_neuron_subclusters$reclustered_patch_matrix_exotic,"_" , clean_recon_sobj_with_neuron_subclusters$interferon_alpha_patch)
clean_recon_sobj_with_neuron_subclusters@meta.data 
Idents(clean_recon_sobj_with_neuron_subclusters) = "interferon_alpha_subclasses"
```

```{r}
table(clean_recon_sobj_with_neuron_subclusters$interferon_alpha_subclasses)
```


```{r}
matrix_roi_markers = FindMarkers(clean_recon_sobj_with_neuron_subclusters, ident.1 = "SPN_matrix_Interferon_Positive", ident.2 = "SPN_matrix_Other", min.pct = 0.1, logfc_threshold = 0.5)
matrix_roi_markers
matrix_roi_markers_sig = subset(matrix_roi_markers, subset = p_val_adj < 0.05)
matrix_roi_markers_sig

write.csv(matrix_roi_markers_sig, "interferon_classes_markers/matrix_roi_markers_sig_new.csv")
```


```{r}
matrix_roi_genes = rownames(matrix_roi_markers_sig)
length(matrix_roi_genes)
matrix_roi_genes = intersect(rownames(clean_recon_sobj_with_neuron_subclusters),matrix_roi_genes)
matrix_roi_genes = intersect(rownames(xdp_cah_put),matrix_roi_genes)
length(matrix_roi_genes)

qsave(matrix_roi_genes, "matrix_roi_genes_new.qs")
```

```{r}
matrix_roi_markers_sig
```


```{r}
up_matrix_roi_genes = subset(matrix_roi_markers_sig, subset = avg_log2FC > 0)
up_matrix_roi_genes
up_matrix_roi_genes = rownames(up_matrix_roi_genes)
length(up_matrix_roi_genes)

up_matrix_roi_genes = intersect(rownames(clean_recon_sobj_with_neuron_subclusters),up_matrix_roi_genes)
up_matrix_roi_genes = intersect(rownames(xdp_cah_put),up_matrix_roi_genes)
length(up_matrix_roi_genes)
```


```{r}
down_matrix_roi_genes = subset(matrix_roi_markers_sig, subset = avg_log2FC < 0)
down_matrix_roi_genes = rownames(down_matrix_roi_genes)
length(down_matrix_roi_genes)
down_matrix_roi_genes = intersect(rownames(clean_recon_sobj_with_neuron_subclusters),down_matrix_roi_genes)
down_matrix_roi_genes = intersect(rownames(xdp_cah_put),down_matrix_roi_genes)
length(down_matrix_roi_genes)

qsave(up_matrix_roi_genes, "up_matrix_roi_genes.qs")
qsave(down_matrix_roi_genes, "down_matrix_roi_genes.qs")
```

```{r}
clean_recon_sobj_with_neuron_subclusters <- AddModuleScore_UCell(clean_recon_sobj_with_neuron_subclusters,
  features = list(ROI_upreg_new = up_matrix_roi_genes),
  name = '_SPN_matrix'
)

clean_recon_sobj_with_neuron_subclusters <- AddModuleScore_UCell(clean_recon_sobj_with_neuron_subclusters,
  features = list(ROI_downreg_new = down_matrix_roi_genes),
  name = '_SPN_matrix'
)


xdp_cah_put <- AddModuleScore_UCell(xdp_cah_put,
  features = list(ROI_upreg_new = up_matrix_roi_genes),
  name = '_SPN_matrix'
)

xdp_cah_put <- AddModuleScore_UCell(xdp_cah_put,
  features = list(ROI_downreg_new = down_matrix_roi_genes),
  name = '_SPN_matrix'
)


```
























```{r}
a1 = intersect(up_astrocyte_roi_genes, up_endo_roi_genes)
a1
a1 = intersect(a1,up_oligo_roi_genes )
a1
a1 = intersect(a1,up_opc_roi_genes )
a1
a1 = intersect(a1,up_neuron_roi_genes )
a1
a1 = intersect(a1,up_micro_roi_genes )
a1
```


```{r}
up_astrocyte_roi_genes
up_endo_roi_genes
up_oligo_roi_genes
up_opc_roi_genes
up_neuron_roi_genes
up_micro_roi_genes

down_astrocyte_roi_genes
down_oligo_roi_genes
down_neuron_roi_genes
down_micro_roi_genes
```





```{r}
subset_sobj = subset(clean_recon_sobj_with_neuron_subclusters, subset = reclustered_patch_matrix_exotic == "SPN_matrix")

```

```{r}
scores <- FetchData(subset_sobj, "ROI_upreg_astrocyte")  # replace with the actual name of the score column
gene_expression <- FetchData(subset_sobj, up_astrocyte_roi_genes)  # gene_list contains your gene names
correlation_results <- cor(gene_expression, scores)
correlation_results

contribution <- as.data.frame(cor(gene_expression, scores))
contribution$genes <- rownames(contribution)
contribution
contribution <- na.omit(contribution)
contribution

a= ggplot(contribution, aes(x = reorder(genes, ROI_upreg_astrocyte), y = ROI_upreg_astrocyte)) +
    geom_bar(stat = "identity") +
    coord_flip() +
    labs(title = "ROI_upreg_astrocyte Gene Contribution to \nUCell Score in SPN_matrix", x = "ROI_upreg_astrocyte Genes", y = "Correlation with ROI_upreg_astrocyte Score")

ggsave(a, filename= "pic.png", width = 5, height =10)
```

```{r}
clean_recon_sobj_with_neuron_subclusters@meta.data
```


```{r}
spatial_scores(clean_recon_sobj_with_neuron_subclusters@meta.data, "ROI_downreg_SPN_matrix", "Downregulated in SPN Matrix")

histograms_by_all_celltype(xdp_cah_put@meta.data, "ROI_downreg_SPN_matrix", "Downreg Interferon Matrix score", 20, 10)
```

```{r}
histograms_by_donor(xdp_cah_put@meta.data, "ROI_downreg_SPN_matrix", "SPN_matrix" ,"Downreg Interferon Matrix score", 25, 15)
```

#SPN Matrix score
```{r}
clean_recon_sobj_with_neuron_subclusters@meta.data$interferon_patch_classes_spn = paste0(clean_recon_sobj_with_neuron_subclusters$reclustered_patch_matrix_exotic,"_" , clean_recon_sobj_with_neuron_subclusters$interferon_patch)
clean_recon_sobj_with_neuron_subclusters@meta.data 
Idents(clean_recon_sobj_with_neuron_subclusters) = "interferon_patch_classes_spn"
```

```{r}
table(clean_recon_sobj_with_neuron_subclusters$interferon_patch_classes_spn)
```




```{r}
matrix_roi_markers = FindMarkers(clean_recon_sobj_with_neuron_subclusters, ident.1 = "SPN_matrix_ROI", ident.2 = "SPN_matrix_Other", min.pct = 0.1, logfc_threshold = 0.5, max.cells.per.ident = 10000)
matrix_roi_markers
matrix_roi_markers_sig = subset(matrix_roi_markers, subset = p_val_adj < 0.05)
matrix_roi_markers_sig

write.csv(matrix_roi_markers_sig, "interferon_classes_markers/matrix_roi_markers_sig.csv")
```


```{r}
matrix_roi_genes = rownames(matrix_roi_markers_sig)
length(matrix_roi_genes)
matrix_roi_genes = intersect(rownames(clean_recon_sobj_with_neuron_subclusters),matrix_roi_genes)
matrix_roi_genes = intersect(rownames(xdp_cah_put),matrix_roi_genes)
length(matrix_roi_genes)

qsave(matrix_roi_genes, "matrix_roi_genes.qs")
```

```{r}
matrix_roi_markers_sig
```


```{r}
up_matrix_roi_genes = subset(matrix_roi_markers_sig, subset = avg_log2FC > 0)
up_matrix_roi_genes
up_matrix_roi_genes = rownames(up_matrix_roi_genes)
length(up_matrix_roi_genes)

up_matrix_roi_genes = intersect(rownames(clean_recon_sobj_with_neuron_subclusters),up_matrix_roi_genes)
up_matrix_roi_genes = intersect(rownames(xdp_cah_put),up_matrix_roi_genes)
length(up_matrix_roi_genes)
```


```{r}
down_matrix_roi_genes = subset(matrix_roi_markers_sig, subset = avg_log2FC < 0)
down_matrix_roi_genes = rownames(down_matrix_roi_genes)
length(down_matrix_roi_genes)
down_matrix_roi_genes = intersect(rownames(clean_recon_sobj_with_neuron_subclusters),down_matrix_roi_genes)
down_matrix_roi_genes = intersect(rownames(xdp_cah_put),down_matrix_roi_genes)
length(down_matrix_roi_genes)

qsave(up_matrix_roi_genes, "up_matrix_roi_genes.qs")
qsave(down_matrix_roi_genes, "down_matrix_roi_genes.qs")
```

```{r}
clean_recon_sobj_with_neuron_subclusters <- AddModuleScore_UCell(clean_recon_sobj_with_neuron_subclusters,
  features = list(ROI_upreg = up_matrix_roi_genes),
  name = '_SPN_matrix'
)

clean_recon_sobj_with_neuron_subclusters <- AddModuleScore_UCell(clean_recon_sobj_with_neuron_subclusters,
  features = list(ROI_downreg = down_matrix_roi_genes),
  name = '_SPN_matrix'
)


xdp_cah_put <- AddModuleScore_UCell(xdp_cah_put,
  features = list(ROI_upreg = up_matrix_roi_genes),
  name = '_SPN_matrix'
)

xdp_cah_put <- AddModuleScore_UCell(xdp_cah_put,
  features = list(ROI_downreg = down_matrix_roi_genes),
  name = '_SPN_matrix'
)


```












split_recon_df$split[split_recon_df$split == "caudate" & split_recon_df$x_um > 11000 & split_recon_df$y_um < 3000] = "caudate_1"
split_recon_df$split[split_recon_df$split == "caudate" & split_recon_df$x_um > 11000 & split_recon_df$y_um <6000] = "caudate_2"
split_recon_df$split[split_recon_df$split == "caudate" & split_recon_df$x_um > 11000 & split_recon_df$y_um <9000] = "caudate_3"
split_recon_df$split[split_recon_df$split == "caudate" & split_recon_df$x_um > 11000 & split_recon_df$y_um <12000] = "caudate_4"

split_recon_df$split[split_recon_df$split == "putamen" & split_recon_df$x_um < 11000 & split_recon_df$y_um < 3000] = "putamen_1"
split_recon_df$split[split_recon_df$split == "putamen" & split_recon_df$x_um < 11000 & split_recon_df$y_um <6000] = "putamen_2"
split_recon_df$split[split_recon_df$split == "putamen" & split_recon_df$x_um < 11000 & split_recon_df$y_um <9000] = "putamen_3"
split_recon_df$split[split_recon_df$split == "putamen" & split_recon_df$x_um < 11000 & split_recon_df$y_um <12000] = "putamen_4"

```{r}
metadata
```


```{r}
ggplot(metadata, aes(x= x_um, y= y_um, color = reclustered_patch_matrix_exotic)) + geom_point(size = 0.1) +  geom_rect(aes(xmin = 12000, xmax = 14000, ymin = 2000, ymax = 5000), 
            color = "red", fill = NA, size = 0.5)+
  geom_rect(aes(xmin = 12000, xmax = 14000, ymin = 5000, ymax = 8000), 
            color = "black", fill = NA, size = 0.5)+
  geom_rect(aes(xmin = 12000, xmax = 14000, ymin = 8000, ymax = 11000), 
            color = "purple", fill = NA, size = 0.5)+
  geom_rect(aes(xmin = 5000, xmax = 7000, ymin = 2000, ymax = 5000), 
            color = "red", fill = NA, size = 0.5)+
  geom_rect(aes(xmin = 5000, xmax = 7000, ymin = 5000, ymax = 8000), 
            color = "black", fill = NA, size = 0.5)+
  geom_rect(aes(xmin = 5000, xmax = 7000, ymin = 8000, ymax = 11000), 
            color = "purple", fill = NA, size = 0.5)+
  facet_wrap(~ reclustered_patch_matrix_exotic) 
```


```{r}
ggplot(metadata, aes(x= x_um, y= y_um, color = final_cell_class)) + geom_point(size = 0.1) +  geom_rect(aes(xmin = 5000, xmax = 7000, ymin = 2000, ymax = 5000), 
            color = "red", fill = NA, size = 0.5)+
  geom_rect(aes(xmin = 5000, xmax = 7000, ymin = 5000, ymax = 8000), 
            color = "black", fill = NA, size = 0.5)+
  geom_rect(aes(xmin = 5000, xmax = 7000, ymin = 8000, ymax = 11000), 
            color = "purple", fill = NA, size = 0.5)+
  facet_wrap(~ final_cell_class) 
```

```{r}
caudate_roi = subset(metadata, subset = x_um >5000 & x_um < 7000 & y_um >8000 & y_um < 11000)
caudate_roi
table(caudate_roi$reclustered_patch_matrix_exotic)
```


```{r}
length(matrix_roi_genes)
length(neuron_roi_genes)
```

```{r}
a = intersect(up_matrix_roi_genes, c_plus_genes)
a
```

```{r}
setdiff(matrix_roi_genes, neuron_roi_genes)
```


```{r}
setdiff(neuron_roi_genes,matrix_roi_genes)
```








```{r}
metadata_zoom = subset(metadata, subset = x_um >11000 & x_um <15000 & y_um >15000 & x_um <19000)
```


```{r}
ggplot(metadata_zoom, aes(x= x_um, y= y_um, color = Scoreinterferon_alpha)) + geom_point(size = 1)+scale_color_viridis_c(option = "magma", name = "Expression", direction = -1) + facet_wrap(~ final_cell_class)+ geom_vline(xintercept = 14000, color = "blue", size = 0.5)+ geom_vline(xintercept = 13000, color = "blue", size = 0.5)+ geom_hline(yintercept = 16250, color = "blue", size = 0.5) + geom_hline(yintercept = 17750, color = "blue", size = 0.5) 
```



```{r}
table(clean_recon_sobj_with_neuron_subclusters$final_cell_class)
```



```{r}
SN_sobj = qread("~/SN_sobj_sct_scores.qs")
SN_sobj
SN_sobj@meta.data
DimPlot(SN_sobj, label = T)
```
```{r}
DFC_sobj = qread("~/DFC_sobj_sct_scores.qs")
DFC_sobj
DFC_sobj@meta.data
DimPlot(DFC_sobj, label = T)
```



```{r}
spatial_gene_exp_plot(gene_of_interest = "DYNC1H1", seurat_object =Recon_neuron_new_clean_final, cluster = "reclustered_patch_matrix_exotic", type = "Neurons", wid = 18, hei = 10)
```


```{r}
spatial_gene_exp_plot(gene_of_interest = "RNF31", seurat_object =recon , cluster = "final_cell_class",type = "All Cell Classes", wid = 13, hei = 10)
```








```{r}
neuron_roi_markers_sig #everything in the roi
neuron_roi_markers_sig_new #only interferon related genes?
```



```{r}
roi_a = rownames(neuron_roi_markers_sig)
length(roi_a)
roi_b = rownames(neuron_roi_markers_sig_new)
length(roi_b)
```


```{r}
a1 = intersect(roi_a, roi_b)
length(a1)
a2 = setdiff(roi_a, roi_b)
length(a2)

head(a2)
```

```{r}
interesting_genes_maybe = subset(neuron_roi_markers_sig, subset = rownames(neuron_roi_markers_sig) %in% a2)
interesting_genes_maybe
```

```{r}
write.csv(interesting_genes_maybe, "interesting_genes_maybe.csv")
```

