---
title: "R Notebook"
output: html_notebook
---

```{r}
plot_overlapping_density_histogram = function(
    df, 
    hist_col,
    fill_col,
    colors = c("blue", "red"),
    alpha=0.5,
    breaks=seq(0, 16, 1),
    title= NULL,
    xlab = NULL,
    fig_filename = NULL

){
    # hist_col is the column you're making a histogram of (e.g. nUMI)
    # fill_col is the column you're coloring by (e.g. cell_class)
    # if fig_filename is not null, the plot will be saved to that file
     
    if (is.null(xlab)){
        xlab = hist_col
    }

    if (is.null(title)){
        title = paste0("Density Histogram of ", xlab, " by ", fill_col)
    }


    p = (
        ggplot(df, aes_string(x=hist_col, fill=fill_col)) 
        + geom_histogram(aes(y=..density..), alpha=alpha, position="identity", breaks=breaks)
        + labs(title=title, x=xlab, y="Density")    
        + theme(
                plot.title = element_text(size=16),
                axis.line = element_line(color = "black"),  # Add axis lines
                axis.ticks = element_line(color = "black"),  # Add axis ticks
                axis.text = element_text(size = 14),  # Increase tick label font size
                axis.title = element_text(size = 15)  # Increase axis label font size
            ) 
        + scale_fill_manual(values=colors)   
    )

    if (!is.null(fig_filename)){
        ggsave(fig_filename, p, width=8, height=6)
    }

    return(p)
}
```

```{r}
library(Seurat)
library(ggplot2)
library(dplyr)
library(tidyr)
library(qs)
library(tibble)

caudate_neurons = qread("Current_subclusters/caudate_neuron.qs")
putamen_neurons= qread("Current_subclusters/putamen_neuron.qs")

caudate_neurons
putamen_neurons
```

```{r}
caudate_neurons@meta.data$sub_class = caudate_neurons@meta.data$cell_class
caudate_neurons$sub_class[caudate_neurons$sub_class == "D1_SPN_1"] = "SPN"
caudate_neurons$sub_class[caudate_neurons$sub_class == "interneuron_2"] = "interneuron"
caudate_neurons$sub_class[caudate_neurons$sub_class == "D1_SPN_matrix_1"] = "SPN"
caudate_neurons$sub_class[caudate_neurons$sub_class == "D2_SPN_matrix_1"] = "SPN"
caudate_neurons$sub_class[caudate_neurons$sub_class == "SPN"] = "SPN"
caudate_neurons$sub_class[caudate_neurons$sub_class == "D1_SPN_2"] = "SPN"
caudate_neurons$sub_class[caudate_neurons$sub_class == "D2_SPN_1"] = "SPN"
caudate_neurons$sub_class[caudate_neurons$sub_class == "interneuron_3"] = "interneuron"
caudate_neurons$sub_class[caudate_neurons$sub_class == "D1_SPN_matrix_2"] = "SPN"
caudate_neurons$sub_class[caudate_neurons$sub_class == "D1_SPN_patch"] = "SPN"
caudate_neurons$sub_class[caudate_neurons$sub_class == "interneuron_1"] = "interneuron"
caudate_neurons$sub_class[caudate_neurons$sub_class == "eSPN"] = "eSPN"
caudate_neurons$sub_class[caudate_neurons$sub_class == "interneuron_5"] = "interneuron"
caudate_neurons$sub_class[caudate_neurons$sub_class == "cholinergic"] = "cholinergic"
caudate_neurons$sub_class[caudate_neurons$sub_class == "interneuron_4"] = "interneuron"
caudate_neurons$sub_class[caudate_neurons$sub_class == "interneuron_6"] = "interneuron"

caudate_neurons@meta.data
```

#putamen
```{r}
putamen_neurons@meta.data$sub_class = putamen_neurons@meta.data$cell_class
putamen_neurons$sub_class[putamen_neurons$sub_class == "D2_SPN_matrix"] = "SPN"
putamen_neurons$sub_class[putamen_neurons$sub_class == "interneuron_2"] = "interneuron"
putamen_neurons$sub_class[putamen_neurons$sub_class == "interneuron_4"] = "interneuron"
putamen_neurons$sub_class[putamen_neurons$sub_class == "D1_SPN_matrix_1"] = "SPN"
putamen_neurons$sub_class[putamen_neurons$sub_class == "glutamatergic"] = "glutamatergic"
putamen_neurons$sub_class[putamen_neurons$sub_class == "interneuron_5"] = "interneuron"
putamen_neurons$sub_class[putamen_neurons$sub_class == "D1_SPN_1"] = "SPN"
putamen_neurons$sub_class[putamen_neurons$sub_class == "D1_SPN_patch"] = "SPN"
putamen_neurons$sub_class[putamen_neurons$sub_class == "D2_SPN_1"] = "SPN"
putamen_neurons$sub_class[putamen_neurons$sub_class == "eSPN"] = "eSPN"
putamen_neurons$sub_class[putamen_neurons$sub_class == "interneuron_3"] = "interneuron"
putamen_neurons$sub_class[putamen_neurons$sub_class == "interneuron_1"] = "interneuron"
putamen_neurons$sub_class[putamen_neurons$sub_class == "D2_SPN_2"] = "SPN"
putamen_neurons$sub_class[putamen_neurons$sub_class == "cholinergic"] = "cholinergic"
putamen_neurons$sub_class[putamen_neurons$sub_class == "interneuron_7"] = "interneuron"
putamen_neurons$sub_class[putamen_neurons$sub_class == "interneuron_6"] = "interneuron"

putamen_neurons@meta.data
```

```{r}
SPN_cah= subset(caudate_neurons, subset = sub_class == "SPN")
non_SPN_cah = subset(caudate_neurons, subset = sub_class != "SPN")

SPN_put= subset(putamen_neurons, subset = sub_class == "SPN")
non_SPN_put= subset(putamen_neurons, subset = sub_class != "SPN")

```

```{r}
SPN_cah
non_SPN_cah

SPN_put
non_SPN_put
```

#SC Transform SPNs and other neurons- did SCtransform on all neurons first before I subsetted
```{r}
library(sctransform)
caudate_neurons_transformed = SCTransform(caudate_neurons, vars.to.regress = "pct_mito", verbose = FALSE)
DefaultAssay(caudate_neurons_transformed) = "SCT"

SPN_cah_transformed= subset(caudate_neurons, subset = sub_class == "SPN")
non_SPN_cah_transformed= subset(caudate_neurons, subset = sub_class != "SPN")
SPN_cah_transformed
non_SPN_cah_transformed

putamen_neurons_transformed = SCTransform(putamen_neurons, vars.to.regress = "pct_mito", verbose = FALSE)
DefaultAssay(putamen_neurons_transformed) = "SCT"

SPN_put_transformed= subset(putamen_neurons, subset = sub_class == "SPN")
non_SPN_put_transformed= subset(putamen_neurons, subset = sub_class != "SPN")

SPN_put_transformed
non_SPN_put_transformed
```

#DE csvs
```{r}
CaH_ALL_SPN_DE = read.csv("~/GSEA/de/SPN_DE/XDP DE - Caudate Neurons - 20240603 - All SPN.csv")
Put_ALL_SPN_DE = read.csv("~/GSEA/de/SPN_DE/XDP DE - Putamen Neurons - 20240603 - All SPN.csv")

CaH_ALL_SPN_DE
Put_ALL_SPN_DE

#significant DE only
CaH_ALL_SPN_DE_sig = CaH_ALL_SPN_DE[CaH_ALL_SPN_DE$adj.P.Val < 0.05,]
Put_ALL_SPN_DE_sig = Put_ALL_SPN_DE[Put_ALL_SPN_DE$adj.P.Val < 0.05,]

CaH_ALL_SPN_DE_sig
Put_ALL_SPN_DE_sig

```


#CaH SPN
```{r}
CaH_ALL_SPN_DE_sig = CaH_ALL_SPN_DE[CaH_ALL_SPN_DE$adj.P.Val < 0.05,]
significant_genes = unique(CaH_ALL_SPN_DE_sig$gene)

exp = FetchData(SPN_cah_transformed, vars = significant_genes)
exp

CaH_ALL_SPN_DE_sig

rownames(CaH_ALL_SPN_DE_sig) = CaH_ALL_SPN_DE_sig$gene
CaH_ALL_SPN_DE_sig$gene = NULL
exp

# Assuming CaH_ALL_SPN_DE_sig and exp are your dataframes
logfc_df <- CaH_ALL_SPN_DE_sig
counts_df <- exp

logfc_df
counts_df

# Check if all column names in counts_df are in the row names of logfc_df
if (!all(colnames(counts_df) %in% rownames(logfc_df))) {
  stop("Not all genes in counts_df are present in logfc_df")
}

# Reorder logfc_df to match the columns in counts_df
logfc_ordered <- logfc_df[colnames(counts_df), , drop = FALSE]
logfc_ordered

# Check for NA values in logfc_ordered
if (any(is.na(logfc_ordered$logFC))) {
  stop("There are NA values in the logFC values for the genes")
}

# Ensure logfc_ordered$logfc is a numeric vector
logfc_vector <- as.numeric(logfc_ordered$logFC)
logfc_vector

#THE ACTUAL
result_df <- sweep(counts_df, 2, logfc_vector, FUN = "*")

# Print the result
print(result_df)

metadata <- SPN_cah_transformed@meta.data
# Assume that metadata has a column 'donor' that contains donor information
# If not, modify this part according to your metadata structure
if(!"donor_id" %in% colnames(metadata)) {
  stop("Metadata does not contain 'donor' information. Please check your metadata structure.")
}

# Create unique identifiers for cell column to avoid conflicts
expr_data <- result_df %>%
  rownames_to_column(var = "cell_id")

metadata <- metadata %>%
  rownames_to_column(var = "cell_id")

# Combine expression data with donor information
expr_data_long <- expr_data %>%
  pivot_longer(cols = -cell_id, names_to = "gene", values_to = "expression") %>%
  left_join(metadata %>% select(cell_id, donor_id, Condition), by = "cell_id")
# Assuming your data frame is named expr_data_long
expr_data_long
# Count the number of occurrences for each gene per donor

summarized_expr_data <- expr_data_long %>%
  group_by(cell_id, donor_id,Condition) %>%
  summarize(total_expression = sum(expression), .groups = 'drop')
CaH_SPN_summary_expr_data = summarized_expr_data
CaH_SPN_summary_expr_data
```


#CaH nonSPN
```{r}
CaH_ALL_SPN_DE_sig = CaH_ALL_SPN_DE[CaH_ALL_SPN_DE$adj.P.Val < 0.05,]
significant_genes = unique(CaH_ALL_SPN_DE_sig$gene)

exp = FetchData(non_SPN_cah_transformed, vars = significant_genes)
exp

CaH_ALL_SPN_DE_sig

rownames(CaH_ALL_SPN_DE_sig) = CaH_ALL_SPN_DE_sig$gene
CaH_ALL_SPN_DE_sig$gene = NULL
exp

# Assuming CaH_ALL_SPN_DE_sig and exp are your dataframes
logfc_df <- CaH_ALL_SPN_DE_sig
counts_df <- exp

logfc_df
counts_df

# Check if all column names in counts_df are in the row names of logfc_df
if (!all(colnames(counts_df) %in% rownames(logfc_df))) {
  stop("Not all genes in counts_df are present in logfc_df")
}

# Reorder logfc_df to match the columns in counts_df
logfc_ordered <- logfc_df[colnames(counts_df), , drop = FALSE]
logfc_ordered

# Check for NA values in logfc_ordered
if (any(is.na(logfc_ordered$logFC))) {
  stop("There are NA values in the logFC values for the genes")
}

# Ensure logfc_ordered$logfc is a numeric vector
logfc_vector <- as.numeric(logfc_ordered$logFC)
logfc_vector

#THE ACTUAL
result_df <- sweep(counts_df, 2, logfc_vector, FUN = "*")

# Print the result
print(result_df)

metadata <- non_SPN_cah_transformed@meta.data
# Assume that metadata has a column 'donor' that contains donor information
# If not, modify this part according to your metadata structure
if(!"donor_id" %in% colnames(metadata)) {
  stop("Metadata does not contain 'donor' information. Please check your metadata structure.")
}

# Create unique identifiers for cell column to avoid conflicts
expr_data <- result_df %>%
  rownames_to_column(var = "cell_id")

metadata <- metadata %>%
  rownames_to_column(var = "cell_id")

# Combine expression data with donor information
expr_data_long <- expr_data %>%
  pivot_longer(cols = -cell_id, names_to = "gene", values_to = "expression") %>%
  left_join(metadata %>% select(cell_id, donor_id, Condition), by = "cell_id")
# Assuming your data frame is named expr_data_long
expr_data_long
# Count the number of occurrences for each gene per donor

summarized_expr_data <- expr_data_long %>%
  group_by(cell_id, donor_id,Condition) %>%
  summarize(total_expression = sum(expression), .groups = 'drop')
CaH_non_SPN_summary_expr_data = summarized_expr_data
CaH_non_SPN_summary_expr_data
```


```{r}
CaH_SPN_summary_expr_data$cell_type = "SPN"
CaH_non_SPN_summary_expr_data$cell_type = "non-SPN"

head(CaH_SPN_summary_expr_data)
head(CaH_non_SPN_summary_expr_data)

CaH_Transcriptional_DE = rbind(CaH_SPN_summary_expr_data, CaH_non_SPN_summary_expr_data)
CaH_Transcriptional_DE
```

```{r}
library(gridExtra)
options(repr.plot.width=24, repr.plot.height=16)
plots = list()

donors = c("PCMC-16-011", "PCMC-16-012", "SCF-18-003", "SCF-18-004", "SCF-18-006","SCF-19-009", "SCF-19-014", "SCF-19-018",  "SCF-20-023",  "SCF_20-024", "SCF-20-025", "SCF-21-030", "SCF_22-043", "SCF_21-037CM2","SCF-22-054CM","SCF-22-058CF","SCF-23-068CM")

for (donor in donors) {
test = CaH_Transcriptional_DE[CaH_Transcriptional_DE$donor_id == donor,]
   plots[[donor]] = plot_overlapping_density_histogram(df = test, 
    hist_col = "total_expression",
    fill_col = "cell_type",
    colors = c("blue", "red"),
    #alpha=0.5,
    breaks=seq(-400,400,10),
    title= paste("CaH SPNs vs non-SPNs: ", donor),
    xlab = "CaH DE Score",
    fig_filename = NULL)
}

layout_matrix <- rbind(
  c(1, 2, 3, 4, 5),
  c(6, 7, 8, 9, 10),
  c(11, 12, 13, NA, NA),
  c(14, 15, 16, 17, NA)
)

# Arrange the plots according to the custom layout
grid_plots <- grid.arrange(grobs = plots, layout_matrix = layout_matrix)

# Save the arranged plots to a PNG file
ggsave(filename = "output/CAH_TranscriptionalDE_donors.png", plot = grid_plots, width = 30, height = 16)

```

```{r}
XDP = CaH_Transcriptional_DE[CaH_Transcriptional_DE$Condition == "XDP",]
plot_overlapping_density_histogram(df = XDP, 
    hist_col = "total_expression",
    fill_col = "cell_type",
    colors = c("blue", "red"),
    #alpha=0.5,
    breaks=seq(-400,400,10),
    title= paste("CaH SPNs vs non-SPNs: XDP Combined"),
    xlab = "CaH DE score",
    fig_filename = NULL)

Control = CaH_Transcriptional_DE[CaH_Transcriptional_DE$Condition == "Control",]
plot_overlapping_density_histogram(df = Control, 
    hist_col = "total_expression",
    fill_col = "cell_type",
    colors = c("blue", "red"),
    #alpha=0.5,
    breaks=seq(-400,400,10),
    title= paste("CaH SPNs vs non-SPNs: Control Combined"),
    xlab = "CaH DE score",
    fig_filename = NULL)
```


```{r}
SPNs = CaH_Transcriptional_DE[CaH_Transcriptional_DE$cell_type == "SPN",]
plot_overlapping_density_histogram(df = SPNs, 
    hist_col = "total_expression",
    fill_col = "Condition",
    colors = c("blue", "red"),
    #alpha=0.5,
    breaks=seq(-400,400,10),
    title= paste("CaH SPNs: XDP vs Control"),
    xlab = "CaH DE score",
    fig_filename = NULL)

nonSPNs = CaH_Transcriptional_DE[CaH_Transcriptional_DE$cell_type == "non-SPN",]
plot_overlapping_density_histogram(df = nonSPNs, 
    hist_col = "total_expression",
    fill_col = "Condition",
    colors = c("blue", "red"),
    #alpha=0.5,
    breaks=seq(-400,400,10),
    title= paste("CaH non-SPNs: XDP vs Control"),
    xlab = "CaH DE score",
    fig_filename = NULL)
```


#PUT TRANSCRIPTIONAL DE

#Put SPN
```{r}
Put_ALL_SPN_DE_sig = Put_ALL_SPN_DE[Put_ALL_SPN_DE$adj.P.Val < 0.05,]
significant_genes = unique(Put_ALL_SPN_DE_sig$gene)

exp = FetchData(SPN_put_transformed, vars = significant_genes)
exp

Put_ALL_SPN_DE_sig

rownames(Put_ALL_SPN_DE_sig) = Put_ALL_SPN_DE_sig$gene
Put_ALL_SPN_DE_sig$gene = NULL
exp

# Assuming Put_ALL_SPN_DE_sig and exp are your dataframes
logfc_df <- Put_ALL_SPN_DE_sig
counts_df <- exp

logfc_df
counts_df

# Check if all column names in counts_df are in the row names of logfc_df
if (!all(colnames(counts_df) %in% rownames(logfc_df))) {
  stop("Not all genes in counts_df are present in logfc_df")
}

# Reorder logfc_df to match the columns in counts_df
logfc_ordered <- logfc_df[colnames(counts_df), , drop = FALSE]
logfc_ordered

# Check for NA values in logfc_ordered
if (any(is.na(logfc_ordered$logFC))) {
  stop("There are NA values in the logFC values for the genes")
}

# Ensure logfc_ordered$logfc is a numeric vector
logfc_vector <- as.numeric(logfc_ordered$logFC)
logfc_vector

#THE ACTUAL
result_df <- sweep(counts_df, 2, logfc_vector, FUN = "*")

# Print the result
print(result_df)

metadata <- SPN_put_transformed@meta.data
# Assume that metadata has a column 'donor' that contains donor information
# If not, modify this part according to your metadata structure
if(!"donor_id" %in% colnames(metadata)) {
  stop("Metadata does not contain 'donor' information. Please check your metadata structure.")
}

# Create unique identifiers for cell column to avoid conflicts
expr_data <- result_df %>%
  rownames_to_column(var = "cell_id")

metadata <- metadata %>%
  rownames_to_column(var = "cell_id")

# Combine expression data with donor information
expr_data_long <- expr_data %>%
  pivot_longer(cols = -cell_id, names_to = "gene", values_to = "expression") %>%
  left_join(metadata %>% select(cell_id, donor_id, Condition), by = "cell_id")
# Assuming your data frame is named expr_data_long
expr_data_long
# Count the number of occurrences for each gene per donor

summarized_expr_data <- expr_data_long %>%
  group_by(cell_id, donor_id,Condition) %>%
  summarize(total_expression = sum(expression), .groups = 'drop')
Put_SPN_summary_expr_data = summarized_expr_data
Put_SPN_summary_expr_data
```


#Put nonSPN
```{r}
Put_ALL_SPN_DE_sig = Put_ALL_SPN_DE[Put_ALL_SPN_DE$adj.P.Val < 0.05,]
significant_genes = unique(Put_ALL_SPN_DE_sig$gene)

exp = FetchData(non_SPN_put_transformed, vars = significant_genes)
exp

Put_ALL_SPN_DE_sig

rownames(Put_ALL_SPN_DE_sig) = Put_ALL_SPN_DE_sig$gene
Put_ALL_SPN_DE_sig$gene = NULL
exp

# Assuming Put_ALL_SPN_DE_sig and exp are your dataframes
logfc_df <- Put_ALL_SPN_DE_sig
counts_df <- exp

logfc_df
counts_df

# Check if all column names in counts_df are in the row names of logfc_df
if (!all(colnames(counts_df) %in% rownames(logfc_df))) {
  stop("Not all genes in counts_df are present in logfc_df")
}

# Reorder logfc_df to match the columns in counts_df
logfc_ordered <- logfc_df[colnames(counts_df), , drop = FALSE]
logfc_ordered

# Check for NA values in logfc_ordered
if (any(is.na(logfc_ordered$logFC))) {
  stop("There are NA values in the logFC values for the genes")
}

# Ensure logfc_ordered$logfc is a numeric vector
logfc_vector <- as.numeric(logfc_ordered$logFC)
logfc_vector

#THE ACTUAL
result_df <- sweep(counts_df, 2, logfc_vector, FUN = "*")

# Print the result
print(result_df)

metadata <- non_SPN_put_transformed@meta.data
# Assume that metadata has a column 'donor' that contains donor information
# If not, modify this part according to your metadata structure
if(!"donor_id" %in% colnames(metadata)) {
  stop("Metadata does not contain 'donor' information. Please check your metadata structure.")
}

# Create unique identifiers for cell column to avoid conflicts
expr_data <- result_df %>%
  rownames_to_column(var = "cell_id")

metadata <- metadata %>%
  rownames_to_column(var = "cell_id")

# Combine expression data with donor information
expr_data_long <- expr_data %>%
  pivot_longer(cols = -cell_id, names_to = "gene", values_to = "expression") %>%
  left_join(metadata %>% select(cell_id, donor_id, Condition), by = "cell_id")
# Assuming your data frame is named expr_data_long
expr_data_long
# Count the number of occurrences for each gene per donor

summarized_expr_data <- expr_data_long %>%
  group_by(cell_id, donor_id,Condition) %>%
  summarize(total_expression = sum(expression), .groups = 'drop')
Put_non_SPN_summary_expr_data = summarized_expr_data
Put_non_SPN_summary_expr_data
```


```{r}
Put_SPN_summary_expr_data$cell_type = "SPN"
Put_non_SPN_summary_expr_data$cell_type = "non-SPN"

head(Put_SPN_summary_expr_data)
head(Put_non_SPN_summary_expr_data)

Put_Transcriptional_DE = rbind(Put_SPN_summary_expr_data, Put_non_SPN_summary_expr_data)
Put_Transcriptional_DE
```

```{r}
library(gridExtra)
options(repr.plot.width=24, repr.plot.height=16)
plots = list()

donors = c("PCMC-16-011", "PCMC-16-012", "SCF-18-003", "SCF-18-004", "SCF-18-006","SCF-19-009", "SCF-19-014", "SCF-19-018",  "SCF-20-023",  "SCF_20-024", "SCF-20-025", "SCF-21-030", "SCF_22-043", "SCF_21-037CM2","SCF-22-054CM","SCF-22-058CF","SCF-23-068CM")

for (donor in donors) {
test = Put_Transcriptional_DE[Put_Transcriptional_DE$donor_id == donor,]
   plots[[donor]] = plot_overlapping_density_histogram(df = test, 
    hist_col = "total_expression",
    fill_col = "cell_type",
    colors = c("blue", "red"),
    #alpha=0.5,
    breaks=seq(-200,200,5),
    title= paste("Put SPNs vs non-SPNs: ", donor),
    xlab = "Put DE Score",
    fig_filename = NULL)
}

layout_matrix <- rbind(
  c(1, 2, 3, 4, 5),
  c(6, 7, 8, 9, 10),
  c(11, 12, 13, NA, NA),
  c(14, 15, 16, 17, NA)
)

# Arrange the plots according to the custom layout
grid_plots <- grid.arrange(grobs = plots, layout_matrix = layout_matrix)

# Save the arranged plots to a PNG file
ggsave(filename = "output/Put_TranscriptionalDE_donors.png", plot = grid_plots, width = 30, height = 16)

```

```{r}
XDP = Put_Transcriptional_DE[Put_Transcriptional_DE$Condition == "XDP",]
plot_overlapping_density_histogram(df = XDP, 
    hist_col = "total_expression",
    fill_col = "cell_type",
    colors = c("blue", "red"),
    #alpha=0.5,
    breaks=seq(-200,200,5),
    title= paste("Put SPNs vs non-SPNs: XDP Combined"),
    xlab = "Put DE score",
    fig_filename = NULL)

Control = Put_Transcriptional_DE[Put_Transcriptional_DE$Condition == "Control",]
plot_overlapping_density_histogram(df = Control, 
    hist_col = "total_expression",
    fill_col = "cell_type",
    colors = c("blue", "red"),
    #alpha=0.5,
    breaks=seq(-200,200,5),
    title= paste("Put SPNs vs non-SPNs: Control Combined"),
    xlab = "Put DE score",
    fig_filename = NULL)
```


```{r}
SPNs = Put_Transcriptional_DE[Put_Transcriptional_DE$cell_type == "SPN",]
plot_overlapping_density_histogram(df = SPNs, 
    hist_col = "total_expression",
    fill_col = "Condition",
    colors = c("blue", "red"),
    #alpha=0.5,
    breaks=seq(-200,200,5),
    title= paste("Put SPNs: XDP vs Control"),
    xlab = "Put DE score",
    fig_filename = NULL)

nonSPNs = Put_Transcriptional_DE[Put_Transcriptional_DE$cell_type == "non-SPN",]
plot_overlapping_density_histogram(df = nonSPNs, 
    hist_col = "total_expression",
    fill_col = "Condition",
    colors = c("blue", "red"),
    #alpha=0.5,
    breaks=seq(-200,200,5),
    title= paste("Put non-SPNs: XDP vs Control"),
    xlab = "Put DE score",
    fig_filename = NULL)
```




#SPN-ness score
#Findallmarkers on controls: SPN vs non-SPNs 
```{r}
caudate_neurons$sub_class[caudate_neurons$sub_class == "interneuron"] = "non-SPN"
caudate_neurons$sub_class[caudate_neurons$sub_class == "eSPN"] = "non-SPN"
caudate_neurons$sub_class[caudate_neurons$sub_class == "cholinergic"] = "non-SPN"

Idents(caudate_neurons) = "sub_class"
DimPlot(caudate_neurons, label = TRUE)
```

```{r}
caudate_controls = subset(caudate_neurons, subset = Condition == "Control")
caudate_controls
Idents(caudate_controls) = "sub_class"
DimPlot(caudate_controls, label = TRUE)
```


```{r}
Idents(caudate_controls) = caudate_controls$sub_class
caudate_controls_markers = FindAllMarkers(caudate_controls)
                                          #, only.pos = TRUE,  min.pct = 0.2, logfc.threshold = 1.25)
caudate_controls_markers

CaH_SPN_markers_from_SPN = subset(caudate_controls_markers, subset= cluster =="SPN")
CaH_SPN_markers_from_SPN1 =subset(CaH_SPN_markers_from_SPN, subset = pct.2 < 0.2)
CaH_SPN_markers_from_SPN2 =subset(CaH_SPN_markers_from_SPN1, subset = pct.1 > 0.6)

CaH_SPN_markers_from_SPN
CaH_SPN_markers_from_SPN1
CaH_SPN_markers_from_SPN2
```


```{r}
CaH_SPN_markers_from_non = subset(caudate_controls_markers, subset= cluster =="non-SPN")
CaH_SPN_markers_from_non1 =subset(CaH_SPN_markers_from_non, subset = pct.2 > 0.6)
CaH_SPN_markers_from_non2 =subset(CaH_SPN_markers_from_non1, subset = pct.1 < 0.2)

CaH_SPN_markers_from_non
CaH_SPN_markers_from_non1
CaH_SPN_markers_from_non2
```



```{r}
caudate_controls_markers[caudate_controls_markers$gene = "PPPRB1B",]
```





```{r}
features <- final$gene
DotPlot(caudate_controls, features = features, dot.scale = 8) + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
nonSPN_cah_transformed= subset(caudate_neurons, subset = sub_class != "SPN")

```

```{r}
SPN_cah_transformed
nonSPN_cah_transformed
```


#CaH SPN
```{r}
significant_genes = unique(final$gene)

exp = FetchData(SPN_cah_transformed, vars = significant_genes)
exp

final

rownames(final) = final$gene
final$gene = NULL
exp

# Assuming final and exp are your dataframes
logfc_df <- final
counts_df <- exp

logfc_df
counts_df

# Check if all column names in counts_df are in the row names of logfc_df
if (!all(colnames(counts_df) %in% rownames(logfc_df))) {
  stop("Not all genes in counts_df are present in logfc_df")
}

# Reorder logfc_df to match the columns in counts_df
logfc_ordered <- logfc_df[colnames(counts_df), , drop = FALSE]
logfc_ordered

# Check for NA values in logfc_ordered
if (any(is.na(logfc_ordered$avg_log2FC))) {
  stop("There are NA values in the logFC values for the genes")
}

# Ensure logfc_ordered$logfc is a numeric vector
logfc_vector <- as.numeric(logfc_ordered$avg_log2FC)
logfc_vector
 #THIS IS FOR LOG TRANSFORMATINO BUT I DONT LIKE IT
 custom_log_transform <- function(x, logFC_value) {
      x * 2^logFC_value
  }
  
# Multiply each column in counts_df by the corresponding logFC value
 result_df <- sweep(counts_df, 2, logfc_vector, FUN = custom_log_transform)


#THE ACTUAL
 #result_df <- sweep(counts_df, 2, logfc_vector, FUN = "*")

# Print the result
print(result_df)

metadata <- SPN_cah_transformed@meta.data
# Assume that metadata has a column 'donor' that contains donor information
# If not, modify this part according to your metadata structure
if(!"donor_id" %in% colnames(metadata)) {
  stop("Metadata does not contain 'donor' information. Please check your metadata structure.")
}

# Create unique identifiers for cell column to avoid conflicts
expr_data <- result_df %>%
  rownames_to_column(var = "cell_id")

metadata <- metadata %>%
  rownames_to_column(var = "cell_id")

# Combine expression data with donor information
expr_data_long <- expr_data %>%
  pivot_longer(cols = -cell_id, names_to = "gene", values_to = "expression") %>%
  left_join(metadata %>% select(cell_id, donor_id), by = "cell_id")
# Assuming your data frame is named expr_data_long
expr_data_long
# Count the number of occurrences for each gene per donor

summarized_expr_data <- expr_data_long %>%
  group_by(donor_id, gene) %>%
  summarize(total_expression = sum(expression), .groups = 'drop')
summarized_expr_data
CaH_SPN_summary_expr_data = summarized_expr_data
CaH_SPN_summary_expr_data
```

#CAH nonSPN
```{r}
significant_genes = unique(final$gene)

exp = FetchData(nonSPN_cah_transformed, vars = significant_genes)
exp

final

rownames(final) = final$gene
final$gene = NULL
exp

# Assuming final and exp are your dataframes
logfc_df <- final
counts_df <- exp

logfc_df
counts_df

# Check if all column names in counts_df are in the row names of logfc_df
if (!all(colnames(counts_df) %in% rownames(logfc_df))) {
  stop("Not all genes in counts_df are present in logfc_df")
}

# Reorder logfc_df to match the columns in counts_df
logfc_ordered <- logfc_df[colnames(counts_df), , drop = FALSE]
logfc_ordered

# Check for NA values in logfc_ordered
if (any(is.na(logfc_ordered$avg_log2FC))) {
  stop("There are NA values in the logFC values for the genes")
}

# Ensure logfc_ordered$logfc is a numeric vector
logfc_vector <- as.numeric(logfc_ordered$avg_log2FC)
logfc_vector
 #THIS IS FOR LOG TRANSFORMATINO BUT I DONT LIKE IT
 custom_log_transform <- function(x, logFC_value) {
      x * 2^logFC_value
  }
  
# Multiply each column in counts_df by the corresponding logFC value
 result_df <- sweep(counts_df, 2, logfc_vector, FUN = custom_log_transform)


#THE ACTUAL
 #result_df <- sweep(counts_df, 2, logfc_vector, FUN = "*")

# Print the result
print(result_df)

metadata <- nonSPN_cah_transformed@meta.data
# Assume that metadata has a column 'donor' that contains donor information
# If not, modify this part according to your metadata structure
if(!"donor_id" %in% colnames(metadata)) {
  stop("Metadata does not contain 'donor' information. Please check your metadata structure.")
}

# Create unique identifiers for cell column to avoid conflicts
expr_data <- result_df %>%
  rownames_to_column(var = "cell_id")

metadata <- metadata %>%
  rownames_to_column(var = "cell_id")

# Combine expression data with donor information
expr_data_long <- expr_data %>%
  pivot_longer(cols = -cell_id, names_to = "gene", values_to = "expression") %>%
  left_join(metadata %>% select(cell_id, donor_id), by = "cell_id")
# Assuming your data frame is named expr_data_long
expr_data_long
# Count the number of occurrences for each gene per donor

summarized_expr_data <- expr_data_long %>%
  group_by(donor_id, gene) %>%
  summarize(total_expression = sum(expression), .groups = 'drop')
summarized_expr_data
CaH_nonSPN_summary_expr_data = summarized_expr_data
CaH_nonSPN_summary_expr_data
```


```{r}
CaH_SPN_summary_expr_data$cell_type = "SPN"
CaH_nonSPN_summary_expr_data$cell_type = "non-SPN" 

head(CaH_SPN_summary_expr_data)
head(CaH_nonSPN_summary_expr_data)
```

```{r}
CaH_Transcriptional_DE = rbind(CaH_SPN_summary_expr_data, CaH_nonSPN_summary_expr_data)
CaH_Transcriptional_DE
```

```{r}
donors = unique(CaH_Transcriptional_DE$donor_id)

for (donor in donors) {
test = CaH_Transcriptional_DE[CaH_Transcriptional_DE$donor_id == donor,]
a = plot_overlapping_density_histogram(df = test, 
    hist_col = "total_expression",
    fill_col = "cell_type",
    colors = c("blue", "red"),
    #alpha=0.5,
    breaks=seq(0,1000,10),
    title= paste("CaH SPNs vs non-SPNs (log transformed): ", donor),
    xlab = "CaH SPN score",
    fig_filename = NULL)

print(a)
}
```

```{r}
CaH_Transcriptional_DE$Condition = ifelse(grepl("SCF_21-037CM2|SCF-23-068CM|SCF-22-058CF|SCF-22-054CM", CaH_Transcriptional_DE$donor_id), "Control", "XDP")
CaH_Transcriptional_DE
```


```{r}
XDP = CaH_Transcriptional_DE[CaH_Transcriptional_DE$Condition == "XDP",]
plot_overlapping_density_histogram(df = XDP, 
    hist_col = "total_expression",
    fill_col = "cell_type",
    colors = c("blue", "red"),
    #alpha=0.5,
    breaks=seq(0,1000,10),
    title= paste("CaH SPNs vs non-SPNs (log transformed): XDP"),
    xlab = "CaH SPN score",
    fig_filename = NULL)

Control = CaH_Transcriptional_DE[CaH_Transcriptional_DE$Condition == "Control",]
plot_overlapping_density_histogram(df = Control, 
    hist_col = "total_expression",
    fill_col = "cell_type",
    colors = c("blue", "red"),
    #alpha=0.5,
    breaks=seq(0,1000,10),
    title= paste("CaH SPNs vs non-SPNs (log transformed): Control "),
    xlab = "CaH SPN score",
    fig_filename = NULL)
```



```{r}
CaH_Transcriptional_DE
```





#plotting pct change on x axis, plotting logfc on y

```{r}
Cah_markers = final
Cah_markers$pct.1_divided_pct.2 = Cah_markers$pct.1/Cah_markers$pct.2
Cah_markers
```

```{r}
significant_genes = unique(rownames(Cah_markers))

CaH_SPN_exp = FetchData(SPN_cah_transformed, vars = significant_genes)
CaH_SPN_exp

metadata <- SPN_cah_transformed@meta.data
if(!"donor_id" %in% colnames(metadata)) {
  stop("Metadata does not contain 'donor' information. Please check your metadata structure.")
}

# Create unique identifiers for cell column to avoid conflicts
CaH_SPN_expr_data <- CaH_SPN_exp %>%
  rownames_to_column(var = "cell_id")

metadata <- metadata %>%
  rownames_to_column(var = "cell_id")

# Combine expression data with donor information
CaH_SPN_expr_data_long <- CaH_SPN_expr_data %>%
  pivot_longer(cols = -cell_id, names_to = "gene", values_to = "expression") %>%
  left_join(metadata %>% select(cell_id, donor_id), by = "cell_id")
# Assuming your data frame is named expr_data_long
CaH_SPN_expr_data_long



CaH_nonSPN_exp = FetchData(nonSPN_cah_transformed, vars = significant_genes)
CaH_nonSPN_exp


metadata <- nonSPN_cah_transformed@meta.data
if(!"donor_id" %in% colnames(metadata)) {
  stop("Metadata does not contain 'donor' information. Please check your metadata structure.")
}

# Create unique identifiers for cell column to avoid conflicts
CaH_nonSPN_expr_data <- CaH_nonSPN_exp %>%
  rownames_to_column(var = "cell_id")

metadata <- metadata %>%
  rownames_to_column(var = "cell_id")

# Combine expression data with donor information
CaH_nonSPN_expr_data_long <- CaH_nonSPN_expr_data %>%
  pivot_longer(cols = -cell_id, names_to = "gene", values_to = "expression") %>%
  left_join(metadata %>% select(cell_id, donor_id), by = "cell_id")
# Assuming your data frame is named expr_data_long
CaH_nonSPN_expr_data_long
```

```{r}
Cah_markers$gene = rownames(Cah_markers)
CaH_SPNs = merge(CaH_SPN_expr_data_long, Cah_markers, by = "gene")
CaH_SPNs$Condition = ifelse(grepl("SCF_21-037CM2|SCF-23-068CM|SCF-22-058CF|SCF-22-054CM", CaH_SPNs$donor_id), "Control", "XDP")
CaH_SPNs
```

```{r}
summarized_expr_data_TEST <- CaH_SPNs %>%
  group_by(donor_id, gene) %>%
  summarize(total_expression = sum(expression), .groups = 'drop')
summarized_expr_data_TEST

summarized_expr_data_TEST = merge(summarized_expr_data_TEST, Cah_markers, by = "gene")
summarized_expr_data_TEST$Condition = ifelse(grepl("SCF_21-037CM2|SCF-23-068CM|SCF-22-058CF|SCF-22-054CM", summarized_expr_data_TEST$donor_id), "Control", "XDP")
summarized_expr_data_TEST
```

```{r}
TEST = subset(summarized_expr_data_TEST, summarized_expr_data_TEST)

ggplot(data=summarized_expr_data_TEST, aes(x= pct.1_divided_pct.2, y= avg_log2FC, color = donor_id)) + geom_point() +  geom_text(aes(label = gene), vjust = -0.5, hjust = 0.5)
```



```{r}
CaH_SPNs_test = subset(CaH_SPNs, CaH_SPNs$donor_id == "SCF-22-054CM")

ggplot(data= CaH_SPNs, aes(x=gene , y=expression, fill = gene)) + geom_boxplot()
```



```{r}
CaH_Transcriptional_DE
```












#PUTAMEN SPN-ness score
#Findallmarkers on controls: SPN vs non-SPNs 
```{r}
putamen_neurons$sub_class[putamen_neurons$sub_class == "interneuron"] = "non-SPN"
putamen_neurons$sub_class[putamen_neurons$sub_class == "eSPN"] = "non-SPN"
putamen_neurons$sub_class[putamen_neurons$sub_class == "cholinergic"] = "non-SPN"
putamen_neurons$sub_class[putamen_neurons$sub_class == "glutamatergic"] = "non-SPN"

Idents(putamen_neurons) = "sub_class"
DimPlot(putamen_neurons, label = TRUE)
```

```{r}
putamen_controls = subset(putamen_neurons, subset = Condition == "Control")
putamen_controls
Idents(putamen_controls) = "sub_class"
DimPlot(putamen_controls, label = TRUE)
```
```{r}
Idents(putamen_controls) = putamen_controls$sub_class
putamen_controls_markers = FindAllMarkers(putamen_controls)
                                          #, only.pos = TRUE,  min.pct = 0.2, logfc.threshold = 1.25)
putamen_controls_markers
```


```{r}
subsetmarkers= subset(putamen_controls_markers, subset = pct.2 < 0.3)
subsetmarkers

newmarker = subset(subsetmarkers, subset = pct.1> 0.7)
newmarker

final = subset(newmarker, subset= cluster =="SPN")
final
```

```{r}
features <- final$gene
DotPlot(putamen_controls, features = features, dot.scale = 8) + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
nonSPN_put_transformed= subset(putamen_neurons, subset = sub_class != "SPN")

```

```{r}
SPN_put_transformed
nonSPN_put_transformed
```


#put SPN
```{r}
final = subset(newmarker, subset= cluster =="SPN")

significant_genes = unique(final$gene)

exp = FetchData(SPN_put_transformed, vars = significant_genes)
exp

final

rownames(final) = final$gene
final$gene = NULL
exp

# Assuming final and exp are your dataframes
logfc_df <- final
counts_df <- exp

logfc_df
counts_df

# Check if all column names in counts_df are in the row names of logfc_df
if (!all(colnames(counts_df) %in% rownames(logfc_df))) {
  stop("Not all genes in counts_df are present in logfc_df")
}

# Reorder logfc_df to match the columns in counts_df
logfc_ordered <- logfc_df[colnames(counts_df), , drop = FALSE]
logfc_ordered

# Check for NA values in logfc_ordered
if (any(is.na(logfc_ordered$avg_log2FC))) {
  stop("There are NA values in the logFC values for the genes")
}

# Ensure logfc_ordered$logfc is a numeric vector
logfc_vector <- as.numeric(logfc_ordered$avg_log2FC)
logfc_vector
 #THIS IS FOR LOG TRANSFORMATINO BUT I DONT LIKE IT
#  custom_log_transform <- function(x, logFC_value) {
#       x * 2^logFC_value
#   }
#   
# # Multiply each column in counts_df by the corresponding logFC value
#  result_df <- sweep(counts_df, 2, logfc_vector, FUN = custom_log_transform)


#THE ACTUAL
 result_df <- sweep(counts_df, 2, logfc_vector, FUN = "*")

# Print the result
print(result_df)

metadata <- SPN_put_transformed@meta.data
# Assume that metadata has a column 'donor' that contains donor information
# If not, modify this part according to your metadata structure
if(!"donor_id" %in% colnames(metadata)) {
  stop("Metadata does not contain 'donor' information. Please check your metadata structure.")
}

# Create unique identifiers for cell column to avoid conflicts
expr_data <- result_df %>%
  rownames_to_column(var = "cell_id")

metadata <- metadata %>%
  rownames_to_column(var = "cell_id")

# Combine expression data with donor information
expr_data_long <- expr_data %>%
  pivot_longer(cols = -cell_id, names_to = "gene", values_to = "expression") %>%
  left_join(metadata %>% select(cell_id, donor_id), by = "cell_id")
# Assuming your data frame is named expr_data_long
expr_data_long
# Count the number of occurrences for each gene per donor

summarized_expr_data <- expr_data_long %>%
  group_by(donor_id, gene) %>%
  summarize(total_expression = sum(expression), .groups = 'drop')
summarized_expr_data
Put_SPN_summary_expr_data = summarized_expr_data
Put_SPN_summary_expr_data
```

#put nonSPN
```{r}
final = subset(newmarker, subset= cluster =="SPN")

significant_genes = unique(final$gene)

exp = FetchData(nonSPN_put_transformed, vars = significant_genes)
exp

final

rownames(final) = final$gene
final$gene = NULL
exp

# Assuming final and exp are your dataframes
logfc_df <- final
counts_df <- exp

logfc_df
counts_df

# Check if all column names in counts_df are in the row names of logfc_df
if (!all(colnames(counts_df) %in% rownames(logfc_df))) {
  stop("Not all genes in counts_df are present in logfc_df")
}

# Reorder logfc_df to match the columns in counts_df
logfc_ordered <- logfc_df[colnames(counts_df), , drop = FALSE]
logfc_ordered

# Check for NA values in logfc_ordered
if (any(is.na(logfc_ordered$avg_log2FC))) {
  stop("There are NA values in the logFC values for the genes")
}

# Ensure logfc_ordered$logfc is a numeric vector
logfc_vector <- as.numeric(logfc_ordered$avg_log2FC)
logfc_vector
 #THIS IS FOR LOG TRANSFORMATINO BUT I DONT LIKE IT
#  custom_log_transform <- function(x, logFC_value) {
#       x * 2^logFC_value
#   }
#   
# # Multiply each column in counts_df by the corresponding logFC value
#  result_df <- sweep(counts_df, 2, logfc_vector, FUN = custom_log_transform)


#THE ACTUAL
 result_df <- sweep(counts_df, 2, logfc_vector, FUN = "*")

# Print the result
print(result_df)

metadata <- nonSPN_put_transformed@meta.data
# Assume that metadata has a column 'donor' that contains donor information
# If not, modify this part according to your metadata structure
if(!"donor_id" %in% colnames(metadata)) {
  stop("Metadata does not contain 'donor' information. Please check your metadata structure.")
}

# Create unique identifiers for cell column to avoid conflicts
expr_data <- result_df %>%
  rownames_to_column(var = "cell_id")

metadata <- metadata %>%
  rownames_to_column(var = "cell_id")

# Combine expression data with donor information
expr_data_long <- expr_data %>%
  pivot_longer(cols = -cell_id, names_to = "gene", values_to = "expression") %>%
  left_join(metadata %>% select(cell_id, donor_id), by = "cell_id")
# Assuming your data frame is named expr_data_long
expr_data_long
# Count the number of occurrences for each gene per donor

summarized_expr_data <- expr_data_long %>%
  group_by(donor_id, gene) %>%
  summarize(total_expression = sum(expression), .groups = 'drop')
summarized_expr_data
Put_nonSPN_summary_expr_data = summarized_expr_data
Put_nonSPN_summary_expr_data
```


```{r}
Put_SPN_summary_expr_data$cell_type = "SPN"
Put_nonSPN_summary_expr_data$cell_type = "non-SPN" 

head(Put_SPN_summary_expr_data)
head(Put_nonSPN_summary_expr_data)
```

```{r}
Put_Transcriptional_DE = rbind(Put_SPN_summary_expr_data, Put_nonSPN_summary_expr_data)
Put_Transcriptional_DE
```

```{r}
donors = unique(Put_Transcriptional_DE$donor_id)

for (donor in donors) {
test = Put_Transcriptional_DE[Put_Transcriptional_DE$donor_id == donor,]
a = plot_overlapping_density_histogram(df = test, 
    hist_col = "total_expression",
    fill_col = "cell_type",
    colors = c("blue", "red"),
    #alpha=0.5,
    breaks=seq(0,1000,10),
    title= paste("Put SPNs vs non-SPNs: ", donor),
    xlab = "Put SPN score",
    fig_filename = NULL)

print(a)
}
```


```{r}
expr_data_long
```

```{r}
Put_Transcriptional_DE$Condition = ifelse(grepl("SCF_21-037CM2|SCF-23-068CM|SCF-22-058CF|SCF-22-054CM", Put_Transcriptional_DE$donor_id), "Control", "XDP")
Put_Transcriptional_DE
```


```{r}
XDP = Put_Transcriptional_DE[Put_Transcriptional_DE$Condition == "XDP",]
plot_overlapping_density_histogram(df = XDP, 
    hist_col = "total_expression",
    fill_col = "cell_type",
    colors = c("blue", "red"),
    #alpha=0.5,
    breaks=seq(0,1000,10),
    title= paste("Put SPNs vs non-SPNs (log transformed): XDP"),
    xlab = "Put SPN score",
    fig_filename = NULL)

Control = Put_Transcriptional_DE[Put_Transcriptional_DE$Condition == "Control",]
plot_overlapping_density_histogram(df = Control, 
    hist_col = "total_expression",
    fill_col = "cell_type",
    colors = c("blue", "red"),
    #alpha=0.5,
    breaks=seq(0,1000,10),
    title= paste("Put SPNs vs non-SPNs (log transformed): Control "),
    xlab = "Put SPN score",
    fig_filename = NULL)
```
