---
title: SPN_Score_compiled
summary: A compilation of all necessary work on creating SPN scores done between July 7-July 28 2024
---
#libraries to load
```{r}
library(Seurat)
library(ggplot2)
library(dplyr)
library(tidyr)
library(qs)
library(tibble)
library(gridExtra)
```

#functions: plot_histogram, 
```{r}
plot_overlapping_density_histogram = function(
    df, 
    hist_col,
    fill_col,
    colors = c("blue", "red"),
    alpha=0.5,
    breaks=seq(0, 16, 1),
    title= NULL,
    xlab = NULL,
    fig_filename = NULL

){
    # hist_col is the column you're making a histogram of (e.g. nUMI)
    # fill_col is the column you're coloring by (e.g. cell_class)
    # if fig_filename is not null, the plot will be saved to that file
     
    if (is.null(xlab)){
        xlab = hist_col
    }

    if (is.null(title)){
        title = paste0("Density Histogram of ", xlab, " by ", fill_col)
    }


    p = (
        ggplot(df, aes_string(x=hist_col, fill=fill_col)) 
        + geom_histogram(aes(y=..density..), alpha=alpha, position="identity", breaks=breaks)
        + labs(title=title, x=xlab, y="Density")    
        + theme(
                plot.title = element_text(size=16),
                axis.line = element_line(color = "black"),  # Add axis lines
                axis.ticks = element_line(color = "black"),  # Add axis ticks
                axis.text = element_text(size = 14),  # Increase tick label font size
                axis.title = element_text(size = 15)  # Increase axis label font size
            ) 
        + scale_fill_manual(values=colors)   
    )

    if (!is.null(fig_filename)){
        ggsave(fig_filename, p, width=8, height=6)
    }

    return(p)
}
```

```{r}
SPN_score_normalized = function(gene_df, SCtransform_counts, logfc_val, method = "multiply"){

significant_genes = unique(gene_df$gene)

exp = FetchData(SCtransform_counts, vars = significant_genes)

rownames(gene_df) = gene_df$gene
gene_df$gene = NULL

# Assuming gene_df and exp are your dataframes
logfc_df <- gene_df
counts_df <- exp

print(logfc_df)
print(counts_df)

# Check if all column names in counts_df are in the row names of logfc_df
if (!all(colnames(counts_df) %in% rownames(logfc_df))) {
  stop("Not all genes in counts_df are present in logfc_df")
}

# Reorder logfc_df to match the columns in counts_df
logfc_ordered <- logfc_df[colnames(counts_df), , drop = FALSE]
logfc_ordered

logfc = logfc_ordered[[logfc_val]]

# Check for NA values in logfc_ordered
if (any(is.na(logfc))) {
  stop("There are NA values in the logFC values for the genes")
}


# Ensure logfc_ordered$logfc is a numeric vector
logfc_vector <- as.numeric(logfc)
logfc_vector

gene_averages <- colMeans(counts_df)
print("Printing gene averages")
print(gene_averages)
if (method == "multiply") {
  
#THE ACTUAL
result_df <- sweep(counts_df, 2, logfc_vector, FUN = "*")
} else if(method == "log"){
  custom_log_transform <- function(x, logFC_value) {
     x * 2^logFC_value
  }
  result_df <- sweep(counts_df, 2, logfc_vector, FUN = custom_log_transform)
} else{
  stop("Calculation is not working")
}
print(result_df)
# Divide each gene's expression by its average expression
result_df <- sweep(result_df, 2, gene_averages, FUN = "/")

# Print the result
print("Printing normalized")
print(result_df)


metadata <- SCtransform_counts@meta.data
# Assume that metadata has a column 'donor' that contains donor information
# If not, modify this part according to your metadata structure
if(!"donor_id" %in% colnames(metadata)) {
  stop("Metadata does not contain 'donor' information. Please check your metadata structure.")
}

# Create unique identifiers for cell column to avoid conflicts
expr_data <- result_df %>%
  rownames_to_column(var = "cell_id")

metadata <- metadata %>%
  rownames_to_column(var = "cell_id")

# Combine expression data with donor information
expr_data_long <- expr_data %>%
  pivot_longer(cols = -cell_id, names_to = "gene", values_to = "expression") %>%
  left_join(metadata %>% select(cell_id, donor_id, Condition, subcluster, sub_class, cell_class, patch_matrix, neuron_type_joint_cluster, pct_mito, nUmi), by = "cell_id")
# Assuming your data frame is named expr_data_long
print(expr_data_long)
# Count the number of occurrences for each gene per donor

summarized_expr_data <- expr_data_long %>%
  group_by(cell_id, donor_id, Condition, subcluster, sub_class, cell_class, patch_matrix, neuron_type_joint_cluster, pct_mito, nUmi) %>%
  summarize(total_expression = sum(expression), .groups = 'drop')
print(summarized_expr_data)

spn_scores = summarized_expr_data$total_expression
mean_spn <- mean(spn_scores, na.rm = TRUE)
sd_spn <- sd(spn_scores, na.rm = TRUE)
# Compute the z-score for each SPN score
summarized_expr_data$z_score  <- (spn_scores - mean_spn) / sd_spn

return(summarized_expr_data)
}
```


```{r}
# Add z-scores to Seurat object metadata
caudate_neurons_transformed <- AddMetaData(caudate_neurons_transformed, metadata = spn_zscores, col.name = "SPN_zscore")

# Visualize SPN z-scores on UMAP
plot <- FeaturePlot(caudate_neurons_transformed, features = "SPN_zscore") +
  ggtitle("SPN Z-Score on UMAP")

# Save the UMAP plot with SPN z-scores as a PNG file
ggsave("SPN_zscore_umap.png", plot = plot, width = 10, height = 8, dpi = 300)

```

```{r}
SPN_score_graphs <- function(SPN_score_output, 
                             donor_order = donors, 
                             total_expression = "total_expression", 
                             fillcol = "subcluster", 
                             color = color1, 
                             donor_graph_title = "CaH SPNs vs non-SPNs", 
                             SCtransformcounts, 
                             SPN_score_col_name) {
  
  # Ensure total_expression is a string
  total_expression <- as.character(total_expression)
  
  # Calculate min, max, and median
  min_SPN_score <- min(SPN_score_output[[total_expression]], na.rm = TRUE)
  max_SPN_score <- max(SPN_score_output[[total_expression]], na.rm = TRUE)
  median_SPN_score <- median(SPN_score_output[[total_expression]], na.rm = TRUE)
  
  print("min, max, median scores")
  print(min_SPN_score)
  print(max_SPN_score)
  print(median_SPN_score)
  
  options(repr.plot.width = 24, repr.plot.height = 16)
  plots <- list()
  
  for (donor in donor_order) {
    test <- SPN_score_output[SPN_score_output$donor_id == donor, ]
    plots[[donor]] <- plot_overlapping_density_histogram(df = test,
                                                         hist_col = total_expression,
                                                         fill_col = fillcol,
                                                         colors = color,
                                                         breaks = seq(min_SPN_score, max_SPN_score, round(max_SPN_score / 50)),
                                                         title = paste(donor_graph_title, ": ", donor),
                                                         xlab = "SPN Score",
                                                         fig_filename = NULL)
  }
  
  layout_matrix <- rbind(
    c(1, 2, 3, 4, 5),
    c(6, 7, 8, 9, 10),
    c(11, 12, 13, NA, NA),
    c(14, 15, 16, 17, NA)
  )
  
  # Arrange the plots according to the custom layout
  grid_plots <- grid.arrange(grobs = plots, layout_matrix = layout_matrix)
  
  # Save the arranged plots to a PNG file
  ggsave(filename = "~/SPN_SCORE_Output/USETEST.png", plot = grid_plots, width = 30, height = 16)
  
  print("donor plot done")
  
  XDP <- SPN_score_output[SPN_score_output$Condition == "XDP", ]
  a <- plot_overlapping_density_histogram(df = XDP, 
                                          hist_col = total_expression,
                                          fill_col = fillcol,
                                          colors = color,
                                          breaks = seq(min_SPN_score, max_SPN_score, round(max_SPN_score / 50)),
                                          title = paste(donor_graph_title, ": ", "XDP"),
                                          xlab = "SPN score",
                                          fig_filename = NULL)
  
  Control <- SPN_score_output[SPN_score_output$Condition == "Control", ]
  b <- plot_overlapping_density_histogram(df = Control, 
                                          hist_col = total_expression,
                                          fill_col = fillcol,
                                          colors = color,
                                          breaks = seq(min_SPN_score, max_SPN_score, round(max_SPN_score / 50)),
                                          title = paste(donor_graph_title, ": ", "Control"),
                                          xlab = "SPN score",
                                          fig_filename = NULL)
  
  print("Condition plot done")
  
    
  SPN <- SPN_score_output[SPN_score_output$subcluster == "SPN", ]
  c <- plot_overlapping_density_histogram(df = SPN, 
                                          hist_col = total_expression,
                                          fill_col = "Condition",
                                          colors = c("XDP" = "red", "Control" = "blue"),
                                          breaks = seq(min_SPN_score, max_SPN_score, round(max_SPN_score / 50)),
                                          title = "XDP vs Control: SPN",
                                          xlab = "SPN score",
                                          fig_filename = NULL)
  
  notSPN <- SPN_score_output[SPN_score_output$subcluster == "non-SPN", ]
  d <- plot_overlapping_density_histogram(df = notSPN, 
                                          hist_col = total_expression,
                                          fill_col = "Condition",
                                          colors = c("XDP" = "red", "Control" = "blue"),
                                          breaks = seq(min_SPN_score, max_SPN_score, round(max_SPN_score / 50)),
                                          title = "XDP vs Control: non-SPN",
                                          xlab = "SPN score",
                                          fig_filename = NULL)
  
  print(a)
  print(b)
  print(c)
  print(d)
  
  print("graphs done")
  
  # Calculate the threshold
  threshold <- quantile(SPN_score_output[[total_expression]][SPN_score_output[["subcluster"]] == "SPN" & SPN_score_output[["Condition"]] == "XDP"], 0.1, na.rm = TRUE)
  print(threshold)
  
  # Subset the data
  SPN_score_output_identity_loss_SPN <- SPN_score_output[SPN_score_output[[total_expression]] < threshold & SPN_score_output[["subcluster"]] == "SPN", ]
  
  print(SPN_score_output_identity_loss_SPN)
  print(table(SPN_score_output_identity_loss_SPN$donor_id))
  
  identity_loss_cells <- unique(SPN_score_output_identity_loss_SPN$cell_id)
  
  SCtransformcounts@meta.data$SPN_identity <- "Other"
  SCtransformcounts$SPN_identity[identity_loss_cells] <- "Losing_Identity"
  SCtransformcounts@meta.data
  
  # Visualize on UMAP
  pic <- DimPlot(SCtransformcounts, reduction = "umap", group.by = "SPN_identity", split.by = "Condition", cols = c("red", "grey"))
  print(pic)
  ggsave("~/SPN_SCORE_Output/SPN_score_10pct_TEST.png", plot = pic, width = 20, height = 8, units = "in", dpi = 300)
  
  spn_scores <- setNames(SPN_score_output[[total_expression]], SPN_score_output$cell_id)
  
  SCtransformcounts <- AddMetaData(SCtransformcounts, metadata = spn_scores, col.name = SPN_score_col_name)
  
  pic2 <- FeaturePlot(SCtransformcounts, features = SPN_score_col_name, split.by = "Condition")
  print(pic2)
  ggsave("~/SPN_SCORE_Output/SPN_UMAP_TEST.png", plot = pic2, width = 20, height = 8, units = "in", dpi = 300)
  
  return(SCtransformcounts)
}

```




#From v8 BICAN- first do everything as normal with the normized version of code. 
#then subset only matrix, run findallmarkers, validate for xdp
#then subset XDP only matrix, run findallmarkers, validate for BICAN

```{r}
BICAN_3P_V8_CaH_neurons_filtered =  qread("~/SOBJ_USE_THESE/BICAN_3P_V8_CaH_neurons_filtered.qs")
BICAN_3P_V8_CaH_neurons_filtered
```

```{r}
BICAN_3P_V8_CaH_neurons_filtered@meta.data

Idents(BICAN_3P_V8_CaH_neurons_filtered) <- "subcluster"
DimPlot(BICAN_3P_V8_CaH_neurons_filtered, label = TRUE)

Idents(BICAN_3P_V8_CaH_neurons_filtered) <- "sub_class"
DimPlot(BICAN_3P_V8_CaH_neurons_filtered, label = TRUE)

Idents(BICAN_3P_V8_CaH_neurons_filtered) <- "SPN_class"
DimPlot(BICAN_3P_V8_CaH_neurons_filtered, label = TRUE)

Idents(BICAN_3P_V8_CaH_neurons_filtered) <- "D1_D2"
DimPlot(BICAN_3P_V8_CaH_neurons_filtered, label = TRUE)

Idents(BICAN_3P_V8_CaH_neurons_filtered) <- "patch_matrix"
DimPlot(BICAN_3P_V8_CaH_neurons_filtered, label = TRUE)
```

```{r}
BICAN_SPN_vs_non_markers= qread("~/SPN_Scores_markers/BICAN_SPN_vs_non_markers.qs")
BICAN_SPN_vs_non_markers #SPN vs nonSPN
final_BICAN_SPN_vs_non_markers = subset(BICAN_SPN_vs_non_markers, subset = p_val_adj < 0.05 & cluster == "SPN")
final_BICAN_SPN_vs_non_markers$pct_subtract = final_BICAN_SPN_vs_non_markers$pct.1 - final_BICAN_SPN_vs_non_markers$pct.2
final_BICAN_SPN_vs_non_markers$weighted_logFC = final_BICAN_SPN_vs_non_markers$avg_log2FC * final_BICAN_SPN_vs_non_markers$pct_subtract
final_BICAN_SPN_vs_non_markers
```

```{r}
merged_xdp = qread("~/SOBJ_USE_THESE/merged_xdp_use_this.qs")
merged_xdp
merged_xdp@meta.data
```

```{r}
library(sctransform)
merged_xdp_transformed = SCTransform(merged_xdp, vars.to.regress = "pct_mito", verbose = FALSE)
DefaultAssay(merged_xdp_transformed) = "SCT"
qsave(merged_xdp_transformed, "~/SOBJ_USE_THESE/merged_xdp_transformed.qs")
```

```{r}
library(sctransform)
BICAN_3P_V8_CaH_neurons_filtered_transformed = SCTransform(BICAN_3P_V8_CaH_neurons_filtered, vars.to.regress = "pct_mt", verbose = FALSE)
DefaultAssay(BICAN_3P_V8_CaH_neurons_filtered_transformed) = "SCT"

#qsave(merged_xdp_transformed, "~/SOBJ_USE_THESE/merged_xdp_transformed.qs")
```


#old 
```{r}
test = SPN_score_normalized(final_BICAN_SPN_vs_non_markers, merged_xdp_transformed, logfc_val = "weighted_logFC", method = "multiply")
```

```{r}

SPN_score_graphs(test, donor_order = donors,   total_expression = "total_expression",  fillcol = "subcluster", 
                             color = color1, 
                             donor_graph_title = "CaH SPNs vs non-SPNs", 
                             merged_xdp_transformed, 
                             SPN_score_col_name = "test") 


```

#normalized
```{r}
test = SPN_score_normalized(final_BICAN_SPN_vs_non_markers, merged_xdp_transformed, logfc_val = "weighted_logFC", method = "multiply")
```

```{r}
SPN_score_graphs(test, donor_order = donors,   total_expression = "total_expression",  fillcol = "subcluster", 
                             color = color1, 
                             donor_graph_title = "CaH SPNs vs non-SPNs", 
                             merged_xdp_transformed, 
                             SPN_score_col_name = "test") 
```

```{r}
SPN_score_graphs(test, donor_order = donors,   total_expression = "total_expression",  fillcol = "sub_class", 
                             color = color3, 
                             donor_graph_title = "CaH SPNs vs non-SPNs", 
                             merged_xdp_transformed, 
                             SPN_score_col_name = "test") 
```

#XDP control Matrix SPN markers only 

```{r}
head(merged_xdp_transformed)
unique(merged_xdp_transformed$neuron_type_joint_cluster)
```

```{r}
Idents(merged_xdp_transformed) = "neuron_type_joint_cluster"
DimPlot(merged_xdp_transformed, label = TRUE)
```
```{r}
merged_xdp_transformed@meta.data$patch_matrix = merged_xdp_transformed@meta.data$neuron_type_joint_cluster
merged_xdp_transformed$patch_matrix[merged_xdp_transformed$patch_matrix == "interneuron"] = "non-SPN"
merged_xdp_transformed$patch_matrix[merged_xdp_transformed$patch_matrix == "other"] = "non-SPN"
merged_xdp_transformed$patch_matrix[merged_xdp_transformed$patch_matrix == "espn"] = "non-SPN"
table(merged_xdp_transformed$patch_matrix)
```

```{r}
Idents(merged_xdp_transformed) = "patch_matrix"
DimPlot(merged_xdp_transformed, label = T)
```
```{r}
merged_xdp_transformed_controls = subset(merged_xdp_transformed, subset = Condition == "Control")
merged_xdp_transformed_controls
DimPlot(merged_xdp_transformed_controls, label = TRUE)
```


```{r}
xdp_matrix_markers =FindMarkers(object = merged_xdp_transformed_controls, ident.1 = "matrix_spn", ident.2 = "non-SPN", only.pos = TRUE)
                                #,  min.pct = 0.2, logfc.threshold = 1.25)
xdp_matrix_markers
```


#Take each gene, divide by average expression first 
```{r}
xdp_matrix_markers_final = subset(xdp_matrix_markers, subset = p_val_adj < 0.05)
xdp_matrix_markers_final$pct_subtract = xdp_matrix_markers_final$pct.1 - xdp_matrix_markers_final$pct.2
xdp_matrix_markers_final$weighted_logFC = xdp_matrix_markers_final$avg_log2FC * xdp_matrix_markers_final$pct_subtract
xdp_matrix_markers_final
```

```{r}
xdp_matrix_markers_final$gene = rownames(xdp_matrix_markers_final)
```


#normalized
```{r}
XDP_matrix_scores = SPN_score_normalized(xdp_matrix_markers_final, merged_xdp_transformed, logfc_val = "weighted_logFC", method = "multiply")
```

```{r}
merged_xdp_transformed@meta.data
```


```{r}
color4 = c("matrix_spn" = "red", "non-SPN" = "blue", "patch_spn" = "green")

SPN_score_graphs(XDP_matrix_scores, donor_order = donors, total_expression = "total_expression",  fillcol = "patch_matrix", 
                             color = color4, 
                             donor_graph_title = "SPN Matrix vs non-SPNs", 
                             merged_xdp_transformed, 
                             SPN_score_col_name = "test") 
```

```{r}
SPN_score_graphs(XDP_matrix_scores, donor_order = donors, total_expression = "total_expression",  fillcol = "subcluster", 
                             color = color1, 
                             donor_graph_title = "Matrix SPNs vs non-SPNs", 
                             merged_xdp_transformed, 
                             SPN_score_col_name = "test")
```


```{r}
matrix <- XDP_matrix_scores[XDP_matrix_scores$patch_matrix == "matrix_spn", ]
  plot_overlapping_density_histogram(df = matrix, 
                                          hist_col = matrix$total_expression,
                                          fill_col = "Condition",
                                          colors = c("XDP" = "red", "Control" = "blue"),
                                          breaks = seq(0, 250, 5),
                                          title = "XDP vs Control: SPN matrix",
                                          xlab = "SPN score",
                                          fig_filename = NULL)
  patch <- XDP_matrix_scores[XDP_matrix_scores$patch_matrix == "patch_spn", ]
  plot_overlapping_density_histogram(df = patch, 
                                          hist_col = patch$total_expression,
                                          fill_col = "Condition",
                                          colors = c("XDP" = "red", "Control" = "blue"),
                                          breaks = seq(0, 250, 5),
                                          title = "XDP vs Control: SPN patch",
                                          xlab = "SPN score",
                                          fig_filename = NULL)
  nonSPN <- XDP_matrix_scores[XDP_matrix_scores$patch_matrix == "non-SPN", ]
  plot_overlapping_density_histogram(df = nonSPN, 
                                          hist_col = nonSPN$total_expression,
                                          fill_col = "Condition",
                                          colors = c("XDP" = "red", "Control" = "blue"),
                                          breaks = seq(0, 250, 5),
                                          title = "XDP vs Control: non-SPN",
                                          xlab = "SPN score",
                                          fig_filename = NULL)
```


#validate with BICAN

```{r}
BICAN_3P_V8_CaH_neurons_filtered@meta.data$cell_id = BICAN_3P_V8_CaH_neurons_filtered@meta.data$cell
BICAN_3P_V8_CaH_neurons_filtered@meta.data
```


```{r}
BICAN_SPN_score_normalized = function(gene_df, SCtransform_counts, logfc_val, method = "multiply"){

significant_genes = unique(gene_df$gene)

exp = FetchData(SCtransform_counts, vars = significant_genes)

rownames(gene_df) = gene_df$gene
gene_df$gene = NULL

# Assuming gene_df and exp are your dataframes
logfc_df <- gene_df
counts_df <- exp

print(logfc_df)
print(counts_df)

# Check if all column names in counts_df are in the row names of logfc_df
if (!all(colnames(counts_df) %in% rownames(logfc_df))) {
  stop("Not all genes in counts_df are present in logfc_df")
}

# Reorder logfc_df to match the columns in counts_df
logfc_ordered <- logfc_df[colnames(counts_df), , drop = FALSE]
logfc_ordered

logfc = logfc_ordered[[logfc_val]]

# Check for NA values in logfc_ordered
if (any(is.na(logfc))) {
  stop("There are NA values in the logFC values for the genes")
}


# Ensure logfc_ordered$logfc is a numeric vector
logfc_vector <- as.numeric(logfc)
logfc_vector

gene_averages <- colMeans(counts_df)
print("Printing gene averages")
print(gene_averages)
if (method == "multiply") {
  
#THE ACTUAL
result_df <- sweep(counts_df, 2, logfc_vector, FUN = "*")
} else if(method == "log"){
  custom_log_transform <- function(x, logFC_value) {
     x * 2^logFC_value
  }
  result_df <- sweep(counts_df, 2, logfc_vector, FUN = custom_log_transform)
} else{
  stop("Calculation is not working")
}
print(result_df)
# Divide each gene's expression by its average expression
result_df <- sweep(result_df, 2, gene_averages, FUN = "/")

# Print the result
print("Printing normalized")
print(result_df)


metadata <- SCtransform_counts@meta.data
# Assume that metadata has a column 'donor' that contains donor information
# If not, modify this part according to your metadata structure
if(!"DONOR" %in% colnames(metadata)) {
  stop("Metadata does not contain 'donor' information. Please check your metadata structure.")
}

# Check if 'cell_id' column already exists and remove it
if ("cell_id" %in% colnames(metadata)) {
  metadata <- metadata %>% select(-cell_id)
}

# Ensure unique row names in result_df
if (any(duplicated(rownames(result_df)))) {
  stop("Row names in result_df are not unique. Please ensure unique row names.")
}


# Create unique identifiers for cell column to avoid conflicts
expr_data <- result_df %>%
  rownames_to_column(var = "cell_id")

metadata <- metadata %>%
  rownames_to_column(var = "cell_id")

# Combine expression data with donor information
expr_data_long <- expr_data %>%
  pivot_longer(cols = -cell_id, names_to = "gene", values_to = "expression") %>%
  left_join(metadata %>% select(cell_id, DONOR, subcluster, sub_class, SPN_class, patch_matrix), by = "cell_id")
# Assuming your data frame is named expr_data_long
print(expr_data_long)
# Count the number of occurrences for each gene per donor

summarized_expr_data <- expr_data_long %>%
  group_by(cell_id, DONOR, subcluster, sub_class, SPN_class, patch_matrix) %>%
  summarize(total_expression = sum(expression), .groups = 'drop')
print(summarized_expr_data)

spn_scores = summarized_expr_data$total_expression
mean_spn <- mean(spn_scores, na.rm = TRUE)
sd_spn <- sd(spn_scores, na.rm = TRUE)
# Compute the z-score for each SPN score
summarized_expr_data$z_score  <- (spn_scores - mean_spn) / sd_spn

return(summarized_expr_data)
}
```



```{r}
BICANdonors = unique(BICAN_3P_V8_CaH_neurons_filtered$DONOR)
BICANdonors
```


```{r}
BICAN_SPN_score_graphs <- function(SPN_score_output, 
                             donor_order = BICANdonors, 
                             total_expression = "total_expression", 
                             fillcol = "subcluster", 
                             color = color1, 
                             donor_graph_title = "CaH SPNs vs non-SPNs", 
                             SCtransformcounts, 
                             SPN_score_col_name) {
  
  # Ensure total_expression is a string
  total_expression <- as.character(total_expression)
  
  # Calculate min, max, and median
  min_SPN_score <- min(SPN_score_output[[total_expression]], na.rm = TRUE)
  max_SPN_score <- max(SPN_score_output[[total_expression]], na.rm = TRUE)
  median_SPN_score <- median(SPN_score_output[[total_expression]], na.rm = TRUE)
  
  print("min, max, median scores")
  print(min_SPN_score)
  print(max_SPN_score)
  print(median_SPN_score)
  
  options(repr.plot.width = 24, repr.plot.height = 16)
  plots <- list()
  
  for (donor in donor_order) {
    test <- SPN_score_output[SPN_score_output$DONOR == donor, ]
    plots[[donor]] <- plot_overlapping_density_histogram(df = test,
                                                         hist_col = total_expression,
                                                         fill_col = fillcol,
                                                         colors = color,
                                                         breaks = seq(min_SPN_score, max_SPN_score, round(max_SPN_score / 50)),
                                                         title = paste(donor_graph_title, ": ", donor),
                                                         xlab = "SPN Score",
                                                         fig_filename = NULL)
  }
  
  layout_matrix <- rbind(
    c(1, 2, 3, 4, 5),
    c(6, 7, 8, 9, 10),
    c(11, 12, 13, 14, 15),
    c(16, 17, 18, 19, NA)
  )
  
  # Arrange the plots according to the custom layout
  grid_plots <- grid.arrange(grobs = plots, layout_matrix = layout_matrix)
  
  # Save the arranged plots to a PNG file
  ggsave(filename = "~/SPN_SCORE_Output/USETEST.png", plot = grid_plots, width = 30, height = 16)
  
  print("donor plot done")
  
  b <- plot_overlapping_density_histogram(df = SPN_score_output, 
                                          hist_col = total_expression,
                                          fill_col = fillcol,
                                          colors = color,
                                          breaks = seq(min_SPN_score, max_SPN_score, round(max_SPN_score / 50)),
                                          title = paste(donor_graph_title, ": ", "Control"),
                                          xlab = "SPN score",
                                          fig_filename = NULL)
  
  print("Condition plot done")
  

  
  # Calculate the threshold
  threshold <- quantile(SPN_score_output[[total_expression]][SPN_score_output[["SPN_class"]] == "SPN"], 0.1, na.rm = TRUE)
  print(threshold)
  
  # Subset the data
  SPN_score_output_identity_loss_SPN <- SPN_score_output[SPN_score_output[[total_expression]] < threshold & SPN_score_output[["SPN_class"]] == "SPN", ]
  
  print(SPN_score_output_identity_loss_SPN)
  print(table(SPN_score_output_identity_loss_SPN$DONOR))
  
  identity_loss_cells <- unique(SPN_score_output_identity_loss_SPN$cell_id)
  
  SCtransformcounts@meta.data$SPN_identity <- "Other"
  SCtransformcounts$SPN_identity[identity_loss_cells] <- "Losing_Identity"
  SCtransformcounts@meta.data
  
  # Visualize on UMAP
  pic <- DimPlot(SCtransformcounts, reduction = "umap", group.by = "SPN_identity", cols = c("red", "grey"))
  print(pic)
  ggsave("~/SPN_SCORE_Output/SPN_score_10pct_TEST.png", plot = pic, width = 20, height = 8, units = "in", dpi = 300)
  
  spn_scores <- setNames(SPN_score_output[[total_expression]], SPN_score_output$cell_id)
  
  SCtransformcounts <- AddMetaData(SCtransformcounts, metadata = spn_scores, col.name = SPN_score_col_name)
  
  pic2 <- FeaturePlot(SCtransformcounts, features = SPN_score_col_name)
  print(pic2)
  ggsave("~/SPN_SCORE_Output/SPN_UMAP_TEST.png", plot = pic2, width = 20, height = 8, units = "in", dpi = 300)
  
  return(SCtransformcounts)
}

```





#BICAN validation 
```{r}
XDP_matrix_scores = BICAN_SPN_score_normalized(xdp_matrix_markers_final, BICAN_3P_V8_CaH_neurons_filtered_transformed, logfc_val = "weighted_logFC", method = "multiply")
```

```{r}
color4 = c("matrix" = "red", "non-SPN" = "blue", "patch" = "green")

BICAN_SPN_score_graphs(XDP_matrix_scores, donor_order = BICANdonors, total_expression = "total_expression",  fillcol = "patch_matrix", 
                             color = color4, 
                             donor_graph_title = "SPNs vs non-SPNs", 
                             BICAN_3P_V8_CaH_neurons_filtered_transformed, 
                             SPN_score_col_name = "BICANtest") 
```


```{r}
SPN_score_graphs(XDP_matrix_scores, donor_order = donors,   total_expression = "total_expression",  fillcol = "sub_class", 
                             color = color1, 
                             donor_graph_title = "SPNs vs non-SPNs", 
                             merged_xdp_transformed, 
                             SPN_score_col_name = "test")
```


```{r}
matrix <- XDP_matrix_scores[XDP_matrix_scores$patch_matrix == "matrix_spn", ]
  plot_overlapping_density_histogram(df = matrix, 
                                          hist_col = matrix$total_expression,
                                          fill_col = "Condition",
                                          colors = c("XDP" = "red", "Control" = "blue"),
                                          breaks = seq(200, 1400, 20),
                                          title = "XDP vs Control: SPN matrix",
                                          xlab = "SPN score",
                                          fig_filename = NULL)
  patch <- XDP_matrix_scores[XDP_matrix_scores$patch_matrix == "patch_spn", ]
  plot_overlapping_density_histogram(df = patch, 
                                          hist_col = patch$total_expression,
                                          fill_col = "Condition",
                                          colors = c("XDP" = "red", "Control" = "blue"),
                                          breaks = seq(200, 1400, 20),
                                          title = "XDP vs Control: SPN patch",
                                          xlab = "SPN score",
                                          fig_filename = NULL)
  nonSPN <- XDP_matrix_scores[XDP_matrix_scores$patch_matrix == "non-SPN", ]
  plot_overlapping_density_histogram(df = nonSPN, 
                                          hist_col = nonSPN$total_expression,
                                          fill_col = "Condition",
                                          colors = c("XDP" = "red", "Control" = "blue"),
                                          breaks = seq(200, 1400, 20),
                                          title = "XDP vs Control: non-SPN",
                                          xlab = "SPN score",
                                          fig_filename = NULL)
```























#From Additional_SPN_scoring
#if you want to plot quantiles or jitter plot look at this file




#FROM SPN_JUL26_continue
```{r}
CaH_summary_expr_data = SPN_score(BICAN_SPN_vs_non_markers_filtered, caudate_neurons_transformed,logfc_val = "avg_log2FC")
```

```{r}
donors = c("PCMC-16-011", "PCMC-16-012", "SCF-18-003", "SCF-18-004", "SCF-18-006","SCF-19-009", "SCF-19-014", "SCF-19-018",  "SCF-20-023",  "SCF_20-024", "SCF-20-025", "SCF-21-030", "SCF_22-043", "SCF_21-037CM2","SCF-22-054CM","SCF-22-058CF","SCF-23-068CM")
```

```{r}
CaH_summary_expr_data = SPN_score(BICAN_SPN_vs_non_markers_filtered, caudate_neurons_transformed,logfc_val = "avg_log2FC")
```

```{r}
color1 = c("SPN" = "red", "non-SPN"= "blue") #for subcluster: SPN vs non-SPNS
color2 = c("red", "blue")
color3 = c("SPN"="red", "eSPN"="yellow", "other"="green", "IN" = "blue") #for subclass
```

```{r}
caudate_neurons_transformed@meta.data
SPN_score_graphs(CaH_summary_expr_data, donor_order = donors, fillcol = "subcluster", color = color1, donor_graph_title = "CaH SPNs vs non-SPNs", caudate_neurons_transformed, "BICAN_SPN_score_52genes")
caudate_neurons_transformed@meta.data
```

```{r}
SPN_score_graphs(CaH_summary_expr_data, donor_order = donors, fillcol = "subcluster", color = color1, donor_graph_title = "CaH SPNs vs non-SPNs", caudate_neurons_transformed, "BICAN_SPN_score_52genes")

```


```{r}
merged_xdp = qread("SPN_Scores/merged_xdp_use_this.qs")
merged_xdp@meta.data
```




```{r}
merged_xdp@meta.data$subcluster = merged_xdp$neuron_type
merged_xdp$subcluster[merged_xdp$subcluster == "IN"] = "non-SPN"
merged_xdp$subcluster[merged_xdp$subcluster == "other"] = "non-SPN"
merged_xdp$subcluster[merged_xdp$subcluster == "eSPN"] = "non-SPN"

merged_xdp@meta.data$sub_class = merged_xdp$neuron_type
```



```{r}
merged_xdp@meta.data
#10%: 50.29607
merged_XDP_summary_expr_data = SPN_score(BICAN_SPN_vs_non_markers_filtered, merged_xdp,logfc_val = "avg_log2FC")

merged_xdp= SPN_score_graphs(merged_XDP_summary_expr_data, donor_order = donors, fillcol = "subcluster", color = color1, donor_graph_title = "SPNs vs non-SPNs", merged_xdp, "BICAN_SPN_Score_52genes")

merged_xdp= SPN_score_graphs(merged_XDP_summary_expr_data, donor_order = donors, fillcol = "sub_class", color = color3, donor_graph_title = "SPNs vs non-SPNs", merged_xdp, "BICAN_SPN_Score_52genes")

merged_xdp@meta.data
```


```{r}
final_BICAN_SPN_vs_non_markers = subset(BICAN_SPN_vs_non_markers, subset = p_val_adj < 0.05 & cluster == "SPN")
final_BICAN_SPN_vs_non_markers$pct_subtract = final_BICAN_SPN_vs_non_markers$pct.1 - final_BICAN_SPN_vs_non_markers$pct.2
final_BICAN_SPN_vs_non_markers$weighted_logFC = final_BICAN_SPN_vs_non_markers$avg_log2FC * final_BICAN_SPN_vs_non_markers$pct_subtract
final_BICAN_SPN_vs_non_markers
```


```{r}
merged_xdp@meta.data
#10%: 300.81
merged_XDP_BICAN_pct = SPN_score(final_BICAN_SPN_vs_non_markers, merged_xdp,logfc_val = "weighted_logFC")

merged_xdp= SPN_score_graphs(merged_XDP_BICAN_pct, donor_order = donors, fillcol = "subcluster", color = color1, donor_graph_title = "SPNs vs non-SPNs", merged_xdp, "BICAN_SPN_Score_pct")

SPN_score_graphs(merged_XDP_BICAN_pct, donor_order = donors, fillcol = "sub_class", color = color3, donor_graph_title = "SPNs vs non-SPNs", merged_xdp, "BICAN_SPN_Score_pct")

merged_xdp@meta.data
```

```{r}
table(BICAN_SPN_classes_vs_non_markers$cluster)
```


```{r}
final_BICAN_SPN_classes_vs_non_markers = subset(BICAN_SPN_classes_vs_non_markers, subset = cluster != "non-SPN")
final_BICAN_SPN_classes_vs_non_markers
table(final_BICAN_SPN_classes_vs_non_markers$cluster)
```



```{r}
final_BICAN_SPN_classes_vs_non_markers = subset(final_BICAN_SPN_classes_vs_non_markers, subset = p_val_adj < 0.05)
final_BICAN_SPN_classes_vs_non_markers_unique <- final_BICAN_SPN_classes_vs_non_markers %>%
  distinct(gene, .keep_all = TRUE)
final_BICAN_SPN_classes_vs_non_markers_unique

table(final_BICAN_SPN_classes_vs_non_markers_unique$cluster) #107 is now lowest
```

```{r}
final_BICAN_SPN_classes_vs_non_markers_unique
```


```{r}
final_final_BICAN_SPN_classes_vs_non_markers <- final_BICAN_SPN_classes_vs_non_markers_unique %>%
  group_by(cluster) %>%
  arrange(p_val_adj) %>%
  slice_head(n = 107) %>%
  ungroup()
final_final_BICAN_SPN_classes_vs_non_markers = as.data.frame(final_final_BICAN_SPN_classes_vs_non_markers)
final_final_BICAN_SPN_classes_vs_non_markers
# View the result
rownames(final_final_BICAN_SPN_classes_vs_non_markers)= final_final_BICAN_SPN_classes_vs_non_markers$gene
final_final_BICAN_SPN_classes_vs_non_markers
```



```{r}
final_final_BICAN_SPN_classes_vs_non_markers$pct_subtract = final_final_BICAN_SPN_classes_vs_non_markers$pct.1 - final_final_BICAN_SPN_classes_vs_non_markers$pct.2
final_final_BICAN_SPN_classes_vs_non_markers$weighted_logFC = final_final_BICAN_SPN_classes_vs_non_markers$avg_log2FC * final_final_BICAN_SPN_classes_vs_non_markers$pct_subtract
final_final_BICAN_SPN_classes_vs_non_markers
```




```{r}
merged_xdp@meta.data

merged_XDP_BICAN_pct_SPNclasses = SPN_score(final_final_BICAN_SPN_classes_vs_non_markers, merged_xdp,logfc_val = "weighted_logFC")

merged_xdp= SPN_score_graphs(merged_XDP_BICAN_pct_SPNclasses, donor_order = donors, fillcol = "subcluster", color = color1, donor_graph_title = "SPNs vs non-SPNs", merged_xdp, "BICAN_SPNclasses_Score_pct")

SPN_score_graphs(merged_XDP_BICAN_pct, donor_order = donors, fillcol = "sub_class", color = color3, donor_graph_title = "SPNs vs non-SPNs", merged_xdp, "BICAN_SPNclasses_Score_pct")

merged_xdp@meta.data
```

```{r}
final_BICAN_SPN_classes_vs_non_markers$pct_subtract = final_BICAN_SPN_classes_vs_non_markers$pct.1 - final_BICAN_SPN_classes_vs_non_markers$pct.2
final_BICAN_SPN_classes_vs_non_markers$weighted_logFC = final_BICAN_SPN_classes_vs_non_markers$avg_log2FC * final_BICAN_SPN_classes_vs_non_markers$pct_subtract
final_BICAN_SPN_classes_vs_non_markers
```

```{r}
merged_xdp@meta.data

merged_XDP_BICAN_pct_SPNclasses = SPN_score(final_BICAN_SPN_classes_vs_non_markers, merged_xdp,logfc_val = "weighted_logFC")

merged_xdp= SPN_score_graphs(merged_XDP_BICAN_pct_SPNclasses, donor_order = donors, fillcol = "subcluster", color = color1, donor_graph_title = "SPNs vs non-SPNs", merged_xdp, "BICAN_SPNclasses_Score_pcttest")

#SPN_score_graphs(merged_XDP_BICAN_pct, donor_order = donors, fillcol = "sub_class", color = color3, donor_graph_title = "SPNs vs non-SPNs", merged_xdp, "BICAN_SPNclasses_Score_pcttest")

merged_xdp@meta.data
```












```{r}
DimPlot(merged_xdp, group.by = "RNA_snn_res.0.2")
DimPlot(merged_xdp, group.by = "RNA_snn_res.0.3")
DimPlot(merged_xdp, group.by = "region")
```
```{r}
FeaturePlot(merged_xdp, features =c("CASZ1","CHAT"))
FeaturePlot(merged_xdp, features =c("DRD1", "DRD2"))
FeaturePlot(merged_xdp, features =c("PPP1R1B", "pct_mito"))
FeaturePlot(merged_xdp, features =c("SST", "VIP"))
FeaturePlot(merged_xdp, features =c("GAD1", "GAD2"))
FeaturePlot(merged_xdp, features = c("EPHA4", "SEMA3E"))
FeaturePlot(merged_xdp, features =c("PVALB", "CALB2"))
FeaturePlot(merged_xdp, features =c("SLC17A7", "SLC17A6"))

```





