---
title: "R Notebook"
output: html_notebook
---

#Goals today: do the Matrix scoring on the integrated one- this will be the one to show tomorrow
#highlight low scores that are in the matrix clusters, ignore patch
#do the 3 graphs to show low numi, high pct_mito, low matrix score
#Also cluster BICAN data based on sct counts so that it is comparable 
#If you have time, that is when you do 


```{r}
library(Seurat)
library(ggplot2)
library(dplyr)
library(tidyr)
library(qs)
library(tibble)
library(gridExtra)
```

#functions: plot_histogram, 
```{r}
plot_overlapping_density_histogram = function(
    df, 
    hist_col,
    fill_col,
    colors = c("blue", "red"),
    alpha=0.5,
    breaks=seq(0, 16, 1),
    title= NULL,
    xlab = NULL,
    fig_filename = NULL

){
    # hist_col is the column you're making a histogram of (e.g. nUMI)
    # fill_col is the column you're coloring by (e.g. cell_class)
    # if fig_filename is not null, the plot will be saved to that file
     
    if (is.null(xlab)){
        xlab = hist_col
    }

    if (is.null(title)){
        title = paste0("Density Histogram of ", xlab, " by ", fill_col)
    }


    p = (
        ggplot(df, aes_string(x=hist_col, fill=fill_col)) 
        + geom_histogram(aes(y=..density..), alpha=alpha, position="identity", breaks=breaks)
        + labs(title=title, x=xlab, y="Density")    
        + theme(
                plot.title = element_text(size=16),
                axis.line = element_line(color = "black"),  # Add axis lines
                axis.ticks = element_line(color = "black"),  # Add axis ticks
                axis.text = element_text(size = 14),  # Increase tick label font size
                axis.title = element_text(size = 15)  # Increase axis label font size
            ) 
        + scale_fill_manual(values=colors)   
    )

    if (!is.null(fig_filename)){
        ggsave(fig_filename, p, width=8, height=6)
    }

    return(p)
}

SPN_score_normalized = function(gene_df, SCtransform_counts, logfc_val, method = "multiply"){

significant_genes = unique(gene_df$gene)

exp = FetchData(SCtransform_counts, vars = significant_genes)

rownames(gene_df) = gene_df$gene
gene_df$gene = NULL

# Assuming gene_df and exp are your dataframes
logfc_df <- gene_df
counts_df <- exp

print(logfc_df)
print(counts_df)

# Check if all column names in counts_df are in the row names of logfc_df
if (!all(colnames(counts_df) %in% rownames(logfc_df))) {
  stop("Not all genes in counts_df are present in logfc_df")
}

# Reorder logfc_df to match the columns in counts_df
logfc_ordered <- logfc_df[colnames(counts_df), , drop = FALSE]
logfc_ordered

logfc = logfc_ordered[[logfc_val]]

# Check for NA values in logfc_ordered
if (any(is.na(logfc))) {
  stop("There are NA values in the logFC values for the genes")
}


# Ensure logfc_ordered$logfc is a numeric vector
logfc_vector <- as.numeric(logfc)
logfc_vector

 gene_averages <- colMeans(counts_df)
 print("Printing gene averages")
 print(gene_averages)
if (method == "multiply") {
  
#THE ACTUAL
result_df <- sweep(counts_df, 2, logfc_vector, FUN = "*")
} else if(method == "log"){
  custom_log_transform <- function(x, logFC_value) {
     x * 2^logFC_value
  }
  result_df <- sweep(counts_df, 2, logfc_vector, FUN = custom_log_transform)
} else{
  stop("Calculation is not working")
}
print(result_df)
# Divide each gene's expression by its average expression
 result_df <- sweep(result_df, 2, gene_averages, FUN = "/")

# Print the result
print("Printing normalized")
print(result_df)


metadata <- SCtransform_counts@meta.data
# Assume that metadata has a column 'donor' that contains donor information
# If not, modify this part according to your metadata structure
if(!"donor_id" %in% colnames(metadata)) {
  stop("Metadata does not contain 'donor' information. Please check your metadata structure.")
}

# Create unique identifiers for cell column to avoid conflicts
expr_data <- result_df %>%
  rownames_to_column(var = "cell_id")

metadata <- metadata %>%
  rownames_to_column(var = "cell_id")

# Combine expression data with donor information
expr_data_long <- expr_data %>%
  pivot_longer(cols = -cell_id, names_to = "gene", values_to = "expression") %>%
  left_join(metadata %>% select(cell_id, donor_id, Condition, cell_class, reclustered_cell_class, reclustered_neuron_type, reclustered_patch_matrix, reclustered_subcluster, reclustered_patch_matrix_exotic, reclustered_neuron_type_joint_cluster, pct_mito, nUmi, region), by = "cell_id")
# Assuming your data frame is named expr_data_long
print(expr_data_long)
# Count the number of occurrences for each gene per donor

summarized_expr_data <- expr_data_long %>%
  group_by(cell_id, donor_id, Condition, cell_class, reclustered_cell_class, reclustered_neuron_type, reclustered_patch_matrix, reclustered_subcluster, reclustered_patch_matrix_exotic, reclustered_neuron_type_joint_cluster, pct_mito, nUmi,region) %>%
  summarize(total_expression = sum(expression), .groups = 'drop')
print(summarized_expr_data)

spn_scores = summarized_expr_data$total_expression
mean_spn <- mean(spn_scores, na.rm = TRUE)
sd_spn <- sd(spn_scores, na.rm = TRUE)
# Compute the z-score for each SPN score
summarized_expr_data$z_score  <- (spn_scores - mean_spn) / sd_spn

return(summarized_expr_data)
}

# # Add z-scores to Seurat object metadata
# caudate_neurons_transformed <- AddMetaData(caudate_neurons_transformed, metadata = spn_zscores, col.name = "SPN_zscore")
# 
# # Visualize SPN z-scores on UMAP
# plot <- FeaturePlot(caudate_neurons_transformed, features = "SPN_zscore") +
#   ggtitle("SPN Z-Score on UMAP")
# 
# # Save the UMAP plot with SPN z-scores as a PNG file
# ggsave("SPN_zscore_umap.png", plot = plot, width = 10, height = 8, dpi = 300)


SPN_score_graphs <- function(SPN_score_output, 
                             donor_order = donors, 
                             total_expression = "total_expression", 
                             fillcol = "reclustered_subcluster", 
                             color = color1, 
                             donor_graph_title = "CaH SPNs vs non-SPNs", 
                             SCtransformcounts, 
                             SPN_score_col_name) {
  
  # Ensure total_expression is a string
  total_expression <- as.character(total_expression)
  
  # Calculate min, max, and median
  min_SPN_score <- min(SPN_score_output[[total_expression]], na.rm = TRUE)
  max_SPN_score <- max(SPN_score_output[[total_expression]], na.rm = TRUE)
  median_SPN_score <- median(SPN_score_output[[total_expression]], na.rm = TRUE)
  
  print("min, max, median scores")
  print(min_SPN_score)
  print(max_SPN_score)
  print(median_SPN_score)
  
  options(repr.plot.width = 24, repr.plot.height = 16)
  plots <- list()
  
  for (donor in donor_order) {
    test <- SPN_score_output[SPN_score_output$donor_id == donor, ]
    plots[[donor]] <- plot_overlapping_density_histogram(df = test,
                                                         hist_col = total_expression,
                                                         fill_col = fillcol,
                                                         colors = color,
                                                         breaks = seq(min_SPN_score, max_SPN_score, round(max_SPN_score / 50)),
                                                         title = paste(donor_graph_title, ": ", donor),
                                                         xlab = "SPN Matrix Score",
                                                         fig_filename = NULL)
  }
  
  layout_matrix <- rbind(
    c(1, 2, 3, 4, 5),
    c(6, 7, 8, 9, 10),
    c(11, 12, 13, NA, NA),
    c(14, 15, 16, 17, NA)
  )
  
  # Arrange the plots according to the custom layout
  grid_plots <- grid.arrange(grobs = plots, layout_matrix = layout_matrix)
  
  # Save the arranged plots to a PNG file
  ggsave(filename = "~/SPN_SCORE_Output/USETEST.png", plot = grid_plots, width = 30, height = 16)
  
  print("donor plot done")
  
  XDP <- SPN_score_output[SPN_score_output$Condition == "XDP", ]
  a <- plot_overlapping_density_histogram(df = XDP, 
                                          hist_col = total_expression,
                                          fill_col = fillcol,
                                          colors = color,
                                          breaks = seq(min_SPN_score, max_SPN_score, round(max_SPN_score / 50)),
                                          title = paste(donor_graph_title, ": ", "XDP"),
                                          xlab = "SPN Matrix score",
                                          fig_filename = NULL)
  
  Control <- SPN_score_output[SPN_score_output$Condition == "Control", ]
  b <- plot_overlapping_density_histogram(df = Control, 
                                          hist_col = total_expression,
                                          fill_col = fillcol,
                                          colors = color,
                                          breaks = seq(min_SPN_score, max_SPN_score, round(max_SPN_score / 50)),
                                          title = paste(donor_graph_title, ": ", "Control"),
                                          xlab = "SPN Matrix score",
                                          fig_filename = NULL)
  
  print("Condition plot done")
  
    
  SPN <- SPN_score_output[SPN_score_output$reclustered_subcluster == "SPN", ]
  c <- plot_overlapping_density_histogram(df = SPN, 
                                          hist_col = total_expression,
                                          fill_col = "Condition",
                                          colors = c("XDP" = "red", "Control" = "blue"),
                                          breaks = seq(min_SPN_score, max_SPN_score, round(max_SPN_score / 50)),
                                          title = "XDP vs Control: SPN",
                                          xlab = "SPN Matrix score",
                                          fig_filename = NULL)
  
  notSPN <- SPN_score_output[SPN_score_output$reclustered_subcluster == "non-SPN", ]
  d <- plot_overlapping_density_histogram(df = notSPN, 
                                          hist_col = total_expression,
                                          fill_col = "Condition",
                                          colors = c("XDP" = "red", "Control" = "blue"),
                                          breaks = seq(min_SPN_score, max_SPN_score, round(max_SPN_score / 50)),
                                          title = "XDP vs Control: non-SPN",
                                          xlab = "SPN Matrix score",
                                          fig_filename = NULL)
  
  print(a)
  print(b)
  print(c)
  print(d)
  
  print("graphs done")
  
  # Calculate the threshold
  threshold <- quantile(SPN_score_output[[total_expression]][SPN_score_output[["reclustered_subcluster"]] == "SPN" & SPN_score_output[["Condition"]] == "XDP"], 0.1, na.rm = TRUE)
  print(threshold)
  
  # Subset the data
  SPN_score_output_identity_loss_SPN <- SPN_score_output[SPN_score_output[[total_expression]] < threshold & SPN_score_output[["reclustered_subcluster"]] == "SPN", ]
  
  print(SPN_score_output_identity_loss_SPN)
  print(table(SPN_score_output_identity_loss_SPN$donor_id))
  
  identity_loss_cells <- unique(SPN_score_output_identity_loss_SPN$cell_id)
  
  SCtransformcounts@meta.data$SPN_identity <- "Other"
  SCtransformcounts$SPN_identity[identity_loss_cells] <- "Losing_Identity"
  SCtransformcounts@meta.data
  
  # Visualize on UMAP
  pic <- DimPlot(SCtransformcounts, reduction = "umap", group.by = "SPN_identity", split.by = "Condition", cols = c("red", "grey"))
  print(pic)
  ggsave("~/SPN_SCORE_Output/SPN_score_10pct_TEST.png", plot = pic, width = 20, height = 8, units = "in", dpi = 300)
  
  spn_scores <- setNames(SPN_score_output[[total_expression]], SPN_score_output$cell_id)
  
  SCtransformcounts <- AddMetaData(SCtransformcounts, metadata = spn_scores, col.name = SPN_score_col_name)
  
  pic2 <- FeaturePlot(SCtransformcounts, features = SPN_score_col_name, split.by = "Condition")
  print(pic2)
  ggsave("~/SPN_SCORE_Output/SPN_UMAP_TEST.png", plot = pic2, width = 20, height = 8, units = "in", dpi = 300)
  
  return(SCtransformcounts)
}

```

#Lost things to make this sobj cause R suddenly crashed/restarted and lost my stuff

```{r}
final_merged_xdp_transformed = qread("~/SOBJ_USE_THESE/final_merged_xdp_transformed_081224.qs")
final_merged_xdp_transformed
final_merged_xdp_transformed@meta.data
DimPlot(final_merged_xdp_transformed, group.by = "reclustered_neuron_type_joint_cluster")
```

#Previously I had been doing matrix(including exotics) with only interneurons. But I should do only matrix vs non-spns (including eSPNs)

#but first I will validate
```{r}
head(final_merged_xdp_transformed)
```


```{r}
merged_xdp_transformed_controls = subset(final_merged_xdp_transformed, subset = Condition == "Control")
merged_xdp_transformed_controls
Idents(merged_xdp_transformed_controls) = "reclustered_neuron_type_joint_cluster"
DimPlot(merged_xdp_transformed_controls, label = TRUE)
```
#I got 804 genes originally but now getting 1024 and I don't udnerstand how!!!! FUCK R
#NOW ITS BACK TO 804????
#welp

```{r}
xdp_matrix_markers =FindMarkers(object = merged_xdp_transformed_controls, ident.1 = "SPN_matrix", ident.2 = "non-SPN", only.pos = TRUE)
                                #,  min.pct = 0.2, logfc.threshold = 1.25)
xdp_matrix_markers

xdp_matrix_markers_final = subset(xdp_matrix_markers, subset = p_val_adj < 0.05)
xdp_matrix_markers_final$pct_subtract = xdp_matrix_markers_final$pct.1 - xdp_matrix_markers_final$pct.2
xdp_matrix_markers_final$weighted_logFC = xdp_matrix_markers_final$avg_log2FC * xdp_matrix_markers_final$pct_subtract
xdp_matrix_markers_final

xdp_matrix_markers_final$gene = rownames(xdp_matrix_markers_final)
```
```{r}
Intersected_xdp_matrix_markers_final = subset(xdp_matrix_markers_final, subset = gene %in% genes_in_both)
Intersected_xdp_matrix_markers_final
```

#xdp integrated Matrix score
```{r}
XDP_matrix_scores = SPN_score_normalized(Intersected_xdp_matrix_markers_final, final_merged_xdp_transformed, logfc_val = "weighted_logFC", method = "multiply")
```


```{r}
donors = c("PCMC-16-011","PCMC-16-012", "SCF-18-003", "SCF-18-004", "SCF-18-006", "SCF-19-009", "SCF-19-014",  "SCF-19-018", "SCF-20-023", "SCF_20-024", "SCF-20-025", "SCF-21-030", "SCF_22-043","SCF_21-037CM2","SCF-22-054CM", "SCF-22-058CF","SCF-23-068CM") 
color6 = c("SPN_matrix" = "red", "non-SPN" = "blue", "SPN_patch" = "orange")

SPN_score_graphs(XDP_matrix_scores, donor_order = donors, total_expression = "total_expression",  fillcol = "reclustered_patch_matrix", 
                             color = color6, 
                             donor_graph_title = "SPN Matrix vs non-SPNs", 
                             final_merged_xdp_transformed, 
                             SPN_score_col_name = "test") 
```

```{r}
color1 = c("SPN" = "red", "non-SPN"= "blue") #for subcluster: SPN vs non-SPNS
color2 = c("red", "blue")
color3 = c("SPN"="red", "eSPN"="yellow", "other"="green", "IN" = "blue") #for subclass

SPN_score_graphs(XDP_matrix_scores, donor_order = donors, total_expression = "total_expression",  fillcol = "reclustered_subcluster", 
                             color = color1, 
                             donor_graph_title = "Matrix SPNs vs non-SPNs", 
                             final_merged_xdp_transformed, 
                             SPN_score_col_name = "test")
```

```{r}
color3 = c("SPN_matrix"="red", "SPN_patch" = "orange", "SPN_exotic" = "yellow" ,"eSPN"="green", "non-SPN" = "blue") 

SPN_score_graphs(XDP_matrix_scores, donor_order = donors, total_expression = "total_expression",  fillcol = "reclustered_patch_matrix_exotic", 
                             color = color3, 
                             donor_graph_title = "Matrix SPNs vs non-SPNs", 
                             final_merged_xdp_transformed, 
                             SPN_score_col_name = "test")
```


```{r}
matrix <- XDP_matrix_scores[XDP_matrix_scores$reclustered_patch_matrix_exotic == "SPN_matrix", ]
  plot_overlapping_density_histogram(df = matrix, 
                                          hist_col = matrix$total_expression,
                                          fill_col = "Condition",
                                          colors = c("XDP" = "red", "Control" = "blue"),
                                          breaks = seq(0, 150, 2),
                                          title = "XDP vs Control: SPN matrix",
                                          xlab = "SPN Matrix score",
                                          fig_filename = NULL)
  patch <- XDP_matrix_scores[XDP_matrix_scores$reclustered_patch_matrix_exotic == "SPN_patch", ]
  plot_overlapping_density_histogram(df = patch, 
                                          hist_col = patch$total_expression,
                                          fill_col = "Condition",
                                          colors = c("XDP" = "red", "Control" = "blue"),
                                          breaks = seq(0, 150, 2),
                                          title = "XDP vs Control: SPN patch",
                                          xlab = "SPN Matrix score",
                                          fig_filename = NULL)
  nonSPN <- XDP_matrix_scores[XDP_matrix_scores$reclustered_patch_matrix_exotic == "non-SPN", ]
  plot_overlapping_density_histogram(df = nonSPN, 
                                          hist_col = nonSPN$total_expression,
                                          fill_col = "Condition",
                                          colors = c("XDP" = "red", "Control" = "blue"),
                                          breaks = seq(0, 150, 2),
                                          title = "XDP vs Control: non-SPN",
                                          xlab = "SPN Matrix score",
                                          fig_filename = NULL)
    exotic <- XDP_matrix_scores[XDP_matrix_scores$reclustered_patch_matrix_exotic == "SPN_exotic", ]
  plot_overlapping_density_histogram(df = exotic, 
                                          hist_col = exotic$total_expression,
                                          fill_col = "Condition",
                                          colors = c("XDP" = "red", "Control" = "blue"),
                                          breaks = seq(0, 150, 2),
                                          title = "XDP vs Control: Exotic SPN",
                                          xlab = "SPN Matrix score",
                                          fig_filename = NULL)
    eSPN <- XDP_matrix_scores[XDP_matrix_scores$reclustered_patch_matrix_exotic == "eSPN", ]
  plot_overlapping_density_histogram(df = eSPN, 
                                          hist_col = eSPN$total_expression,
                                          fill_col = "Condition",
                                          colors = c("XDP" = "red", "Control" = "blue"),
                                          breaks = seq(0, 150, 2),
                                          title = "XDP vs Control: eSPN",
                                          xlab = "SPN Matrix score",
                                          fig_filename = NULL)
```
```{r}
SPN_score_output
```


```{r}
 options(repr.plot.width = 24, repr.plot.height = 16)
  plots <- list()
   SPN_score_output=  XDP_matrix_scores 
  for (donor in donors) {
   test <- SPN_score_output[SPN_score_output$donor_id == donor, ]
   threshold <- quantile(test[["total_expression"]][test[["reclustered_patch_matrix"]] == "SPN_matrix"], 0.1, na.rm = TRUE)
  print(threshold)
    
  SPN_score_output_identity_loss_SPN <- test[test[["total_expression"]] < threshold & test[["reclustered_patch_matrix"]] == "SPN_matrix", ]
   
  print(SPN_score_output_identity_loss_SPN)
  identity_loss_cells <- unique(SPN_score_output_identity_loss_SPN$cell_id)
  
  SCtransformcounts = subset(final_merged_xdp_transformed, subset = donor_id == donor)
  
  SCtransformcounts@meta.data$SPN_identity <- "Other"
  SCtransformcounts$SPN_identity[identity_loss_cells] <- "Bottom 10% of Matrix SPNs"
  SCtransformcounts@meta.data
  
  # Visualize on UMAP
  plots[[donor]] <- DimPlot(SCtransformcounts, reduction = "umap", group.by = "SPN_identity", cols = c("red", "grey")) +ggtitle(paste(donor, ", 10% threshold:", sprintf("%.2f", threshold)))
}
  
  
  layout_matrix <- rbind(
    c(1, 2, 3, 4, 5),
    c(6, 7, 8, 9, 10),
    c(11, 12, 13, NA, NA),
    c(14, 15, 16, 17, NA)
  )
  
  # Arrange the plots according to the custom layout
  grid_plots <- grid.arrange(grobs = plots, layout_matrix = layout_matrix)
  
  # Save the arranged plots to a PNG file
  ggsave(filename = "~/SPN_SCORE_Output/tenten.png", plot = grid_plots, width = 36, height = 16)

```

```{r}
SPN_score_output=  XDP_matrix_scores 
   threshold <- quantile(SPN_score_output[["total_expression"]][SPN_score_output[["reclustered_patch_matrix"]] == "SPN_matrix"], 0.1, na.rm = TRUE)
  print(threshold)
    
  SPN_score_output_identity_loss_SPN <- SPN_score_output[SPN_score_output[["total_expression"]] < threshold & SPN_score_output[["reclustered_patch_matrix"]] == "SPN_matrix", ]
   
  print(SPN_score_output_identity_loss_SPN)
  identity_loss_cells <- unique(SPN_score_output_identity_loss_SPN$cell_id)
  
  SCtransformcounts = final_merged_xdp_transformed
  
  SCtransformcounts@meta.data$SPN_identity <- "Other"
  SCtransformcounts$SPN_identity[identity_loss_cells] <- "Bottom 10% of Matrix SPNs"
  SCtransformcounts@meta.data
  
  # Visualize on UMAP
DimPlot(SCtransformcounts, reduction = "umap", group.by = "SPN_identity", cols = c("red", "grey"), split.by = "Condition")

```



#validate with BICAN
#first redo BICAN V8 clustering: 


#BICAN validation 
```{r}
BICAN_V8_CaH = qread("~/SOBJ_USE_THESE/BICAN_3P_V8_CaH.qs")
BICAN_V8_CaH
BICAN_V8_CaH@meta.data
DimPlot(BICAN_V8_CaH, label = T)
```
```{r}
BICAN_V8_CaH = subset(BICAN_V8_CaH, subset = predClass == "SPN" | predClass == "interneuron")
BICAN_V8_CaH
BICAN_V8_CaH@meta.data
DimPlot(BICAN_V8_CaH, label = T)
```

```{r}
BICAN_V8_CaH <- subset(BICAN_V8_CaH, subset = pct_mt < 10 & pct_intronic >= 0.25)
BICAN_V8_CaH <- subset(BICAN_V8_CaH, subset = pct_mt < 10 & pct_intronic >= 0.25)
BICAN_V8_CaH
```

```{r}
normalizeScalePcaClusterUmap = function(
    sobj,
    subset_col="donor_id",
    n_hvgs_orig=2500,
    n_dims_use=20,
    resolutions=c(0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1)){
    # this function takes a Seurat object, normalizes the data, scales the data, runs PCA, finds neighbors, clusters, and runs UMAP
    # it then returns the Seurat object
    hvgs = getSeuratVarFeatureIntersectByCol(sobj, subset_col=subset_col, original_nfeatures=n_hvgs_orig)
    sobj = (sobj
        %>% NormalizeData() # log normalizes raw counts
        %>% ScaleData(features=hvgs, split.by=subset_col) # within each cell, for each gene scale the data (i.e. subtract the donor's mean and divide by the donor's standard deviation of that gene)
        %>% RunPCA(features=hvgs, npcs=n_dims_use) # Reduce the dimensions to the n_dims_use dimensions that best explain the data https://en.wikipedia.org/wiki/Principal_component_analysis
        %>% FindNeighbors(dims = 1:n_dims_use) # Finds every cells closest neighbors in the PCA space
    )
    for(res in resolutions){
        sobj = sobj %>% FindClusters(resolution = res) # finds clusters at a variety of resolutions
    }
    sobj = sobj %>% RunUMAP(dims = 1:n_dims_use) # projects the PCA onto 2 dimensions, such that both "global structure" and "local structure" are optimally preserved
    return(sobj)
}
```


```{r}
DimPlot(BICAN_V8_CaH, label =TRUE)
BICAN_V8_CaH_neurons = normalizeScalePcaClusterUmap(BICAN_V8_CaH, subset_col = "DONOR", resolutions = c(0.01, 0.05, 0.1, 0.15, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1))

Idents(BICAN_V8_CaH_neurons) = BICAN_V8_CaH_neurons@meta.data$predClass
DimPlot(BICAN_V8_CaH_neurons, label =TRUE)
```

```{r}
table(BICAN_V8_CaH_neurons$predClass)
head(BICAN_V8_CaH_neurons)
```

```{r}
Idents(BICAN_V8_CaH_neurons) = BICAN_V8_CaH_neurons@meta.data$predClass
DimPlot(BICAN_V8_CaH_neurons, group.by = "RNA_snn_res.0.2")
FeaturePlot(BICAN_V8_CaH_neurons, features = c("CASZ1", "CHAT"))
```

```{r}
FeaturePlot(BICAN_V8_CaH_neurons, features = c("DRD1", "DRD2"))
FeaturePlot(BICAN_V8_CaH_neurons, features = c("EPHA4", "SEMA3E"))
FeaturePlot(BICAN_V8_CaH_neurons, features =c("CASZ1","CHAT"))
FeaturePlot(BICAN_V8_CaH_neurons, features =c("C1QA"))
FeaturePlot(BICAN_V8_CaH_neurons, features = c("PPP1R1B", "pct_mt"))
FeaturePlot(BICAN_V8_CaH_neurons, features = c("GAD1", "GAD2"))
FeaturePlot(BICAN_V8_CaH_neurons, features = c("VIP", "CALB2"))
FeaturePlot(BICAN_V8_CaH_neurons, features = c("SLC17A7", "SLC17A6"))

FeaturePlot(BICAN_V8_CaH_neurons, features = c("DRD1", "DRD2"))
#Patch
FeaturePlot(BICAN_V8_CaH_neurons, features = c("SEMA5B", "MFGE8")) 
FeaturePlot(BICAN_V8_CaH_neurons, features = c("KREMEN1", "PDE1C")) 
FeaturePlot(BICAN_V8_CaH_neurons, features = c("LYPD1")) 
#matrix
FeaturePlot(BICAN_V8_CaH_neurons, features = c("ID4", "SGK1")) 
FeaturePlot(BICAN_V8_CaH_neurons, features = c("EPHA4", "SV2B")) 
 #Coexpression is exopatch
FeaturePlot(BICAN_V8_CaH_neurons, features = c("ID4", "OPRM1"))
#first 2 is D1 patch,
FeaturePlot(BICAN_V8_CaH_neurons, features = c("NECAB1", "COL11A1")) 
#next is D2 patch
FeaturePlot(BICAN_V8_CaH_neurons, features = c("ASIC4")) 

FeaturePlot(BICAN_V8_CaH_neurons, features = c("TSHZ1", "TAC1", "NNAT")) #patch

FeaturePlot(BICAN_V8_CaH_neurons, features = c("pct_mt"))
FeaturePlot(BICAN_V8_CaH_neurons, features = c("PPP1R1B", "CASZ1")) 
```

```{r}
subcluster = c("D1_SPN_matrix","D2_SPN_matrix", "D1_SPN_patch", "D2_SPN_patch", "interneuron_1", "eSPN", "SPN-junk", "interneuron_2", "interneuron_3","D2_SPN_patch","cholinergic", "D1_SPN_matrix")

BICAN_V8_CaH_neurons= assignCellClasses(BICAN_V8_CaH_neurons, classes=subcluster, cluster_col="RNA_snn_res.0.2", class_col = "subcluster")

Idents(BICAN_V8_CaH_neurons) <- "subcluster"
DimPlot(BICAN_V8_CaH_neurons, label = TRUE)
```

#Now sctransform, recluster, integrate, relabel

```{r}
library(sctransform)
BICAN_V8_CaH_neurons_transformed = SCTransform(BICAN_V8_CaH_neurons, vars.to.regress = "pct_mt", verbose = FALSE)
DefaultAssay(BICAN_V8_CaH_neurons_transformed) = "SCT"
BICAN_V8_CaH_neurons_transformed
```

```{r}
BICAN_V8_CaH_neurons_transformed[["umap_rna"]] = BICAN_V8_CaH_neurons_transformed[["umap"]]
BICAN_V8_CaH_neurons_transformed
```

#Actual Clustering Code- BICAN
```{r}
hvgs = getSeuratVarFeatureIntersectByCol(BICAN_V8_CaH_neurons_transformed, subset_col="DONOR", original_nfeatures=2500)


n_dims_use=20


BICAN_V8_CaH_neurons_transformed_reclustered = (BICAN_V8_CaH_neurons_transformed
   %>% NormalizeData() # log normalizes raw counts
   %>% ScaleData(features=hvgs, split.by="DONOR") # within each cell, for each gene scale the data (i.e. subtract the donor's mean and divide by the donor's standard deviation of that gene)
   %>% RunPCA(features=hvgs, npcs=n_dims_use) # Reduce the dimensions to the n_dims_use dimensions that best explain the data https://en.wikipedia.org/wiki/Principal_component_analysis
   %>% FindNeighbors(dims = 1:n_dims_use) # Finds every cells closest neighbors in the PCA space
   %>% FindClusters(resolution = 0.2) # finds clusters at a variety of resolutions
   %>% FindClusters(resolution = 0.3) # after clustering, the cluster labels are accessible via merged_xdp_transformed$RNA_snn_res.{resolution}
   %>% FindClusters(resolution = 0.4) # e.g.: merged_xdp_transformed$RNA_snn_res.0.4 here
   %>% FindClusters(resolution = 0.5)
   %>% FindClusters(resolution = 0.6)
   %>% FindClusters(resolution = 0.7)
   %>% FindClusters(resolution = 0.8)
   %>% FindClusters(resolution = 0.9)
   %>% FindClusters(resolution = 1)
   %>% RunUMAP(dims = 1:n_dims_use) # projects the PCA onto 2 dimensions, such that both "global structure" and "local structure" are optimally preserved
)


# Now we can plot the data and just see what it looks like


setwh = function(w, h){
   # set the width of the ensuing figures to w, height to h
   options(repr.plot.width = w, repr.plot.height = h)
}

library(harmony)
BICAN_V8_CaH_neurons_transformed_reclustered_integrated = (BICAN_V8_CaH_neurons_transformed_reclustered
    %>% RunHarmony(
        group.by = "DONOR")
    %>% FindNeighbors(reduction='harmony', dims=1:20)
    %>% FindClusters(res=0.2)
    %>% FindClusters(res=0.3)
    %>% FindClusters(res=0.4)
    %>% FindClusters(res=0.5)
    %>% RunUMAP(reduction="harmony", dims=1:20)
)
BICAN_V8_CaH_neurons_transformed_reclustered_integrated
BICAN_V8_CaH_neurons_transformed_reclustered_integrated@meta.data
```

```{r}
qsave(BICAN_V8_CaH_neurons_transformed_reclustered_integrated, "~/BICAN_V8_CaH_neurons_transformed_reclustered_integrated.qs")
```


```{r}
DimPlot(BICAN_V8_CaH_neurons_transformed_reclustered_integrated, group.by = "predClass")
DimPlot(BICAN_V8_CaH_neurons_transformed_reclustered_integrated, group.by = "DONOR")
DimPlot(BICAN_V8_CaH_neurons_transformed_reclustered_integrated, group.by = "library")
```


```{r}
DimPlot(BICAN_V8_CaH_neurons_transformed_reclustered_integrated, group.by = "SCT_snn_res.0.2")
DimPlot(BICAN_V8_CaH_neurons_transformed_reclustered_integrated, group.by = "SCT_snn_res.0.3")
DimPlot(BICAN_V8_CaH_neurons_transformed_reclustered_integrated, group.by = "SCT_snn_res.0.4")
DimPlot(BICAN_V8_CaH_neurons_transformed_reclustered_integrated, group.by = "SCT_snn_res.0.5")

FeaturePlot(BICAN_V8_CaH_neurons_transformed_reclustered_integrated, features = c("DRD1", "DRD2"))
FeaturePlot(BICAN_V8_CaH_neurons_transformed_reclustered_integrated, features = c("EPHA4", "SEMA3E"))
FeaturePlot(BICAN_V8_CaH_neurons_transformed_reclustered_integrated, features =c("CASZ1","CHAT"))
FeaturePlot(BICAN_V8_CaH_neurons_transformed_reclustered_integrated, features =c("C1QA"))
FeaturePlot(BICAN_V8_CaH_neurons_transformed_reclustered_integrated, features = c("PPP1R1B", "pct_mt"))
FeaturePlot(BICAN_V8_CaH_neurons_transformed_reclustered_integrated, features = c("GAD1", "GAD2"))
FeaturePlot(BICAN_V8_CaH_neurons_transformed_reclustered_integrated, features = c("VIP", "CALB2"))
FeaturePlot(BICAN_V8_CaH_neurons_transformed_reclustered_integrated, features = c("SLC17A7", "SLC17A6"))
```

```{r}
FeaturePlot(BICAN_V8_CaH_neurons_transformed_reclustered_integrated, features = c("SEMA5B", "MTGE8", "KREMEN1", "PDE1C"))
FeaturePlot(BICAN_V8_CaH_neurons_transformed_reclustered_integrated, features = c("TSHZ1", "TAC1", "NNAT"))
FeaturePlot(BICAN_V8_CaH_neurons_transformed_reclustered_integrated, features = c("ID4", "OPRM1"))
FeaturePlot(BICAN_V8_CaH_neurons_transformed_reclustered_integrated, features = c("ID4", "SGK1", "EPHA4", "SV2B"))
FeaturePlot(BICAN_V8_CaH_neurons_transformed_reclustered_integrated, features = c("NECAB1", "CO11A1"))
FeaturePlot(BICAN_V8_CaH_neurons_transformed_reclustered_integrated, features = c("ASIC4"))
FeaturePlot(BICAN_V8_CaH_neurons_transformed_reclustered_integrated, features = c("CASZ1", "CACNG5", "PCDH8"))
FeaturePlot(BICAN_V8_CaH_neurons_transformed_reclustered_integrated, features = c("pct_mt", "NUM_TRANSCRIPTS"))
```
```{r}
DimPlot(BICAN_V8_CaH_neurons_transformed_reclustered_integrated, group.by = "predClass", label = T)
DimPlot(BICAN_V8_CaH_neurons_transformed_reclustered_integrated, group.by = "subcluster",label = T)
```

```{r}
assignCellClasses = function(
   obj,
   classes,
   cluster_col="seurat_clusters",
   class_col="cell_class") {
   obj@meta.data[[class_col]] = NA  # Assign NA to the entire column in meta.data


   for (i in 1:length(classes)) {
       # Find cells that match the current cluster number
       cells = which(as.numeric(obj@meta.data[[cluster_col]]) == i)
     
       # Assign the class label to the cells in the cluster
       obj@meta.data[cells, class_col] = classes[[i]]
   }
  
   # Return the modified object
   return(obj)
}
```

```{r}
a=FindAllMarkers(object = BICAN_V8_CaH_neurons_transformed_reclustered_integrated, only.pos = TRUE)
a
s= subset(a, subset = cluster =="D2_patch_1" | cluster == "D2_patch_2")
s
```

```{r}
classes = c("D1_matrix", "D2_matrix", "D1_patch", "D2_patch", "interneuron_1", "eSPN", "interneuron_2", "SPN_junk", "interneuron_3", "interneuron_4", "D2_patch", "cholinergic")

BICAN_V8_CaH_neurons_transformed_reclustered_integrated = assignCellClasses(BICAN_V8_CaH_neurons_transformed_reclustered_integrated, classes=classes, cluster_col="SCT_snn_res.0.2", class_col = "reclustered_cell_class")
Idents(BICAN_V8_CaH_neurons_transformed_reclustered_integrated) = "reclustered_cell_class"
DimPlot(BICAN_V8_CaH_neurons_transformed_reclustered_integrated, label = T)
```
```{r}
final_V8_BICAN_neurons = subset(BICAN_V8_CaH_neurons_transformed_reclustered_integrated, subset = reclustered_cell_class != "SPN_junk")
unique(final_V8_BICAN_neurons$reclustered_cell_class)
```
```{r}
final_V8_BICAN_neurons@meta.data$reclustered_subcluster = final_V8_BICAN_neurons@meta.data$reclustered_cell_class
final_V8_BICAN_neurons$reclustered_subcluster[final_V8_BICAN_neurons$reclustered_subcluster == "D1_matrix"] = "SPN"
final_V8_BICAN_neurons$reclustered_subcluster[final_V8_BICAN_neurons$reclustered_subcluster == "D2_matrix"] = "SPN"
final_V8_BICAN_neurons$reclustered_subcluster[final_V8_BICAN_neurons$reclustered_subcluster == "D1_patch"] = "SPN"
final_V8_BICAN_neurons$reclustered_subcluster[final_V8_BICAN_neurons$reclustered_subcluster == "D2_patch"] = "SPN"
final_V8_BICAN_neurons$reclustered_subcluster[final_V8_BICAN_neurons$reclustered_subcluster == "interneuron_1"] = "non-SPN"
final_V8_BICAN_neurons$reclustered_subcluster[final_V8_BICAN_neurons$reclustered_subcluster == "eSPN"] = "non-SPN"
final_V8_BICAN_neurons$reclustered_subcluster[final_V8_BICAN_neurons$reclustered_subcluster == "interneuron_2"] = "non-SPN"
final_V8_BICAN_neurons$reclustered_subcluster[final_V8_BICAN_neurons$reclustered_subcluster == "interneuron_3"] = "non-SPN"
final_V8_BICAN_neurons$reclustered_subcluster[final_V8_BICAN_neurons$reclustered_subcluster == "interneuron_4"] = "non-SPN"
final_V8_BICAN_neurons$reclustered_subcluster[final_V8_BICAN_neurons$reclustered_subcluster == "cholinergic"] = "non-SPN"

final_V8_BICAN_neurons@meta.data$reclustered_neuron_type = final_V8_BICAN_neurons$reclustered_subcluster
final_V8_BICAN_neurons$reclustered_neuron_type[final_V8_BICAN_neurons$reclustered_cell_class == "eSPN"] = "eSPN"

final_V8_BICAN_neurons@meta.data$reclustered_patch_matrix = final_V8_BICAN_neurons$reclustered_cell_class
final_V8_BICAN_neurons$reclustered_patch_matrix[final_V8_BICAN_neurons$reclustered_patch_matrix == "D1_matrix"] = "SPN_matrix"
final_V8_BICAN_neurons$reclustered_patch_matrix[final_V8_BICAN_neurons$reclustered_patch_matrix == "D2_matrix"] = "SPN_matrix"
final_V8_BICAN_neurons$reclustered_patch_matrix[final_V8_BICAN_neurons$reclustered_patch_matrix == "D1_patch"] = "SPN_patch"
final_V8_BICAN_neurons$reclustered_patch_matrix[final_V8_BICAN_neurons$reclustered_patch_matrix == "D2_patch"] = "SPN_patch"
final_V8_BICAN_neurons$reclustered_patch_matrix[final_V8_BICAN_neurons$reclustered_patch_matrix == "interneuron_1"] = "non-SPN"
final_V8_BICAN_neurons$reclustered_patch_matrix[final_V8_BICAN_neurons$reclustered_patch_matrix == "eSPN"] = "non-SPN"
final_V8_BICAN_neurons$reclustered_patch_matrix[final_V8_BICAN_neurons$reclustered_patch_matrix == "interneuron_2"] = "non-SPN"
final_V8_BICAN_neurons$reclustered_patch_matrix[final_V8_BICAN_neurons$reclustered_patch_matrix == "interneuron_3"] = "non-SPN"
final_V8_BICAN_neurons$reclustered_patch_matrix[final_V8_BICAN_neurons$reclustered_patch_matrix == "interneuron_4"] = "non-SPN"
final_V8_BICAN_neurons$reclustered_patch_matrix[final_V8_BICAN_neurons$reclustered_patch_matrix == "cholinergic"] = "non-SPN"

final_V8_BICAN_neurons@meta.data$reclustered_neuron_joint_type = final_V8_BICAN_neurons$reclustered_patch_matrix
final_V8_BICAN_neurons$reclustered_neuron_joint_type[final_V8_BICAN_neurons$reclustered_cell_class == "eSPN"] = "eSPN"
final_V8_BICAN_neurons@meta.data
```

```{r}
table(final_V8_BICAN_neurons$reclustered_cell_class)
table(final_V8_BICAN_neurons$reclustered_subcluster)
table(final_V8_BICAN_neurons$reclustered_neuron_type)
table(final_V8_BICAN_neurons$reclustered_patch_matrix)
table(final_V8_BICAN_neurons$reclustered_neuron_joint_type)
```
```{r}
qsave(final_V8_BICAN_neurons, "~/SOBJ_USE_THESE/final_V8_BICAN_neurons.qs")

#final_V8_BICAN_neurons = qread("~/SOBJ_USE_THESE/final_V8_BICAN_neurons.qs")
```

```{r}
DimPlot(final_V8_BICAN_neurons, group.by = "reclustered_cell_class" , label = T)
DimPlot(final_V8_BICAN_neurons, group.by = "reclustered_subcluster" , label = T)
DimPlot(final_V8_BICAN_neurons, group.by = "reclustered_neuron_type" , label = T)
DimPlot(final_V8_BICAN_neurons, group.by = "reclustered_patch_matrix", label = T)
DimPlot(final_V8_BICAN_neurons, group.by = "reclustered_neuron_joint_type" , label = T)
```


```{r}
BICAN_SPN_score_normalized = function(gene_df, SCtransform_counts, logfc_val, method = "multiply"){

significant_genes = unique(gene_df$gene)

exp = FetchData(SCtransform_counts, vars = significant_genes)

rownames(gene_df) = gene_df$gene
gene_df$gene = NULL

# Assuming gene_df and exp are your dataframes
logfc_df <- gene_df
counts_df <- exp

print(logfc_df)
print(counts_df)

# Check if all column names in counts_df are in the row names of logfc_df
if (!all(colnames(counts_df) %in% rownames(logfc_df))) {
  stop("Not all genes in counts_df are present in logfc_df")
}

# Reorder logfc_df to match the columns in counts_df
logfc_ordered <- logfc_df[colnames(counts_df), , drop = FALSE]
logfc_ordered

logfc = logfc_ordered[[logfc_val]]

# Check for NA values in logfc_ordered
if (any(is.na(logfc))) {
  stop("There are NA values in the logFC values for the genes")
}


# Ensure logfc_ordered$logfc is a numeric vector
logfc_vector <- as.numeric(logfc)
logfc_vector

gene_averages <- colMeans(counts_df)
print("Printing gene averages")
print(gene_averages)
if (method == "multiply") {
  
#THE ACTUAL
result_df <- sweep(counts_df, 2, logfc_vector, FUN = "*")
} else if(method == "log"){
  custom_log_transform <- function(x, logFC_value) {
     x * 2^logFC_value
  }
  result_df <- sweep(counts_df, 2, logfc_vector, FUN = custom_log_transform)
} else{
  stop("Calculation is not working")
}
print(result_df)
# Divide each gene's expression by its average expression
result_df <- sweep(result_df, 2, gene_averages, FUN = "/")

# Print the result
print("Printing normalized")
print(result_df)


metadata <- SCtransform_counts@meta.data
# Assume that metadata has a column 'donor' that contains donor information
# If not, modify this part according to your metadata structure
if(!"DONOR" %in% colnames(metadata)) {
  stop("Metadata does not contain 'donor' information. Please check your metadata structure.")
}

# Check if 'cell_id' column already exists and remove it
if ("cell_id" %in% colnames(metadata)) {
  metadata <- metadata %>% select(-cell_id)
}

# Ensure unique row names in result_df
if (any(duplicated(rownames(result_df)))) {
  stop("Row names in result_df are not unique. Please ensure unique row names.")
}


# Create unique identifiers for cell column to avoid conflicts
expr_data <- result_df %>%
  rownames_to_column(var = "cell_id")

metadata <- metadata %>%
  rownames_to_column(var = "cell_id")

# Combine expression data with donor information
expr_data_long <- expr_data %>%
  pivot_longer(cols = -cell_id, names_to = "gene", values_to = "expression") %>%
  left_join(metadata %>% select(cell_id, DONOR, subcluster, pct_mt, NUM_TRANSCRIPTS, reclustered_cell_class, reclustered_subcluster,reclustered_neuron_type,reclustered_patch_matrix,reclustered_neuron_joint_type), by = "cell_id")

# Assuming your data frame is named expr_data_long
print(expr_data_long)
# Count the number of occurrences for each gene per donor

summarized_expr_data <- expr_data_long %>%
  group_by(cell_id, DONOR, subcluster, pct_mt, NUM_TRANSCRIPTS, reclustered_cell_class, reclustered_subcluster,reclustered_neuron_type,reclustered_patch_matrix,reclustered_neuron_joint_type) %>%
  summarize(total_expression = sum(expression), .groups = 'drop')
print(summarized_expr_data)

spn_scores = summarized_expr_data$total_expression
mean_spn <- mean(spn_scores, na.rm = TRUE)
sd_spn <- sd(spn_scores, na.rm = TRUE)
# Compute the z-score for each SPN score
summarized_expr_data$z_score  <- (spn_scores - mean_spn) / sd_spn

return(summarized_expr_data)
}
```


```{r}
BICANdonors = unique(final_V8_BICAN_neurons$DONOR)
BICANdonors
```

```{r}
BICAN_SPN_score_graphs <- function(SPN_score_output, 
                             donor_order = BICANdonors, 
                             total_expression = "total_expression", 
                             fillcol = "subcluster", 
                             color = color1, 
                             donor_graph_title = "BICAN SPN Matrix vs non-SPNs", 
                             SCtransformcounts, 
                             SPN_score_col_name) {
  
  # Ensure total_expression is a string
  total_expression <- as.character(total_expression)
  
  # Calculate min, max, and median
  min_SPN_score <- min(SPN_score_output[[total_expression]], na.rm = TRUE)
  max_SPN_score <- max(SPN_score_output[[total_expression]], na.rm = TRUE)
  median_SPN_score <- median(SPN_score_output[[total_expression]], na.rm = TRUE)
  
  print("min, max, median scores")
  print(min_SPN_score)
  print(max_SPN_score)
  print(median_SPN_score)
  
  options(repr.plot.width = 24, repr.plot.height = 16)
  plots <- list()
  
  for (donor in donor_order) {
    test <- SPN_score_output[SPN_score_output$DONOR == donor, ]
    plots[[donor]] <- plot_overlapping_density_histogram(df = test,
                                                         hist_col = total_expression,
                                                         fill_col = fillcol,
                                                         colors = color,
                                                         breaks = seq(min_SPN_score, max_SPN_score, round(max_SPN_score/50)),
                                                         title = paste(donor_graph_title, ": ", donor),
                                                         xlab = "Matrix SPN Score",
                                                         fig_filename = NULL)
  }
  
  layout_matrix <- rbind(
    c(1, 2, 3, 4, 5),
    c(6, 7, 8, 9, 10),
    c(11, 12, 13, 14, 15),
    c(16, 17, 18, 19, NA)
  )
  
  # Arrange the plots according to the custom layout
  grid_plots <- grid.arrange(grobs = plots, layout_matrix = layout_matrix)
  
  # Save the arranged plots to a PNG file
  ggsave(filename = "~/SPN_SCORE_Output/USETEST.png", plot = grid_plots, width = 30, height = 16)
  
  print("donor plot done")
  
  b <- plot_overlapping_density_histogram(df = SPN_score_output, 
                                          hist_col = total_expression,
                                          fill_col = fillcol,
                                          colors = color,
                                          breaks = seq(min_SPN_score, max_SPN_score, round(max_SPN_score/50)),
                                          title = paste(donor_graph_title, ": ", "Control"),
                                          xlab = "SPN score",
                                          fig_filename = NULL)
  
  print("Condition plot done")
  

  
  # Calculate the threshold
  threshold <- quantile(SPN_score_output[[total_expression]][SPN_score_output[["reclustered_neuron_type"]] == "SPN"], 0.1, na.rm = TRUE)
  print(threshold)
  
  # Subset the data
  SPN_score_output_identity_loss_SPN <- SPN_score_output[SPN_score_output[[total_expression]] < threshold & SPN_score_output[["reclustered_neuron_type"]] == "SPN", ]
  
  print(SPN_score_output_identity_loss_SPN)
  print(table(SPN_score_output_identity_loss_SPN$DONOR))
  
  identity_loss_cells <- unique(SPN_score_output_identity_loss_SPN$cell_id)
  
  SCtransformcounts@meta.data$SPN_identity <- "Other"
  SCtransformcounts$SPN_identity[identity_loss_cells] <- "Bottom 10%"
  SCtransformcounts@meta.data
  
  # Visualize on UMAP
  pic <- DimPlot(SCtransformcounts, reduction = "umap", group.by = "SPN_identity", cols = c("red", "grey")) 
  print(pic)
  ggsave("~/SPN_SCORE_Output/SPN_score_10pct_TEST.png", plot = pic, width = 20, height = 8, units = "in", dpi = 300)
  
  spn_scores <- setNames(SPN_score_output[[total_expression]], SPN_score_output$cell_id)
  
  SCtransformcounts <- AddMetaData(SCtransformcounts, metadata = spn_scores, col.name = SPN_score_col_name)
  
  pic2 <- FeaturePlot(SCtransformcounts, features = SPN_score_col_name)
  print(pic2)
  ggsave("~/SPN_SCORE_Output/SPN_UMAP_TEST.png", plot = pic2, width = 20, height = 8, units = "in", dpi = 300)
  
  return(SCtransformcounts)
}

```


#BICAN validation 

```{r}
final_V8_BICAN_neurons@meta.data$cell_id = rownames(final_V8_BICAN_neurons@meta.data)
head(final_V8_BICAN_neurons)
```


```{r}
#this is the list of genes to us
genes_in_both = colnames(counts_df)
```

```{r}
XDP_matrix_scores = BICAN_SPN_score_normalized(Intersected_xdp_matrix_markers_final, final_V8_BICAN_neurons, logfc_val = "weighted_logFC", method = "multiply")
```

```{r}
color4 = c("SPN_matrix" = "red", "non-SPN" = "blue", "SPN_patch" = "orange")

BICAN_SPN_score_graphs(XDP_matrix_scores, donor_order = BICANdonors, total_expression = "total_expression",  fillcol = "reclustered_patch_matrix", 
                             color = color4, 
                             donor_graph_title = "Matrix SPNs vs non-SPNs", 
                             final_V8_BICAN_neurons, 
                             SPN_score_col_name = "BICANtest") 
```

```{r}
color1 = c("SPN" = "red", "non-SPN"= "blue") #for subcluster: SPN vs non-SPNS
color2 = c("red", "blue")
color3 = c("SPN"="red", "eSPN"="yellow", "other"="green", "IN" = "blue") #for subclass

BICAN_SPN_score_graphs(XDP_matrix_scores, donor_order = BICANdonors, total_expression = "total_expression",  fillcol = "reclustered_subcluster", 
                             color = color1, 
                             donor_graph_title = "SPNs vs non-SPNs", 
                             final_V8_BICAN_neurons, 
                             SPN_score_col_name = "BICANtest") 
```


```{r}
color5 = c("SPN_matrix"="red", "SPN_patch" = "orange","eSPN"="green", "non-SPN" = "blue") 

BICAN_SPN_score_graphs(XDP_matrix_scores, donor_order = BICANdonors, total_expression = "total_expression",  fillcol = "reclustered_neuron_joint_type", 
                             color = color5, 
                             donor_graph_title = "Matrix SPNs vs non-SPNs", 
                             final_V8_BICAN_neurons, 
                             SPN_score_col_name = "BICANtest") 
```

```{r}
matrix <- XDP_matrix_scores[XDP_matrix_scores$reclustered_neuron_joint_type == "SPN_matrix", ]
  plot_overlapping_density_histogram(df = matrix, 
                                          hist_col = matrix$total_expression,
                                          fill_col = "reclustered_neuron_joint_type",
                                          colors = c("SPN_matrix" = "red"),
                                          breaks = seq(0, 150, 2),
                                          title = "XDP vs Control: SPN matrix",
                                          xlab = "Matrix SPN score",
                                          fig_filename = NULL)
  patch <- XDP_matrix_scores[XDP_matrix_scores$reclustered_neuron_joint_type == "SPN_patch", ]
  plot_overlapping_density_histogram(df = patch, 
                                          hist_col = patch$total_expression,
                                          fill_col = "reclustered_neuron_joint_type",
                                          colors = c("SPN_patch" = "orange"),
                                          breaks = seq(0, 150, 2),
                                          title = "XDP vs Control: SPN patch",
                                          xlab = "Matrix SPN score",
                                          fig_filename = NULL)
  nonSPN <- XDP_matrix_scores[XDP_matrix_scores$reclustered_neuron_joint_type == "non-SPN", ]
  plot_overlapping_density_histogram(df = nonSPN, 
                                          hist_col = nonSPN$total_expression,
                                          fill_col = "reclustered_neuron_joint_type",
                                          colors = c("non-SPN" = "blue"),
                                          breaks = seq(0, 150, 2),
                                          title = "XDP vs Control: non-SPN",
                                          xlab = "Matrix SPN score",
                                          fig_filename = NULL)
    eSPN <- XDP_matrix_scores[XDP_matrix_scores$reclustered_neuron_joint_type == "eSPN", ]
  plot_overlapping_density_histogram(df = eSPN, 
                                          hist_col = eSPN$total_expression,
                                          fill_col = "reclustered_neuron_joint_type",
                                          colors = c("eSPN" = "green"),
                                          breaks = seq(0, 150, 2),
                                          title = "XDP vs Control: eSPN",
                                          xlab = "Matrix SPN score",
                                          fig_filename = NULL)
  
```
```{r}
XDP_matrix_scores
```


```{r}
 options(repr.plot.width = 24, repr.plot.height = 16)
  plots <- list()
   SPN_score_output=  XDP_matrix_scores 
  for (donor in BICANdonors) {
   test <- SPN_score_output[SPN_score_output$DONOR == donor, ]
   threshold <- quantile(test[["total_expression"]][test[["reclustered_patch_matrix"]] == "SPN_matrix"], 0.1, na.rm = TRUE)
  print(threshold)
    
  SPN_score_output_identity_loss_SPN <- test[test[["total_expression"]] < threshold & test[["reclustered_patch_matrix"]] == "SPN_matrix", ]
   
  print(SPN_score_output_identity_loss_SPN)
  identity_loss_cells <- unique(SPN_score_output_identity_loss_SPN$cell_id)
  
  SCtransformcounts = subset(final_V8_BICAN_neurons, subset = DONOR == donor)
  
  SCtransformcounts@meta.data$SPN_identity <- "Other"
  SCtransformcounts$SPN_identity[identity_loss_cells] <- "Bottom 10% of Matrix SPNs"
  SCtransformcounts@meta.data
  
  # Visualize on UMAP
  plots[[donor]] <- DimPlot(SCtransformcounts, reduction = "umap", group.by = "SPN_identity", cols = c("red", "grey")) +ggtitle(paste(donor, ", 10% threshold:", sprintf("%.2f", threshold)))
}
  
  
  layout_matrix <- rbind(
    c(1, 2, 3, 4, 5),
    c(6, 7, 8, 9, 10),
    c(11, 12, 13, 14, 15),
    c(16, 17, 18, 19, NA)
  )
  
  # Arrange the plots according to the custom layout
  grid_plots <- grid.arrange(grobs = plots, layout_matrix = layout_matrix)
  
  # Save the arranged plots to a PNG file
  ggsave(filename = "~/SPN_SCORE_Output/tenten.png", plot = grid_plots, width = 36, height = 16)

```

```{r}
SPN_score_output=  XDP_matrix_scores 
   threshold <- quantile(SPN_score_output[["total_expression"]][SPN_score_output[["reclustered_patch_matrix"]] == "SPN_matrix"], 0.1, na.rm = TRUE)
  print(threshold)
    
  SPN_score_output_identity_loss_SPN <- SPN_score_output[SPN_score_output[["total_expression"]] < threshold & SPN_score_output[["reclustered_patch_matrix"]] == "SPN_matrix", ]
   
  print(SPN_score_output_identity_loss_SPN)
  identity_loss_cells <- unique(SPN_score_output_identity_loss_SPN$cell_id)
  
  SCtransformcounts = final_V8_BICAN_neurons
  
  SCtransformcounts@meta.data$SPN_identity <- "Other"
  SCtransformcounts$SPN_identity[identity_loss_cells] <- "Bottom 10% of Matrix SPNs"
  SCtransformcounts@meta.data
  
  # Visualize on UMAP
DimPlot(SCtransformcounts, reduction = "umap", group.by = "SPN_identity", cols = c("red", "grey"))

```












