---
title: "R Notebook"
output: html_notebook
---

```{r}
counts_df
```


```{r}
BICAN_SPN_score_normalized = function(gene_df, SCtransform_counts, logfc_val, method = "multiply"){

significant_genes = unique(gene_df$gene)

exp = FetchData(SCtransform_counts, vars = significant_genes)

rownames(gene_df) = gene_df$gene
gene_df$gene = NULL

# Assuming gene_df and exp are your dataframes
logfc_df <- gene_df
counts_df <- exp

print(logfc_df)
print(counts_df)

# Check if all column names in counts_df are in the row names of logfc_df
if (!all(colnames(counts_df) %in% rownames(logfc_df))) {
  stop("Not all genes in counts_df are present in logfc_df")
}

# Reorder logfc_df to match the columns in counts_df
logfc_ordered <- logfc_df[colnames(counts_df), , drop = FALSE]
logfc_ordered

logfc = logfc_ordered[[logfc_val]]

# Check for NA values in logfc_ordered
if (any(is.na(logfc))) {
  stop("There are NA values in the logFC values for the genes")
}


# Ensure logfc_ordered$logfc is a numeric vector
logfc_vector <- as.numeric(logfc)
logfc_vector

gene_averages <- colMeans(counts_df)
print("Printing gene averages")
print(gene_averages)
if (method == "multiply") {
  
#THE ACTUAL
result_df <- sweep(counts_df, 2, logfc_vector, FUN = "*")
} else if(method == "log"){
  custom_log_transform <- function(x, logFC_value) {
     x * 2^logFC_value
  }
  result_df <- sweep(counts_df, 2, logfc_vector, FUN = custom_log_transform)
} else{
  stop("Calculation is not working")
}
print(result_df)
# Divide each gene's expression by its average expression
result_df <- sweep(result_df, 2, gene_averages, FUN = "/")

# Print the result
print("Printing normalized")
print(result_df)


metadata <- SCtransform_counts@meta.data
# Assume that metadata has a column 'donor' that contains donor information
# If not, modify this part according to your metadata structure
if(!"DONOR" %in% colnames(metadata)) {
  stop("Metadata does not contain 'donor' information. Please check your metadata structure.")
}

# Check if 'cell_id' column already exists and remove it
if ("cell_id" %in% colnames(metadata)) {
  metadata <- metadata %>% select(-cell_id)
}

# Ensure unique row names in result_df
if (any(duplicated(rownames(result_df)))) {
  stop("Row names in result_df are not unique. Please ensure unique row names.")
}


# Create unique identifiers for cell column to avoid conflicts
expr_data <- result_df %>%
  rownames_to_column(var = "cell_id")

metadata <- metadata %>%
  rownames_to_column(var = "cell_id")

# Combine expression data with donor information
expr_data_long <- expr_data %>%
  pivot_longer(cols = -cell_id, names_to = "gene", values_to = "expression") %>%
  left_join(metadata %>% select(cell_id, DONOR, subcluster, pct_mt, NUM_TRANSCRIPTS, reclustered_cell_class, reclustered_subcluster,reclustered_neuron_type,reclustered_patch_matrix,reclustered_neuron_joint_type), by = "cell_id")

# Assuming your data frame is named expr_data_long
print(expr_data_long)
# Count the number of occurrences for each gene per donor

summarized_expr_data <- expr_data_long %>%
  group_by(cell_id, DONOR, subcluster, pct_mt, NUM_TRANSCRIPTS, reclustered_cell_class, reclustered_subcluster,reclustered_neuron_type,reclustered_patch_matrix,reclustered_neuron_joint_type) %>%
  summarize(total_expression = sum(expression), .groups = 'drop')
print(summarized_expr_data)


return(summarized_expr_data)
}

SPN_score_normalized = function(gene_df, SCtransform_counts, logfc_val, method = "multiply"){

significant_genes = unique(gene_df$gene)

exp = FetchData(SCtransform_counts, vars = significant_genes)

rownames(gene_df) = gene_df$gene
gene_df$gene = NULL

# Assuming gene_df and exp are your dataframes
logfc_df <- gene_df
counts_df <- exp

print(logfc_df)
print(counts_df)

# Check if all column names in counts_df are in the row names of logfc_df
if (!all(colnames(counts_df) %in% rownames(logfc_df))) {
  stop("Not all genes in counts_df are present in logfc_df")
}

# Reorder logfc_df to match the columns in counts_df
logfc_ordered <- logfc_df[colnames(counts_df), , drop = FALSE]
logfc_ordered

logfc = logfc_ordered[[logfc_val]]

# Check for NA values in logfc_ordered
if (any(is.na(logfc))) {
  stop("There are NA values in the logFC values for the genes")
}


# Ensure logfc_ordered$logfc is a numeric vector
logfc_vector <- as.numeric(logfc)
logfc_vector

 gene_averages <- colMeans(counts_df)
 print("Printing gene averages")
 print(gene_averages)
if (method == "multiply") {
  
#THE ACTUAL
result_df <- sweep(counts_df, 2, logfc_vector, FUN = "*")
} else if(method == "log"){
  custom_log_transform <- function(x, logFC_value) {
     x * 2^logFC_value
  }
  result_df <- sweep(counts_df, 2, logfc_vector, FUN = custom_log_transform)
} else{
  stop("Calculation is not working")
}
print(result_df)
# Divide each gene's expression by its average expression
 result_df <- sweep(result_df, 2, gene_averages, FUN = "/")

# Print the result
print("Printing normalized")
print(result_df)


metadata <- SCtransform_counts@meta.data
# Assume that metadata has a column 'donor' that contains donor information
# If not, modify this part according to your metadata structure
if(!"donor_id" %in% colnames(metadata)) {
  stop("Metadata does not contain 'donor' information. Please check your metadata structure.")
}

# Create unique identifiers for cell column to avoid conflicts
expr_data <- result_df %>%
  rownames_to_column(var = "cell_id")

metadata <- metadata %>%
  rownames_to_column(var = "cell_id")

# Combine expression data with donor information
expr_data_long <- expr_data %>%
  pivot_longer(cols = -cell_id, names_to = "gene", values_to = "expression") %>%
  left_join(metadata %>% select(cell_id, donor_id, Condition, cell_class, reclustered_cell_class, reclustered_neuron_type, reclustered_patch_matrix, reclustered_subcluster, reclustered_patch_matrix_exotic, reclustered_neuron_type_joint_cluster, pct_mito, nUmi, region), by = "cell_id")
# Assuming your data frame is named expr_data_long
print(expr_data_long)
# Count the number of occurrences for each gene per donor

summarized_expr_data <- expr_data_long %>%
  group_by(cell_id, donor_id, Condition, cell_class, reclustered_cell_class, reclustered_neuron_type, reclustered_patch_matrix, reclustered_subcluster, reclustered_patch_matrix_exotic, reclustered_neuron_type_joint_cluster, pct_mito, nUmi,region) %>%
  summarize(total_expression = sum(expression), .groups = 'drop')
print(summarized_expr_data)

return(summarized_expr_data)
}

```

#BICAN V17 markers

```{r}
final_V17_BICAN_neurons@meta.data$reclustered_patch_matrix_exotic = final_V17_BICAN_neurons$reclustered_neuron_joint_type
final_V17_BICAN_neurons$reclustered_patch_matrix_exotic[final_V17_BICAN_neurons$reclustered_cell_class == "D1_patch_exotic" | final_V17_BICAN_neurons$reclustered_cell_class == "D2_patch_exotic"] = "SPN_exotic"
table(final_V17_BICAN_neurons$reclustered_patch_matrix_exotic)

qsave(final_V17_BICAN_neurons, "~/SOBJ_USE_THESE/final_V17_BICAN_neurons.qs")
final_V17_BICAN_neurons@meta.data
```


```{r}
Idents(final_V17_BICAN_neurons) = "reclustered_patch_matrix"
DimPlot(final_V17_BICAN_neurons, label = T)
```


```{r}
BICAN_V17_matrix_markers =FindMarkers(object = final_V17_BICAN_neurons, ident.1 = "SPN_matrix", ident.2 = "non-SPN", only.pos = TRUE)
                                #,  min.pct = 0.2, logfc.threshold = 1.25)
BICAN_V17_matrix_markers
```


```{r}
BICAN_V17_matrix_markers_final = subset(BICAN_V17_matrix_markers, subset = p_val_adj < 0.05)
BICAN_V17_matrix_markers_final$pct_subtract = BICAN_V17_matrix_markers_final$pct.1 - BICAN_V17_matrix_markers_final$pct.2
BICAN_V17_matrix_markers_final$weighted_logFC = BICAN_V17_matrix_markers_final$avg_log2FC * BICAN_V17_matrix_markers_final$pct_subtract
BICAN_V17_matrix_markers_final

BICAN_V17_matrix_markers_final$gene = rownames(BICAN_V17_matrix_markers_final)
```

```{r}
gene_df = BICAN_V17_matrix_markers_final
SCtransform_counts = final_merged_xdp_transformed
logfc_val = "weighted_logFC"
method = "multiply"

significant_genes = unique(gene_df$gene)

exp = FetchData(SCtransform_counts, vars = significant_genes)

rownames(gene_df) = gene_df$gene
gene_df$gene = NULL

# Assuming gene_df and exp are your dataframes
logfc_df <- gene_df
counts_df <- exp
counts_df
```

```{r}
#this is the list of genes to us
genes_in_both = colnames(counts_df)

```

```{r}
 Intersected_BICAN_V17_matrix_markers_final = subset(BICAN_V17_matrix_markers_final, subset = gene %in% genes_in_both)
 Intersected_BICAN_V17_matrix_markers_final
```

```{r}
Intersected_BICAN_V17_matrix_markers_final_filtered =subset(Intersected_BICAN_V17_matrix_markers_final, subset = pct_subtract > 0)
Intersected_BICAN_V17_matrix_markers_final_filtered
```



```{r}
Intersected_BICAN_matrix_markers_final_filtered
```

```{r}
write.csv(Intersected_BICAN_matrix_markers_final_filtered, "~/genes.csv")
```

```{r}
Intersected_BICAN_matrix_markers_final_filtered$new_weight = Intersected_BICAN_matrix_markers_final_filtered$avg_log2FC * (Intersected_BICAN_matrix_markers_final_filtered$pct.1/(Intersected_BICAN_matrix_markers_final_filtered$pct.1 + Intersected_BICAN_matrix_markers_final_filtered$pct.2))


Intersected_BICAN_matrix_markers_final_filtered$relative_pct =Intersected_BICAN_matrix_markers_final_filtered$pct.1/(Intersected_BICAN_matrix_markers_final_filtered$pct.1 + Intersected_BICAN_matrix_markers_final_filtered$pct.2)
```

```{r}
ggplot(data = Intersected_BICAN_matrix_markers_final_filtered, aes(x = pct.1, y = pct.2)) + geom_point(alpha = 0.4, color = "red")
ggplot(data = Intersected_BICAN_matrix_markers_final_filtered, aes(x = pct.1, y = pct_subtract)) + geom_point(alpha = 0.4, color = "red")
ggplot(data = Intersected_BICAN_matrix_markers_final_filtered, aes(x = pct.2, y = pct_subtract)) + geom_point(alpha = 0.4, color = "red")
```



```{r}
ggplot(data = Intersected_BICAN_matrix_markers_final_filtered, aes(x = avg_log2FC, y = pct.1)) + geom_point(alpha = 0.4, color = "red") 
ggplot(data = Intersected_BICAN_matrix_markers_final_filtered, aes(x = avg_log2FC, y = pct.2)) + geom_point(alpha = 0.4, color = "red") 
ggplot(data = Intersected_BICAN_matrix_markers_final_filtered, aes(x = avg_log2FC, y = new_weight)) + geom_point(alpha = 0.4, color = "red") 
ggplot(data = Intersected_BICAN_matrix_markers_final_filtered, aes(x = avg_log2FC, y = weighted_logFC)) + geom_point(alpha = 0.4, color = "red") 
ggplot(data = Intersected_BICAN_matrix_markers_final_filtered, aes(x = weighted_logFC, y = new_weight)) + geom_point(alpha = 0.4, color = "red")
ggplot(data = Intersected_BICAN_matrix_markers_final_filtered, aes(x = pct_subtract, y = relative_pct)) + geom_point(alpha = 0.4, color = "red")
```
```{r}
cor.test(Intersected_BICAN_matrix_markers_final_filtered$avg_log2FC, Intersected_BICAN_matrix_markers_final_filtered$weighted_logFC, method = "spearman", exact = FALSE )

cor.test(Intersected_BICAN_matrix_markers_final_filtered$avg_log2FC, Intersected_BICAN_matrix_markers_final_filtered$new_weight, method = "spearman", exact = FALSE )

cor.test(Intersected_BICAN_matrix_markers_final_filtered$weighted_logFC, Intersected_BICAN_matrix_markers_final_filtered$new_weight, method = "spearman", exact = FALSE )
```



```{r}
cortest1 <- cor.test(Intersected_BICAN_matrix_markers_final_filtered$avg_log2FC, Intersected_BICAN_matrix_markers_final_filtered$weighted_logFC, method = "spearman", exact = FALSE )
cortest1

cor1 <- cortest1$estimate
pvalue1 <- cortest1$p.value
subtitl1 <- paste("Spearman correlation: ", sprintf("%0.3f", cor1), " R^2: ", sprintf("%0.3f", cor1^2), " p-value:", sprintf("%0.3f", pvalue1^2))

b <- ggplot(data = Intersected_BICAN_matrix_markers_final_filtered, aes(x =Intersected_BICAN_matrix_markers_final_filtered$avg_log2FC, y = Intersected_BICAN_matrix_markers_final_filtered$weighted_logFC, color = Intersected_BICAN_matrix_markers_final_filtered$donor_id)) + 
  geom_point() + labs(fill = "Condition") + xlab("SPN Score") + ylab("MMR Score") + geom_smooth(method = "lm", se = FALSE, color = "black") + ggtitle("CB: GP RQS vs GP DV200", subtitle = subtitl1) + scale_color_discrete(name = "Sample ID")

b
```

```{r}
final_V8_BICAN_neurons_final@meta.data

table(final_V8_BICAN_neurons_final$reclustered_patch_matrix)
```

```{r}
final_V8_BICAN_neurons_final
```

```{r}

```


```{r}
A <- FetchData(final_V8_BICAN_neurons, vars = c("GPC5", "reclustered_patch_matrix"))
print(A)

mean1 <- mean(A$GPC5[A$reclustered_patch_matrix == "SPN_matrix"])

mean2 <- mean(A$GPC5[A$reclustered_patch_matrix == "non-SPN"])

log2(mean1/mean2)

mean1 <- mean(A$GPC5[A$reclustered_patch_matrix == "SPN_matrix"]) + 1
mean2 <- mean(A$GPC5[A$reclustered_patch_matrix == "non-SPN"]) + 1
log2FC <- log2(mean1 / mean2)
log2FC
```


```{r}
FeaturePlot(final_V8_BICAN_neurons_final, features = c("GPC5"))
```
```{r}
BICAN_matrix_markers =FindMarkers(object = final_V8_BICAN_neurons, ident.1 = "SPN_matrix", ident.2 = "non-SPN", only.pos = TRUE,  min.pct = 0.2, logfc.threshold = 3)
```
```{r}
# Fetch data
A <- FetchData(final_V8_BICAN_neurons_final, vars = c("COCH", "reclustered_patch_matrix"))

# Calculate mean expression for each group with pseudo-count
pseudocount <- 1
mean1 <- mean(exp(A$COCH[A$reclustered_patch_matrix == "SPN_matrix"]) - 1) + pseudocount
mean2 <- mean(exp(A$COCH[A$reclustered_patch_matrix == "non-SPN"]) - 1) + pseudocount

mean1
mean2

# Compute log fold change
log2FC <- log2(mean1 / mean2)
log2FC
```

```{r}
log2(mean1/mean2) * (mean1- mean2)
```

```{r}
Intersected_BICAN_matrix_markers_final_filtered
```

```{r}
filtered_BICAN = subset(Intersected_BICAN_matrix_markers_final_filtered, gene %in% genes_to_keep)
filtered_BICAN
```

```{r}
final_V17_BICAN_neurons = qread("~/SOBJ_USE_THESE/final_V17_BICAN_neurons.qs")
```

```{r}
#BICAN_filtered_test = subset(Intersected_BICAN_matrix_markers_final_filtered, subset = avg_log2FC > 1)
BICAN_filtered_test = subset(Intersected_BICAN_matrix_markers_final_filtered, subset = relative_pct > 0.6)
BICAN_filtered_test
```




```{r}
XDP_matrix_scores = SPN_score_normalized(Intersected_BICAN_V17_matrix_markers_final_filtered, final_merged_xdp_transformed, logfc_val = "weighted_logFC", method = "multiply")

XDP_matrix_scores$Village = "XDP"

V8_BICAN_scores = BICAN_SPN_score_normalized(Intersected_BICAN_V17_matrix_markers_final_filtered, final_V8_BICAN_neurons, logfc_val = "weighted_logFC", method = "multiply")
V8_BICAN_scores$Village = "BICAN_V8"
V8_BICAN_scores

V17_BICAN_scores = BICAN_SPN_score_normalized(Intersected_BICAN_V17_matrix_markers_final_filtered, final_V17_BICAN_neurons, logfc_val = "weighted_logFC", method = "multiply")
V17_BICAN_scores$Village = "BICAN_V17"
V17_BICAN_scores
```

```{r}
 V17_BICAN_scores$reclustered_patch_matrix_exotic = V17_BICAN_scores$reclustered_neuron_joint_type
 V17_BICAN_scores$reclustered_patch_matrix_exotic[V17_BICAN_scores$reclustered_cell_class == "D1_patch_exotic" | V17_BICAN_scores$reclustered_cell_class == "D2_patch_exotic"] = "SPN_exotic"
table(V17_BICAN_scores$reclustered_patch_matrix_exotic)

library(dplyr)
XDP_matrix_scores_df = XDP_matrix_scores %>% select(cell_id, donor_id, Condition, reclustered_patch_matrix_exotic, reclustered_patch_matrix, reclustered_neuron_type_joint_cluster,  pct_mito, nUmi, total_expression , z_score, Village)

BICAN_matrix_scores_df_V8 = V8_BICAN_scores %>% select(cell_id, DONOR, reclustered_patch_matrix, reclustered_neuron_joint_type,  pct_mt, NUM_TRANSCRIPTS, total_expression , z_score, Village) 

BICAN_matrix_scores_df_V17 = V17_BICAN_scores %>% select(cell_id, DONOR, reclustered_patch_matrix, reclustered_neuron_joint_type,reclustered_patch_matrix_exotic,  pct_mt, NUM_TRANSCRIPTS, total_expression , z_score, Village) 

BICAN_matrix_scores_df_V8$Condition = "BICAN Control V8"
BICAN_matrix_scores_df_V8$reclustered_patch_matrix_exotic = BICAN_matrix_scores_df_V8$reclustered_neuron_joint_type
BICAN_matrix_scores_df_V8$reclustered_neuron_type_joint_cluster = BICAN_matrix_scores_df_V8$reclustered_neuron_joint_type
BICAN_matrix_scores_df_V8$reclustered_neuron_joint_type = NULL
BICAN_matrix_scores_df_V8$pct_mito = BICAN_matrix_scores_df_V8$pct_mt 
BICAN_matrix_scores_df_V8$pct_mt  = NULL
BICAN_matrix_scores_df_V8$nUmi = BICAN_matrix_scores_df_V8$NUM_TRANSCRIPTS 
BICAN_matrix_scores_df_V8$NUM_TRANSCRIPTS  = NULL
BICAN_matrix_scores_df_V8$donor_id = BICAN_matrix_scores_df_V8$DONOR 
BICAN_matrix_scores_df_V8$DONOR  = NULL

BICAN_matrix_scores_df_V17$Condition = "BICAN Control V17"
BICAN_matrix_scores_df_V17$reclustered_neuron_type_joint_cluster = BICAN_matrix_scores_df_V17$reclustered_neuron_joint_type
BICAN_matrix_scores_df_V17$reclustered_neuron_joint_type = NULL
BICAN_matrix_scores_df_V17$pct_mito = BICAN_matrix_scores_df_V17$pct_mt 
BICAN_matrix_scores_df_V17$pct_mt  = NULL
BICAN_matrix_scores_df_V17$nUmi = BICAN_matrix_scores_df_V17$NUM_TRANSCRIPTS 
BICAN_matrix_scores_df_V17$NUM_TRANSCRIPTS  = NULL
BICAN_matrix_scores_df_V17$donor_id = BICAN_matrix_scores_df_V17$DONOR 
BICAN_matrix_scores_df_V17$DONOR  = NULL

XDP_matrix_scores_df
BICAN_matrix_scores_df_V8
BICAN_matrix_scores_df_V17

Combined_scores_df = rbind(XDP_matrix_scores_df, BICAN_matrix_scores_df_V8, BICAN_matrix_scores_df_V17)
Combined_scores_df

table(Combined_scores_df$Condition, Combined_scores_df$reclustered_patch_matrix_exotic)
table(Combined_scores_df$Condition, Combined_scores_df$reclustered_neuron_type_joint_cluster)

min(Combined_scores_df$total_expression)
max(Combined_scores_df$total_expression)
```

```{r}
matrix <- Combined_scores_df[Combined_scores_df$reclustered_patch_matrix_exotic == "SPN_matrix", ]
  plot_overlapping_density_histogram(df = matrix, 
                                          hist_col = matrix$total_expression,
                                          fill_col = "Condition",
                                          colors = c("XDP" = "red", "Control" = "blue", "BICAN Control V8" = "green", "BICAN Control V17" = "orange"),
                                          breaks = seq(0, 120, 2),
                                          title = "XDP vs Control: SPN matrix",
                                          xlab = "SPN Matrix score",
                                          fig_filename = NULL)
  patch <- Combined_scores_df[Combined_scores_df$reclustered_patch_matrix_exotic == "SPN_patch", ]
  plot_overlapping_density_histogram(df = patch, 
                                          hist_col = patch$total_expression,
                                          fill_col = "Condition",
                                          colors = c("XDP" = "red", "Control" = "blue", "BICAN Control V8" = "green", "BICAN Control V17" = "orange"),
                                          breaks = seq(0, 60, 1),
                                          title = "XDP vs Control: SPN patch",
                                          xlab = "SPN Matrix score",
                                          fig_filename = NULL)
  nonSPN <- Combined_scores_df[Combined_scores_df$reclustered_patch_matrix_exotic == "non-SPN", ]
  plot_overlapping_density_histogram(df = nonSPN, 
                                          hist_col = nonSPN$total_expression,
                                          fill_col = "Condition",
                                          colors = c("XDP" = "red", "Control" = "blue","BICAN Control V8" = "green", "BICAN Control V17" = "orange"),
                                          breaks = seq(0, 60, 1),
                                          title = "XDP vs Control: non-SPN",
                                          xlab = "SPN Matrix score",
                                          fig_filename = NULL)
    exotic <- Combined_scores_df[Combined_scores_df$reclustered_patch_matrix_exotic == "SPN_exotic", ]
  plot_overlapping_density_histogram(df = exotic, 
                                          hist_col = exotic$total_expression,
                                          fill_col = "Condition",
                                          colors = c("XDP" = "red", "Control" = "blue","BICAN Control V8" = "green", "BICAN Control V17" = "orange"),
                                          breaks = seq(0, 60, 1),
                                          title = "XDP vs Control: Exotic SPN",
                                          xlab = "SPN Matrix score",
                                          fig_filename = NULL)
    eSPN <- Combined_scores_df[Combined_scores_df$reclustered_patch_matrix_exotic == "eSPN", ]
  plot_overlapping_density_histogram(df = eSPN, 
                                          hist_col = eSPN$total_expression,
                                          fill_col = "Condition",
                                          colors = c("XDP" = "red", "Control" = "blue","BICAN Control V8" = "green", "BICAN Control V17" = "orange"),
                                          breaks = seq(0, 60, 1),
                                          title = "XDP vs Control: eSPN",
                                          xlab = "SPN Matrix score",
                                          fig_filename = NULL)
```

```{r}
XDP_weighted = Combined_scores_df
XDP_weighted
```
```{r}
XDP_weighted
Combined_scores_df$total_expression1 = Combined_scores_df$total_expression
Combined_scores_df
```
```{r}
new_weights = Combined_scores_df %>% select(cell_id, total_expression1)
new_weights

combo = merge(XDP_weighted, new_weights, by = "cell_id")
combo
```
```{r}
classes = unique(combo$reclustered_patch_matrix_exotic)

 for (class in classes) {
   test = subset(combo, subset = reclustered_patch_matrix_exotic == class)
   
   
cortest1 <- cor.test(test$total_expression, test$total_expression1, method = "spearman", exact = FALSE )
print(cortest1)

cor1 <- cortest1$estimate
pvalue1 <- cortest1$p.value
subtitl1 <- paste("Spearman correlation: ", sprintf("%0.3f", cor1), " R^2: ", sprintf("%0.3f", cor1^2), " p-value:", sprintf("%0.3f", pvalue1^2))

   
   
   
  a = ggplot(test, aes(x = total_expression, y = total_expression1, color = Condition)) + geom_point(alpha = 0.4) + xlab("Weighted_logFC") +ylab("New_weighted_logFC") +  scale_color_manual(values = c("XDP" = "red", 
                                "Control" = "blue", 
                                "BICAN Control V8" = "green", 
                                "BICAN Control V17" = "orange")) + geom_smooth(method = "lm", se = FALSE, color = "black")+ ggtitle(paste(class), subtitle = subtitl1)
  print(a)
 }

```




```{r}
SCtransform_counts@assays$SCT
SCtransform_counts@assays$RNA
```



#filter
```{r}
gene_df = Intersected_BICAN_matrix_markers_final_filtered
SCtransform_counts = final_V8_BICAN_neurons_final


significant_genes = unique(gene_df$gene)

exp = FetchData(SCtransform_counts, vars = significant_genes)
exp
```


```{r}
rownames(gene_df) = gene_df$gene
gene_df$gene = NULL

# Assuming gene_df and exp are your dataframes
logfc_df <- gene_df
counts_df <- exp

print(logfc_df)
print(counts_df)
```


```{r}
# Assuming 'data' is your expression matrix with genes as rows and cells as columns
avg_expression <- as.data.frame(colMeans(counts_df))
avg_expression 

avg_expression <- colMeans(counts_df)

# Filter out the lowest 1% of genes by average expression
threshold_avg <- quantile(avg_expression, 0.01)
threshold_avg

avg_expression_keep = avg_expression > threshold
avg_expression_keep = as.data.frame(avg_expression_keep)
avg_expression_keep

avg_expression_keep = subset(avg_expression_keep, avg_expression_keep == TRUE)
avg_expression_keep

genes_to_keep <- rownames(avg_expression_keep)
genes_to_keep

```


#remove genes with < 10 counts overall (i cant), remove genes expressed in < 1% of cells)

```{r}
counts_df
```



```{r}
# Calculate the number of cells expressing each gene
expressing_cells <- rowSums(counts_df > 0)

# Calculate the percentage of cells expressing each gene
percent_expressed <- expressing_cells / ncol(counts_df)

# Filter out genes expressed in < 1% of cells
filtered_data <- counts_df[percent_expressed >= 0.01, ]
filtered_data
```



```{r}
merged_xdp_transformed_controls = subset(final_merged_xdp_transformed, subset = Condition == "Control")
merged_xdp_transformed_controls
```


```{r}
Idents(final_merged_xdp_transformed_controls) = "reclustered_neuron_type_joint_cluster"
DimPlot(final_merged_xdp_transformed_controls, label = TRUE)

xdp_matrix_markers =FindMarkers(object = final_merged_xdp_transformed_controls, ident.1 = "SPN_matrix", ident.2 = "non-SPN", only.pos = TRUE)
                                #,  min.pct = 0.2, logfc.threshold = 1.25)
xdp_matrix_markers

xdp_matrix_markers_final = subset(xdp_matrix_markers, subset = p_val_adj < 0.05)
xdp_matrix_markers_final$pct_subtract = xdp_matrix_markers_final$pct.1 - xdp_matrix_markers_final$pct.2
xdp_matrix_markers_final$weighted_logFC = xdp_matrix_markers_final$avg_log2FC * xdp_matrix_markers_final$pct_subtract
xdp_matrix_markers_final

xdp_matrix_markers_final$gene = rownames(xdp_matrix_markers_final)
```

```{r}
gene_df = xdp_matrix_markers_final
SCtransform_counts = final_V8_BICAN_neurons

significant_genes = unique(gene_df$gene)

exp = FetchData(SCtransform_counts, vars = significant_genes)
exp
```


```{r}
rownames(gene_df) = gene_df$gene
gene_df$gene = NULL

# Assuming gene_df and exp are your dataframes
logfc_df <- gene_df
counts_df <- exp

print(logfc_df)
print(counts_df)
```

```{r}
genes_in_both = colnames(counts_df)
genes_in_both
```


```{r}
Intersected_xdp_matrix_markers_final = subset(xdp_matrix_markers_final, subset = gene %in% genes_in_both)
Intersected_xdp_matrix_markers_final$new_weight = Intersected_xdp_matrix_markers_final$avg_log2FC * (Intersected_xdp_matrix_markers_final$pct.1/(Intersected_xdp_matrix_markers_final$pct.1 + Intersected_xdp_matrix_markers_final$pct.2))
Intersected_xdp_matrix_markers_final
```
```{r}
Intersected_xdp_matrix_markers_final_filtered = subset(Intersected_xdp_matrix_markers_final, subset = pct_subtract > 0)
Intersected_xdp_matrix_markers_final_filtered
```


















```{r}
BICAN_matrix_markers =FindMarkers(object = final_V8_BICAN_neurons, ident.1 = "SPN_matrix", ident.2 = "non-SPN", only.pos = TRUE)
                                #,  min.pct = 0.2, logfc.threshold = 1.25)
BICAN_matrix_markers
```


```{r}
BICAN_matrix_markers_final = subset(BICAN_matrix_markers, subset = p_val_adj < 0.05)
BICAN_matrix_markers_final$pct_subtract = BICAN_matrix_markers_final$pct.1 - BICAN_matrix_markers_final$pct.2
BICAN_matrix_markers_final$weighted_logFC = BICAN_matrix_markers_final$avg_log2FC * BICAN_matrix_markers_final$pct_subtract
BICAN_matrix_markers_final

BICAN_matrix_markers_final$gene = rownames(BICAN_matrix_markers_final)
```




```{r}
# Assuming 'data' is your expression matrix with genes as rows and cells as columns
avg_expression <- rowMeans(data)

# Calculate the threshold for the lowest 1%
threshold <- quantile(avg_expression, 0.01)

# Filter out genes below the threshold
filtered_data <- data[avg_expression > threshold, ]

```


