---
title: "R Notebook"
output: html_notebook
---

```{r}
library(qs)
library(Seurat)
library(ggplot2)
library(dplyr)
library(tidyr)
```

```{r}
BICAN_3P_V8_CaH_neurons_filtered =  qread("SPN_Scores/BICAN_3P_V8_CaH_neurons_filtered.qs")
```

```{r}
BICAN_3P_V8_CaH_neurons_filtered@meta.data

Idents(BICAN_3P_V8_CaH_neurons_filtered) <- "subcluster"
DimPlot(BICAN_3P_V8_CaH_neurons_filtered, label = TRUE)

Idents(BICAN_3P_V8_CaH_neurons_filtered) <- "sub_class"
DimPlot(BICAN_3P_V8_CaH_neurons_filtered, label = TRUE)

Idents(BICAN_3P_V8_CaH_neurons_filtered) <- "SPN_class"
DimPlot(BICAN_3P_V8_CaH_neurons_filtered, label = TRUE)

Idents(BICAN_3P_V8_CaH_neurons_filtered) <- "D1_D2"
DimPlot(BICAN_3P_V8_CaH_neurons_filtered, label = TRUE)

Idents(BICAN_3P_V8_CaH_neurons_filtered) <- "patch_matrix"
DimPlot(BICAN_3P_V8_CaH_neurons_filtered, label = TRUE)
```

```{r}
BICAN_D1_D2_vs_non_markers =  qread("SPN_Scores/BICAN_D1_D2_vs_non_markers")
BICAN_patch_matrix_vs_non_markers =  qread("SPN_Scores/BICAN_patch_matrix_vs_non_markers")
BICAN_SPN_vs_non_markers= qread("SPN_Scores/BICAN_SPN_vs_non_markers.qs")
BICAN_SPN_classes_vs_non_markers =  qread("SPN_Scores/BICAN_SPN_classes_vs_non_markers.qs")


BICAN_SPN_vs_non_markers #SPN vs nonSPN
BICAN_SPN_classes_vs_non_markers #D1 matrix, D1 patch, D2 matrix, D2 patch vs nonSPN
BICAN_D1_D2_vs_non_markers #D1, D2 vs non SPNS
BICAN_patch_matrix_vs_non_markers# patch, matrix vs nonSPNs
```

```{r}
BICAN_SPN_vs_non_markers_filtered = subset(BICAN_SPN_vs_non_markers, subset = pct.1 >0.7 & pct.2 <0.2 & cluster != "non-SPN")
BICAN_SPN_classes_vs_non_markers_filtered= subset(BICAN_SPN_classes_vs_non_markers, subset = pct.1 >0.7 & pct.2 <0.2 & cluster != "non-SPN")
BICAN_D1_D2_vs_non_markers_filtered = subset(BICAN_D1_D2_vs_non_markers, subset = pct.1 >0.7 & pct.2 <0.2 & cluster != "non-SPN")
BICAN_patch_matrix_vs_non_markers_filtered= subset(BICAN_patch_matrix_vs_non_markers, subset = pct.1 >0.7 & pct.2 <0.2 & cluster != "non-SPN")

BICAN_SPN_vs_non_markers_filtered
BICAN_SPN_classes_vs_non_markers_filtered
BICAN_D1_D2_vs_non_markers_filtered
BICAN_patch_matrix_vs_non_markers_filtered
```

```{r}
D1_matrix_markers= qread("SPN_Scores/D1_matrix_markers.qs")
D1_patch_markers= qread("SPN_Scores/D1_patch_markers.qs")
D2_matrix_markers= qread("SPN_Scores/D2_matrix_markers.qs")
D2_patch_markers= qread("SPN_Scores/D2_patch_markers.qs") 

D1_markers= qread("SPN_Scores/D1_markers.qs")
D2_markers =qread("SPN_Scores/D2_markers.qs")
patch_markers= qread("SPN_Scores/patch_markers.qs")
matrix_markers= qread("SPN_Scores/matrix_markers.qs")

D1_matrix_markers
D1_patch_markers
D2_matrix_markers
D2_patch_markers

D1_markers
D2_markers
patch_markers
matrix_markers
```

```{r}
D1_matrix_markers_filtered = subset(D1_matrix_markers, subset = pct.1 >0.7 & pct.2 <0.2)
D1_patch_markers_filtered= subset(D1_patch_markers, subset = pct.1 >0.7 & pct.2 <0.2)
D2_matrix_markers_filtered= subset(D2_matrix_markers, subset = pct.1 >0.7 & pct.2 <0.2)
D2_patch_markers_filtered= subset(D2_patch_markers, subset = pct.1 >0.7 & pct.2 <0.2)

D1_markers_filtered= subset(D1_markers, subset = pct.1 >0.7 & pct.2 <0.2)
D2_markers_filtered= subset(D2_markers, subset = pct.1 >0.7 & pct.2 <0.2)
patch_markers_filtered= subset(patch_markers, subset = pct.1 >0.7 & pct.2 <0.2)
matrix_markers_filtered= subset(matrix_markers, subset = pct.1 >0.7 & pct.2 <0.2)

D1_matrix_markers_filtered
D1_patch_markers_filtered
D2_matrix_markers_filtered
D2_patch_markers_filtered

D1_markers_filtered
D2_markers_filtered
patch_markers_filtered
matrix_markers_filtered
```


#SPN SCORING

```{r}
caudate_neurons = qread("Current_subclusters/caudate_neuron.qs")
putamen_neurons= qread("Current_subclusters/putamen_neuron.qs")

caudate_neurons
putamen_neurons
```


```{r}
caudate_neurons@meta.data$sub_class = caudate_neurons@meta.data$cell_class
caudate_neurons$sub_class[caudate_neurons$sub_class == "D1_SPN_1"] = "SPN"
caudate_neurons$sub_class[caudate_neurons$sub_class == "interneuron_2"] = "interneuron"
caudate_neurons$sub_class[caudate_neurons$sub_class == "D1_SPN_matrix_1"] = "SPN"
caudate_neurons$sub_class[caudate_neurons$sub_class == "D2_SPN_matrix_1"] = "SPN"
caudate_neurons$sub_class[caudate_neurons$sub_class == "SPN"] = "SPN"
caudate_neurons$sub_class[caudate_neurons$sub_class == "D1_SPN_2"] = "SPN"
caudate_neurons$sub_class[caudate_neurons$sub_class == "D2_SPN_1"] = "SPN"
caudate_neurons$sub_class[caudate_neurons$sub_class == "interneuron_3"] = "interneuron"
caudate_neurons$sub_class[caudate_neurons$sub_class == "D1_SPN_matrix_2"] = "SPN"
caudate_neurons$sub_class[caudate_neurons$sub_class == "D1_SPN_patch"] = "SPN"
caudate_neurons$sub_class[caudate_neurons$sub_class == "interneuron_1"] = "interneuron"
caudate_neurons$sub_class[caudate_neurons$sub_class == "eSPN"] = "eSPN"
caudate_neurons$sub_class[caudate_neurons$sub_class == "interneuron_5"] = "interneuron"
caudate_neurons$sub_class[caudate_neurons$sub_class == "cholinergic"] = "cholinergic"
caudate_neurons$sub_class[caudate_neurons$sub_class == "interneuron_4"] = "interneuron"
caudate_neurons$sub_class[caudate_neurons$sub_class == "interneuron_6"] = "interneuron"

caudate_neurons@meta.data

putamen_neurons@meta.data$sub_class = putamen_neurons@meta.data$cell_class
putamen_neurons$sub_class[putamen_neurons$sub_class == "D2_SPN_matrix"] = "SPN"
putamen_neurons$sub_class[putamen_neurons$sub_class == "interneuron_2"] = "interneuron"
putamen_neurons$sub_class[putamen_neurons$sub_class == "interneuron_4"] = "interneuron"
putamen_neurons$sub_class[putamen_neurons$sub_class == "D1_SPN_matrix_1"] = "SPN"
putamen_neurons$sub_class[putamen_neurons$sub_class == "glutamatergic"] = "glutamatergic"
putamen_neurons$sub_class[putamen_neurons$sub_class == "interneuron_5"] = "interneuron"
putamen_neurons$sub_class[putamen_neurons$sub_class == "D1_SPN_1"] = "SPN"
putamen_neurons$sub_class[putamen_neurons$sub_class == "D1_SPN_patch"] = "SPN"
putamen_neurons$sub_class[putamen_neurons$sub_class == "D2_SPN_1"] = "SPN"
putamen_neurons$sub_class[putamen_neurons$sub_class == "eSPN"] = "eSPN"
putamen_neurons$sub_class[putamen_neurons$sub_class == "interneuron_3"] = "interneuron"
putamen_neurons$sub_class[putamen_neurons$sub_class == "interneuron_1"] = "interneuron"
putamen_neurons$sub_class[putamen_neurons$sub_class == "D2_SPN_2"] = "SPN"
putamen_neurons$sub_class[putamen_neurons$sub_class == "cholinergic"] = "cholinergic"
putamen_neurons$sub_class[putamen_neurons$sub_class == "interneuron_7"] = "interneuron"
putamen_neurons$sub_class[putamen_neurons$sub_class == "interneuron_6"] = "interneuron"

putamen_neurons@meta.data
```



#SC Transform SPNs and other neurons- did SCtransform on all neurons first before I subsetted
```{r}
library(sctransform)
caudate_neurons_transformed = SCTransform(caudate_neurons, vars.to.regress = "pct_mito", verbose = FALSE)
DefaultAssay(caudate_neurons_transformed) = "SCT"


putamen_neurons_transformed = SCTransform(putamen_neurons, vars.to.regress = "pct_mito", verbose = FALSE)
DefaultAssay(putamen_neurons_transformed) = "SCT"

SPN_put_transformed= subset(putamen_neurons_transformed, subset = sub_class == "SPN")
non_SPN_put_transformed= subset(putamen_neurons_transformed, subset = sub_class != "SPN")

SPN_put_transformed
non_SPN_put_transformed
```

#SPN-ness score
#BICAN- SPN vs nonSPNs
```{r}
caudate_neurons_transformed@meta.data$subcluster = caudate_neurons_transformed@meta.data$sub_class
caudate_neurons_transformed$subcluster[caudate_neurons_transformed$subcluster == "interneuron"] = "non-SPN"
caudate_neurons_transformed$subcluster[caudate_neurons_transformed$subcluster == "eSPN"] = "non-SPN"
caudate_neurons_transformed$subcluster[caudate_neurons_transformed$subcluster == "cholinergic"] = "non-SPN"

caudate_neurons_transformed@meta.data

SPN_cah_transformed= subset(caudate_neurons_transformed, subset = sub_class == "SPN")
non_SPN_cah_transformed= subset(caudate_neurons_transformed, subset = sub_class != "SPN")
SPN_cah_transformed
non_SPN_cah_transformed

```

```{r}
qsave(caudate_neurons_transformed, "SPN_Scores/caudate_neurons_transformed.qs")
qsave(putamen_neurons_transformed, "SPN_Scores/putamen_neurons_transformed.qs")
```


```{r}
BICAN_SPN_vs_non_markers_filtered
```


```{r}
features = BICAN_SPN_vs_non_markers_filtered$gene
Idents(caudate_neurons_transformed) = "subcluster"
DotPlot(caudate_neurons_transformed, features = features, dot.scale = 8) + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


```{r}
CaH_SPN_summary_expr_data = SPN_score(BICAN_SPN_vs_non_markers_filtered, SPN_cah_transformed)

CaH_non_SPN_summary_expr_data = SPN_score(BICAN_SPN_vs_non_markers_filtered, non_SPN_cah_transformed)
```

```{r}
CaH_SPN_summary_expr_data
CaH_non_SPN_summary_expr_data
```

```{r}
CaH_SPN_summary_expr_data$cell_type = "SPN"
CaH_non_SPN_summary_expr_data$cell_type = "non-SPN" 

head(CaH_SPN_summary_expr_data)
head(CaH_non_SPN_summary_expr_data)

CaH_Transcriptional_DE = rbind(CaH_SPN_summary_expr_data, CaH_non_SPN_summary_expr_data)
CaH_Transcriptional_DE
```

```{r}
min(CaH_Transcriptional_DE$total_expression)
max(CaH_Transcriptional_DE$total_expression)
median(CaH_Transcriptional_DE$total_expression)

```

```{r}
library(gridExtra)
options(repr.plot.width=24, repr.plot.height=16)
plots = list()
donors = c("PCMC-16-011", "PCMC-16-012", "SCF-18-003", "SCF-18-004", "SCF-18-006","SCF-19-009", "SCF-19-014", "SCF-19-018",  "SCF-20-023",  "SCF_20-024", "SCF-20-025", "SCF-21-030", "SCF_22-043", "SCF_21-037CM2","SCF-22-054CM","SCF-22-058CF","SCF-23-068CM")
for (donor in donors) {
test = CaH_Transcriptional_DE[CaH_Transcriptional_DE$donor_id == donor,]
plots[[donor]] = plot_overlapping_density_histogram(df = test,
hist_col = "total_expression",
fill_col = "cell_type",
colors = c("red", "blue"),
#alpha=0.5,
breaks=seq(0,100,2),
title= paste("CaH SPNs vs non-SPNs: ", donor),
xlab = "CaH SPN Score",
fig_filename = NULL)
}
layout_matrix <- rbind(
c(1, 2, 3, 4, 5),
c(6, 7, 8, 9, 10),
c(11, 12, 13, NA, NA),
c(14, 15, 16, 17, NA)
)
# Arrange the plots according to the custom layout
grid_plots <- grid.arrange(grobs = plots, layout_matrix = layout_matrix)
# Save the arranged plots to a PNG file
ggsave(filename = "SPN_SCORE_Output/CaH_BICAN_SPNvnonSPN_donors.png", plot = grid_plots, width = 30, height = 16)

```



```{r}
CaH_Transcriptional_DE$Condition = ifelse(grepl("SCF_21-037CM2|SCF-23-068CM|SCF-22-058CF|SCF-22-054CM", CaH_Transcriptional_DE$donor_id), "Control", "XDP")
CaH_Transcriptional_DE

XDP = CaH_Transcriptional_DE[CaH_Transcriptional_DE$Condition == "XDP",]
plot_overlapping_density_histogram(df = XDP, 
    hist_col = "total_expression",
    fill_col = "cell_type",
    colors = c("blue", "red"),
    #alpha=0.5,
    breaks=seq(0,100,2),
    title= paste("CaH SPNs vs non-SPNs: XDP"),
    xlab = "CaH SPN score",
    fig_filename = NULL)

Control = CaH_Transcriptional_DE[CaH_Transcriptional_DE$Condition == "Control",]
plot_overlapping_density_histogram(df = Control, 
    hist_col = "total_expression",
    fill_col = "cell_type",
    colors = c("blue", "red"),
    #alpha=0.5,
    breaks=seq(0,100,2),
    title= paste("CaH SPNs vs non-SPNs: Control "),
    xlab = "CaH SPN score",
    fig_filename = NULL)
```

```{r}
metadata = caudate_neurons_transformed@meta.data
metadata$cell_id = rownames(metadata)
SPN_Score = CaH_Transcriptional_DE %>% select(cell_id, total_expression)
metadata = metadata %>% left_join(SPN_Score, by = "cell_id")
caudate_neurons_transformed@meta.data = metadata
caudate_neurons_transformed@meta.data
```

#GO BACK TO THIS- showing total expression on a umap
```{r}
FeaturePlot(caudate_neurons_transformed, features = c("total_expression"))
```


```{r}
# Check if the feature exists and what values it contains
summary(caudate_neurons_transformed@assays$SCT@data["total_expression", ])
```







`




#SPN-ness score
#BICAN- SPN vs nonSPNs
```{r}
putamen_neurons_transformed@meta.data$subcluster = putamen_neurons_transformed@meta.data$sub_class
putamen_neurons_transformed$subcluster[putamen_neurons_transformed$subcluster == "interneuron"] = "non-SPN"
putamen_neurons_transformed$subcluster[putamen_neurons_transformed$subcluster == "eSPN"] = "non-SPN"
putamen_neurons_transformed$subcluster[putamen_neurons_transformed$subcluster == "cholinergic"] = "non-SPN"
putamen_neurons_transformed$subcluster[putamen_neurons_transformed$subcluster == "glutamatergic"] = "non-SPN"

putamen_neurons_transformed@meta.data

SPN_put_transformed= subset(putamen_neurons_transformed, subset = sub_class == "SPN")
non_SPN_put_transformed= subset(putamen_neurons_transformed, subset = sub_class != "SPN")
SPN_put_transformed
non_SPN_put_transformed

```

```{r}
BICAN_SPN_vs_non_markers_filtered
```


```{r}
features = BICAN_SPN_vs_non_markers_filtered$gene
Idents(putamen_neurons_transformed) = "subcluster"
DotPlot(putamen_neurons_transformed, features = features, dot.scale = 8) + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


```{r}
Put_SPN_summary_expr_data = SPN_score(BICAN_SPN_vs_non_markers_filtered, SPN_put_transformed)

Put_non_SPN_summary_expr_data = SPN_score(BICAN_SPN_vs_non_markers_filtered, non_SPN_put_transformed)
```

```{r}
Put_SPN_summary_expr_data
Put_non_SPN_summary_expr_data
```

```{r}
Put_SPN_summary_expr_data$cell_type = "SPN"
Put_non_SPN_summary_expr_data$cell_type = "non-SPN" 

head(Put_SPN_summary_expr_data)
head(Put_non_SPN_summary_expr_data)

Put_Transcriptional_DE = rbind(Put_SPN_summary_expr_data, Put_non_SPN_summary_expr_data)
Put_Transcriptional_DE
```

```{r}
min(Put_Transcriptional_DE$total_expression)
max(Put_Transcriptional_DE$total_expression)
median(Put_Transcriptional_DE$total_expression)

```

```{r}
library(gridExtra)
options(repr.plot.width=24, repr.plot.height=16)
plots = list()
donors = c("PCMC-16-011", "PCMC-16-012", "SCF-18-003", "SCF-18-004", "SCF-18-006","SCF-19-009", "SCF-19-014", "SCF-19-018",  "SCF-20-023",  "SCF_20-024", "SCF-20-025", "SCF-21-030", "SCF_22-043", "SCF_21-037CM2","SCF-22-054CM","SCF-22-058CF","SCF-23-068CM")
for (donor in donors) {
test = Put_Transcriptional_DE[Put_Transcriptional_DE$donor_id == donor,]
plots[[donor]] = plot_overlapping_density_histogram(df = test,
hist_col = "total_expression",
fill_col = "cell_type",
colors = c("red", "blue"),
#alpha=0.5,
breaks=seq(0,100,2),
title= paste("Put SPNs vs non-SPNs: ", donor),
xlab = "Put SPN Score",
fig_filename = NULL)
}
layout_matrix <- rbind(
c(1, 2, 3, 4, 5),
c(6, 7, 8, 9, 10),
c(11, 12, 13, NA, NA),
c(14, 15, 16, 17, NA)
)
# Arrange the plots according to the custom layout
grid_plots <- grid.arrange(grobs = plots, layout_matrix = layout_matrix)
# Save the arranged plots to a PNG file
ggsave(filename = "SPN_SCORE_Output/Put_BICAN_SPNvnonSPN_donors.png", plot = grid_plots, width = 30, height = 16)

```



```{r}
Put_Transcriptional_DE$Condition = ifelse(grepl("SCF_21-037CM2|SCF-23-068CM|SCF-22-058CF|SCF-22-054CM", Put_Transcriptional_DE$donor_id), "Control", "XDP")
Put_Transcriptional_DE

XDP = Put_Transcriptional_DE[Put_Transcriptional_DE$Condition == "XDP",]
plot_overlapping_density_histogram(df = XDP, 
    hist_col = "total_expression",
    fill_col = "cell_type",
    colors = c("blue", "red"),
    #alpha=0.5,
    breaks=seq(0,100,2),
    title= paste("Put SPNs vs non-SPNs: XDP"),
    xlab = "Put SPN score",
    fig_filename = NULL)

Control = Put_Transcriptional_DE[Put_Transcriptional_DE$Condition == "Control",]
plot_overlapping_density_histogram(df = Control, 
    hist_col = "total_expression",
    fill_col = "cell_type",
    colors = c("blue", "red"),
    #alpha=0.5,
    breaks=seq(0,100,2),
    title= paste("Put SPNs vs non-SPNs: Control "),
    xlab = "Put SPN score",
    fig_filename = NULL)
```





































```{r}
BICAN_3P_V8_CaH_neurons_filtered@meta.data
```


#SC Transform SPNs and other neurons- did SCtransform on all neurons first before I subsetted
```{r}
BICAN_3P_V8_CaH_neurons_filtered_transformed = SCTransform(BICAN_3P_V8_CaH_neurons_filtered, vars.to.regress = "pct_mt", verbose = FALSE)
DefaultAssay(BICAN_3P_V8_CaH_neurons_filtered_transformed) = "SCT"

BICAN_SPN_cah_transformed= subset(BICAN_3P_V8_CaH_neurons_filtered_transformed, subset = SPN_class == "SPN")
BICAN_non_SPN_cah_transformed= subset(BICAN_3P_V8_CaH_neurons_filtered_transformed, subset = SPN_class != "SPN")
BICAN_SPN_cah_transformed
BICAN_non_SPN_cah_transformed

BICAN_SPN_cah_transformed$donor_id = BICAN_SPN_cah_transformed$DONOR
BICAN_non_SPN_cah_transformed$donor_id = BICAN_non_SPN_cah_transformed$DONOR
```



```{r}
BICAN_CaH_SPN_summary_expr_data = SPN_score(BICAN_SPN_vs_non_markers_filtered, BICAN_SPN_cah_transformed)

BICAN_CaH_non_SPN_summary_expr_data = SPN_score(BICAN_SPN_vs_non_markers_filtered, BICAN_non_SPN_cah_transformed)
```

```{r}
BICAN_CaH_SPN_summary_expr_data
BICAN_CaH_non_SPN_summary_expr_data
```

```{r}
BICAN_CaH_SPN_summary_expr_data$cell_type = "SPN"
BICAN_CaH_non_SPN_summary_expr_data$cell_type = "non-SPN" 

head(BICAN_CaH_SPN_summary_expr_data)
head(BICAN_CaH_non_SPN_summary_expr_data)

CaH_Transcriptional_DE = rbind(BICAN_CaH_SPN_summary_expr_data, BICAN_CaH_non_SPN_summary_expr_data)
CaH_Transcriptional_DE
```

```{r}
min(CaH_Transcriptional_DE$total_expression)
max(CaH_Transcriptional_DE$total_expression)
median(CaH_Transcriptional_DE$total_expression)
```


```{r}
library(gridExtra)
options(repr.plot.width=24, repr.plot.height=16)
plots = list()
donors = unique(CaH_Transcriptional_DE$donor_id)
for (donor in donors) {
test = CaH_Transcriptional_DE[CaH_Transcriptional_DE$donor_id == donor,]
plots[[donor]] = plot_overlapping_density_histogram(df = test,
hist_col = "total_expression",
fill_col = "cell_type",
colors = c("red", "blue"),
#alpha=0.5,
breaks=seq(0,180,2),
title= paste("BICAN CaH SPNs vs non-SPNs: ", donor),
xlab = "CaH SPN Score",
fig_filename = NULL)
}
layout_matrix <- rbind(
c(1, 2, 3, 4, 5),
c(6, 7, 8, 9, 10),
c(11, 12, 13, 14, 15),
c(16, 17, 18, 19, NA)
)
# Arrange the plots according to the custom layout
grid_plots <- grid.arrange(grobs = plots, layout_matrix = layout_matrix)
# Save the arranged plots to a PNG file
ggsave(filename = "SPN_SCORE_Output/BICAN_CONTROLS_CaH_BICAN_SPNvnonSPN_donors.png", plot = grid_plots, width = 30, height = 16)

```



```{r}
plot_overlapping_density_histogram(df = CaH_Transcriptional_DE, 
    hist_col = "total_expression",
    fill_col = "cell_type",
    colors = c("blue", "red"),
    #alpha=0.5,
    breaks=seq(0,180,2),
    title= paste("CaH SPNs vs non-SPNs: BICAN Controls "),
    xlab = "CaH SPN score",
    fig_filename = NULL)
```




```{r}
BICAN_3P_V8_CaH_neurons_filtered
```







#PUTAMEN SPN-ness score
#Findallmarkers on controls: SPN vs non-SPNs 
```{r}
putamen_neurons$sub_class[putamen_neurons$sub_class == "interneuron"] = "non-SPN"
putamen_neurons$sub_class[putamen_neurons$sub_class == "eSPN"] = "non-SPN"
putamen_neurons$sub_class[putamen_neurons$sub_class == "cholinergic"] = "non-SPN"
putamen_neurons$sub_class[putamen_neurons$sub_class == "glutamatergic"] = "non-SPN"

Idents(putamen_neurons) = "sub_class"
DimPlot(putamen_neurons, label = TRUE)
```

```{r}
putamen_controls = subset(putamen_neurons, subset = Condition == "Control")
putamen_controls
Idents(putamen_controls) = "sub_class"
DimPlot(putamen_controls, label = TRUE)
```
```{r}
Idents(putamen_controls) = putamen_controls$sub_class
putamen_controls_markers = FindAllMarkers(putamen_controls)
                                          #, only.pos = TRUE,  min.pct = 0.2, logfc.threshold = 1.25)
putamen_controls_markers
```


```{r}
subsetmarkers= subset(putamen_controls_markers, subset = pct.2 < 0.3)
subsetmarkers

newmarker = subset(subsetmarkers, subset = pct.1> 0.7)
newmarker

final = subset(newmarker, subset= cluster =="SPN")
final
```

```{r}
features <- final$gene
DotPlot(putamen_controls, features = features, dot.scale = 8) + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
nonSPN_put_transformed= subset(putamen_neurons, subset = sub_class != "SPN")

```

```{r}
SPN_put_transformed
nonSPN_put_transformed
```


#put SPN
```{r}
final = subset(newmarker, subset= cluster =="SPN")

significant_genes = unique(final$gene)

exp = FetchData(SPN_put_transformed, vars = significant_genes)
exp

final

rownames(final) = final$gene
final$gene = NULL
exp

# Assuming final and exp are your dataframes
logfc_df <- final
counts_df <- exp

logfc_df
counts_df

# Check if all column names in counts_df are in the row names of logfc_df
if (!all(colnames(counts_df) %in% rownames(logfc_df))) {
  stop("Not all genes in counts_df are present in logfc_df")
}

# Reorder logfc_df to match the columns in counts_df
logfc_ordered <- logfc_df[colnames(counts_df), , drop = FALSE]
logfc_ordered

# Check for NA values in logfc_ordered
if (any(is.na(logfc_ordered$avg_log2FC))) {
  stop("There are NA values in the logFC values for the genes")
}

# Ensure logfc_ordered$logfc is a numeric vector
logfc_vector <- as.numeric(logfc_ordered$avg_log2FC)
logfc_vector
 #THIS IS FOR LOG TRANSFORMATINO BUT I DONT LIKE IT
#  custom_log_transform <- function(x, logFC_value) {
#       x * 2^logFC_value
#   }
#   
# # Multiply each column in counts_df by the corresponding logFC value
#  result_df <- sweep(counts_df, 2, logfc_vector, FUN = custom_log_transform)


#THE ACTUAL
 result_df <- sweep(counts_df, 2, logfc_vector, FUN = "*")

# Print the result
print(result_df)

metadata <- SPN_put_transformed@meta.data
# Assume that metadata has a column 'donor' that contains donor information
# If not, modify this part according to your metadata structure
if(!"donor_id" %in% colnames(metadata)) {
  stop("Metadata does not contain 'donor' information. Please check your metadata structure.")
}

# Create unique identifiers for cell column to avoid conflicts
expr_data <- result_df %>%
  rownames_to_column(var = "cell_id")

metadata <- metadata %>%
  rownames_to_column(var = "cell_id")

# Combine expression data with donor information
expr_data_long <- expr_data %>%
  pivot_longer(cols = -cell_id, names_to = "gene", values_to = "expression") %>%
  left_join(metadata %>% select(cell_id, donor_id), by = "cell_id")
# Assuming your data frame is named expr_data_long
expr_data_long
# Count the number of occurrences for each gene per donor

summarized_expr_data <- expr_data_long %>%
  group_by(donor_id, gene) %>%
  summarize(total_expression = sum(expression), .groups = 'drop')
summarized_expr_data
Put_SPN_summary_expr_data = summarized_expr_data
Put_SPN_summary_expr_data
```

#put nonSPN
```{r}
final = subset(newmarker, subset= cluster =="SPN")

significant_genes = unique(final$gene)

exp = FetchData(nonSPN_put_transformed, vars = significant_genes)
exp

final

rownames(final) = final$gene
final$gene = NULL
exp

# Assuming final and exp are your dataframes
logfc_df <- final
counts_df <- exp

logfc_df
counts_df

# Check if all column names in counts_df are in the row names of logfc_df
if (!all(colnames(counts_df) %in% rownames(logfc_df))) {
  stop("Not all genes in counts_df are present in logfc_df")
}

# Reorder logfc_df to match the columns in counts_df
logfc_ordered <- logfc_df[colnames(counts_df), , drop = FALSE]
logfc_ordered

# Check for NA values in logfc_ordered
if (any(is.na(logfc_ordered$avg_log2FC))) {
  stop("There are NA values in the logFC values for the genes")
}

# Ensure logfc_ordered$logfc is a numeric vector
logfc_vector <- as.numeric(logfc_ordered$avg_log2FC)
logfc_vector
 #THIS IS FOR LOG TRANSFORMATINO BUT I DONT LIKE IT
#  custom_log_transform <- function(x, logFC_value) {
#       x * 2^logFC_value
#   }
#   
# # Multiply each column in counts_df by the corresponding logFC value
#  result_df <- sweep(counts_df, 2, logfc_vector, FUN = custom_log_transform)


#THE ACTUAL
 result_df <- sweep(counts_df, 2, logfc_vector, FUN = "*")

# Print the result
print(result_df)

metadata <- nonSPN_put_transformed@meta.data
# Assume that metadata has a column 'donor' that contains donor information
# If not, modify this part according to your metadata structure
if(!"donor_id" %in% colnames(metadata)) {
  stop("Metadata does not contain 'donor' information. Please check your metadata structure.")
}

# Create unique identifiers for cell column to avoid conflicts
expr_data <- result_df %>%
  rownames_to_column(var = "cell_id")

metadata <- metadata %>%
  rownames_to_column(var = "cell_id")

# Combine expression data with donor information
expr_data_long <- expr_data %>%
  pivot_longer(cols = -cell_id, names_to = "gene", values_to = "expression") %>%
  left_join(metadata %>% select(cell_id, donor_id), by = "cell_id")
# Assuming your data frame is named expr_data_long
expr_data_long
# Count the number of occurrences for each gene per donor

summarized_expr_data <- expr_data_long %>%
  group_by(donor_id, gene) %>%
  summarize(total_expression = sum(expression), .groups = 'drop')
summarized_expr_data
Put_nonSPN_summary_expr_data = summarized_expr_data
Put_nonSPN_summary_expr_data
```


```{r}
Put_SPN_summary_expr_data$cell_type = "SPN"
Put_nonSPN_summary_expr_data$cell_type = "non-SPN" 

head(Put_SPN_summary_expr_data)
head(Put_nonSPN_summary_expr_data)
```

```{r}
Put_Transcriptional_DE = rbind(Put_SPN_summary_expr_data, Put_nonSPN_summary_expr_data)
Put_Transcriptional_DE
```

```{r}
donors = unique(Put_Transcriptional_DE$donor_id)

for (donor in donors) {
test = Put_Transcriptional_DE[Put_Transcriptional_DE$donor_id == donor,]
a = plot_overlapping_density_histogram(df = test, 
    hist_col = "total_expression",
    fill_col = "cell_type",
    colors = c("blue", "red"),
    #alpha=0.5,
    breaks=seq(0,1000,10),
    title= paste("Put SPNs vs non-SPNs: ", donor),
    xlab = "Put SPN score",
    fig_filename = NULL)

print(a)
}
```


```{r}
expr_data_long
```

```{r}
Put_Transcriptional_DE$Condition = ifelse(grepl("SCF_21-037CM2|SCF-23-068CM|SCF-22-058CF|SCF-22-054CM", Put_Transcriptional_DE$donor_id), "Control", "XDP")
Put_Transcriptional_DE
```


```{r}
XDP = Put_Transcriptional_DE[Put_Transcriptional_DE$Condition == "XDP",]
plot_overlapping_density_histogram(df = XDP, 
    hist_col = "total_expression",
    fill_col = "cell_type",
    colors = c("blue", "red"),
    #alpha=0.5,
    breaks=seq(0,1000,10),
    title= paste("Put SPNs vs non-SPNs (log transformed): XDP"),
    xlab = "Put SPN score",
    fig_filename = NULL)

Control = Put_Transcriptional_DE[Put_Transcriptional_DE$Condition == "Control",]
plot_overlapping_density_histogram(df = Control, 
    hist_col = "total_expression",
    fill_col = "cell_type",
    colors = c("blue", "red"),
    #alpha=0.5,
    breaks=seq(0,1000,10),
    title= paste("Put SPNs vs non-SPNs (log transformed): Control "),
    xlab = "Put SPN score",
    fig_filename = NULL)
```









```{r}
metadata
metadata_subset = metadata[, c("donor_id", "Condition", "rowname")]
metadata_subset
SPN_df_new = merge(SPN_df, metadata_subset, by.x = "rowname", by.y = "rowname")
SPN_df_new
sct_counts_nonSPN <- FetchData(non_SPN_put_transformed, vars = significant_genes)
sct_counts_nonSPN
# Calculate z-scores for the main dataset using control statistics
z_scores <- sweep(sct_counts_nonSPN, 2, means, "-")
z_scores <- sweep(z_scores, 2, sds, "/")
# Sum z-scores across genes for each cell
sum_z_scores <- rowSums(z_scores)
sum_z_scores
Idents(non_SPN_put_transformed) = "sub_class"
# Create a dataframe with summed z-scores and sample identities
nonSPNdf <- data.frame(Sample = Idents(non_SPN_put_transformed), ZScoreSum = sum_z_scores)
nonSPNdf$rowname <- rownames(nonSPNdf)
nonSPNdf
metadata <- non_SPN_put_transformed@meta.data
metadata$rowname <- rownames(metadata)
metadata
metadata_subset = metadata[, c("donor_id", "Condition", "rowname")]
metadata_subset
nonSPN_df_new = merge(nonSPNdf, metadata_subset, by.x = "rowname", by.y = "rowname")
nonSPN_df_new
nonSPN_df_new$Sample = "non-SPN"
nonSPN_df_new
final_df = rbind(SPN_df_new, nonSPN_df_new)
final_df
final_df$cell_type = final_df$Sample
final_df$Sample = NULL
final_df
library(gridExtra)
options(repr.plot.width=24, repr.plot.height=16)
plots = list()
donors = c("PCMC-16-011", "PCMC-16-012", "SCF-18-003", "SCF-18-004", "SCF-18-006","SCF-19-009", "SCF-19-014", "SCF-19-018",  "SCF-20-023",  "SCF_20-024", "SCF-20-025", "SCF-21-030", "SCF_22-043", "SCF_21-037CM2","SCF-22-054CM","SCF-22-058CF","SCF-23-068CM")
for (donor in donors) {
test = final_df[final_df$donor_id == donor,]
plots[[donor]] = plot_overlapping_density_histogram(df = test,
hist_col = "ZScoreSum",
fill_col = "cell_type",
colors = c("red", "blue"),
#alpha=0.5,
breaks=seq(-60,60,2),
title= paste("Put SPNs vs non-SPNs: ", donor),
xlab = "Put SPN Z-Score",
fig_filename = NULL)
}
layout_matrix <- rbind(
c(1, 2, 3, 4, 5),
c(6, 7, 8, 9, 10),
c(11, 12, 13, NA, NA),
c(14, 15, 16, 17, NA)
)
# Arrange the plots according to the custom layout
grid_plots <- grid.arrange(grobs = plots, layout_matrix = layout_matrix)
# Save the arranged plots to a PNG file
ggsave(filename = "output/Put_ZSCORE_donors.png", plot = grid_plots, width = 30, height = 16)

XDP = final_df[final_df$Condition == "XDP",]
plot_overlapping_density_histogram(df = XDP,
hist_col = "ZScoreSum",
fill_col = "cell_type",
colors = c("red", "blue"),
#alpha=0.5,
breaks=seq(-60,60,2),
title= paste("Put SPNs vs non-SPNs: XDP Combined"),
xlab = "Put SPN Z-Score",
fig_filename = NULL)
Control = final_df[final_df$Condition == "Control",]
plot_overlapping_density_histogram(df = Control,
hist_col = "ZScoreSum",
fill_col = "cell_type",
colors = c("red", "blue"),
#alpha=0.5,
breaks=seq(-60,60,2),
title= paste("Put SPNs vs non-SPNs: Control Combined"),
xlab = "Put SPN Z-Score",
fig_filename = NULL)
SPNs = final_df[final_df$cell_type == "SPN",]
plot_overlapping_density_histogram(df = SPNs,
hist_col = "ZScoreSum",
fill_col = "Condition",
colors = c("red", "blue"),
#alpha=0.5,
breaks=seq(-60,60,2),
title= paste("Put SPNs: XDP vs Control"),
xlab = "Put SPN Z-Score",
fig_filename = NULL)
nonSPNs = final_df[final_df$cell_type == "non-SPN",]
plot_overlapping_density_histogram(df = nonSPNs,
hist_col = "ZScoreSum",
fill_col = "Condition",
colors = c("red", "blue"),
#alpha=0.5,
breaks=seq(-60,60,2),
title= paste("Put non-SPNs: XDP vs Control"),
xlab = "Put SPN Z-Score",
fig_filename = NULL)
SPNs = final_df[final_df$cell_type == "SPN",]
plot_overlapping_density_histogram(df = SPNs,
hist_col = "ZScoreSum",
fill_col = "Condition",
colors = c("blue", "red"),
#alpha=0.5,
breaks=seq(-60,60,2),
title= paste("Put SPNs: XDP vs Control"),
xlab = "Put SPN Z-Score",
fig_filename = NULL)
nonSPNs = final_df[final_df$cell_type == "non-SPN",]
plot_overlapping_density_histogram(df = nonSPNs,
hist_col = "ZScoreSum",
fill_col = "Condition",
colors = c("blue", "red"),
#alpha=0.5,
breaks=seq(-60,60,2),
title= paste("Put non-SPNs: XDP vs Control"),
xlab = "Put SPN Z-Score",
fig_filename = NULL)
# Extract SCT counts for C-minus genes from the control dataset
CaH_SPN_sct_counts_control <- FetchData(SPN_cah_transformed_controls, vars = significant_genes)
CaH_SPN_sct_counts_control
# Calculate mean and standard deviation for each gene in the control dataset
means <- colMeans(sct_counts_control)
sds <- apply(sct_counts_control, 2, sd)
# Extract SCT counts for C-minus genes from the main dataset
sct_counts_SPN <- FetchData(SPN_cah_transformed, vars = significant_genes)
sct_counts_SPN
# Calculate z-scores for the main dataset using control statistics
z_scores <- sweep(sct_counts_SPN, 2, means, "-")
z_scores <- sweep(z_scores, 2, sds, "/")
# Sum z-scores across genes for each cell
sum_z_scores <- rowSums(z_scores)
sum_z_scores
Idents(SPN_cah_transformed) = "sub_class"
# Create a dataframe with summed z-scores and sample identities
SPN_df <- data.frame(Sample = Idents(SPN_cah_transformed), ZScoreSum = sum_z_scores)
SPN_df$rowname <- rownames(SPN_df)
SPN_df
metadata <- SPN_cah_transformed@meta.data
metadata$rowname <- rownames(metadata)
metadata
metadata_subset = metadata[, c("donor_id", "Condition", "rowname")]
metadata_subset
SPN_df_new = merge(SPN_df, metadata_subset, by.x = "rowname", by.y = "rowname")
SPN_df_new
sct_counts_nonSPN <- FetchData(non_SPN_cah_transformed, vars = significant_genes)
sct_counts_nonSPN
# Calculate z-scores for the main dataset using control statistics
z_scores <- sweep(sct_counts_nonSPN, 2, means, "-")
z_scores <- sweep(z_scores, 2, sds, "/")
# Sum z-scores across genes for each cell
sum_z_scores <- rowSums(z_scores)
sum_z_scores
Idents(non_SPN_cah_transformed) = "sub_class"
# Create a dataframe with summed z-scores and sample identities
nonSPNdf <- data.frame(Sample = Idents(non_SPN_cah_transformed), ZScoreSum = sum_z_scores)
nonSPNdf$rowname <- rownames(nonSPNdf)
nonSPNdf
metadata <- non_SPN_cah_transformed@meta.data
metadata$rowname <- rownames(metadata)
metadata
metadata_subset = metadata[, c("donor_id", "Condition", "rowname")]
metadata_subset
nonSPN_df_new = merge(nonSPNdf, metadata_subset, by.x = "rowname", by.y = "rowname")
nonSPN_df_new
nonSPN_df_new$Sample = "non-SPN"
nonSPN_df_new
final_df = rbind(SPN_df_new, nonSPN_df_new)
final_df
final_df$cell_type = final_df$Sample
final_df$Sample = NULL
final_df
plot_overlapping_density_histogram(df = XDP,
hist_col = "ZScoreSum",
fill_col = "cell_type",
colors = c("red", "blue"),
#alpha=0.5,
breaks=seq(-60,60,2),
title= paste("CaH SPNs vs non-SPNs: XDP Combined"),
xlab = "CaH SPN Z-Score",
fig_filename = NULL)
Control = final_df[final_df$Condition == "Control",]
plot_overlapping_density_histogram(df = Control,
hist_col = "ZScoreSum",
fill_col = "cell_type",
colors = c("red", "blue"),
#alpha=0.5,
breaks=seq(-60,60,2),
title= paste("CaH SPNs vs non-SPNs: Control Combined"),
xlab = "CaH SPN Z-Score",
fig_filename = NULL)
XDP = final_df[final_df$Condition == "XDP",]
plot_overlapping_density_histogram(df = XDP,
hist_col = "ZScoreSum",
fill_col = "cell_type",
colors = c("red", "blue"),
#alpha=0.5,
breaks=seq(-60,60,2),
title= paste("CaH SPNs vs non-SPNs: XDP Combined"),
xlab = "CaH SPN Z-Score",
fig_filename = NULL)
Control = final_df[final_df$Condition == "Control",]
plot_overlapping_density_histogram(df = Control,
hist_col = "ZScoreSum",
fill_col = "cell_type",
colors = c("red", "blue"),
#alpha=0.5,
breaks=seq(-60,60,2),
title= paste("CaH SPNs vs non-SPNs: Control Combined"),
xlab = "CaH SPN Z-Score",
fig_filename = NULL)
final =subset(CaH_SPN_markers_from_SPN1, subset = pct.1 > 0.5)
CaH_SPN_markers_from_SPN = subset(caudate_controls_markers, subset= cluster =="SPN")
Idents(caudate_controls) = caudate_controls$sub_class

```



#attempt to merge caudate and putamen but doesn;t look right
```{r}
caudate_neurons <- AddMetaData(caudate_neurons, metadata = "Caudate", col.name = "Region")
putamen_neurons <- AddMetaData(putamen_neurons, metadata = "Putamen", col.name = "Region")
caudate_neurons@meta.data
putamen_neurons@meta.data
```

```{r}
list = c(caudate_neurons, putamen_neurons)
Combined_CaH_Put_neurons = mergeSeuratListWithMetadata(list)

Combined_CaH_Put_neurons
Combined_CaH_Put_neurons@meta.data

```

```{r}
# Counts of super lowly expressed genes are essentially random
# Also, we don't really care about housekeeping genes that have almost the same expression in every cell
# Before clustering, we want to find "highly variable genes" (hvgs) that tend to be both highly expressed and vary a lot from cell to cell.
# we do this by finding hvgs in each participant individually, and collating the genes
# that are highly variable in at least half of donors.


# this code should take 3-10 minutes
hvgs = getSeuratVarFeatureIntersectByCol(Combined_CaH_Put_neurons, subset_col="donor_id", original_nfeatures=2500)


n_dims_use=20


# now cluster
# the "%>%" lines are a little piece of syntax from the dyplr library,
# short for "apply the following transformation to the object"
# doing this line-by-line allows us to apply a sequence of transformations
Combined_CaH_Put_neurons = (Combined_CaH_Put_neurons
   %>% NormalizeData() # log normalizes raw counts
   %>% ScaleData(features=hvgs, split.by="donor_id") # within each cell, for each gene scale the data (i.e. subtract the donor's mean and divide by the donor's standard deviation of that gene)
   %>% RunPCA(features=hvgs, npcs=n_dims_use) # Reduce the dimensions to the n_dims_use dimensions that best explain the data https://en.wikipedia.org/wiki/Principal_component_analysis
   %>% FindNeighbors(dims = 1:n_dims_use) # Finds every cells closest neighbors in the PCA space
   %>% FindClusters(resolution = 0.2) # finds clusters at a variety of resolutions
   %>% FindClusters(resolution = 0.3) # after clustering, the cluster labels are accessible via BICAN_3P_V8_CaH$RNA_snn_res.{resolution}
   %>% FindClusters(resolution = 0.4) # e.g.: BICAN_3P_V8_CaH$RNA_snn_res.0.4 here
   %>% FindClusters(resolution = 0.5)
   %>% FindClusters(resolution = 0.6)
   %>% FindClusters(resolution = 0.7)
   %>% FindClusters(resolution = 0.8)
   %>% FindClusters(resolution = 0.9)
   %>% FindClusters(resolution = 1)
   %>% RunUMAP(dims = 1:n_dims_use) # projects the PCA onto 2 dimensions, such that both "global structure" and "local structure" are optimally preserved
)


# Now we can plot the data and just see what it looks like


setwh = function(w, h){
   # set the width of the ensuing figures to w, height to h
   options(repr.plot.width = w, repr.plot.height = h)
}
```

```{r}
DimPlot(Combined_CaH_Put_neurons, group.by = "library", label=T) # are there library batch effects?
DimPlot(Combined_CaH_Put_neurons, group.by = "donor_id", label=T) # are there donor batch effects?
DimPlot(Combined_CaH_Put_neurons, group.by = "cell_class") # change col name to whatever your case control metadata column is
DimPlot(Combined_CaH_Put_neurons, group.by = "RNA_snn_res.0.5", label=T) # make a few of these for different resolutions
DimPlot(Combined_CaH_Put_neurons, group.by = "sub_class", label=T) # make a few of these for different resolutions


#qsave(filtered_merged_caudate, "somewhere/cozy")

```

```{r}
library(harmony)
Combined_CaH_Put_neurons = (Combined_CaH_Put_neurons
    %>% RunHarmony(
        group.by = "donor_id")
    %>% FindNeighbors(reduction='harmony', dims=1:20)
    %>% FindClusters(res=0.2)
    %>% FindClusters(res=0.3)
    %>% FindClusters(res=0.4)
    %>% FindClusters(res=0.5)
    %>% RunUMAP(reduction="harmony", dims=1:20)
)

DimPlot(Combined_CaH_Put_neurons, group.by = "library", label=T) # are there library batch effects?
DimPlot(Combined_CaH_Put_neurons, group.by = "donor_id", label=T) # are there donor batch effects?
DimPlot(Combined_CaH_Put_neurons, group.by = "cell_class") # change col name to whatever your case control metadata column is
DimPlot(Combined_CaH_Put_neurons, group.by = "RNA_snn_res.0.5", label=T) # make a few of these for different resolutions
DimPlot(Combined_CaH_Put_neurons, group.by = "sub_class", label=T) # make a few of these for different resolutions
DimPlot(Combined_CaH_Put_neurons, group.by = "Condition", label=T) # make a few of these for different resolutions
```
